### 2018-10-29 Paper/sync.tex
Updates from ShareLaTeX

\pdfoutput=1
\documentclass[twocolumn]{aastex61}
%%\documentclass[]{emulateapj}

%Accepted/received/... %%

\received{xxx}
\revised{yyy}
\accepted{zzz}

%% Command to document which AAS Journal the manuscript was submitted to.
\submitjournal{AAS Journals}

%% Short title/authors

\shorttitle{Reproducing Kepler Stellar Rotation Periods}
\shortauthors{Fleming et al.}

%% Begin document, title, packages %%
\usepackage{hyperref}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[caption=false]{subfig}

%% Custom commands
\def\mearth{{\rm\,M_\oplus}}
\def\rearth{{\rm\,R_\oplus}}
\def\msun{{\rm\,M_\odot}}
\def\rsun{{\rm\,R_\odot}}
\def\lsun{{\rm\,L_\odot}}
\def\gsim{~\rlap{$>$}{\lower 1.0ex\hbox{$\sim$}}}
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}

\newcommand{\vplanet}[0]{\texttt{VPLANET}\xspace}
\newcommand{\eqtide}[0]{\texttt{EQTIDE}\xspace}
\newcommand{\stellar}[0]{\texttt{STELLAR}\xspace}

%% Begin doc %%

\begin{document}

\title{Reproducing Stellar Rotation Periods in the Kepler Field via Magnetic Braking and Tidal Torques}

%% AUTHORS %%

%%\correspondingauthor{David P. Fleming}
%%\email{dflemin3@uw.edu}

%%\author[0000-0001-9293-4043]{David P. Fleming}
\author{David P. Fleming}
\affil{Astronomy Department, University of Washington \\
Box 951580, Seattle, WA 98195}

\author{Rory Barnes}
\affiliation{Astronomy Department, University of Washington \\
Box 951580, Seattle, WA 98195}

\author{James R. A. Davenport}
\affiliation{Astronomy Department, University of Washington \\
Box 951580, Seattle, WA 98195}

\author{Rodrigo Luger}
\affiliation{Center for Computational Astrophysics, Flatiron Institute \\
New York, NY 10010}

%% ABSTRACT %%

\begin{abstract}

Stellar rotation period (P$_{rot}$) studies have produced numerous insights into the angular momentum evolution of low-mass stars, but previous investigations have focused on single stars even though roughly half of Sun-like stars are in stellar binaries. We examine the impact of unresolved stellar binaries on the stellar P$_{rot}$ distribution of the \textit{Kepler} field by performing Monte Carlo simulations of a population composed of single and double stars by modeling stellar evolution, magnetic braking, and tidal forces. We find that tides modify stellar rotations in all binaries with orbital periods up to 70 days.  At short orbital periods, most stellar binaries tidally-lock into synchronous rotation, naturally explaining the observed population of fast rotators in the \textit{Kepler} field that cannot be reproduced by single-star-only models.  Many binaries with longer orbital periods, some up to orbital periods of 70 days, tidally-lock, synchronizing the stellar P$_{rot}$ with the binary orbital period, causing P$_{rot}$ to not be a valid proxy for age in all cases, i.e. gyrochronology methods must be applied carefully. Our assumed flat stellar age distribution over $1-4$ Gyr underpredicts the number of fast rotators in the \textit{Kepler} field. If instead we assume a bimodal stellar age distribution in the \textit{Kepler} field composed of a young ($100$ Myr$-1$ Gyr old) and an old ($1-4$ Gyr) population, we successfully reproduce the \textit{Kepler} P$_{rot}$ distribution.

\end{abstract}

%% KEYWORDS %%

\keywords{binaries: close}

%% INTRO %%

\section{Introduction} \label{sec:intro}

Measuring stellar rotation periods via star spot modulation is a critical tool for understanding stellar angular momentum evolution in low-mass stars (M$\lsim 1.2$ M$_{\odot}$). As stars evolve, angular momentum conservation dictates that their evolving moments of inertia necessarily change their P$_{rot}$, e.g. stars spin-up as their radii contract, whereas magnetic braking, the coupling of the corotating stellar wind with the stellar magnetic field, removes angular momentum from the star, causing the observed long-term spin-down of stars \citep{Parker1958,Skumanich1972}.

Extensive measurements of ${>}10,000$ stellar rotation periods in the \textit{Kepler} field \citep{Borucki2010,McQuillan2013,Reinhold2013,McQuillan2014,Reinhold2015,Lurie2017} have provided exquisite observational constraints for theorists to model long-term stellar angular momentum evolution. Various models have been developed to explain the observed P$_{rot}$ distribution in the \textit{Kepler} field \citep[e.g.][]{Matt2015,vanSaders2018} and to link the long-term spin-down of low-mass stars to their ages, a method known as gyrochronology \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}. These models, however, have failed to reproduce the large observed population of rapid rotators (P$_{rot} {\lsim} 7.5$ days) across all stellar masses in the \textit{Kepler} field \citep{McQuillan2013,McQuillan2014}.  Both gyrochonology relations \citep{McQuillan2014} and a modern magnetic braking model \citep{Matt2015} suggest that these fast rotators could be at most 1 Gyr old, in tension with the expected older ages of \textit{Kepler} field stars \citep{Chaplin2014,Angus2015}.

The aforementioned models have two limitations.  First, these models assumed stellar populations composed solely of single stars, whereas roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}. Binaries are difficult to detect in photometric surveys because of low transit probabilities, so typically they are isolated using magnitude cuts \citep[e.g.][]{Davenport2018,Simonian2018}. However, 1000s of eclipsing binaries are known in the \textit{Kepler} field \citep{Kirk2016}, with likely many more unresolved binaries present in the data. In short-period binaries, tidal torques modify stellar rotations by driving the binary towards the tidally-locked state \citep{Zahn1989,Fleming2018}. Tidal forces in binaries are commonly understood to act in binaries with orbital periods (P$_{orb}$) ${\lsim} 10$ days \citep[e.g.][]{Zahn1989,Meibom2005}, however \citet{Lurie2017} measured 100s of \textit{Kepler} eclipsing binary stellar rotation periods using starspot variability and found that tides can synchronize stellar binary rotation periods up to P$_{orb} \approx 45$ days.  Recently \citet{Simonian2018} found that most rapid rotators in the \textit{Kepler} field are likely non-eclipsing, tidally-synchronized short-period binaries. Clearly, unresolved stellar binaries impart a significant signal in P$_{rot}$ distributions, particularly in the \textit{Kepler} field at short P$_{orb}$, but can also influence P$_{rot}$ measurements at larger binary P$_{orb}$.

Second, no previous study has simulated how a bimodal age distribution in the \textit{Kepler} field, suggested by \citet{McQuillan2014}, \citet{Matt2015}, and \citet{Davenport2017}, could explain the observed population of rapid rotators.  If the age distribution is in fact bimodal, a young, fast rotating population could explain the observed P$_{rot}$ distribution.  Here, we perform the first forward modeling of the angular momentum evolution of a realistic stellar population of single and double stars, accounting for stellar evolution, magnetic braking, tidal torques in stellar binaries, and testing the impact of a bimodal age distribution to reproduce the observed P$_{rot}$ distribution in the \textit{Kepler} field.

%Here, we examine the role of unresolved, tidally-interacting stellar binaries in the \textit{Kepler} field using a stellar evolution model, with magnetic braking, coupled with a secular equilibrium tidal model for stellar binaries. to see if their presence can explain features in the \textit{Kepler} P$_{rot}$ distribution. In $\S$~\ref{sec:methods} and $\S$~\ref{sec:simulations}, we describe our computational methods and simulation set-up, and discuss our results and their implications in $\S$~\ref{sec:results} and $\S$~\ref{sec:discussion}, respectively.

%% Maybe just focus on Kepler Prot observations?
% in clusters \citep[e.g. Praesepe;][]{Scholz2007,Scholz2011,Agueros2011,Delorme2011,Kovacs2014,Douglas2017} and extensive P$_{rot}$ measurements

%% SECTION : Methods %%

\section{Methods} \label{sec:methods}

\subsection{Stellar Evolution} \label{sec:methods:stellar}

We simulate stellar rotation evolution for both single stars and stellar binaries using the coupled stellar-tidal evolution model presented in \citet{Fleming2018} and implemented in the open-source code VPLanet\footnote{VPLanet is publically available
at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep[][Barnes et al., in prep]{Barnes2016}. We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models. We model stellar spin-down due to magnetic braking using the model derived in \citet{Matt2015} since it has been shown to successfully reproduce many features in the P$_{rot}$ distribution of the \textit{Kepler} field, notably the upper and lower envelopes of the latter distribution and the observed ``dip" in the distribution around M${\approx}0.6$ M$_{\odot}$. We integrate the following equation to model the net change in the stellar rotation rate due to stellar evolution and magnetic braking:

\begin{equation} \label{eqn:stellar_rot_rate_dt}
\dot{\omega} = \frac{\dot{J}_{mb}}{I} - \frac{2 \dot{R} \omega}{R} - \frac{2 \dot{r_g} \omega}{r_g}
\end{equation}
where the moment of inertia $I = M r_g^2 R^2$ and $\dot{J}_{mb}$ is the angular momentum loss due to magnetic braking.

\subsection{Tidal Evolution} \label{sec:methods:eqtide}

For stellar binaries, we assume that they tidally interact via the ``Constant Phase Lag" (CPL) version \citep[][]{FerrazMello2008,Heller2011} of the commonly used equilibrium tidal model \citep{Darwin1880}.  The CPL model tracks the secular evolution of the binary's semi-major axis, $a$, eccentricity, $e$, and stellar rotation rates, $\omega$, due to tidal torques.  The CPL model assumes that the tidal torque on one star due to its companion arises from a linear combination of uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass and is analogous to a driven, damped harmonic oscillator \citep{Greenberg2009}. We use the \eqtide implementation of the CPL model in VPLanet described in \citet{Fleming2018}, their equations 8-16, and simultaneously integrate those equations with Eq.~(\ref{eqn:stellar_rot_rate_dt}) for stellar binaries.  All equations are integrated using a $4^{th}$ order Runge-Kutta scheme.

%Following \citet{Fleming2018}, when one or both binary stars are tidally locked, we assume that tidal forces prevent magnetic braking from spinning down the tidally locked star(s), but instead, any angular momentum lost is lost at the expense of the binary orbit, decreasing the orbital semi-major axis as a result.  Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify Eqs. (18) and (20) from \citet{Fleming2018} to account for $r_g$ evolution for the cases when one or both stars tidally lock under conservation of angular momentum, respectively:
%\small
%\begin{equation} \label{eqn:tidal_locked_one}
%\begin{split}
%\dot{a}_{coupled}^{(1)} = \frac{-\dot{J_{mb}} - 2 \omega \left( m_1 r_{g,1}^2 R_1 \dot{R_1} - m_1 r_{g,1} \dot{r}_{g,1} R_1^2 \right)}
%{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} m_1 r_{g,1}^2 R_1^2}
%\end{split}
%\end{equation}
%\normalsize
% and
% \small
%\begin{equation} \label{eqn:tidal_locked_two}
%\begin{split}
%\dot{a}_{coupled}^{(2)} = \frac{-\dot{J_{mb}} - 2 \omega \left( \sum_{i=1}^{2} m_i r_{g,i}^2 R_i \dot{R_i} + m_i r_{g,i} \dot{r}_{g,i} R_i^2 \right)}
%{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} \left( m_1 r_{g,1}^2 R_1^2 + m_2 r_{g,2}^2 R_2^2 \right)}
%\end{split}
%\end{equation}
%\normalsize
%where $J_{orb}$ is the orbital angular momentum.

%% SECTION: Simulations + Initial conditions %%

\subsection{Simulations} \label{sec:simulations}

We simulate the P$_{rot}$ distribution of the \textit{Kepler} field by performing Monte Carlo study consisting of a population of 10,000 stellar systems composed of single stars and binaries using VPLanet. For each simulation, we sample the stellar mass from a Gaussian fit to the stellar mass distribution in \citet{McQuillan2014} with $\mu=0.885$ M$_{\odot}$ and $\sigma=0.201$ M$_{\odot}$ to approximate the \textit{Kepler} mission's intended selection bias of monitoring FGK stars for Earth-like planets. We restrict stellar masses to be within [$0.1,1.2$] M$_{\odot}$ to limit the influence of subgiants since more massive stars evolve off the main sequence after $\gsim 4$ Gyr, our age limit, and because more massive stars do not experience magnetic braking due to their shallow convective envelopes \citep{Kraft1967}. Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of stars in the ${\sim}2$ Myr old ONC.  We randomly assign stars a stellar companion according to stellar multiplicity fractions from \citet{Duchene2013}. In binaries, we compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010}.We uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$] and stellar tidal Qs are sampled from a log-uniform distribution over $[10^5,10^8]$. For the initial binary $P_{orb}$, we test two models. In Model 1, $10\%$ of binaries have their initial P$_{orb}$ uniformly sampled over [$3,80$] days, consistent with P$_{orb}$ distributions of field stars \citep[e.g.][]{Raghavan2010}, and the remaining $90\%$ have P$_{orb} > 80$ d and are considered ``single" stars since for P$_{orb} > 80$ d, tidal torques are sufficiently weak such that the stars' evolution is effectively independent of their companion.  In Model 2, we consider an enhanced population of short period binaries by sampling all initial P$_{orb}$ uniformly over [$3,80$] days.

In Models 1 and 2, we simulate each system to an age uniformly sampled over [$1,4$] Gyr, the approximate age range of \textit{Kepler} field stars \citep{McQuillan2014,Angus2015}, with the upper-limit suggested by \citet{McQuillan2014} and \citet{Matt2015}. We also consider a ``Two Population" model that uses the initial conditions of Model 2 and assumes that $40\%$ of stars comprise a ``young" population with ages uniformly sampled over [$0.1,1$] Gyr, while the remainder comprise the ``old" population with ages uniformly sampled over [$1,4$] Gyr. We track the P$_{rot}$ evolution for each single star and the primary, i.e. more massive star, in binaries as it is observationally easier to measure a P$_{rot}$ due to starspot modulations on the more massive, and hence brighter, star. We compare our simulation results with the \textit{Kepler} P$_{rot}$ distribution measured by \citet{McQuillan2014}.

All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/binGyro}{https://github.com/dflemin3/binGyro}.} with accompanying descriptions and documentation.

%  For simulations of binaries, given the inherent uncertainty in and complexity of the formation of binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration.

%Our goal is to determine how the distribution of $P_{rot}$ for a population stars composed solely of single stars differs, if at all, from a population of both single stars and binaries.  If there is a substantial difference between the two distributions, e.g. tidally-locked fast rotators, then any study examining the distribution of stellar $P_{rot}$ should therefore account for the impact of tidally-interacting stellar binaries.  We approach this problem with a ...

%% SECTION: VPLANET RESULTS %%

\section{Results} \label{sec:results}

\subsection{Tidal Locking in Stellar Binaries} \label{sec:results:locking}

Here we examine the tidal locking in binaries using the simulations from Model 2 (see $\S$~\ref{sec:results:kepler2}) by plotting P$_{rot}$ as a function of mass and age for binaries in Fig.~\ref{fig:locking}. We find that $35.5\%$ of binaries tidally lock before the age of the system, typically for tidal $Q \lsim 10^7$. Tidally-locked binaries tend to rotate more rapidly than unlocked binaries, reflected in the increased density for P$_{rot} \lsim 20$ d in the marginal distribution for the locked population and in the distribution medians (locked median P$_{rot} = 19.8$ d compared to unlocked median P$_{rot} = 27.5$ d). Binaries readily tidally-lock at short P$_{orb}$ due to the tidal force's strong semi-major axis dependence, increasing the strength of tidal torques at shorter orbital separations. When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} = $P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.

Tidal forces modify the spins of unlocked binaries as well, tending to spin-up stars against the effects of magnetic braking towards orbital synchronization, reflected in unlocked binaries' median P$_{rot} = 27.5$ d, compared to the single star median P$_{rot} = 30.6$ d. Tidal locking is not limited to P$_{orb} \lsim 10$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 70$ d, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars (compare to Fig.~\ref{fig:kepler2}). This behavior is consistent with observations of P$_{rot}$ in \textit{Kepler} eclipsing binaries by \citet{Lurie2017} who find that binaries can tidally lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d. The observed age distribution of tidally-locked binaries has little dependence on P$_{rot}$, with both young and old systems tidally-locking at long and short P$_{rot}$, causing the predictions of gyrochronology models to fail in such systems.

\begin{figure}[h]
	\includegraphics[width=0.45\textwidth]{../Plots/locked_kepler.pdf}
   \caption{Simulated distribution of P$_{rot}$ for tidally-locked (cross) and unlocked (circle) stellar binaries, colored by age for Model 2 (see $\S$~\ref{sec:results:kepler2}). Left: P$_{rot}$ as a function of stellar mass.  Right: Marginalized P$_{rot}$ distribution for locked (blue) and unlocked (orange) binaries.  Tidally-locked binaries rotate more rapidly than unlocked binaries, displaying an increased density in the P$_{rot}$ distribution for P$_{rot} {\lsim} 20$ d, and with a median P$_{rot} = 19.8$ d, compared to the unlocked median P$_{rot} = 27.5$ d. Unlocked binaries on average rotate more quickly than single stars (compare to single distribution in Fig.~\ref{fig:kepler2}) due to the influence of tidal torques.}
    \label{fig:locking}
\end{figure}

\subsection{Comparison to \textit{Kepler}}

\subsubsection{Model 1} \label{sec:results:kepler1}

In Fig.~\ref{fig:kepler1}, we display our simulated P$_{rot}$ for Model 1, compared to the observed \textit{Kepler} distribution \citep{McQuillan2014}. Single stars, defined as stars without a companion, or with a companion with a P$_{orb} > 80$ d where tides do not modify the orbit or stellar rotations, rotate more slowly with age due to magnetic braking, with more massive stars rotating more quickly than low mass stars, consistent with \textit{Kepler} P$_{rot}$ observations.  There is a larger spread in P$_{rot}$ for stars with M$\lsim 1$ M$_{\odot}$ with the spread increasing with decreasing stellar mass, due to the mass-dependence magnetic braking torque in the \citet{Matt2015} model.
%in which lower mass stars tend to spin down more quickly than higher-mass stars once they enter the "unsaturated" regime (see \citet{Matt2015} equations 6 and 7).

The inclusion stellar binaries naturally produces a population of fast rotators (P$_{rot} < 7.5$ d) across all masses as binaries readily tidally-lock at shorter P$_{orb}$, with ${\sim}90\%$ of stars in this P$_{rot}$ regime being tidally-locked binaries. Single-star only models fail to reproduce this feature.  The fast rotator population only comprises $0.25\%$ of all simulated stars, compared to $16\%$ of \textit{Kepler} stars in the same regime, and is therefore insufficient to explain the observed population of fast rotators. Our simulated P$_{rot}$ distribution is systematically biased towards longer P$_{rot}$ than the \textit{Kepler} distribution with a total median P$_{rot} = 30.3$ d compared to \textit{Kepler} median P$_{rot} = 17.3$ d. The inclusion of binaries does little to shift the P$_{rot}$ distribution towards shorter P$_{rot}$ as Model 1's assumed binary P$_{orb}$ distribution, based on observations of field stars \citep{Raghavan2010}, predicts that only ${\sim}4\%$ of all stars will be in stellar binaries with P$_{orb} < 80$ d, short enough for tides to appreciably modify stellar rotations.

\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{../Plots/kepler_box_dist_corr.pdf}
   \caption{Simulated distribution of P$_{rot}$ for single (circle) and double (cross, defined as stellar binaries with P$_{orb} < 80$ d) stars for Model 1. Left: P$_{rot}$ as a function of stellar mass, colored by age.  Center: Marginalized P$_{rot}$ distributions for single stars (orange), binary stars (green), and the total sample (blue), and \textit{Kepler} \citep{McQuillan2014} (red). Except for a few low-mass (${\lsim}0.3$M$_{\odot}$) single stars, regardless of age, rotate more slowly (single median P$_{rot} = 30.5$ d) than \textit{Kepler} stars (\textit{Kepler} median P$_{rot} = 17.3$ d) and fail to explain the observed population of fast rotators.  Binaries, however, readily tidally-lock at shorter P$_{orb}$, naturally producing a population of fast rotators (binary median P$_{rot} = 25.9$ d). Right: box-and-whisker plot of the P$_{rot}$ distribution for the simulations seen in the left panel in 0.1 M$_{\odot}$ bins. The shaded regions cover the distributions' interquartile range, the white lines indicate the medians, and the whiskers do not indicate error bars, but rather the spread of the distribution, extending to the 5th and 95th percentiles. The numbers over the boxes indicate the percent difference between the single (orange) and total (blue) distribution medians with the \citet{McQuillan2014} (red) sample. Even with the inclusion of binaries that tidally-lock at short P$_{rot}$, Model 1 produces a more slowly rotating stellar population than is observed in the \textit{Kepler} field, indicating that our assumed stellar population is incorrect.}%
    \label{fig:kepler1}%
\end{figure*}

\subsubsection{Model 2: Enhanced Binary Population} \label{sec:results:kepler2}

In Model 2, we attempt to increase the number of fast rotators by increasing the number of systems that could tidally-lock at short P$_{rot}$.  We assume all stellar binaries have a P$_{orb}$ uniformly sampled over [$3,80$] d, as opposed to just $10\%$ of binaries in Model 1.  Although optimistic, our assumption is reasonable as an analysis of fast rotators in the \textit{Kepler} field by \citet{Simonian2018} found that $\sim 59\%$ of stars with $1.5 <$ P$_{rot} < 7$ d are likely unresolved, tidally-locked binaries, suggesting that a substantial short-period rapidly rotating binary population does exist in the \textit{Kepler} field.

As in Model 1, stellar binaries tidally-lock at short P$_{rot}$ producing a population of fast rotators below the single star sequence with binaries comprising ${\sim}99\%$ of fast rotators.  The inclusion of additional short P$_{orb}$ binaries noticeably modifies the simulated P$_{rot}$ distribution, shifting the median P$_{rot}$ for the total sample down to $28.6$ d, compared to $30.6$ d for the single-star-only distribution.  The impact of binaries is evident across all masses in the right panel of Fig.~\ref{fig:kepler2} as the total P$_{rot}$ distribution, which includes binaries, better matches the observed \textit{Kepler} distribution than the single star only distribution, due to the fast rotating tidally-locked binary population.  The enhanced binary population of Model 2 still predicts a more slowly rotating stellar P$_{rot}$ distribution that is observed, specifically in the fast rotator regime, with Model 2 predicting $3.4\%$ of our simulated stars are fast rotators, compared to $16\%$ of \textit{Kepler} stars. As noted in $\S$~\ref{sec:results:locking}, tidally-locked binaries fix P$_{rot} = $ P$_{orb}$ as tidal torques overpower the torque due to magnetic braking, causing P$_{rot}$ in binaries to not correlate with age.

\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{../Plots/kepler_box_dist.pdf}
   \caption{Identical format to Fig.~\ref{fig:kepler1}, but displaying simulated data from Model 2 (see $\S$~\ref{sec:results:kepler2}). Left: Single stars, regardless of age, rotate too slowly (single median P$_{rot} = 30.6$ d) to explain the population of fast rotators in the \textit{Kepler} field (\textit{Kepler} median P$_{rot} = 17.3$ d).  The enhanced population of short P$_{orb}$ binaries, however, naturally produces a significant population of tidally-locked binaries at P$_{rot} \lsim 20$ d (binary median P$_{rot} = 25.4$ d) that is reflected in the observed total P$_{rot}$ distribution (total median P$_{rot} = 28.6$ d). Right: For most masses, the total distribution, which includes binaries, can better reproduce the \textit{Kepler} medians and fast rotator population compared to the single-star only distribution.  Even with the addition of a large population of short-period binaries, Model 2 still underpredicts the number of fast rotators across all stellar masses.}%
    \label{fig:kepler2}%
\end{figure*}

\subsubsection{Two Population Model} \label{sec:results:kepler_two_pop}

Observations \citep{McQuillan2014,Davenport2017} and theoretical work \citep{Matt2015} suggest that the \textit{Kepler} field contains a substantial population of young stars with ages $< 1$ Gyr.  Given the inability of Models 1 and 2 to reproduce the \textit{Kepler} P$_{rot}$ distribution, especially at short P$_{rot}$, we test the hypothesize of a bimodal \textit{Kepler} stellar age distribution. We use the enhanced binary population of Model 2 and assume that $40\%$ of stars comprise the ``young" population, with ages uniformly sampled over [$0.1,1$] Gyr, while the remainder comprise the ``old" population, with ages uniformly sampled over [$1,4$] Gyr.  We display the results of these simulations in Fig.~\ref{fig:kepler_two_pop}.

We find that the inclusion of the young population substantially improves our simulated P$_{rot}$ distribution's fit to the \textit{Kepler} distribution with the total median P$_{rot} = 19.6$ d, a close match to the median P$_{rot} = 17.3$ d in the \textit{Kepler} field. In the right panel of Fig.~\ref{fig:kepler_two_pop} for $ 0.4 \textrm{M}_{\odot} \lsim \textrm{M} \lsim 1.2 \textrm{M}_{\odot}$, the inclusion of both a young population and tidally-interacting binaries in the total (blue) sample leads to an excellent match with the \textit{Kepler} distribution (red) compared to the single-star only (orange) model, emphasizing the importance of including stellar binaries in P$_{rot}$ forward models. The combination of a fast-rotating young stellar population and enhanced binary population produces a population of fast rotators (P$_{rot} < 7.5$ d) comprising $15.5\%$ of simulated stellar systems, remarkably similar to the \textit{Kepler} field's value of $16\%$.  Moreover, we find that $52\%$ of fast rotators are in stellar binaries, in excellent agreement with the suggested value of $59\%$ for likely \textit{Kepler} photometric binaries estimated by \citet{Simonian2018}.  Our simulations suggest that not only is there likely a substantial population of young stars in the \textit{Kepler} field comprising $\sim 40\%$ of all stars, but also that a substantial fractions of fast rotators in the \textit{Kepler} field are tidally-interacting binaries. Our simulations still predict a more slowly-rotating population that what is observed, suggesting that our simple bimodal age distribution should take on a different form, e.g. two Gaussians, or that stellar angular momentum evolution might proceed differently for older stars as proposed by \citet[e.g.][]{vanSaders2016,vanSaders2018}.

\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{../Plots/kepler_two_pop_box_dist.pdf}
   \caption{Identical format to Fig.~\ref{fig:kepler1}, but displaying simulated data from our two population model (see $\S$~\ref{sec:results:kepler_two_pop}). Our two population model reproduces the observed \textit{Kepler} P$_{rot}$ distribution, especially the fast rotator population, with a total median P$_{rot} = 19.6$ d compared to the \textit{Kepler} median P$_{rot} = 17.3$ d, a close match. In the right panel for $ 0.4 \textrm{M}_{\odot} \lsim \textrm{M} \lsim 1.2 \textrm{M}_{\odot}$, the inclusion of tidally-interacting binaries in the total (blue) sample leads to an excellent match with the \textit{Kepler} distribution (red) compared to the single-star only (orange) model, emphasizing the importance of including stellar binaries in P$_{rot}$ forward models.}%
    \label{fig:kepler_two_pop}%
\end{figure*}

%% Old stuff %%

% Stellar evolution figure caption
% Stellar evolution for 1 M$_{\odot}$ (blue), 0.7 M$_{\odot}$, (orange), and 0.2 M$_{\odot}$ (green) stars according to our interpolation of the \citet{Baraffe2015} stellar models, magnetic braking, and assuming conservation of angular momentum according to \stellar.  We display the radius (left panel), $r_g$ (middle panel), and rotation rate (right panel) evolution for each star using two magnetic braking models from \citet{Reiners2012} (solid curve) and \citet{Repetto2014} (dashed curve).  The dot on each curve denotes the approximate time that each star reaches the main sequence.  At early times, radius contract and $r_g$ decreasing spins up the stars, while at late times, magnetic braking efficiently spins down stars.

%\begin{figure*}[ht]
%	\includegraphics[width=\textwidth]{../Plots/Qrot.pdf}
%   \caption{Rotation period as a function of time for a solar mass star (No Tides, black dot-dashed curve) and solar twin binaries with initial orbital periods of 20 days (solid curves) and 50 days (dashed curves).  The dotted lines indicate rotation periods of 20 and 50 days, respectively.  The different colors of the binaries' curves denote the assumed tidal Qs, which range from $10^4$ to $10^7$.  For the first ${\sim}5 \times 10^7$ years, stellar radius contraction and $r_g$ evolution on the pre-main sequence spin up the stars.  Once the stars reach the main sequence, magnetic braking, and tides in the binary systems, slow the rotation rates.  In the long term, stellar rotation rates in binaries with lower tidal Qs slow more rapidly, and eventually all binary stars tidally lock into synchronous rotation, whereas the single star in the No Tides continues to spin down under the influence of magnetic braking.  Once the stellar binaries tidally lock, their spins undergo little evolution, with most slightly speeding up as binary orbital decay, caused by magnetic braking, slowly decreases their orbital periods.}
%    \label{fig:Qrot}
%\end{figure*}

%\subsection{P$_{rot}$ Distribution Over Time} \label{sec:results:prot_ages}

% XXX put stats discussion in results, reference them here to shorten caption
%\begin{figure}[h]
%	\includegraphics[width=0.4\textwidth]{../Plots/kepler_prot_dist_ensemble.pdf}
%  \caption{Simulated P$_{rot}$ distributions for synthetic clusters ages of 100, 300, 600 Myr, 1 Gyr, and the final age of each system, 1-4 Gyr.  For each cluster age, we gauge the impact of binarity on the P$_{rot}$ distribution by performing the two-sample Kolmogorov–Smirnov test, testing the null hypothesis that the total (blue) and single (orange) sample are drawn from the same parent distribution for a significane level $\alpha = 0.005$, or an effective $\alpha = 0.001$ after applying the Bonferroni correction \citep{Dunn1961} for testing multiple hypotheses. At 100 Myr, many stars in the sample are still contracting on the pre-main sequence, and hence the P$_{rot}$ distributions are shifted towards faster P$_{rot}$ compared to the initial distribution (black). As the stars age, the P$_{rot}$ distributions shift towards slower P$_{rot}$ due to magnetic braking.  For clusters older than 600 Myr, we can reject the null hypothesis that the single and total distributions are drawn from the same parent distribution, suggesting that the inclusion of binaries modifies the observed P$_{rot}$ distribution. This feature can be seen in the 1 Gyr old cluster as an enhancement of fast rotating stellar binaries, a signature of tidal-locking. For older field stars, the simulated distributions the impact of binaries is significant, reflected by a population of fast rotators not seen in the single population, and causes the total distribution to better reflect the \textit{Kepler} distribution.}
%    \label{fig:kepler_dist_ensemble}
%\end{figure}

%% SECTION: DISCUSSION %%

\section{Discussion} \label{sec:discussion}

For the first time, we model the angular momentum evolution of a realistic population of single and double stars, accounting for stellar evolution, magnetic braking, and tidal evolution in binaries, in order to reproduce the P$_{rot}$ distribution of \textit{Kepler} field stars.  We show that tidal forces drive stellar binaries to the tidally-locked state, naturally produce a population of fast rotators that single star models fail to produce.  We find that binaries can tidally-lock up to $P_{orb} \approx 70$ d, causing P$_{rot}$ in binaries to not be a valid proxy for age.  We show that reproducing the \textit{Kepler} P$_{rot}$ distribution requires an enhanced number of short-period binaries that comprise most of the fast rotating population, consistent with the findings of \citet{Simonian2018}, and a bimodal age distribution for \textit{Kepler} field stars, with $\approx 40\%$ of stars having ages between 100 Myr and 1 Gyr, as suggested by observations \citep{McQuillan2014,Davenport2017} and theory \citep{Matt2015}. The hypothesis of an age bimodality in the \textit{Kepler} field can be tested by measuring astroseimic ages \citep[e.g.][]{Ulrich1986,Silva2015}, for example, but also requires modeling tidal torques in binaries as P$_{rot}$ measured for unresolved, tidally-locked binaries does not correlate with age, complicating inference.

On-going observations by the extended Kepler mission \citep[K2,][]{Howell2014}, the Transiting Exoplanet Survey Satellite \citep{Ricker2014}, and Gaia \citep{Gaia2016,Lanzafame2018} will significantly increase the number of P$_{rot}$ observation for low-mass stars, providing an excellent test for models of stellar angular momentum evolution in both the field and in clusters.  We emphasize that modelling these stellar populations requires modeling tidally-interacting binaries, especially for fast rotators.

%% ACKNOWLEDGEMENTS %%
\acknowledgments
This work was facilitated though the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system and funded by the STF at the University of Washington. This work was supported by NASA Headquarters under the NASA Earth and Space Science Fellowship Program - Grant 80NSSC17K0482.  DPF and RB acknowledge that this work was supported by the NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement number NNA13AA93A.

%% SOFTWARE %%
%\software{matplotlib: \citet{Hunter2007}, numpy: \citet{vanderWalt2011}, pandas: \citet{Mckinney2010}}

%% BIBLIOGRAPHY %%
\bibliography{binGyro}

\end{document}

%% End document %%

### 2018-10-29 Analysis/tidalExample.py
added tidal evolution example

@@ -1,0 +1,70 @@
+"""
+
+@author: David P. Fleming, University of Washington, Oct 2018
+@email: dflemin3 (at) uw (dot) edu
+
+This script plots tidal evolution of
+
+"""
+
+import numpy as np
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
+mpl.rc('text', usetex=True)
+
+# Path to sim results
+cplDir = "../Sims/TidalExample/CPL"
+ctlDir = "../Sims/TidalExample/CTL"
+
+# When each system tidally locked in yr
+cplLock = 1.061738e+08
+ctlLock = 4.400000e+07
+
+# Load data
+# saOutputOrder	Time -TotEn -TotAngMom -Semim -RotPer Ecce -OrbPer -SurfEnFluxTotal
+cpl = np.genfromtxt(os.path.join(cplDir,"bintides.secondary.forward"))
+ctl = np.genfromtxt(os.path.join(ctlDir,"bintides.secondary.forward"))
+
+# Plot e, Porb, Prot
+fig, axes = plt.subplots(nrows=3, figsize=(9,24), sharex=True)
+
+# Left: e
+axes[0].plot(cpl[:,0], cpl[:,5], lw=3, ls="-", color="C0", label="CPL")
+axes[0].plot(ctl[:,0], ctl[:,5], lw=3, ls="-", color="C1", label="CTL")
+
+axes[0].set_ylabel("Eccentricity")
+axes[0].legend(loc="best", framealpha=0)
+
+# Middle: Porb
+axes[1].plot(cpl[:,0], cpl[:,6], lw=3, ls="-", color="C0", label="CPL")
+axes[1].plot(ctl[:,0], ctl[:,6], lw=3, ls="-", color="C1", label="CTL")
+
+axes[1].set_ylabel("Orbital Period [d]")
+
+# Right: Prot
+axes[2].plot(cpl[:,0], cpl[:,4], lw=3, ls="-", color="C0", label="CPL")
+axes[2].plot(ctl[:,0], ctl[:,4], lw=3, ls="-", color="C1", label="CTL")
+
[…]

### 2018-10-29 Paper/sync.tex
Updates from ShareLaTeX

@@
 
-\shorttitle{Reproducing Kepler Stellar Rotation Periods}
-\shortauthors{Fleming et al.}
+\shorttitle{Tidal Locking in Stellar Binaries}
+\shortauthors{Fleming \& Barnes}
 
@@
 
-\newcommand{\vplanet}[0]{\texttt{VPLANET}\xspace}
+\newcommand{\vplanet}[0]{\texttt{VPLanet}\xspace}
 \newcommand{\eqtide}[0]{\texttt{EQTIDE}\xspace}
 \newcommand{\stellar}[0]{\texttt{STELLAR}\xspace}
+\newcommand{\kepler}[0]{\textit{Kepler}\xspace}
 
@@
 
-\title{Reproducing Stellar Rotation Periods in the Kepler Field via Magnetic Braking and Tidal Torques}
+\title{Tidal Locking in Low-Mass Stellar Binary Stars}
 
@@
 
-\author{James R. A. Davenport}
-\affiliation{Astronomy Department, University of Washington \\
-Box 951580, Seattle, WA 98195}
-
-\author{Rodrigo Luger}
-\affiliation{Center for Computational Astrophysics, Flatiron Institute \\
-New York, NY 10010}
-
 %% ABSTRACT %%
@@
 
-Stellar rotation period (P$_{rot}$) studies have produced numerous insights into the angular momentum evolution of low-mass stars, but previous investigations have focused on single stars even though roughly half of Sun-like stars are in stellar binaries.
-We examine the impact of unresolved stellar binaries on the stellar P$_{rot}$ distribution of the \textit{Kepler} field by performing Monte Carlo simulations of a population composed of single and double stars by modeling stellar evolution, magnetic braking, and tidal forces.
-We find that tides modify stellar rotations in all binaries with orbital periods up to 70 days.
-At short orbital periods, most stellar binaries tidally-lock into synchronous rotation, naturally explaining the observed population of fast rotators in the \textit{Kepler} field that cannot be reproduced by single-star-only models.
-Many binaries with longer orbital periods, some up to orbital periods of 70 days, tidally-lock, synchronizing the stellar P$_{rot}$ with the binary orbital period, causing P$_{rot}$ to not be a valid proxy for age in all cases, i.e.
-gyrochronology methods must be applied carefully.
-Our assumed flat stellar age distribution over $1-4$ Gyr underpredicts the number of fast rotators in the \textit{Kepler} field.
-If instead we assume a bimodal stellar age distribution in the \textit{Kepler} field composed of a young ($100$ Myr$-1$ Gyr old) and an old ($1-4$ Gyr) population, we successfully reproduce the \textit{Kepler} P$_{rot}$ distribution.
+Abstract
 
@@
 
-Measuring stellar rotation periods via star spot modulation is a critical tool for understanding stellar angular momentum evolution in low-mass stars (M$\lsim 1.2$ M$_{\odot}$).
-As stars evolve, angular momentum conservation dictates that their evolving moments of inertia necessarily change their P$_{rot}$, e.g.
-stars spin-up as their radii contract, whereas magnetic braking, the coupling of the corotating stellar wind with the stellar magnetic field, removes angular momentum from the star, causing the observed long-term spin-down of stars \citep{Parker1958,Skumanich1972}.
+XXX
 
-Extensive measurements of ${>}10,000$ stellar rotation periods in the \textit{Kepler} field \citep{Borucki2010,McQuillan2013,Reinhold2013,McQuillan2014,Reinhold2015,Lurie2017} have provided exquisite observational constraints for theorists to model long-term stellar angular momentum evolution.
-Various models have been developed to explain the observed P$_{rot}$ distribution in the \textit{Kepler} field \citep[e.g.][]{Matt2015,vanSaders2018} and to link the long-term spin-down of low-mass stars to their ages, a method known as gyrochronology \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}.
-These models, however, have failed to reproduce the large observed population of rapid rotators (P$_{rot} {\lsim} 7.5$ days) across all stellar masses in the \textit{Kepler} field \citep{McQuillan2013,McQuillan2014}.
-Both gyrochonology relations \citep{McQuillan2014} and a modern magnetic braking model \citep{Matt2015} suggest that these fast rotators could be at most 1 Gyr old, in tension with the expected older ages of \textit{Kepler} field stars \citep{Chaplin2014,Angus2015}.
+tidal locking assumed for short period binaries, tsync << tcirc eg Mazeh 2008, so some have speculated that low-mass binaries can synchronize out to longer orbital periods.
+tentative evidence for this in Lurie 2017 data, discuss pseudosyncronization, synchronization, \citet{Counselman1973} tidal end states
 
-The aforementioned models have two limitations.
-First, these models assumed stellar populations composed solely of single stars, whereas roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}.
-Binaries are difficult to detect in photometric surveys because of low transit probabilities, so typically they are isolated using magnitude cuts \citep[e.g.][]{Davenport2018,Simonian2018}.
-However, 1000s of eclipsing binaries are known in the \textit{Kepler} field \citep{Kirk2016}, with likely many more unresolved binaries present in the data.
-In short-period binaries, tidal torques modify stellar rotations by driving the binary towards the tidally-locked state \citep{Zahn1989,Fleming2018}.
-Tidal forces in binaries are commonly understood to act in binaries with orbital periods (P$_{orb}$) ${\lsim} 10$ days \citep[e.g.][]{Zahn1989,Meibom2005}, however \citet{Lurie2017} measured 100s of \textit{Kepler} eclipsing binary stellar rotation periods using starspot variability and found that tides can synchronize stellar binary rotation periods up to P$_{orb} \approx 45$ days.
-Recently \citet{Simonian2018} found that most rapid rotators in the \textit{Kepler} field are likely non-eclipsing, tidally-synchronized short-period binaries.
-Clearly, unresolved stellar binaries impart a significant signal in P$_{rot}$ distributions, particularly in the \textit{Kepler} field at short P$_{orb}$, but can also influence P$_{rot}$ measurements at larger binary P$_{orb}$.
+we give a shit about sync binaries as they are fast rotators, e.g.
+what \citet{Simonian2018} might be seeing in \kepler field, can have enhanced magnetic activity (short Prot -> enhanced Lxuv), impact gyrochronology predictions, it's a fundamental evolutionary process for roughly half of sun like stars.
 
-Second, no previous study has simulated how a bimodal age distribution in the \textit{Kepler} field, suggested by \citet{McQuillan2014}, \citet{Matt2015}, and \citet{Davenport2017}, could explain the observed population of rapid rotators.
-If the age distribution is in fact bimodal, a young, fast rotating population could explain the observed P$_{rot}$ distribution.
-Here, we perform the first forward modeling of the angular momentum evolution of a realistic stellar population of single and double stars, accounting for stellar evolution, magnetic braking, tidal torques in stellar binaries, and testing the impact of a bimodal age distribution to reproduce the observed P$_{rot}$ distribution in the \textit{Kepler} field.
-
-%Here, we examine the role of unresolved, tidally-interacting stellar binaries in the \textit{Kepler} field using a stellar evolution model, with magnetic braking, coupled with a secular equilibrium tidal model for stellar binaries.
-to see if their presence can explain features in the \textit{Kepler} P$_{rot}$ distribution.
-In $\S$~\ref{sec:methods} and $\S$~\ref{sec:simulations}, we describe our computational methods and simulation set-up, and discuss our results and their implications in $\S$~\ref{sec:results} and $\S$~\ref{sec:discussion}, respectively.
-
-%% Maybe just focus on Kepler Prot observations?
-% in clusters \citep[e.g. Praesepe;][]{Scholz2007,Scholz2011,Agueros2011,Delorme2011,Kovacs2014,Douglas2017} and extensive P$_{rot}$ measurements
+dank plot showing average tsync for various orbital period ranges for CPL vs CTL marginalizing over all other parameters. do it with small multiples
 
@@
 
+XXX
+
 We simulate stellar rotation evolution for both single stars and stellar binaries using the coupled stellar-tidal evolution model presented in \citet{Fleming2018} and implemented in the open-source code VPLanet\footnote{VPLanet is publically available
-at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep[][Barnes et al., in prep]{Barnes2016}.
+at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep[][Barnes et al., in prep]{Barnes2016,vplanet2018}.
 We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
@@
 
-For stellar binaries, we assume that they tidally interact via the ``Constant Phase Lag" (CPL) version \citep[][]{FerrazMello2008,Heller2011} of the commonly used equilibrium tidal model \citep{Darwin1880}.
-The CPL model tracks the secular evolution of the binary's semi-major axis, $a$, eccentricity, $e$, and stellar rotation rates, $\omega$, due to tidal torques.
-The CPL model assumes that the tidal torque on one star due to its companion arises from a linear combination of uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass and is analogous to a driven, damped harmonic oscillator \citep{Greenberg2009}.
-We use the \eqtide implementation of the CPL model in VPLanet described in \citet{Fleming2018}, their equations 8-16, and simultaneously integrate those equations with Eq.~(\ref{eqn:stellar_rot_rate_dt}) for stellar binaries.
-All equations are integrated using a $4^{th}$ order Runge-Kutta scheme.
+ Equilibrium tidal models, first introduced by \citep{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega$, and obliquities $\psi$, of the tidally-interacting bodies subject to tidal torques.
+Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that are offset from the line connecting the both bodies' centers of mass, driving torques that exchange angular momentum between the orbit and both bodies' spins.
+Equilibrium tidal models are linear since the tidal waves that comprise the tidal bulge on a body are assumed to be uncoupled.
+Under these assumptions, the tidal evolution is analogous to a driven, damped harmonic oscillator \citep{Greenberg2009}.
+We refer the reader to \citet{Barnes2017} for an in-depth discussion of the assumptions and limitations of equilibrium tidal models.
+Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both exoplanets \citet[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Repetto2014,Fleming2018}.
+Here, we consider two equilibrium tidal models to study the secular spin-orbital evolution of low-mass stellar binaries.
+
+\subsubsection{Constant Phase Lag (CPL) Model}
+
+The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass.
+We use the \eqtide implementation of the CPL model in VPLanet following the derivation of \citet{FerrazMello2008}.
+The equations that govern the secular change in $e$ and $a$ are as follows:
+
+\begin{equation} \label{eqn:cpl_e}
+\frac{de}{dt} = -\frac{ae}{8 G m_1 m_2} \sum_{i=1}^2 Z_{i,\mathrm{CPL}} \left( 2 \varepsilon_{0,i} - \frac{49}{2} \varepsilon_{1,i} + \frac{1}{2} \varepsilon_{2,i} + 3 \varepsilon_{5,i} \right)
+\end{equation}
+\begin{equation} \label{eqn:cpl_a}
+\frac{da}{dt} = \sum_{i=1}^2 \frac{da_i}{dt}
+\end{equation}
+where if the $i^{th}$ body is tidally-locked in a synchronous orbit,
+\begin{equation} \label{eqn:cpl_dadt_locked}
+\frac{da_{i,sync}}{dt} = -\frac{a^2}{G m_1 m_2} Z_{i,\mathrm{CPL}} \left( 7 e^2 + \sin^2 (\psi_i) \right) \varepsilon_{2,i},
+\end{equation}
+otherwise
+\begin{equation}
+\begin{split}
+\frac{da_i}{dt} & = \frac{a^2}{4 G m_1 m_2} Z_{i,\mathrm{CPL}} \left( 4 \varepsilon_{0,i} + e^2 \left[ -20 \varepsilon_{0,i} + \frac{147}{2} \varepsilon_{1,i} \right. \right. \\
+&  + \left. \left. \frac{1}{2} \varepsilon_{2,i} - 3 \varepsilon_{5,i} \right] - 4 \sin^2 (\psi_i) \left[ \varepsilon_{0,i} - \varepsilon_{8,i} \right] \right).
+\end{split}
+\end{equation}
+The CPL equations for $\psi$ and $\omega$ evolution are
+\begin{equation} \label{eqn:cpl_psi}
+\frac{d\psi_i}{dt} = \frac{Z_{i,\mathrm{CPL}} \sin(\psi_i)}{4 m_i r_{g,i}^2 R_i^2 n \omega_i} \left( [1-\xi_i] \varepsilon_{0,i} + [1+\xi_i](\varepsilon_{8,i} - \varepsilon_{9,i}) \right)
+\end{equation}
+\begin{equation} \label{eqn:cpl_omega}
+\begin{split}
+\frac{d\omega_i}{dt}& = -\frac{Z_{i,\mathrm{CPL}}}{8m_i r_{g,i}^2 R_i^2 n} \left(4 \varepsilon_{0,i} + e^2\left[-20\varepsilon_{0,i} + 49\varepsilon_{1,i} + \varepsilon_{2,i} \right] \right. \\
+& \left. + 2 \sin^2(\psi_i) \left[ -2 \varepsilon_{0,i} + \varepsilon_{8,i} + \varepsilon_{9,i} \right] \right)
+\end{split}
+\end{equation}
+where $G$ is Newton's gravitational constant, $n$ is the binary's mean motion, and the index $i$ denotes that $i^{th}$ body.
+The tidal phase lags signs, $\varepsilon$, for the $i^{th}$ body are given by
+\begin{equation} \label{eqn:cpl_eps}
+\begin{split}
+\varepsilon_{0,i} & = \Sigma(2 \omega_i - 2n) \\
+\varepsilon_{1,i} & = \Sigma(2 \omega_i - 3n) \\
+\varepsilon_{2,i} & = \Sigma(2 \omega_i - n) \\
+\varepsilon_{5,i} & = \Sigma(n) \\
+\varepsilon_{8,i} & = \Sigma(\omega_i - 2n) \\
+\varepsilon_{9,i} & = \Sigma(\omega_i)
+\end{split}
+\end{equation}
+where the function $\Sigma(x)$ returns $1$ for positive $x$, $-1$ for negative $x$, and $0$ otherwise.
+
+The intermediate variable $Z_{\mathrm{CPL},i}$ is given by
+\begin{equation} \label{eqn:cpl_z}
+Z_{i,\mathrm{CPL}} = 3 G^2 k_{2,i} m_j^2 (m_1 + m_2) \frac{R_i^5}{a^9} \frac{1}{n Q_i}
+\end{equation}
+where the $j^{th}$ body is the $i^{th}$ body's companion, $k_{2}$ is the body's Love number of degree 2, and $Q$ is the tidal quality factor (``tidal Q").
+The tidal Q parameterizes the energy dissipation due to tidal evolution, with lower tidal Qs, i.e.
+larger phase differences between the tidal bulges, driving more rapid tidal evolution.
+
+$\xi_i$ is defined as
+\begin{equation}\label{eqn:cpl_chi}
+\xi_i = \frac{r_{\mathrm{g},i}^2 R_i^2 \omega_i a n }{ G M_j}.
+\end{equation}
 
@@
 
+\subsubsection{Constant Time Lag (CTL) Model}
+
+The ``Constant Time Lag" (CTL) \citep[][]{Hut1981,Leconte2010} equilibrium tidal model assumes a constant time interval between the body's tidal bulge and the passage of the tidally-interacting companion.
+In this formalism, unlike the CPL model, the CTL model is continuous over a range of tidal wave frequencies.
+However if the assumption of linearity is relaxed, i.e.
+frequencies associate with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}.
+We use the \eqtide implementation of the CTL model in VPLanet following the derivation of \citet{Leconte2010}.
+The equations that govern the secular change in $e$, $a$, $\omega$, and $\psi$ are as follows:
+
+\begin{equation} \label{eqn:ctl_e}
+  \frac{de}{dt} = \frac{11 ae}{2 G M_1 M_2}
+  \sum_{i = 1}^2 Z_{\mathrm{CTL},i} \left( \cos(\psi_i) \frac{f_4(e)}{\beta^{10}(e)}  \frac{\omega_i}{n} -\frac{18}{11} \frac{f_3(e)}{\beta^{13}(e)}\right),
+\end{equation}
+
+\begin{equation}\label{eqn:ctl_a}
+  \frac{da}{dt} \ = \  \frac{2 a^2}{G M_1 M_2}
+  \sum\limits_{i = 1}^2 Z_{\mathrm{CTL},i} \left( \cos(\psi_i) \frac{f_2(e)}{\beta^{12}(e)} \frac{\omega_i}{n} - \frac{f_1(e)}{\beta^{15}(e)}\right),
+\end{equation}
+
+\begin{equation}\label{eqn:ctl_o}
+  \frac{d\omega_i}{dt} \ = \ \frac{Z_{\mathrm{CTL},i}}{2 M_i r_{g,i}^2 
+R_i^2 n} \left( 2 \cos(\psi_i) \frac{f_2(e)}{\beta^{12}(e)} - \left[ 1+\cos^2(\psi)
+ \right] \frac{f_5(e)}{\beta^9(e)} 
+\frac{\omega_i}{n} \right),  
+\end{equation}
+and
+\begin{equation}\label{eqn:ctl_psi}
+  \frac{d\psi_i}{dt} = \frac{Z_{\mathrm{CTL},i} \sin(\psi_i)}{2 M_i r_{g,i}^2 R_i^2 n \omega_i}\left( \left[ \cos(\psi_i) - \frac{\xi_i}{ \beta} \right] \frac{f_5(e)}{\beta^9(e)} \frac{\omega_i}{n} - 2 \frac{f_2(e)}{\beta^{12}(e)} \right).
+\end{equation}
+where the intermediate variables are given by 
+\begin{equation}\label{eqn:ctl_z}
+ Z_{i,\mathrm{CTL}} = 3 G^2 k_{2,i} M_j^2 (M_i+M_j) \frac{R_i^5}{a^9} \tau_i ,
+\end{equation}
+and 
+\begin{equation}\label{eqn:ctl_f_e}
+\begin{array}{l}
+\beta(e) = \sqrt{1-e^2},\\
+f_1(e) = 1 + \frac{31}{2} e^2 + \frac{255}{8} e^4 + \frac{185}{16} e^6 + \frac{25}{
+64} e^8,\\
+f_2(e) = 1 + \frac{15}{2} e^2 + \frac{45}{8} e^4 + \frac{5}{16} e^6,\\
+f_3(e) = 1 + \frac{15}{4} e^2 + \frac{15}{8} e^4 + \frac{5}{64} e^6,\\
+f_4(e) = 1 + \frac{3}{2} e^2 + \frac{1}{8} e^4,\\
+f_5(e) = 1 + 3 e^2 + \frac{3}{8} e^4.
+\end{array}
+\end{equation}
+
+In both the CPL and CTL model, We assume $k_2 = 0.5$.
+This choice of $k_2$ does not impact our results as $k_2$ is degenerate with Q in the CPL model, e.g.
+the $k_2/Q$ scaling in Eq.~(\ref{eqn:cpl_z}), and with $\tau$ in the CTL model, e.g.
+$k_2 \tau$ scaling in Eq.~(\ref{eqn:ctl_z}), so we instead examine how our results scale with $Q$ and $\tau$.
+
+\subsubsection{Tidal Locking and Numerical Integration}
+
+XXX discuss integration specifics, equilbrium rotation rates, timestepping, $4^{th}$ order Runge-Kutta scheme.
+
+\subsubsection{Example Tidal Evolution}
+
+Here in Fig.~\ref{fig:tidalExample} we plot the example tidal evolution for $a$, $e$, and $\omega$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively.
+
+\begin{figure*}[h]
+	\includegraphics[width=\textwidth]{../Plots/tidalExample.pdf}
+   \caption{Spin-orbital tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary. XXX}%
+    \label{fig:tidalExample}%
+\end{figure*}
+
 %% SECTION: Simulations + Initial conditions %%
 
-\subsection{Simulations} \label{sec:simulations}
+\section{Simulations} \label{sec:simulations}
 
-We simulate the P$_{rot}$ distribution of the \textit{Kepler} field by performing Monte Carlo study consisting of a population of 10,000 stellar systems composed of single stars and binaries using VPLanet.
-For each simulation, we sample the stellar mass from a Gaussian fit to the stellar mass distribution in \citet{McQuillan2014} with $\mu=0.885$ M$_{\odot}$ and $\sigma=0.201$ M$_{\odot}$ to approximate the \textit{Kepler} mission's intended selection bias of monitoring FGK stars for Earth-like planets.
-We restrict stellar masses to be within [$0.1,1.2$] M$_{\odot}$ to limit the influence of subgiants since more massive stars evolve off the main sequence after $\gsim 4$ Gyr, our age limit, and because more massive stars do not experience magnetic braking due to their shallow convective envelopes \citep{Kraft1967}.
-Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of stars in the ${\sim}2$ Myr old ONC.
-We randomly assign stars a stellar companion according to stellar multiplicity fractions from \citet{Duchene2013}.
-In binaries, we compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010}.We uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$] and stellar tidal Qs are sampled from a log-uniform distribution over $[10^5,10^8]$.
-For the initial binary $P_{orb}$, we test two models.
-In Model 1, $10\%$ of binaries have their initial P$_{orb}$ uniformly sampled over [$3,80$] days, consistent with P$_{orb}$ distributions of field stars \citep[e.g.][]{Raghavan2010}, and the remaining $90\%$ have P$_{orb} > 80$ d and are considered ``single" stars since for P$_{orb} > 80$ d, tidal torques are sufficiently weak such that the stars' evolution is effectively independent of their companion.
-In Model 2, we consider an enhanced population of short period binaries by sampling all initial P$_{orb}$ uniformly over [$3,80$] days.
+XXX
 
-In Models 1 and 2, we simulate each system to an age uniformly sampled over [$1,4$] Gyr, the approximate age range of \textit{Kepler} field stars \citep{McQuillan2014,Angus2015}, with the upper-limit suggested by \citet{McQuillan2014} and \citet{Matt2015}.
-We also consider a ``Two Population" model that uses the initial conditions of Model 2 and assumes that $40\%$ of stars comprise a ``young" population with ages uniformly sampled over [$0.1,1$] Gyr, while the remainder comprise the ``old" population with ages uniformly sampled over [$1,4$] Gyr.
-We track the P$_{rot}$ evolution for each single star and the primary, i.e.
-more massive star, in binaries as it is observationally easier to measure a P$_{rot}$ due to starspot modulations on the more massive, and hence brighter, star.
-We compare our simulation results with the \textit{Kepler} P$_{rot}$ distribution measured by \citet{McQuillan2014}.
-
-All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/binGyro}{https://github.com/dflemin3/binGyro}.} with accompanying descriptions and documentation.
-
-%  For simulations of binaries, given the inherent uncertainty in and complexity of the formation of binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration.
-
-%Our goal is to determine how the distribution of $P_{rot}$ for a population stars composed solely of single stars differs, if at all, from a population of both single stars and binaries.
-If there is a substantial difference between the two distributions, e.g.
-tidally-locked fast rotators, then any study examining the distribution of stellar $P_{rot}$ should therefore account for the impact of tidally-interacting stellar binaries.
-We approach this problem with a ...
+- we consider eccentricities up to 0.3 as the CPL model fails above that.  CTL model can handle it, but at high eccentricity, dynamical tide, e.g. dissipation via waves and stuff, becomes important
 
@@
 
-\subsection{Tidal Locking in Stellar Binaries} \label{sec:results:locking}
+XXX
 
-Here we examine the tidal locking in binaries using the simulations from Model 2 (see $\S$~\ref{sec:results:kepler2}) by plotting P$_{rot}$ as a function of mass and age for binaries in Fig.~\ref{fig:locking}.
-We find that $35.5\%$ of binaries tidally lock before the age of the system, typically for tidal $Q \lsim 10^7$.
-Tidally-locked binaries tend to rotate more rapidly than unlocked binaries, reflected in the increased density for P$_{rot} \lsim 20$ d in the marginal distribution for the locked population and in the distribution medians (locked median P$_{rot} = 19.8$ d compared to unlocked median P$_{rot} = 27.5$ d).
-Binaries readily tidally-lock at short P$_{orb}$ due to the tidal force's strong semi-major axis dependence, increasing the strength of tidal torques at shorter orbital separations.
-When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} = $P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.
-
-Tidal forces modify the spins of unlocked binaries as well, tending to spin-up stars against the effects of magnetic braking towards orbital synchronization, reflected in unlocked binaries' median P$_{rot} = 27.5$ d, compared to the single star median P$_{rot} = 30.6$ d.
-Tidal locking is not limited to P$_{orb} \lsim 10$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 70$ d, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars (compare to Fig.~\ref{fig:kepler2}).
-This behavior is consistent with observations of P$_{rot}$ in \textit{Kepler} eclipsing binaries by \citet{Lurie2017} who find that binaries can tidally lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-The observed age distribution of tidally-locked binaries has little dependence on P$_{rot}$, with both young and old systems tidally-locking at long and short P$_{rot}$, causing the predictions of gyrochronology models to fail in such systems.
-
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/locked_kepler.pdf}
-   \caption{Simulated distribution of P$_{rot}$ for tidally-locked (cross) and unlocked (circle) stellar binaries, colored by age for Model 2 (see $\S$~\ref{sec:results:kepler2}).
-Left: P$_{rot}$ as a function of stellar mass.
-Right: Marginalized P$_{rot}$ distribution for locked (blue) and unlocked (orange) binaries.
-Tidally-locked binaries rotate more rapidly than unlocked binaries, displaying an increased density in the P$_{rot}$ distribution for P$_{rot} {\lsim} 20$ d, and with a median P$_{rot} = 19.8$ d, compared to the unlocked median P$_{rot} = 27.5$ d.
-Unlocked binaries on average rotate more quickly than single stars (compare to single distribution in Fig.~\ref{fig:kepler2}) due to the influence of tidal torques.}
-    \label{fig:locking}
-\end{figure}
-
-\subsection{Comparison to \textit{Kepler}}
-
-\subsubsection{Model 1} \label{sec:results:kepler1}
-
-In Fig.~\ref{fig:kepler1}, we display our simulated P$_{rot}$ for Model 1, compared to the observed \textit{Kepler} distribution \citep{McQuillan2014}.
-Single stars, defined as stars without a companion, or with a companion with a P$_{orb} > 80$ d where tides do not modify the orbit or stellar rotations, rotate more slowly with age due to magnetic braking, with more massive stars rotating more quickly than low mass stars, consistent with \textit{Kepler} P$_{rot}$ observations.
-There is a larger spread in P$_{rot}$ for stars with M$\lsim 1$ M$_{\odot}$ with the spread increasing with decreasing stellar mass, due to the mass-dependence magnetic braking torque in the \citet{Matt2015} model.
-%in which lower mass stars tend to spin down more quickly than higher-mass stars once they enter the "unsaturated" regime (see \citet{Matt2015} equations 6 and 7).
-
-The inclusion stellar binaries naturally produces a population of fast rotators (P$_{rot} < 7.5$ d) across all masses as binaries readily tidally-lock at shorter P$_{orb}$, with ${\sim}90\%$ of stars in this P$_{rot}$ regime being tidally-locked binaries.
-Single-star only models fail to reproduce this feature.
-The fast rotator population only comprises $0.25\%$ of all simulated stars, compared to $16\%$ of \textit{Kepler} stars in the same regime, and is therefore insufficient to explain the observed population of fast rotators.
-Our simulated P$_{rot}$ distribution is systematically biased towards longer P$_{rot}$ than the \textit{Kepler} distribution with a total median P$_{rot} = 30.3$ d compared to \textit{Kepler} median P$_{rot} = 17.3$ d.
-The inclusion of binaries does little to shift the P$_{rot}$ distribution towards shorter P$_{rot}$ as Model 1's assumed binary P$_{orb}$ distribution, based on observations of field stars \citep{Raghavan2010}, predicts that only ${\sim}4\%$ of all stars will be in stellar binaries with P$_{orb} < 80$ d, short enough for tides to appreciably modify stellar rotations.
-
-\begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/kepler_box_dist_corr.pdf}
-   \caption{Simulated distribution of P$_{rot}$ for single (circle) and double (cross, defined as stellar binaries with P$_{orb} < 80$ d) stars for Model 1.
-Left: P$_{rot}$ as a function of stellar mass, colored by age.
-Center: Marginalized P$_{rot}$ distributions for single stars (orange), binary stars (green), and the total sample (blue), and \textit{Kepler} \citep{McQuillan2014} (red).
-Except for a few low-mass (${\lsim}0.3$M$_{\odot}$) single stars, regardless of age, rotate more slowly (single median P$_{rot} = 30.5$ d) than \textit{Kepler} stars (\textit{Kepler} median P$_{rot} = 17.3$ d) and fail to explain the observed population of fast rotators.
-Binaries, however, readily tidally-lock at shorter P$_{orb}$, naturally producing a population of fast rotators (binary median P$_{rot} = 25.9$ d).
-Right: box-and-whisker plot of the P$_{rot}$ distribution for the simulations seen in the left panel in 0.1 M$_{\odot}$ bins.
-The shaded regions cover the distributions' interquartile range, the white lines indicate the medians, and the whiskers do not indicate error bars, but rather the spread of the distribution, extending to the 5th and 95th percentiles.
-The numbers over the boxes indicate the percent difference between the single (orange) and total (blue) distribution medians with the \citet{McQuillan2014} (red) sample.
-Even with the inclusion of binaries that tidally-lock at short P$_{rot}$, Model 1 produces a more slowly rotating stellar population than is observed in the \textit{Kepler} field, indicating that our assumed stellar population is incorrect.}%
-    \label{fig:kepler1}%
-\end{figure*}
-
-\subsubsection{Model 2: Enhanced Binary Population} \label{sec:results:kepler2}
-
-In Model 2, we attempt to increase the number of fast rotators by increasing the number of systems that could tidally-lock at short P$_{rot}$.
-We assume all stellar binaries have a P$_{orb}$ uniformly sampled over [$3,80$] d, as opposed to just $10\%$ of binaries in Model 1.
-Although optimistic, our assumption is reasonable as an analysis of fast rotators in the \textit{Kepler} field by \citet{Simonian2018} found that $\sim 59\%$ of stars with $1.5 <$ P$_{rot} < 7$ d are likely unresolved, tidally-locked binaries, suggesting that a substantial short-period rapidly rotating binary population does exist in the \textit{Kepler} field.
-
-As in Model 1, stellar binaries tidally-lock at short P$_{rot}$ producing a population of fast rotators below the single star sequence with binaries comprising ${\sim}99\%$ of fast rotators.
-The inclusion of additional short P$_{orb}$ binaries noticeably modifies the simulated P$_{rot}$ distribution, shifting the median P$_{rot}$ for the total sample down to $28.6$ d, compared to $30.6$ d for the single-star-only distribution.
-The impact of binaries is evident across all masses in the right panel of Fig.~\ref{fig:kepler2} as the total P$_{rot}$ distribution, which includes binaries, better matches the observed \textit{Kepler} distribution than the single star only distribution, due to the fast rotating tidally-locked binary population.
-The enhanced binary population of Model 2 still predicts a more slowly rotating stellar P$_{rot}$ distribution that is observed, specifically in the fast rotator regime, with Model 2 predicting $3.4\%$ of our simulated stars are fast rotators, compared to $16\%$ of \textit{Kepler} stars.
-As noted in $\S$~\ref{sec:results:locking}, tidally-locked binaries fix P$_{rot} = $ P$_{orb}$ as tidal torques overpower the torque due to magnetic braking, causing P$_{rot}$ in binaries to not correlate with age.
-
-\begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/kepler_box_dist.pdf}
-   \caption{Identical format to Fig.~\ref{fig:kepler1}, but displaying simulated data from Model 2 (see $\S$~\ref{sec:results:kepler2}).
-Left: Single stars, regardless of age, rotate too slowly (single median P$_{rot} = 30.6$ d) to explain the population of fast rotators in the \textit{Kepler} field (\textit{Kepler} median P$_{rot} = 17.3$ d).
-The enhanced population of short P$_{orb}$ binaries, however, naturally produces a significant population of tidally-locked binaries at P$_{rot} \lsim 20$ d (binary median P$_{rot} = 25.4$ d) that is reflected in the observed total P$_{rot}$ distribution (total median P$_{rot} = 28.6$ d).
-Right: For most masses, the total distribution, which includes binaries, can better reproduce the \textit{Kepler} medians and fast rotator population compared to the single-star only distribution.
-Even with the addition of a large population of short-period binaries, Model 2 still underpredicts the number of fast rotators across all stellar masses.}%
-    \label{fig:kepler2}%
-\end{figure*}
-
-\subsubsection{Two Population Model} \label{sec:results:kepler_two_pop}
-
-Observations \citep{McQuillan2014,Davenport2017} and theoretical work \citep{Matt2015} suggest that the \textit{Kepler} field contains a substantial population of young stars with ages $< 1$ Gyr.
-Given the inability of Models 1 and 2 to reproduce the \textit{Kepler} P$_{rot}$ distribution, especially at short P$_{rot}$, we test the hypothesize of a bimodal \textit{Kepler} stellar age distribution.
-We use the enhanced binary population of Model 2 and assume that $40\%$ of stars comprise the ``young" population, with ages uniformly sampled over [$0.1,1$] Gyr, while the remainder comprise the ``old" population, with ages uniformly sampled over [$1,4$] Gyr.
-We display the results of these simulations in Fig.~\ref{fig:kepler_two_pop}.
-
-We find that the inclusion of the young population substantially improves our simulated P$_{rot}$ distribution's fit to the \textit{Kepler} distribution with the total median P$_{rot} = 19.6$ d, a close match to the median P$_{rot} = 17.3$ d in the \textit{Kepler} field.
-In the right panel of Fig.~\ref{fig:kepler_two_pop} for $ 0.4 \textrm{M}_{\odot} \lsim \textrm{M} \lsim 1.2 \textrm{M}_{\odot}$, the inclusion of both a young population and tidally-interacting binaries in the total (blue) sample leads to an excellent match with the \textit{Kepler} distribution (red) compared to the single-star only (orange) model, emphasizing the importance of including stellar binaries in P$_{rot}$ forward models.
-The combination of a fast-rotating young stellar population and enhanced binary population produces a population of fast rotators (P$_{rot} < 7.5$ d) comprising $15.5\%$ of simulated stellar systems, remarkably similar to the \textit{Kepler} field's value of $16\%$.
-Moreover, we find that $52\%$ of fast rotators are in stellar binaries, in excellent agreement with the suggested value of $59\%$ for likely \textit{Kepler} photometric binaries estimated by \citet{Simonian2018}.
-Our simulations suggest that not only is there likely a substantial population of young stars in the \textit{Kepler} field comprising $\sim 40\%$ of all stars, but also that a substantial fractions of fast rotators in the \textit{Kepler} field are tidally-interacting binaries.
-Our simulations still predict a more slowly-rotating population that what is observed, suggesting that our simple bimodal age distribution should take on a different form, e.g.
-two Gaussians, or that stellar angular momentum evolution might proceed differently for older stars as proposed by \citet[e.g.][]{vanSaders2016,vanSaders2018}.
-
-\begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/kepler_two_pop_box_dist.pdf}
-   \caption{Identical format to Fig.~\ref{fig:kepler1}, but displaying simulated data from our two population model (see $\S$~\ref{sec:results:kepler_two_pop}).
+%\begin{figure*}[ht]
+%	\includegraphics[width=\textwidth]{../Plots/kepler_two_pop_box_dist.pdf}
+%   \caption{Identical format to Fig.~\ref{fig:kepler1}, but displaying simulated data from our two population model (see $\S$~\ref{sec:results:kepler_two_pop}).
 Our two population model reproduces the observed \textit{Kepler} P$_{rot}$ distribution, especially the fast rotator population, with a total median P$_{rot} = 19.6$ d compared to the \textit{Kepler} median P$_{rot} = 17.3$ d, a close match.
 In the right panel for $ 0.4 \textrm{M}_{\odot} \lsim \textrm{M} \lsim 1.2 \textrm{M}_{\odot}$, the inclusion of tidally-interacting binaries in the total (blue) sample leads to an excellent match with the \textit{Kepler} distribution (red) compared to the single-star only (orange) model, emphasizing the importance of including stellar binaries in P$_{rot}$ forward models.}%
-    \label{fig:kepler_two_pop}%
-\end{figure*}
-
-%% Old stuff %%
-
-% Stellar evolution figure caption
-% Stellar evolution for 1 M$_{\odot}$ (blue), 0.7 M$_{\odot}$, (orange), and 0.2 M$_{\odot}$ (green) stars according to our interpolation of the \citet{Baraffe2015} stellar models, magnetic braking, and assuming conservation of angular momentum according to \stellar.
-We display the radius (left panel), $r_g$ (middle panel), and rotation rate (right panel) evolution for each star using two magnetic braking models from \citet{Reiners2012} (solid curve) and \citet{Repetto2014} (dashed curve).
-The dot on each curve denotes the approximate time that each star reaches the main sequence.
-At early times, radius contract and $r_g$ decreasing spins up the stars, while at late times, magnetic braking efficiently spins down stars.
-
-%\begin{figure*}[ht]
-%	\includegraphics[width=\textwidth]{../Plots/Qrot.pdf}
-%   \caption{Rotation period as a function of time for a solar mass star (No Tides, black dot-dashed curve) and solar twin binaries with initial orbital periods of 20 days (solid curves) and 50 days (dashed curves).
-The dotted lines indicate rotation periods of 20 and 50 days, respectively.
-The different colors of the binaries' curves denote the assumed tidal Qs, which range from $10^4$ to $10^7$.
-For the first ${\sim}5 \times 10^7$ years, stellar radius contraction and $r_g$ evolution on the pre-main sequence spin up the stars.
-Once the stars reach the main sequence, magnetic braking, and tides in the binary systems, slow the rotation rates.
-In the long term, stellar rotation rates in binaries with lower tidal Qs slow more rapidly, and eventually all binary stars tidally lock into synchronous rotation, whereas the single star in the No Tides continues to spin down under the influence of magnetic braking.
-Once the stellar binaries tidally lock, their spins undergo little evolution, with most slightly speeding up as binary orbital decay, caused by magnetic braking, slowly decreases their orbital periods.}
-%    \label{fig:Qrot}
+%    \label{fig:kepler_two_pop}%
 %\end{figure*}
-
-%\subsection{P$_{rot}$ Distribution Over Time} \label{sec:results:prot_ages}
-
-% XXX put stats discussion in results, reference them here to shorten caption
-%\begin{figure}[h]
-%	\includegraphics[width=0.4\textwidth]{../Plots/kepler_prot_dist_ensemble.pdf}
-%  \caption{Simulated P$_{rot}$ distributions for synthetic clusters ages of 100, 300, 600 Myr, 1 Gyr, and the final age of each system, 1-4 Gyr.
-For each cluster age, we gauge the impact of binarity on the P$_{rot}$ distribution by performing the two-sample Kolmogorov–Smirnov test, testing the null hypothesis that the total (blue) and single (orange) sample are drawn from the same parent distribution for a significane level $\alpha = 0.005$, or an effective $\alpha = 0.001$ after applying the Bonferroni correction \citep{Dunn1961} for testing multiple hypotheses.
-At 100 Myr, many stars in the sample are still contracting on the pre-main sequence, and hence the P$_{rot}$ distributions are shifted towards faster P$_{rot}$ compared to the initial distribution (black).
-As the stars age, the P$_{rot}$ distributions shift towards slower P$_{rot}$ due to magnetic braking.
-For clusters older than 600 Myr, we can reject the null hypothesis that the single and total distributions are drawn from the same parent distribution, suggesting that the inclusion of binaries modifies the observed P$_{rot}$ distribution.
-This feature can be seen in the 1 Gyr old cluster as an enhancement of fast rotating stellar binaries, a signature of tidal-locking.
-For older field stars, the simulated distributions the impact of binaries is significant, reflected by a population of fast rotators not seen in the single population, and causes the total distribution to better reflect the \textit{Kepler} distribution.}
-%    \label{fig:kepler_dist_ensemble}
-%\end{figure}
 
@@
 
-For the first time, we model the angular momentum evolution of a realistic population of single and double stars, accounting for stellar evolution, magnetic braking, and tidal evolution in binaries, in order to reproduce the P$_{rot}$ distribution of \textit{Kepler} field stars.
-We show that tidal forces drive stellar binaries to the tidally-locked state, naturally produce a population of fast rotators that single star models fail to produce.
-We find that binaries can tidally-lock up to $P_{orb} \approx 70$ d, causing P$_{rot}$ in binaries to not be a valid proxy for age.
-We show that reproducing the \textit{Kepler} P$_{rot}$ distribution requires an enhanced number of short-period binaries that comprise most of the fast rotating population, consistent with the findings of \citet{Simonian2018}, and a bimodal age distribution for \textit{Kepler} field stars, with $\approx 40\%$ of stars having ages between 100 Myr and 1 Gyr, as suggested by observations \citep{McQuillan2014,Davenport2017} and theory \citep{Matt2015}.
-The hypothesis of an age bimodality in the \textit{Kepler} field can be tested by measuring astroseimic ages \citep[e.g.][]{Ulrich1986,Silva2015}, for example, but also requires modeling tidal torques in binaries as P$_{rot}$ measured for unresolved, tidally-locked binaries does not correlate with age, complicating inference.
-
-On-going observations by the extended Kepler mission \citep[K2,][]{Howell2014}, the Transiting Exoplanet Survey Satellite \citep{Ricker2014}, and Gaia \citep{Gaia2016,Lanzafame2018} will significantly increase the number of P$_{rot}$ observation for low-mass stars, providing an excellent test for models of stellar angular momentum evolution in both the field and in clusters.
-We emphasize that modelling these stellar populations requires modeling tidally-interacting binaries, especially for fast rotators.
+XXX
 
@@
 %% BIBLIOGRAPHY %%
-\bibliography{binGyro}
+\bibliography{sync}
 

### 2018-10-29 Analysis/tidalExample.py
tweaked tidal example figure

@@ -34,7 +34,7 @@ cpl = np.genfromtxt(os.path.join(cplDir,"bintides.secondary.forward"))
 ctl = np.genfromtxt(os.path.join(ctlDir,"bintides.secondary.forward"))
 
 # Plot e, Porb, Prot
-fig, axes = plt.subplots(nrows=3, figsize=(9,24), sharex=True)
+fig, axes = plt.subplots(nrows=3, figsize=(6,16), sharex=True)
 
 # Left: e
 axes[0].plot(cpl[:,0], cpl[:,5], lw=3, ls="-", color="C0", label="CPL")

### 2018-10-30 Analysis/stellar.py
added stellar evolution example

@@ -1,0 +1,101 @@
+
+"""
+
+@author: David P Fleming, University of Washington, Oct 2018
+@email dflemin3 (at) uw (dot) edu
+
+This script examines our stellar evolution model, a bicubic interpolation of the
+Baraffe+2015 stellar evolution model, coupled with the Matt+2015 magnetic
+braking model.
+
+"""
+
+import numpy as np
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 25.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
+mpl.rc('text', usetex=True)
+
+# Read in output files
+path = "../Sims/StellarEvolution/"
+
+# saOutputOrder Time -Radius -RotPer RadGyra
+g = np.genfromtxt(os.path.join(path,"stellar.g.forward"))
+k = np.genfromtxt(os.path.join(path,"stellar.k.forward"))
+m = np.genfromtxt(os.path.join(path,"stellar.m.forward"))
+
+# All on same time grid
+time = g[:,0]
+
+# Find time indices of approximate ZAMS times from Henny Lamers'
+# stellar evolution notes
+ind_g = np.argmin(np.fabs(time-6.2e7))
+ind_k = np.argmin(np.fabs(time-1.0e8))
+ind_m = np.argmin(np.fabs(time-3.0e8))
+
+# Plot!
+ncols = 3
+fig, ax = plt.subplots(ncols=ncols, figsize=(9*ncols,8))
+
+# Left panel: Stellar radius evolution
+ax[0].plot(time, g[:,1], lw=2.5, color="C0", label=r"$1$ M$_{\odot}$")
+ax[0].plot(time, k[:,1], lw=2.5, color="C1", label=r"$0.7$ M$_{\odot}$")
+ax[0].plot(time, m[:,1], lw=2.5, color="C2", label=r"$0.2$ M$_{\odot}$")
+
+# Plot points to indicate ZAMS
+ax[0].scatter(time[ind_g], g[ind_g,1], s=75, color="C0")
[…]

### 2018-10-30 Paper/sync.tex
Updates from ShareLaTeX

@@
 
-tidal locking assumed for short period binaries, tsync << tcirc eg Mazeh 2008, so some have speculated that low-mass binaries can synchronize out to longer orbital periods.
+tidal locking assumed for short period binaries, tsync < tcirc eg Mazeh 2008, so some have speculated that low-mass binaries can synchronize out to longer orbital periods.
 tentative evidence for this in Lurie 2017 data, discuss pseudosyncronization, synchronization, \citet{Counselman1973} tidal end states
@@
 
-dank plot showing average tsync for various orbital period ranges for CPL vs CTL marginalizing over all other parameters. do it with small multiples
+Plots:
+
+-average tsync for various orbital period ranges for CPL vs CTL marginalizing over all other parameters. do it with small multiples
+-example plots for stellar, eqtide, both
+-Prot as a function of mass for locked (unlocked) binaries of random ages for gyrochronology results, naturally produces fast rotators
+-Plot showing diff between binary and single star Prot evolution for various Porbs
+-tsync contours over Q/tau and Porb
 
@@
 
+We simulate coupled stellar-tidal evolution for stellar binaries using the coupled stellar-tidal evolution model presented in \citet{Fleming2018} and implemented in the open-source code VPLanet\footnote{VPLanet is publically available
+at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep[][Barnes et al., in prep]{Barnes2016,vplanet2018}.
+We integrate all model equations (see $\S$~\ref{sec:methods:stellar} and $\S$~\ref{sec:methods:eqtide}) using a $4^{th}$ order Runge-Kutta scheme with adaptive timestepping described in \citet{Fleming2018}.
+
 \subsection{Stellar Evolution} \label{sec:methods:stellar}
 
-XXX
-
-We simulate stellar rotation evolution for both single stars and stellar binaries using the coupled stellar-tidal evolution model presented in \citet{Fleming2018} and implemented in the open-source code VPLanet\footnote{VPLanet is publically available
-at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep[][Barnes et al., in prep]{Barnes2016,vplanet2018}.
 We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
-We model stellar spin-down due to magnetic braking using the model derived in \citet{Matt2015} since it has been shown to successfully reproduce many features in the P$_{rot}$ distribution of the \textit{Kepler} field, notably the upper and lower envelopes of the latter distribution and the observed ``dip" in the distribution around M${\approx}0.6$ M$_{\odot}$.
-We integrate the following equation to model the net change in the stellar rotation rate due to stellar evolution and magnetic braking:
-
+We model stellar spin-down due to magnetic braking using the model derived in \citet{Matt2015} since it has been shown to successfully reproduce many features in the P$_{rot}$ distribution of low-mass stars in the \textit{Kepler} field, notably the upper and lower envelopes of the latter distribution and the observed ``dip" in the distribution around M${\approx}0.6$ M$_{\odot}$.
+We model the net change in the stellar rotation rate due to stellar evolution and magnetic braking via the following equation
 \begin{equation} \label{eqn:stellar_rot_rate_dt}
@@
 where the moment of inertia $I = M r_g^2 R^2$ and $\dot{J}_{mb}$ is the angular momentum loss due to magnetic braking.
+
+\subsubsection{Example Stellar Evolution} \label{sec:methods:stellarExample}
+
+In Fig.~\ref{fig:stellarExample}, we plot the evolution of the radius, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars computed according to our stellar evolution model, \stellar, assuming all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr.
+
+\begin{figure}[h]
+	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
+   \caption{Caption.}%
+    \label{fig:stellarExample}%
+\end{figure}
 
@@
 
-\begin{equation} \label{eqn:cpl_e}
+\begin{equation} \label{eqn:cpl:e}
 \frac{de}{dt} = -\frac{ae}{8 G m_1 m_2} \sum_{i=1}^2 Z_{i,\mathrm{CPL}} \left( 2 \varepsilon_{0,i} - \frac{49}{2} \varepsilon_{1,i} + \frac{1}{2} \varepsilon_{2,i} + 3 \varepsilon_{5,i} \right)
 \end{equation}
-\begin{equation} \label{eqn:cpl_a}
+\begin{equation} \label{eqn:cpl:a}
 \frac{da}{dt} = \sum_{i=1}^2 \frac{da_i}{dt}
@@
 where if the $i^{th}$ body is tidally-locked in a synchronous orbit,
-\begin{equation} \label{eqn:cpl_dadt_locked}
+\begin{equation} \label{eqn:cpl:dadt_locked}
 \frac{da_{i,sync}}{dt} = -\frac{a^2}{G m_1 m_2} Z_{i,\mathrm{CPL}} \left( 7 e^2 + \sin^2 (\psi_i) \right) \varepsilon_{2,i},
@@
 The CPL equations for $\psi$ and $\omega$ evolution are
-\begin{equation} \label{eqn:cpl_psi}
+\begin{equation} \label{eqn:cpl:psi}
 \frac{d\psi_i}{dt} = \frac{Z_{i,\mathrm{CPL}} \sin(\psi_i)}{4 m_i r_{g,i}^2 R_i^2 n \omega_i} \left( [1-\xi_i] \varepsilon_{0,i} + [1+\xi_i](\varepsilon_{8,i} - \varepsilon_{9,i}) \right)
 \end{equation}
-\begin{equation} \label{eqn:cpl_omega}
+\begin{equation} \label{eqn:cpl:omega}
 \begin{split}
@@
 The tidal phase lags signs, $\varepsilon$, for the $i^{th}$ body are given by
-\begin{equation} \label{eqn:cpl_eps}
+\begin{equation} \label{eqn:cpl:eps}
 \begin{split}
@@
 The intermediate variable $Z_{\mathrm{CPL},i}$ is given by
-\begin{equation} \label{eqn:cpl_z}
+\begin{equation} \label{eqn:cpl:z}
 Z_{i,\mathrm{CPL}} = 3 G^2 k_{2,i} m_j^2 (m_1 + m_2) \frac{R_i^5}{a^9} \frac{1}{n Q_i}
@@
 $\xi_i$ is defined as
-\begin{equation}\label{eqn:cpl_chi}
+\begin{equation}\label{eqn:cpl:chi}
 \xi_i = \frac{r_{\mathrm{g},i}^2 R_i^2 \omega_i a n }{ G M_j}.
 \end{equation}
-
-%Following \citet{Fleming2018}, when one or both binary stars are tidally locked, we assume that tidal forces prevent magnetic braking from spinning down the tidally locked star(s), but instead, any angular momentum lost is lost at the expense of the binary orbit, decreasing the orbital semi-major axis as a result.
-Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify Eqs.
-(18) and (20) from \citet{Fleming2018} to account for $r_g$ evolution for the cases when one or both stars tidally lock under conservation of angular momentum, respectively:
-%\small
-%\begin{equation} \label{eqn:tidal_locked_one}
-%\begin{split}
-%\dot{a}_{coupled}^{(1)} = \frac{-\dot{J_{mb}} - 2 \omega \left( m_1 r_{g,1}^2 R_1 \dot{R_1} - m_1 r_{g,1} \dot{r}_{g,1} R_1^2 \right)}
-%{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} m_1 r_{g,1}^2 R_1^2}
-%\end{split}
-%\end{equation}
-%\normalsize
-% and
-% \small
-%\begin{equation} \label{eqn:tidal_locked_two}
-%\begin{split}
-%\dot{a}_{coupled}^{(2)} = \frac{-\dot{J_{mb}} - 2 \omega \left( \sum_{i=1}^{2} m_i r_{g,i}^2 R_i \dot{R_i} + m_i r_{g,i} \dot{r}_{g,i} R_i^2 \right)}
-%{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} \left( m_1 r_{g,1}^2 R_1^2 + m_2 r_{g,2}^2 R_2^2 \right)}
-%\end{split}
-%\end{equation}
-%\normalsize
-%where $J_{orb}$ is the orbital angular momentum.
 
@@
 
-\begin{equation} \label{eqn:ctl_e}
+\begin{equation} \label{eqn:ctl:e}
   \frac{de}{dt} = \frac{11 ae}{2 G M_1 M_2}
@@
 
-\begin{equation}\label{eqn:ctl_a}
+\begin{equation}\label{eqn:ctl:a}
   \frac{da}{dt} \ = \  \frac{2 a^2}{G M_1 M_2}
@@
 
-\begin{equation}\label{eqn:ctl_o}
+\begin{equation}\label{eqn:ctl:omega}
   \frac{d\omega_i}{dt} \ = \ \frac{Z_{\mathrm{CTL},i}}{2 M_i r_{g,i}^2 
@@
 and
-\begin{equation}\label{eqn:ctl_psi}
+\begin{equation}\label{eqn:ctl:psi}
   \frac{d\psi_i}{dt} = \frac{Z_{\mathrm{CTL},i} \sin(\psi_i)}{2 M_i r_{g,i}^2 R_i^2 n \omega_i}\left( \left[ \cos(\psi_i) - \frac{\xi_i}{ \beta} \right] \frac{f_5(e)}{\beta^9(e)} \frac{\omega_i}{n} - 2 \frac{f_2(e)}{\beta^{12}(e)} \right).
@@
 where the intermediate variables are given by 
-\begin{equation}\label{eqn:ctl_z}
+\begin{equation}\label{eqn:ctl:z}
  Z_{i,\mathrm{CTL}} = 3 G^2 k_{2,i} M_j^2 (M_i+M_j) \frac{R_i^5}{a^9} \tau_i ,
@@
 and 
-\begin{equation}\label{eqn:ctl_f_e}
+\begin{equation}\label{eqn:ctl:f_e}
 \begin{array}{l}
@@
 This choice of $k_2$ does not impact our results as $k_2$ is degenerate with Q in the CPL model, e.g.
-the $k_2/Q$ scaling in Eq.~(\ref{eqn:cpl_z}), and with $\tau$ in the CTL model, e.g.
-$k_2 \tau$ scaling in Eq.~(\ref{eqn:ctl_z}), so we instead examine how our results scale with $Q$ and $\tau$.
+the $k_2/Q$ scaling in Eq.~(\ref{eqn:cpl:z}), and with $\tau$ in the CTL model, e.g.
+$k_2 \tau$ scaling in Eq.~(\ref{eqn:ctl:z}), so we instead examine how our results scale with $Q$ and $\tau$.
 
-\subsubsection{Tidal Locking and Numerical Integration}
+\subsubsection{Tidal Locking}
 
-XXX discuss integration specifics, equilbrium rotation rates, timestepping, $4^{th}$ order Runge-Kutta scheme.
+When a body tidally locks, tidal torques fix P$_{rot}$ to the equilibrium P$_{rot}$, P$_{eq}$.
+Typically, tidal locking is understood in the context of a synchronized rotator, e.g.
+when P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
+Both the CPL and CTL model, however, predict that tidally-interacting bodies on sufficiently eccentric orbits undergo pseudosynchronous rotation, e.g.
+Mercury's 3:2 spin-orbit resonance \citep[P$_{rot} = 2/3$ P$_{orb}$,][]{GoldreichPeale1966}.
 
-\subsubsection{Example Tidal Evolution}
+The CPL model, owing to its assumption of a finite number of discrete tidal lags, only permits a 1:1 and 3:2 spin-orbit state where, following \citet{Barnes2017}, the CPL P$_{eq}$ is given by
+\begin{equation} \label{eqn:cpl:eqPer}
+P^{\mathrm{CPL}}_{eq} = 
+\begin{cases}
+P_{orb} & \text{if } e < \sqrt{1/19}\\
+\frac{2}{3}P_{orb} & \text{if } e \geq \sqrt{1/19}.
+\end{cases}
+\end{equation}
+Therefore, the CPL model predicts synchronous rotation for $e \lsim 0.23$, and a supersychronous 3:2 spin-orbit state otherwise.
+
+The CTL model, however, is continuous over a range of tidal frequencies and therefore predicts a P$_{eq}$ that is a continuous function of both $e$ and $\psi$.
+Following \citet{Barnes2017}, we define the CTL P$_{eq}$ by
+\begin{equation} \label{eqn:ctl:eqPer}
+P^{\mathrm{CTL}}_{eq} = P\frac{\beta^3 f_5(e) 1 + \cos^2(\psi)}{2f_2(e) \cos(\psi)}.
+\end{equation}
+The CTL model in general predicts that bodies on eccentric orbits tidally-lock into supersyncronous rotation, and only bodies on exactly circular orbits are synchronous rotators.
+
+Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, the numerical solution can produce unphysical evolution.
+We follow \citet{Barnes2013} and \citet{Fleming2018} and address this issue by fixing P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL model, respectively, when P$_{rot}$ is within $1\%$ of P$_{eq}$.
+We find that this scheme produces physically and numerically accurate results.
+
+\subsubsection{Example Tidal Evolution} \label{sec:methods:eqtideExample}
 
 Here in Fig.~\ref{fig:tidalExample} we plot the example tidal evolution for $a$, $e$, and $\omega$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively.
+Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increasing as tides forces the spins toward the tidally-locked state, transfering rotational angular momentum into the orbit in the process.
+At late times, both the CPL and CTL drive the binaries towards orbital circularization, with tidal dissipation decreasing P$_{orb}$.
+The predictions of the CPL and CTL model, differ, however, when the binaries tidally-lock.
+Under the CPL model, the binary tidally-locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
+Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence (see Eq.~(\ref{eqn:ctl:eqPer})).
 
-\begin{figure*}[h]
-	\includegraphics[width=\textwidth]{../Plots/tidalExample.pdf}
-   \caption{Spin-orbital tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary. XXX}%
+\begin{figure}[h]
+	\includegraphics[width=0.4\textwidth]{../Plots/tidalExample.pdf}
+   \caption{Tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary's $e$ (top), P$_{orb}$ (middle) and P$_{rot}$ (bottom) for the CPL (blue) and CTL (orange) model.
+The blue (CPL) and orange (CTL) vertical dashed lines denote when the stellar binary tidally-locks.
+Both the CPL and CTL model predict the same qualitative evolution.
+The rotational evolution differs, however, as under the CPL model, the binary tidally-locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
+Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence (see Eq.~(\ref{eqn:ctl:eqPer})).}%
     \label{fig:tidalExample}%
-\end{figure*}
+\end{figure}
+
+\subsection{Coupled Stellar-Tidal Evolution} \label{sec:coupled}
+
+Following \citet{Fleming2018}, when one or both binary stars are tidally locked, we assume that tidal forces prevent magnetic braking from spinning down the tidally locked star(s), but instead, any angular momentum lost is lost at the expense of the binary orbit, decreasing the $a$ as a result \citep{Verbunt1981}.
+Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify Eqs.
+(18) and (20) from \citet{Fleming2018} to account for $r_g$ evolution for the cases when one or both stars tidally lock under conservation of angular momentum, respectively:
+\small
+\begin{equation} \label{eqn:tidal_locked_one}
+\begin{split}
+\dot{a}_{coupled}^{(1)} = \frac{-\dot{J}_{mb} - 2 \omega \left( m_1 r_{g,1}^2 R_1 \dot{R_1} - m_1 r_{g,1} \dot{r}_{g,1} R_1^2 \right)}
+{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} m_1 r_{g,1}^2 R_1^2}
+\end{split}
+\end{equation}
+\normalsize
+and
+\small
+\begin{equation} \label{eqn:tidal_locked_two}
+\begin{split}
+\dot{a}_{coupled}^{(2)} = \frac{-\dot{J}_{mb} - 2 \omega \left( \sum_{i=1}^{2} m_i r_{g,i}^2 R_i \dot{R_i} + m_i r_{g,i} \dot{r}_{g,i} R_i^2 \right)}
+{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} \left( m_1 r_{g,1}^2 R_1^2 + m_2 r_{g,2}^2 R_2^2 \right)}.
+\end{split}
+\end{equation}
+\normalsize
+where $J_{orb}$ is the orbital angular momentum.
 

### 2018-11-01 Analysis/lockTimeDist.py
improved plot

@@ -71,10 +71,10 @@ porbBinEdges = [0, 20, 40, 60, 80, 100]
 
 fig = plt.figure(figsize=(5, 15))
 
-gs = GridSpec(9, 1, height_ratios=[1, 0.05,
-                                   1, 0.05,
-                                   1, 0.05,
-                                   1, 0.05, 1])
+gs = GridSpec(9, 1, height_ratios=[1, 0.1,
+                                   1, 0.1,
+                                   1, 0.1,
+                                   1, 0.1, 1])
 
 # Loop over gridspec
 for ii in range(9):
@@ -104,16 +104,16 @@ for ii in range(9):
         med = np.median(cpl["Pri_LockTime"][cplMask].values/1.0e9)
         up = np.percentile(cpl["Pri_LockTime"][cplMask].values/1.0e9, 75) - med
         down = med - np.percentile(cpl["Pri_LockTime"][cplMask].values/1.0e9, 25)
-        ax.text(0.76, 0.9, ("CPL: $%.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up, down)),
-                ha="center", va="center", size=14, color="C0", zorder=100,
+        ax.text(0.75, 0.9, ("CPL: $%.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up, down)),
+                ha="center", va="center", size=15, color="C0", zorder=100,
                 bbox=dict(boxstyle="square", fc="white", ec="white", alpha=0.0),
                 transform=ax.transAxes)
 
         med = np.median(ctl["Pri_LockTime"][ctlMask].values/1.0e9)
         up = np.percentile(ctl["Pri_LockTime"][ctlMask].values/1.0e9, 75) - med
         down = med - np.percentile(ctl["Pri_LockTime"][ctlMask].values/1.0e9, 25)
-        ax.text(0.76, 0.76, ("CTL: $%.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up, down)),
-                ha="center", va="center", size=14, color="C1", zorder=100,
+        ax.text(0.75, 0.75, ("CTL: $%.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up, down)),
+                ha="center", va="center", size=15, color="C1", zorder=100,
                 bbox=dict(boxstyle="square", fc="white", ec="white", alpha=0.0),
                 transform=ax.transAxes)
 
@@ -124,7 +124,7 @@ for ii in range(9):
[…]

### 2018-11-01 Paper/sync.tex

@@
 \shorttitle{Tidal Locking in Stellar Binaries}
-\shortauthors{Fleming \& Barnes}
+\shortauthors{Fleming et al.}
 
@@
 %%\author[0000-0001-9293-4043]{David P. Fleming}
-\author{David P. Fleming}
+\author{David P. Fleming et al}
 \affil{Astronomy Department, University of Washington \\
-Box 951580, Seattle, WA 98195}
-
-\author{Rory Barnes}
-\affiliation{Astronomy Department, University of Washington \\
 Box 951580, Seattle, WA 98195}
@@
 
-XXX
+Angular momentum evolution in low-mass (M$\lsim 1$ M$_{\odot}$) short-period (P$_{orb} \lsim 10$ d) stellar binaries is dominated by tides.
+Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}.
+Orbital circularization in ubiquitous in short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
+For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} finds that the orbit of solar twin binaries circularizes during the stellar pre-main sequence.
+Observations by \citet{Meibom2005} support this picture as they find short-period binaries in the ${\sim}150$ Myr old cluster, M35, tend to have circular orbits.
 
-tidal locking assumed for short period binaries, tsync < tcirc eg Mazeh 2008, so some have speculated that low-mass binaries can synchronize out to longer orbital periods.
-tentative evidence for this in Lurie 2017 data, discuss pseudosyncronization, synchronization, \citet{Counselman1973} tidal end states
+Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially so for the stellar spins.
+Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale since the stellar spins harbor much less angular momentum than the orbit \citep{Zahn1989,Witte2002,Mazeh2008}.
+Tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
+A large sample of $816$ rotation period measurements derived using starspot modulations on primary stars of \kepler eclipsing binaries by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked, but also find tentative evidence for tidal-locking in binaries with P$_{orb}$ up to 45 d, their detection limit.
+Observations by the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will contain low-mass, short-period eclipsing binaries, permitting additional tests for models of stellar rotation evolution due to tidal torques.
 
-we give a shit about sync binaries as they are fast rotators, e.g.
-what \citet{Simonian2018} might be seeing in \kepler field, can have enhanced magnetic activity (short Prot -> enhanced Lxuv), impact gyrochronology predictions, it's a fundamental evolutionary process for roughly half of sun like stars.
+Rotation period distributions measured in clusters and field stars are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys due to low transit probabilities.
+In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P${rot}$ distributions.
+Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}.
+Ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that link the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, will likely be incorrect, however no previous study has quantified this effect.
+Moreover, no previous study has quantified how tidal-locking proceeds in low-mass stellar binaries using a realistic treatment of stellar evolution and magnetic braking.
 
-Plots:
+XXX mention main sequence lifetimes
 
--average tsync for various orbital period ranges for CPL vs CTL marginalizing over all other parameters. do it with small multiples
--example plots for stellar, eqtide, both
--Prot as a function of mass for locked (unlocked) binaries of random ages for gyrochronology results, naturally produces fast rotators
--Plot showing diff between binary and single star Prot evolution for various Porbs
--tsync contours over Q/tau and Porb
+Here, we examine how tidal torques, coupled with a modern stellar evolution and magnetic braking models, shape the angular momentum evolution of low-mass binary stars.
+We investigate under what conditions tidal-locking occurs, and what the typical tidal-locking timescale is as a function of stellar mass and tidal dissipation parameters for two widely-used equilibrium tidal models.
+We show how tidal torques in binaries cause stellar rotation periods to not strongly correlate with age, causing the predictions of gyrochronlogy models to fail in such systems.
+We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
+We present our results in $\S$~\ref{sec:results} and discuss their implications in $\S$~\ref{sec:discussion}.
 
@@
 
-We simulate coupled stellar-tidal evolution for stellar binaries using the coupled stellar-tidal evolution model presented in \citet{Fleming2018} and implemented in the open-source code VPLanet\footnote{VPLanet is publically available
+We simulate coupled stellar-tidal evolution for stellar binaries using an improved version of the coupled stellar-tidal evolution model presented in \citet{Fleming2018}.
+We implement our model in the open-source code VPLanet\footnote{VPLanet is publicly available
 at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep[][Barnes et al., in prep]{Barnes2016,vplanet2018}.
-We integrate all model equations (see $\S$~\ref{sec:methods:stellar} and $\S$~\ref{sec:methods:eqtide}) using a $4^{th}$ order Runge-Kutta scheme with adaptive timestepping described in \citet{Fleming2018}.
+We integrate all model equations (see $\S$~\ref{sec:methods:stellar} and $\S$~\ref{sec:methods:eqtide}) using using the $4^{th}$ order Runge-Kutta scheme with adaptive timestepping described in \citet{Fleming2018}.
 
@@
 We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
-We model stellar spin-down due to magnetic braking using the model derived in \citet{Matt2015} since it has been shown to successfully reproduce many features in the P$_{rot}$ distribution of low-mass stars in the \textit{Kepler} field, notably the upper and lower envelopes of the latter distribution and the observed ``dip" in the distribution around M${\approx}0.6$ M$_{\odot}$.
+We model the long-term stellar spin-down due to magnetic braking using the model derived in \citet{Matt2015} since it has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
 We model the net change in the stellar rotation rate due to stellar evolution and magnetic braking via the following equation
@@
 \end{equation}
-where the moment of inertia $I = M r_g^2 R^2$ and $\dot{J}_{mb}$ is the angular momentum loss due to magnetic braking.
+where the moment of inertia $I = M r_g^2 R^2$, $\dot{J}_{mb}$ is the angular momentum loss due to magnetic braking, and the time derivatives of the stellar radius and $r_g$ are computed numerically using our interpolation of the \citet{Baraffe2015} stellar evolution grids.
 
@@
 
-In Fig.~\ref{fig:stellarExample}, we plot the evolution of the radius, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars computed according to our stellar evolution model, \stellar, assuming all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr.
+In Fig.~\ref{fig:stellarExample}, we plot the evolution of the radius, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar.
+We assume all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr.
+All stars' radii contract along the pre-main sequence, spinning the stars up (right panel).
+Once the stars reach the main sequence, their radii slowly grow and magnetic braking dominates the stellar angular momentum evolution, significantly spinning-down the stars over long timescales.
+The $r_g$ evolution noticably differs between the stars as the M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
 
-\begin{figure}[h]
+\begin{figure*}[ht]
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{Caption.}%
+   \caption{Stellar radius (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}).
+Each dot denotes the approximate time when each star reaches the main sequence.}%
     \label{fig:stellarExample}%
-\end{figure}
+\end{figure*}
 
@@
 
- Equilibrium tidal models, first introduced by \citep{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega$, and obliquities $\psi$, of the tidally-interacting bodies subject to tidal torques.
-Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that are offset from the line connecting the both bodies' centers of mass, driving torques that exchange angular momentum between the orbit and both bodies' spins.
-Equilibrium tidal models are linear since the tidal waves that comprise the tidal bulge on a body are assumed to be uncoupled.
+ Equilibrium tidal models, first introduced by \citep{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega$, and obliquities $\psi$, of the both gravitating bodies due to tidal torques.
+Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the both bodies' centers of mass due to friction within each body.
+The tidal bulges cause torques that drive the exchange of angular momentum between the orbit and both bodies' spins.
+Equilibrium tidal models are linear since they assume that the tidal waves that comprise the tidal bulge raised on a body are uncoupled.
 Under these assumptions, the tidal evolution is analogous to a driven, damped harmonic oscillator \citep{Greenberg2009}.
 We refer the reader to \citet{Barnes2017} for an in-depth discussion of the assumptions and limitations of equilibrium tidal models.
-Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both exoplanets \citet[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Repetto2014,Fleming2018}.
+Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citet[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Repetto2014,Fleming2018}.
 Here, we consider two equilibrium tidal models to study the secular spin-orbital evolution of low-mass stellar binaries.
 
-\subsubsection{Constant Phase Lag (CPL) Model}
+\subsubsection{Constant Phase Lag Model}
 
@@
 \begin{equation} \label{eqn:cpl:z}
-Z_{i,\mathrm{CPL}} = 3 G^2 k_{2,i} m_j^2 (m_1 + m_2) \frac{R_i^5}{a^9} \frac{1}{n Q_i}
+Z_{i,\mathrm{CPL}} = 3 G^2 k_{2,i} M_j^2 (M_i + M_j) \frac{R_i^5}{a^9} \frac{1}{n Q_i}
 \end{equation}
@@
 
-$\xi_i$ is defined as
+The other intermediate variable, $\xi_i$, is defined as
 \begin{equation}\label{eqn:cpl:chi}
@@
 
-\subsubsection{Constant Time Lag (CTL) Model}
+\subsubsection{Constant Time Lag Model}
 
@@
 However if the assumption of linearity is relaxed, i.e.
-frequencies associate with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}.
+frequencies associate with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}, hence we do not make this assumption and hold the tidal time lag fixed.
 We use the \eqtide implementation of the CTL model in VPLanet following the derivation of \citet{Leconte2010}.
-The equations that govern the secular change in $e$, $a$, $\omega$, and $\psi$ are as follows:
+The equations that govern the secular changes in $e$, $a$, $\omega$, and $\psi$ are as follows:
 
@@
 when P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
-Both the CPL and CTL model, however, predict that tidally-interacting bodies on sufficiently eccentric orbits undergo pseudosynchronous rotation, e.g.
+Both the CPL and CTL model, however, predict that tidally-interacting bodies on sufficiently eccentric orbits undergo pseudosynchronous, or supersynchronous rotation, e.g.
 Mercury's 3:2 spin-orbit resonance \citep[P$_{rot} = 2/3$ P$_{orb}$,][]{GoldreichPeale1966}.
@@
 \begin{equation} \label{eqn:ctl:eqPer}
-P^{\mathrm{CTL}}_{eq} = P\frac{\beta^3 f_5(e) 1 + \cos^2(\psi)}{2f_2(e) \cos(\psi)}.
+P^{\mathrm{CTL}}_{eq} = P_{orb} \frac{\beta^3 f_5(e) (1 + \cos^2(\psi))}{2f_2(e) \cos(\psi)}.
 \end{equation}
-The CTL model in general predicts that bodies on eccentric orbits tidally-lock into supersyncronous rotation, and only bodies on exactly circular orbits are synchronous rotators.
+The CTL model predicts that bodies on eccentric orbits tidally-lock into supersyncronous rotation, and only bodies with aligned spins on circular orbits are synchronous rotators.
 
-Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, the numerical solution can produce unphysical evolution.
-We follow \citet{Barnes2013} and \citet{Fleming2018} and address this issue by fixing P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL model, respectively, when P$_{rot}$ is within $1\%$ of P$_{eq}$.
+Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, numerical solutions for the CPL and CTL models can produce unphysical evolution.
+We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively, when P$_{rot}$ is within $1\%$ of P$_{eq}$.
 We find that this scheme produces physically and numerically accurate results.
@@
 
-Here in Fig.~\ref{fig:tidalExample} we plot the example tidal evolution for $a$, $e$, and $\omega$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively.
-Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increasing as tides forces the spins toward the tidally-locked state, transfering rotational angular momentum into the orbit in the process.
+We plot the tidal evolution for $a$, $e$, and $\omega$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively, in Fig.~\ref{fig:tidalExample}.
+Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increasing as tides force the spins toward the tidally-locked state, transfering rotational angular momentum into the orbit in the process.
 At late times, both the CPL and CTL drive the binaries towards orbital circularization, with tidal dissipation decreasing P$_{orb}$.
@@
 Under the CPL model, the binary tidally-locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
-Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence (see Eq.~(\ref{eqn:ctl:eqPer})).
+Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence, e.g.
+Eq.~(\ref{eqn:ctl:eqPer}).
 
@@
 
-Following \citet{Fleming2018}, when one or both binary stars are tidally locked, we assume that tidal forces prevent magnetic braking from spinning down the tidally locked star(s), but instead, any angular momentum lost is lost at the expense of the binary orbit, decreasing the $a$ as a result \citep{Verbunt1981}.
-Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify Eqs.
-(18) and (20) from \citet{Fleming2018} to account for $r_g$ evolution for the cases when one or both stars tidally lock under conservation of angular momentum, respectively:
+Following \citet{Fleming2018}, when one or both binary stars are tidally locked, we assume that tidal forces prevent magnetic braking from spinning down the tidally-locked star(s), but instead, any angular momentum lost is lost at the expense of the binary orbit, decreasing $a$ as a result \citep{Verbunt1981}.
+Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify the $a$ decay equations due to stellar evolution and magnetic braking in tidally-locked binaries from \citet{Fleming2018}, their Eqs.
+(18) and (20), to additionally account for $r_g$ evolution when one or both stars tidally-lock, respectively, assuming conservation of angular momentum:
 \small
@@
 
-XXX
+We examine stellar angular momentum evolution in low-mass binaries by performing a Monte Carlo study of two sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism.
+We simulate both stars' spin evolution but mainly consider the P$_{rot}$ evolution for the primary, i.e.
+more massive star in binaries, as it is observationally easier to measure a P$_{rot}$ on the more massive, and hence brighter, star \citep[e.g.][]{Meibom2006,Lurie2017}.
+For each simulation, we sample the primary's mass uniformly over $[0.1, 1]$ M$_{\odot}$.
+Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of young stars in the ${\sim}2$ Myr old ONC.
+We compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010,Moe2018}.
+Given the inherent uncertainty in and complexity of the formation of binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$] and uniformly sampling the initial P$_{orb}$ over [$3,100$] d.
+Altough the CTL model is applicable for $e > 0.3$ while the CPL model is not, we restrict $e \leq 0.3$ to allow us to compare both models.
 
-- we consider eccentricities up to 0.3 as the CPL model fails above that.  CTL model can handle it, but at high eccentricity, dynamical tide, e.g. dissipation via waves and stuff, becomes important
+Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain, likely depend on complex viscous evolution within the stars \citep{Ogilvie2007}, differ for stars of the same spectral clase \citep{Barker2009}, and can vary as a function of stellar age \citep{Bolmont2016}.
+Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2009}, however a wide range of values exist in the literature.
+Therefore, we consider a wide range of tidal parameters and sample stellar tidal Qs loguniformly over $[10^4,10^8]$ and $\tau$ loguniformly over $[10^{-3},1]$ s.
 
-%% SECTION: VPLANET RESULTS %%
+We supplement our Monte Carlo study with simulations of interesting cases and simulations over grids of parameters, such as P$_{orb}$ and $Q$ or $\tau$, respectively, using the parameter ranges justified above.
+We describe those simulation setups below in $\S$~\ref{sec:results}.
+
+All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.} with accompanying descriptions and documentation.
+
+%% SECTION: RESULTS %%
 
@@
 
+we can explain fast rotation observed in \citet{Meibom2007} to be due to tidal torques speeding up stellar Prots
+
+Plots:
+
+-Porb vs Q/tau grid for e=0, and vs mass ratio
+
+-average tsync for various orbital period ranges for CPL vs CTL marginalizing over all other parameters. do it with small multiples
+
+-Prot as a function of mass for locked (unlocked) binaries of random ages for gyrochronology results, naturally produces fast rotators
+
+-Plot showing diff between binary relative to single star Prot evolution for various Porbs
+
+-tsync contours over Q/tau and Porb
+
+\subsection{Tidal-Locking Over $Q,\tau$ and P$_{orb}$}
+
 XXX
 
-%\begin{figure*}[ht]
-%	\includegraphics[width=\textwidth]{../Plots/kepler_two_pop_box_dist.pdf}
-%   \caption{Identical format to Fig.~\ref{fig:kepler1}, but displaying simulated data from our two population model (see $\S$~\ref{sec:results:kepler_two_pop}).
-Our two population model reproduces the observed \textit{Kepler} P$_{rot}$ distribution, especially the fast rotator population, with a total median P$_{rot} = 19.6$ d compared to the \textit{Kepler} median P$_{rot} = 17.3$ d, a close match.
-In the right panel for $ 0.4 \textrm{M}_{\odot} \lsim \textrm{M} \lsim 1.2 \textrm{M}_{\odot}$, the inclusion of tidally-interacting binaries in the total (blue) sample leads to an excellent match with the \textit{Kepler} distribution (red) compared to the single-star only (orange) model, emphasizing the importance of including stellar binaries in P$_{rot}$ forward models.}%
-%    \label{fig:kepler_two_pop}%
-%\end{figure*}
+\subsection{Tidal-Locking as a Function of Age} \label{sec:results:age}
+
+XXX
+
+\subsubsection{Tidal-Locking as a Function of P$_{orb}$} \label{sec:results:porb}
+
+XXX
+
+\begin{figure}[h]
+	\includegraphics[width=0.4\textwidth]{../Plots/tidalExample.pdf}
+   \caption{Tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary's $e$ (top), P$_{orb}$ (middle) and P$_{rot}$ (bottom) for the CPL (blue) and CTL (orange) model.
+The blue (CPL) and orange (CTL) vertical dashed lines denote when the stellar binary tidally-locks.
+Both the CPL and CTL model predict the same qualitative evolution.
+The rotational evolution differs, however, as under the CPL model, the binary tidally-locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
+Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence (see Eq.~(\ref{eqn:ctl:eqPer})).}%
+    \label{fig:tidalExample}%
+\end{figure}
+
+\subsubsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology}
+
+Here we examine the tidal locking in binaries using the simulations from XXX by plotting P$_{rot}$ as a function of mass and age for binaries in XXX.
+We find that $35.5\%$ of binaries tidally lock before the age of the system, typically for tidal $Q \lsim 10^7$.
+Tidally-locked binaries tend to rotate more rapidly than unlocked binaries, reflected in the increased density for P$_{rot} \lsim 20$ d in the marginal distribution for the locked population and in the distribution medians (locked median P$_{rot} = 19.8$ d compared to unlocked median P$_{rot} = 27.5$ d).
+Binaries readily tidally-lock at short P$_{orb}$ due to the tidal force's strong semi-major axis dependence, increasing the strength of tidal torques at shorter orbital separations.
+When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} = $P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.
+
+Tidal forces modify the spins of unlocked binaries as well, tending to spin-up stars against the effects of magnetic braking towards orbital synchronization, reflected in unlocked binaries' median P$_{rot} = 27.5$ d, compared to the single star median P$_{rot} = 30.6$ d.
+Tidal locking is not limited to P$_{orb} \lsim 10$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 70$ d, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars (compare to XXX).
+This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find that binaries can tidally lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
+The observed age distribution of tidally-locked binaries has little dependence on P$_{rot}$, with both young and old systems tidally-locking at long and short P$_{rot}$, causing the predictions of gyrochronology models to fail in such systems.
+
+%\begin{figure}[h]
+%	\includegraphics[width=0.45\textwidth]{../Plots/locked_kepler.pdf}
+%   \caption{Simulated distribution of P$_{rot}$ for tidally-locked (cross) and unlocked (circle) stellar binaries, colored by age for Model 2 (see $\S$~\ref{sec:results:kepler2}).
+Left: P$_{rot}$ as a function of stellar mass.
+Right: Marginalized P$_{rot}$ distribution for locked (blue) and unlocked (orange) binaries.
+Tidally-locked binaries rotate more rapidly than unlocked binaries, displaying an increased density in the P$_{rot}$ distribution for P$_{rot} {\lsim} 20$ d, and with a median P$_{rot} = 19.8$ d, compared to the unlocked median P$_{rot} = 27.5$ d.
+Unlocked binaries on average rotate more quickly than single stars (compare to single distribution in Fig.~\ref{fig:kepler2}) due to the influence of tidal torques.}
+%    \label{fig:locking}
+%\end{figure}
 

### 2018-11-02 Analysis/locking.py
adding tidal locking plots

@@ -1,0 +1,170 @@
+"""
+
+@author: David P Fleming, University of Washington, Nov 2018
+@email dflemin3 (at) uw (dot) edu
+
+Plot to show which binaries are tidally locked in mass-Prot space by the age of
+the system for both the CPL and CTL models.  We randomly assigned ages to systems
+from a uniform distribution over 1-7 Gyr, broadly consistent with stellar ages
+in the Kepler field.
+
+Script Output:
+
+CPL locked Prot median: 28.989571500000004
+CPL unlocked Prot median: 40.65482
+CPL fraction of locked binaries: 0.5742
+
+CTL locked Prot median: 25.438067999999998
+CTL unlocked Prot median: 40.002216
+CTL fraction of locked binaries: 0.5357
+
+"""
+
+import numpy as np
+import os
+import pandas as pd
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+ylims = [0, 100]
+bins = "auto"
+num = 2500
+seed = 42
+
+np.random.seed(seed)
+
+# Load in data
+cpl = pd.read_csv("../Data/mcCPL.csv")
+ctl = pd.read_csv("../Data/mcCTL.csv")
+
+# Create indices for a random sample of num points to make scatterplot legible
+inds = np.random.choice(np.arange(len(cpl)), size=num, replace=False)
+
+# Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
+cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
+cpl["Sec_LockTime"][cpl["Sec_LockTime"] < 0] = 7.0e9
+ctl["Pri_LockTime"][ctl["Pri_LockTime"] < 0] = 7.0e9
+ctl["Sec_LockTime"][ctl["Sec_LockTime"] < 0] = 7.0e9
+
+# Make flag for binaries who are locked by the age of the system
[…]

### 2018-11-02 Analysis/lurie.py

@@ -95,4 +95,28 @@ cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Eccentricity")
 
 fig.savefig("../Plots/lurie.pdf", bbox_inches="tight", dpi=600)
+
+### Now make marginal distribution plot of Porb/Prot for CPL, CTL, and Lurie
+fig, ax = plt.subplots()
+
+bins = np.logspace(np.log10(0.5), np.log10(5), 25)
+
+ax.hist(cpl["Age_Porb"]/cpl["Pri_ProtAge"], bins=bins, histtype="step", lw=2.5,
+        color="C0", density=True, label="CPL",
+        range=[np.log10(0.5), np.log10(5)])
+ax.hist(ctl["Age_Porb"]/ctl["Pri_ProtAge"], bins=bins, histtype="step", lw=2.5,
+        color="C1", density=True, label="CTL",
+        range=[np.log10(0.5), np.log10(5)])
+ax.hist(lurie["Porb"]/lurie["Prot"], bins=bins, histtype="step", lw=2.5,
+        color="red", density=True, label="Lurie et al. (2017)",
+        range=[np.log10(0.5), np.log10(5)])
+
+# Format, save
+ax.set_ylabel("Normalized Counts", fontsize=20)
+ax.set_xlabel("P$_{orb}$ / P$_{rot}$", fontsize=25)
+ax.set_xscale("log")
+ax.legend(loc="best", framealpha=0, fontsize=18)
+fig.tight_layout()
+fig.savefig("../Plots/lurieHist.pdf", bbox_inches="tight", dpi=600)
+
 # Done!

### 2018-11-02 Analysis/qTauPorbHist.py

@@ -19,7 +19,7 @@ mpl.rcParams['font.size'] = 20.0
 cmap = "inferno_r"
 
 ## for Palatino and other serif fonts use:
-mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
+mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data

### 2018-11-02 Analysis/ratioPorbHist.py

@@ -20,7 +20,7 @@ mpl.rcParams['font.size'] = 20.0
 cmap = "inferno_r"
 
 ## for Palatino and other serif fonts use:
-mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
+mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data

### 2018-11-02 Paper/sync.tex
Updates from ShareLaTeX

@@
 
-Abstract
+-tidally locking before 5 Gyr occurs for most binaries below orbital periods of 60-70 days regardless of tidal parameter
+
+-compute cumulative fraction of tidal-locking times in Porb bins plot so I can say x fraction are stars with Porb $< 60$d are tidally locked.
+
+-make locking.py plot from previous paper with new data
+
+- 5 points with Porb < 10 days and Porb/Prot > 10 in lurie plot are likely young stars that have or continue to spin up via angular momentum conservation due to pre main sequence contraction.
+evidence for young stars in the Kepler field, e.g.
+McQuillan+2014, Matt+2015, Davenport+2017,2018 support this theory
+
+- Lurie has some binaries with Porb/Prot = 0.9, subsynchronous rotators.  cite his explaination for them
+
+- Lurie finds tons of short Porb EBs because transit observational biases favor bodies with short orbital periods for a finite observing window, eg Kepler's lifetime
 
@@
 We refer the reader to \citet{Barnes2017} for an in-depth discussion of the assumptions and limitations of equilibrium tidal models.
-Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citet[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Repetto2014,Fleming2018}.
+Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citet[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
 Here, we consider two equilibrium tidal models to study the secular spin-orbital evolution of low-mass stellar binaries.
@@
 \begin{figure}[h]
-	\includegraphics[width=0.4\textwidth]{../Plots/tidalExample.pdf}
-   \caption{Tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary's $e$ (top), P$_{orb}$ (middle) and P$_{rot}$ (bottom) for the CPL (blue) and CTL (orange) model.
-The blue (CPL) and orange (CTL) vertical dashed lines denote when the stellar binary tidally-locks.
-Both the CPL and CTL model predict the same qualitative evolution.
-The rotational evolution differs, however, as under the CPL model, the binary tidally-locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
-Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence (see Eq.~(\ref{eqn:ctl:eqPer})).}%
-    \label{fig:tidalExample}%
+	\includegraphics[width=0.4\textwidth]{../Plots/lockTimePorbHist.pdf}
+   \caption{Caption.}%
+    \label{fig:lockHistPorb}%
 \end{figure}

### 2018-11-09 Analysis/TidalQTests.py
adding lots of plots and scripts

@@ -1,0 +1,95 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+import os
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
+mpl.rc('text', usetex=True)
+
+# Read in Data
+path = "../Sims/TidalLocking"
+
+dirs = [""]
+
+# Note the output order
+# saOutputOrder   Time -RotPer -OrbPer Ecce -TotEn -TotAngMom
+tides_50_dirs = ["50_4","50_5","50_6","50_7"]
+labels = ["$Q = 10^4$", "$Q = 10^5$", "$Q = 10^6$", "$Q = 10^7$"]
+colors = ["C%d" % ii for ii in range(len(tides_50_dirs))]
+
+# Extract data
+notides = np.genfromtxt(os.path.join(path,"NoTides","notides.secondary.forward"))
+
+tides_50 = []
+
+for direct in tides_50_dirs:
+    tides_50.append(np.genfromtxt(os.path.join(path,direct,"tides.secondary.forward")))
+
+# Path to sim results
+cplDir = "../Sims/StellarTidalExample/CPL"
+ctlDir = "../Sims/StellarTidalExample/CTL"
+
+# When each system tidally locked in yr
+cplLock = 2.855941e+07
+ctlLock = 3.126642e+07
+
+# Load data
+# saOutputOrder	Time -TotEn -TotAngMom -Semim -Radius -RotPer Ecce -RotRate -MeanMotion -OrbPer RadGyra -SurfEnFluxTotal
+cpl = np.genfromtxt(os.path.join(cplDir,"bintide.secondary.forward"))
+ctl = np.genfromtxt(os.path.join(ctlDir,"bintide.secondary.forward"))
+
+fig = plt.figure(figsize=(19,6))
+gs = GridSpec(1, 5, width_ratios=[1, 0.01, 1, 0.2, 1])
+
+# Left panel: Ecc
+ax0 = fig.add_subplot(gs[0])
+
+# Left: e
[…]

### 2018-11-09 Analysis/cmap.py

@@ -1,0 +1,60 @@
+"""
+Shifted colormap code from Paul H. on Stack Overflow
+https://stackoverflow.com/questions/7404116/defining-the-midpoint-of-a-colormap-in-matplotlib
+"""
+
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import AxesGrid
+
+def shiftedColorMap(cmap, start=0, midpoint=0.5, stop=1.0, name='shiftedcmap'):
+    '''
+    Function to offset the "center" of a colormap. Useful for
+    data with a negative min and positive max and you want the
+    middle of the colormap's dynamic range to be at zero.
+
+    Input
+    -----
+      cmap : The matplotlib colormap to be altered
+      start : Offset from lowest point in the colormap's range.
+          Defaults to 0.0 (no lower offset). Should be between
+          0.0 and `midpoint`.
+      midpoint : The new center of the colormap. Defaults to
+          0.5 (no shift). Should be between 0.0 and 1.0. In
+          general, this should be  1 - vmax / (vmax + abs(vmin))
+          For example if your data range from -15.0 to +5.0 and
+          you want the center of the colormap at 0.0, `midpoint`
+          should be set to  1 - 5/(5 + 15)) or 0.75
+      stop : Offset from highest point in the colormap's range.
+          Defaults to 1.0 (no upper offset). Should be between
+          `midpoint` and 1.0.
+    '''
+    cdict = {
+        'red': [],
+        'green': [],
+        'blue': [],
+        'alpha': []
+    }
+
+    # regular index to compute the colors
+    reg_index = np.linspace(start, stop, 257)
+
+    # shifted index to match the data
+    shift_index = np.hstack([
+        np.linspace(0.0, midpoint, 128, endpoint=False),
+        np.linspace(midpoint, 1.0, 129, endpoint=True)
+    ])
+
+    for ri, si in zip(reg_index, shift_index):
+        r, g, b, a = cmap(ri)
+
+        cdict['red'].append((si, r, r))
+        cdict['green'].append((si, g, g))
+        cdict['blue'].append((si, b, b))
+        cdict['alpha'].append((si, a, a))
+
+    newcmap = matplotlib.colors.LinearSegmentedColormap(name, cdict)
+    plt.register_cmap(cmap=newcmap)
+
+    return newcmap

### 2018-11-09 Analysis/eccPorbHist.py

@@ -16,15 +16,15 @@ from matplotlib.gridspec import GridSpec
 
 #Typical plot parameters that make for pretty plots
 mpl.rcParams['font.size'] = 20.0
-cmap = "inferno_r"
+cmap = "RdBu"
 
 ## for Palatino and other serif fonts use:
 mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPL.csv")
-ctl = pd.read_csv("../Data/mcCTL.csv")
+cpl = pd.read_csv("../Data/mcCPLTorque.csv")
+ctl = pd.read_csv("../Data/mcCTLTorque.csv")
 
 # Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
 cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
@@ -38,12 +38,12 @@ porbBinEdges = np.linspace(0, 100, num)
 eccBinEdges = np.linspace(0, 0.3, num)
 
 # Array for joint locking time distribution
-cplJoint = np.zeros((num-1, num-1))
-ctlJoint = np.zeros_like(cplJoint)
+cplJointPeq = np.zeros((num-1, num-1))
+ctlJointPeq = np.zeros_like(cplJointPeq)
 
 # Array for joint Prot/Porb distribution
-cplJointPeq = np.zeros_like(cplJoint)
-ctlJointPeq = np.zeros_like(cplJoint)
+cplJointRatio = np.zeros_like(cplJointPeq)
+ctlJointRatio = np.zeros_like(cplJointPeq)
 
 # Loop over bins, compute median tidal locking time in joing Porb/ecc bins
 for ii in range(num-1):
@@ -58,12 +58,12 @@ for ii in range(num-1):
         ctlMask = ctlMask & (ctl["Final_Ecc"] > eccBinEdges[jj])
         ctlMask = ctlMask & (ctl["Final_Ecc"] < eccBinEdges[jj + 1])
 
-        # Populate array elements
-        cplJoint[ii, jj] = np.median(cpl["Pri_LockTime"][cplMask].values/1.0e9)
-        ctlJoint[ii, jj] = np.median(ctl["Pri_LockTime"][ctlMask].values/1.0e9)
+        # Populate array elements ctl["Pri_ProtFinal"][ctlMask]
[…]

### 2018-11-09 Analysis/eqPer.py

@@ -1,0 +1,145 @@
+"""
+
+@author: David P. Fleming, University of Washington, Oct 2018
+@email: dflemin3 (at) uw (dot) edu
+
+This script plots stellar-tidal evolution
+
+"""
+
+import numpy as np
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+# Path to sim results
+cplDir = "../Sims/CPLTest"
+ctlDir = "../Sims/CTLTest"
+logfile_name = "bintides.log"
+dirs = ["5", "10", "20", "30", "40", "50", "60"]
+colors = ["C0", "C1", "C2", "C3", "C4", "C5", "C6"]
+YEARSEC = 3.154e+7 # Seconds per year
+DAYSEC = 24.0*3600.0 # Seconds per day
+
+### Two Panel Plot ###
+
+fig, axes = plt.subplots(ncols=2, figsize=(14, 6))
+
+for ii, dir in enumerate(dirs):
+    # Load data
+    # saOutputOrder	Time -TotEn -TotAngMom -Radius -RotPer -EqRotPer DRotPerDtEqtide DRotPerDtStellar Ecce -OrbPer RadGyra -SurfEnFluxTotal
+    cpl = np.genfromtxt(os.path.join(cplDir, dir, "bintides.secondary.forward"))
+    ctl = np.genfromtxt(os.path.join(ctlDir, dir, "bintides.secondary.forward"))
+
+    # Pull LockTimes out of logfile
+    with open(os.path.join(os.path.join(cplDir, dir, logfile_name)), 'r') as f:
+        logfile_in = f.read()
+
+    matches = []
+    for line in logfile_in.split("\n"):
+        if line.startswith("(LockTime)"):
+            matches.append((line.split()[-1]))
+
+    # Convert lock times to years
+    dLockTimeCPL = float(matches[3])/YEARSEC
+
+    # Pull LockTimes out of logfile
[…]

### 2018-11-09 Analysis/lockTest.py

@@ -1,0 +1,75 @@
+"""
+
+@author: David P. Fleming, University of Washington, Oct 2018
+@email: dflemin3 (at) uw (dot) edu
+
+Examing our tidal locking prescription in the weak-tides regime (large Porb
+and large Q/small tau).
+
+"""
+
+import numpy as np
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+dir = "../Sims/LockingTest"
+
+# saOutputOrder	Time -Radius -RotPer -RotRate -LostAngMom RadGyra -SurfEnFluxTotal
+output = "bintide.primary.forward"
+
+fig, ax = plt.subplots()
+
+### CPL PLOTS ###
+
+# Plot Qs
+data = np.genfromtxt(os.path.join(dir, "q7", output))
+ax.plot(data[:,0]/1.0e9, data[:,2], lw=3, ls="-", color="C0", label="$Q = 10^7$")
+
+data = np.genfromtxt(os.path.join(dir, "q8", output))
+ax.plot(data[:,0]/1.0e9, data[:,2], lw=3, ls="-", color="C1", label="$Q = 10^8$")
+
+# Plot Qs: No Locking
+data = np.genfromtxt(os.path.join(dir, "q7NoLock", output))
+ax.plot(data[:,0]/1.0e9, data[:,2], lw=3, ls="--", color="C0",
+        label="$Q = 10^7$, No Lock")
+
+data = np.genfromtxt(os.path.join(dir, "q8NoLock", output))
+ax.plot(data[:,0]/1.0e9, data[:,2], lw=3, ls="--", color="C1",
+        label="$Q = 10^8$, No Lock")
+
+# Plot taus
+data = np.genfromtxt(os.path.join(dir, "tau0_1", output))
+ax.plot(data[:,0]/1.0e9, data[:,2], lw=3, ls="-", color="C2", label=r"$\tau = 0.1$ s")
+
+data = np.genfromtxt(os.path.join(dir, "tau0_01", output))
[…]

### 2018-11-09 Analysis/locking.py

@@ -10,13 +10,14 @@ in the Kepler field.
 
 Script Output:
 
-CPL locked Prot median: 28.989571500000004
-CPL unlocked Prot median: 40.65482
-CPL fraction of locked binaries: 0.5742
+CPL locked Prot median: 27.049087
+CPL unlocked Prot median: 42.744551
+CPL fraction of locked binaries: 0.2518
 
-CTL locked Prot median: 25.438067999999998
-CTL unlocked Prot median: 40.002216
-CTL fraction of locked binaries: 0.5357
+CTL locked Prot median: 4.683883
+CTL unlocked Prot median: 42.3556655
+CTL fraction of locked binaries: 0.0202
+
 
 """
 
@@ -43,8 +44,8 @@ seed = 42
 np.random.seed(seed)
 
 # Load in data
-cpl = pd.read_csv("../Data/mcCPL.csv")
-ctl = pd.read_csv("../Data/mcCTL.csv")
+cpl = pd.read_csv("../Data/mcCPLTorque.csv")
+ctl = pd.read_csv("../Data/mcCTLTorque.csv")
 
 # Create indices for a random sample of num points to make scatterplot legible
 inds = np.random.choice(np.arange(len(cpl)), size=num, replace=False)

### 2018-11-09 Analysis/lurie.py

@@ -21,9 +21,11 @@ mpl.rcParams['font.size'] = 22.0
 mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
+plotLurie = True
+
 # Load data
-cpl = pd.read_csv("../Data/mcCPL.csv")
-ctl = pd.read_csv("../Data/mcCTL.csv")
+cpl = pd.read_csv("../Data/mcCPLTorque.csv")
+ctl = pd.read_csv("../Data/mcCTLTorque.csv")
 
 lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
                     names=["Porb", "Prot", "Ecc"])
@@ -33,6 +35,8 @@ cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
 cpl["Sec_LockTime"][cpl["Sec_LockTime"] < 0] = 7.0e9
 ctl["Pri_LockTime"][ctl["Pri_LockTime"] < 0] = 7.0e9
 ctl["Sec_LockTime"][ctl["Sec_LockTime"] < 0] = 7.0e9
+
+# 1st Fig modeled after Fig. 6, upper panel from Lurie+2017
 
 fig = plt.figure(figsize=(19, 8))
 gs = GridSpec(1, 5, width_ratios=[1, 0.05, 1, 0.01, 0.075], wspace=0.05)
@@ -45,12 +49,13 @@ im = ax0.scatter(cpl["Age_Porb"], cpl["Age_Porb"]/cpl["Pri_ProtAge"],
                  c=cpl["Age_Ecc"].values, cmap="viridis", zorder=1,
                  s=40, marker="o", vmin=0, vmax=0.3, label="Simulated")
 
-# Plot Lurie+2017 data
-ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=50, zorder=2,
-            marker="x", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
+if plotLurie:
+    # Plot Lurie+2017 data
+    ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=100, zorder=2,
+                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
-# Plot Lurie+2017 detection limit of 45 d
-ax0.plot([1.0, 100.0], [1.0/45.0, 100.0/45.0], ls="-", lw="5", color="k", zorder=3)
[…]

### 2018-11-09 Analysis/qTauPorbHist.py

@@ -13,18 +13,22 @@ import pandas as pd
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 from matplotlib.gridspec import GridSpec
+from cmap import shiftedColorMap
 
 #Typical plot parameters that make for pretty plots
 mpl.rcParams['font.size'] = 20.0
-cmap = "inferno_r"
+
+# Define a shifted colormap so 1 is at the center
+midpoint = 1 - 1.2614218627451428 / (1.2614218627451428 + 0.2518443643281677)
+cmap = shiftedColorMap(mpl.cm.RdBu, midpoint=midpoint)
 
 ## for Palatino and other serif fonts use:
 mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPL.csv")
-ctl = pd.read_csv("../Data/mcCTL.csv")
+cpl = pd.read_csv("../Data/mcCPLTorque.csv")
+ctl = pd.read_csv("../Data/mcCTLTorque.csv")
 
 # Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
 cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
@@ -37,15 +41,15 @@ cpl["Pri_dTidalQ"] = pd.Series(cpl["Pri_dTidaLQ"].values, index=cpl.index)
 num = 11
 porbBinEdges = np.linspace(0, 100, num)
 qBinEdges = np.logspace(4, 8, num)
-tauBinEdges = np.logspace(-3, 0, num)
+tauBinEdges = np.logspace(-3, 0, num)[::-1]
 
 # Array for joint locking time distribution
-cplJoint = np.zeros((num-1, num-1))
-ctlJoint = np.zeros_like(cplJoint)
+cplJointPeq = np.zeros((num-1, num-1))
+ctlJointPeq = np.zeros_like(cplJointPeq)
 
 # Array for joint Prot/Porb distribution
-cplJointPeq = np.zeros_like(cplJoint)
-ctlJointPeq = np.zeros_like(cplJoint)
+cplJointRatio = np.zeros_like(cplJointPeq)
+ctlJointRatio = np.zeros_like(cplJointPeq)
 
 # Loop over bins, compute median tidal locking time in joing Porb/ecc bins
 for ii in range(num-1):
[…]

### 2018-11-09 Analysis/ratioPorbHist.py

@@ -17,15 +17,15 @@ from matplotlib.gridspec import GridSpec
 
 #Typical plot parameters that make for pretty plots
 mpl.rcParams['font.size'] = 20.0
-cmap = "inferno_r"
+cmap = "RdBu"
 
 ## for Palatino and other serif fonts use:
 mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPL.csv")
-ctl = pd.read_csv("../Data/mcCTL.csv")
+cpl = pd.read_csv("../Data/mcCPLTorque.csv")
+ctl = pd.read_csv("../Data/mcCTLTorque.csv")
 
 # Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
 cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
@@ -43,12 +43,12 @@ porbBinEdges = np.linspace(0, 100, num)
 ratioBinEdges = np.linspace(0.1, 1.0, num)
 
 # Array for joint locking time distribution
-cplJoint = np.zeros((num-1, num-1))
-ctlJoint = np.zeros_like(cplJoint)
+cplJointPeq = np.zeros((num-1, num-1))
+ctlJointPeq = np.zeros_like(cplJointPeq)
 
 # Array for joint Prot/Porb distribution
-cplJointPeq = np.zeros_like(cplJoint)
-ctlJointPeq = np.zeros_like(cplJoint)
+cplJointRatio = np.zeros_like(cplJointPeq)
+ctlJointRatio = np.zeros_like(cplJointPeq)
 
 # Loop over bins, compute median tidal locking time in joing Porb/mu bins
 for ii in range(num-1):
@@ -63,12 +63,12 @@ for ii in range(num-1):
         ctlMask = ctlMask & (ctl["MassRatio"] > ratioBinEdges[jj])
         ctlMask = ctlMask & (ctl["MassRatio"] < ratioBinEdges[jj + 1])
 
-        # Populate array elements
-        cplJoint[ii, jj] = np.median(cpl["Pri_LockTime"][cplMask].values/1.0e9)
-        ctlJoint[ii, jj] = np.median(ctl["Pri_LockTime"][ctlMask].values/1.0e9)
+        # Populate array elements ctl["Pri_ProtFinal"][ctlMask]
[…]

### 2018-11-09 Analysis/stellar.py

@@ -17,10 +17,10 @@ import matplotlib.pyplot as plt
 
 #Typical plot parameters that make for pretty plots
 mpl.rcParams['figure.figsize'] = (9,8)
-mpl.rcParams['font.size'] = 25.0
+mpl.rcParams['font.size'] = 30.0
 
 ## for Palatino and other serif fonts use:
-mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
+mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Read in output files
@@ -30,6 +30,9 @@ path = "../Sims/StellarEvolution/"
 g = np.genfromtxt(os.path.join(path,"stellar.g.forward"))
 k = np.genfromtxt(os.path.join(path,"stellar.k.forward"))
 m = np.genfromtxt(os.path.join(path,"stellar.m.forward"))
+gr = np.genfromtxt(os.path.join(path,"stellar.gr.forward"))
+kr = np.genfromtxt(os.path.join(path,"stellar.kr.forward"))
+mr = np.genfromtxt(os.path.join(path,"stellar.mr.forward"))
 
 # All on same time grid
 time = g[:,0]
@@ -55,7 +58,7 @@ ax[0].scatter(time[ind_k], k[ind_k,1], s=75, color="C1")
 ax[0].scatter(time[ind_m], m[ind_m,1], s=75, color="C2")
 
 # Format
-ax[0].legend()
+ax[0].legend(loc="best", framealpha=0, fontsize=25)
 ax[0].set_ylabel("Radius [R$_{\odot}$]")
 ax[0].set_xlabel("$t - t_0$ [yr]")
 
@@ -88,6 +91,18 @@ ax[2].scatter(time[ind_g], g[ind_g,2], s=75, color="C0")
 ax[2].scatter(time[ind_k], k[ind_k,2], s=75, color="C1")
 ax[2].scatter(time[ind_m], m[ind_m,2], s=75, color="C2")
 
+ax[2].plot(time, gr[:,2], lw=2.5, color="C0", ls="--",)
+ax[2].plot(time, kr[:,2], lw=2.5, color="C1", ls="--",)
+ax[2].plot(time, mr[:,2], lw=2.5, color="C2", ls="--",)
+
+# Plot points to indicate ZAMS
+ax[2].scatter(time[ind_g], gr[ind_g,2], s=75, color="C0")
[…]

### 2018-11-09 Analysis/stellarTidalExample.py

@@ -1,0 +1,69 @@
+"""
+
+@author: David P. Fleming, University of Washington, Oct 2018
+@email: dflemin3 (at) uw (dot) edu
+
+This script plots stellar-tidal evolution
+
+"""
+
+import numpy as np
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+# Path to sim results
+cplDir = "../Sims/StellarTidalExample/CPL"
+ctlDir = "../Sims/StellarTidalExample/CTL"
+
+# When each system tidally locked in yr
+cplLock = 2.855941e+07
+ctlLock = 3.126642e+07
+
+# Load data
+# saOutputOrder	Time -TotEn -TotAngMom -Semim -Radius -RotPer Ecce -RotRate -MeanMotion -OrbPer RadGyra -SurfEnFluxTotal
+cpl = np.genfromtxt(os.path.join(cplDir,"bintide.secondary.forward"))
+ctl = np.genfromtxt(os.path.join(ctlDir,"bintide.secondary.forward"))
+
+# Plot e, Porb, Prot
+fig, axes = plt.subplots(ncols=3, figsize=(16,6), sharex=True)
+
+# Left: e
+axes[0].plot(cpl[:,0], cpl[:,6], lw=3, ls="-", color="C0", label="CPL")
+axes[0].plot(ctl[:,0], ctl[:,6], lw=3, ls="-", color="C1", label="CTL")
+
+axes[0].set_ylabel("Eccentricity")
+axes[0].legend(loc="best", framealpha=0)
+
+# Middle: Porb
+axes[1].plot(cpl[:,0], cpl[:,9], lw=3, ls="-", color="C0", label="CPL")
+axes[1].plot(ctl[:,0], ctl[:,9], lw=3, ls="-", color="C1", label="CTL")
+
+axes[1].set_ylabel("Orbital Period [d]")
+
+# Right: Prot
+axes[2].plot(cpl[:,0], cpl[:,5], lw=3, ls="-", color="C0", label="CPL")
[…]

### 2018-11-09 Analysis/subsync.py

@@ -1,0 +1,146 @@
+"""
+
+@author: David P. Fleming, University of Washington, Nov 2018
+@email: dflemin3 (at) uw (dot) edu
+
+Comparison of simulations with Lurie+2017 Porb <= 10 d population, with a focus
+on the subsynchronous rotators.
+
+Script output:
+
+Median subsync tidal tau: 0.024179478659999996
+Median supersync tidal tau: 0.0412718247
+Median subsync ecc: 0.13616799999999998
+Median supersync ecc: 0.1525495
+Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.44291609353507566
+Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.08665749656121045
+
+"""
+
+import numpy as np
+import os
+import pandas as pd
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+plotLurie = True
+
+# Load data
+cpl = pd.read_csv("../Data/mcCPLTorque.csv")
+ctl = pd.read_csv("../Data/mcCTLTorque.csv")
+
+lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
+                    names=["Porb", "Prot", "Ecc"])
+
+# Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
+cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
+cpl["Sec_LockTime"][cpl["Sec_LockTime"] < 0] = 7.0e9
+ctl["Pri_LockTime"][ctl["Pri_LockTime"] < 0] = 7.0e9
+ctl["Sec_LockTime"][ctl["Sec_LockTime"] < 0] = 7.0e9
+
+### 1st Fig modeled after Fig. 7 from Lurie+2017, sub zoomed in on subsync population ###
+
+fig = plt.figure(figsize=(9, 8))
+gs = GridSpec(1, 3, width_ratios=[1, 0.01, 0.075], wspace=0.05)
+
[…]

### 2018-11-09 Analysis/tauToQ.py

@@ -1,0 +1,70 @@
+"""
+
+@author: David P. Fleming, University of Washington, Oct 2018
+@email: dflemin3 (at) uw (dot) edu
+
+Determine approximate tidal Q for circular, non-synchronous orbit using Eqn. 18
+from Leconte+2010 and Eqn. 19 for nearly synchronous orbits
+
+"""
+
+import numpy as np
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+
+DAYSEC = 86400.0 # Seconds in a day
+
+def leconte18(w, tau, n=(2.0*np.pi/50.0)):
+    """
+    w : float
+        rotation rate in 1/days
+    tau : float
+        tidal tau in s
+    """
+    return 1.0/(2.0*tau/DAYSEC*np.fabs(w-n))
+# end function
+
+def leconte19(tau, n=(2.0*np.pi/50.0)):
+    """
+    tau : float
+        tidal tau in s
+    """
+    return 1.0/(tau/DAYSEC*n)
+# end function
+
+dir = "../Sims/LockingTest"
+
+# saOutputOrder	Time -Radius -RotPer -RotRate -LostAngMom RadGyra -SurfEnFluxTotal
+output = "bintide.primary.forward"
+
+fig, ax = plt.subplots()
+
+data = np.genfromtxt(os.path.join(dir, "tau0_1NoLock", output))
+qeffNoSync = leconte18(data[:,3], 0.1)
+qeffSync = leconte19(0.1)*np.ones(len(data))
+ax.plot(data[:,0], qeffNoSync, lw=3, ls="-", color="C0", label=r"$\tau = 0.1$ s Non-Sync")
+ax.plot(data[:,0], qeffSync, lw=3, ls="--", color="C0", label=r"$\tau = 0.1$ s Near Sync")
+
+data = np.genfromtxt(os.path.join(dir, "tau0_01NoLock", output))
+qeffNoSync = leconte18(data[:,3], 0.01)
+qeffSync = leconte19(0.01)*np.ones(len(data))
[…]

### 2018-11-09 Scripts/processMCCPL.py

@@ -1,0 +1,163 @@
+# -*- coding: utf-8 -*-
+"""
+
+@author: David P. Fleming, University of Washington, Seattle
+@email: dflemin3 (at) uw (dot) edu
+Oct 2018
+
+This script parses data from an ensemble of VPLanet CPL/CTL coupled stellar-tidal
+simulations.
+
+"""
+
+import os
+import re
+import numpy as np
+import pandas as pd
+
+# Constants and names
+primary_name = "primary.in"
+secondary_name = "secondary.in"
+logfile_name = "bintides.log"
+primary_out_name = "bintides.primary.forward"
+secondary_out_name = "bintides.secondary.forward"
+YEARSEC = 3.154e+7 # Seconds in a year
+dLowAge = 1.0e9 # Low age limit of 1 Gyr
+dHighAge = 7.0e9 # High limit of 7 Gyr
+PATH = os.path.dirname(os.path.realpath(__file__)) # Run in directory where sim dirs live
+dirs = [] # Container to hold directory names
+ind = None
+
+# Find all simulation directories
+with os.scandir(PATH) as it:
+    print("Finding all simulation directories...")
+    for entry in it:
+        # Only consider simulation directories, ignoring hidden directories
+        if not entry.name.startswith(".") and os.path.isdir(entry.name):
+            dirs.append(entry.name)
+
+# List to store data rows
+table = []
+
+# Extract data in directory
+for ii, directory in enumerate(dirs):
+    print("Parsing directory %d..." % ii)
+
+    # List to hold this sim's data (row in df)
+    tmp = []
+
+    # Read in mass, tidal Q from primary input file
+    with open(os.path.join(directory, primary_name), 'r') as f:
+        primary_in = f.read()
+
+        # Find the line where dMass lives, remove whitespace, make it a float
+        dMass = re.findall('%s(.*?)#' % 'dMass', primary_in)[0]
+        dMass = float("".join(dMass.split()))
+        tmp.append(dMass)
+
+        # Find the line where dTidalQ lives, remove whitespace, make it a float
+        dTidalQ = re.findall('%s(.*?)#' % 'dTidalQ', primary_in)[0]
+        dTidalQ = float("".join(dTidalQ.split()))
+        tmp.append(dTidalQ)
+
+    # Read in mass from secondary input file
[…]

### 2018-11-09 Scripts/processMCCTL.py

@@ -46,7 +46,7 @@ for ii, directory in enumerate(dirs):
     # List to hold this sim's data (row in df)
     tmp = []
 
-    # Read in mass, tidal Q from primary input file
+    # Read in mass, tidal tau from primary input file
     with open(os.path.join(directory, primary_name), 'r') as f:
         primary_in = f.read()
 
@@ -55,10 +55,10 @@ for ii, directory in enumerate(dirs):
         dMass = float("".join(dMass.split()))
         tmp.append(dMass)
 
-        # Find the line where dTidalQ lives, remove whitespace, make it a float
-        dTidalQ = re.findall('%s(.*?)#' % 'dTidalQ', primary_in)[0]
-        dTidalQ = float("".join(dTidalQ.split()))
-        tmp.append(dTidalQ)
+        # Find the line where dTidalTau lives, remove whitespace, make it a float
+        dTidalTau = re.findall('%s(.*?)#' % 'dTidalTau', primary_in)[0]
+        dTidalTau = float("".join(dTidalTau.split()))*YEARSEC
+        tmp.append(dTidalTau)
 
     # Read in mass from secondary input file
     with open(os.path.join(directory, secondary_name), 'r') as f:
@@ -69,10 +69,10 @@ for ii, directory in enumerate(dirs):
         dMass2 = float("".join(dMass2.split()))
         tmp.append(dMass2)
 
-        # Find the line where dTidalQ lives, remove whitespace, make it a float
-        dTidalQ2 = re.findall('%s(.*?)#' % 'dTidalQ', secondary_in)[0]
-        dTidalQ2 = float("".join(dTidalQ2.split()))
-        tmp.append(dTidalQ2)
+        # Find the line where dTidalTau lives, remove whitespace, make it a float
+        dTidalTau2 = re.findall('%s(.*?)#' % 'dTidalTau', secondary_in)[0]
+        dTidalTau2 = float("".join(dTidalTau2.split()))*YEARSEC
+        tmp.append(dTidalTau2)
 
     # Pull LockTimes out of logfile
     with open(os.path.join(os.path.join(directory, logfile_name)), 'r') as f:
@@ -113,7 +113,8 @@ for ii, directory in enumerate(dirs):
     tmp.append(data[-1,2])
 
[…]

### 2018-11-09 Scripts/setupMCCPL.py

@@ -1,0 +1,214 @@
+"""
+
+@author: David P. Fleming, University of Washington, Seattle
+@email: dflemin3 (at) uw (dot) edu
+Nov. 2018
+
+This script produces a population of tidally-interacting binaries stars.
+
+Assumptions:
+- Template files live directory where this script exists.
+
+"""
+
+import numpy as np
+import pandas as pd
+import re
+import os
+from datetime import datetime
+import stat
+import sys
+
+# Constants/Control Flags
+num = 10000 # Number of sets of initial conditions to generate
+PATH = os.path.dirname(os.path.realpath(__file__))
+save_dist = True # Save initial condition distributions?
+show_plots = True # visualize initial condition distributions?
+write_infiles = True # Write the vplanet input files, or just sample from distributions if False
+seed = int(os.getpid()*datetime.now().microsecond/100) # Random seed
+
+# Set RNG seed
+np.random.seed(seed)
+
+# File names
+sys_name = "vpl.in"
+primary_name = "primary.in"
+secondary_name = "secondary.in"
+
+# Lists to hold stuff
+pri_rot = []
+pri_q = []
+pri_mass = []
+
+sec_rot = []
+sec_q = []
+sec_mass = []
+sec_e = []
+sec_orb = []
+
+runfile_names = []
+
+### Helper functions ###
+def lognuniform(low=0, high=1, size=None, base=10.0):
+    return np.power(base, np.random.uniform(low, high, size))
+# end function
+
+
+### Make the simulation initial conditions! ###
+for ii in range(num):
+
+    # Create a directory for the simulation to live in
+    if write_infiles:
+        directory = "simulation_" + str(ii)
+        if not os.path.exists(os.path.join(PATH,directory)):
+            os.makedirs(os.path.join(PATH,directory))
+
+    ### Populate the primary input file ###
+
+    # Read template input file
+    with open(os.path.join(PATH, 'primary.in'), 'r') as f:
+        primary_in = f.read()
+
+    # Sample from priors for initial conditions
+
+    # Loguniform initial rotation period prior over [0.8, 15] days (Matt+2015)
+    dRotPeriod = -lognuniform(low=np.log10(0.8), high=np.log10(15.0)) # negative -> days
+
+    # Log uniform prior for 10^4 - 10^7 for tidal Q
[…]

### 2018-11-09 Scripts/setupMCCTL.py

@@ -1,0 +1,214 @@
+"""
+
+@author: David P. Fleming, University of Washington, Seattle
+@email: dflemin3 (at) uw (dot) edu
+Nov. 2018
+
+This script produces a population of tidally-interacting binaries stars.
+
+Assumptions:
+- Template files live directory where this script exists.
+
+"""
+
+import numpy as np
+import pandas as pd
+import re
+import os
+from datetime import datetime
+import stat
+import sys
+
+# Constants/Control Flags
+num = 10000 # Number of sets of initial conditions to generate
+PATH = os.path.dirname(os.path.realpath(__file__))
+save_dist = True # Save initial condition distributions?
+show_plots = True # visualize initial condition distributions?
+write_infiles = True # Write the vplanet input files, or just sample from distributions if False
+seed = int(os.getpid()*datetime.now().microsecond/100) # Random seed
+
+# Set RNG seed
+np.random.seed(seed)
+
+# File names
+sys_name = "vpl.in"
+primary_name = "primary.in"
+secondary_name = "secondary.in"
+
+# Lists to hold stuff
+pri_rot = []
+pri_q = []
+pri_mass = []
+
+sec_rot = []
+sec_q = []
+sec_mass = []
+sec_e = []
+sec_orb = []
+
+runfile_names = []
+
+### Helper functions ###
+def loguniform(low=0, high=1, size=None, base=10.0):
+    return np.power(base, np.random.uniform(low, high, size))
+# end function
+
+
+### Make the simulation initial conditions! ###
+for ii in range(num):
+
+    # Create a directory for the simulation to live in
+    if write_infiles:
+        directory = "simulation_" + str(ii)
+        if not os.path.exists(os.path.join(PATH,directory)):
+            os.makedirs(os.path.join(PATH,directory))
+
+    ### Populate the primary input file ###
+
+    # Read template input file
+    with open(os.path.join(PATH, 'primary.in'), 'r') as f:
+        primary_in = f.read()
+
+    # Sample from priors for initial conditions
+
+    # Loguniform initial rotation period prior over [0.8, 15] days (Matt+2015)
+    dRotPeriod = -loguniform(low=np.log10(0.8), high=np.log10(15.0)) # negative -> days
+
+    # Log uniform prior for 1.0e-2 - 1s for tidal tau
[…]

### 2018-11-09 Sims/CPLTest/run.py

@@ -1,0 +1,15 @@
+# Run all vplanet simulations
+import subprocess
+import sys
+import os
+
+# Find all subdirectories
+dir_path = os.path.dirname(os.path.realpath(__file__))
+dirs = ["5", "10", "20", "30", "40", "50", "60"]
+
+# Run simulations
+for dir in dirs:
+    os.chdir(os.path.join(dir_path,dir))
+
+    # Run simulation
+    subprocess.call(['vplanet', 'vpl.in'])

### 2018-11-09 Sims/LockingTest/run.py

@@ -1,0 +1,15 @@
+# Run all vplanet simulations
+import subprocess
+import sys
+import os
+
+# Find all subdirectories
+dir_path = os.path.dirname(os.path.realpath(__file__))
+dirs = ["single", "q7", "q8", "tau0_01NoLock", "tau0_1NoLock", "q7NoLock", "q8NoLock", "tau0_01", "tau0_1"]
+
+# Run simulations
+for dir in dirs:
+    os.chdir(os.path.join(dir_path,dir))
+
+    # Run simulation
+    subprocess.call(['vplanet', 'vpl.in'])

### 2018-11-09 Sims/TidalLocking/runQ.py

@@ -1,0 +1,15 @@
+# Run all vplanet simulations
+import subprocess
+import sys
+import os
+
+# Find all subdirectories
+dir_path = os.path.dirname(os.path.realpath(__file__))
+dirs = ["50_4", "50_5", "50_6", "50_7"]
+
+# Run simulations
+for dir in dirs:
+    os.chdir(os.path.join(dir_path,dir))
+
+    # Run simulation
+    subprocess.call(['vplanet', 'vpl.in'])

### 2018-11-09 Paper/sync.tex

@@
 
-Plots:
+\subsection{Tidal-Locking Over $Q,\tau$ and P$_{orb}$}
 
--Porb vs Q/tau grid for e=0, and vs mass ratio
-
--average tsync for various orbital period ranges for CPL vs CTL marginalizing over all other parameters. do it with small multiples
-
--Prot as a function of mass for locked (unlocked) binaries of random ages for gyrochronology results, naturally produces fast rotators
-
--Plot showing diff between binary relative to single star Prot evolution for various Porbs
-
--tsync contours over Q/tau and Porb
-
-\subsection{Tidal-Locking Over $Q,\tau$ and P$_{orb}$}
+\begin{figure*}[ht]
+	\includegraphics[width=\textwidth]{../Plots/qTauPorbLockHist.pdf}
+   \caption{Caption}%
+    \label{fig:qTauLock}%
+\end{figure*}
 
@@
 
-\subsection{Tidal-Locking as a Function of Age} \label{sec:results:age}
+\subsection{Tidal-Locking as a Function of Eccentricity} \label{sec:results:ecc}
+
+\begin{figure*}[ht]
+	\includegraphics[width=\textwidth]{../Plots/eccPorbLockHist.pdf}
+   \caption{Caption.}%
+    \label{fig:eccLock}%
+\end{figure*}
+
+\begin{figure*}[ht]
+	\includegraphics[width=\textwidth]{../Plots/eccPorbRatioHist.pdf}
+   \caption{Caption.}%
+    \label{fig:eccRatio}%
+\end{figure*}
 
@@
 
+\subsection{Comparison with \kepler} \label{sec:kepler}
+
+\begin{figure*}[ht]
+	\includegraphics[width=\textwidth]{../Plots/lurie.pdf}
+   \caption{Caption.}
+    \label{fig:lurie}
+\end{figure*}
+
 \subsubsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology}
+
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
+   \caption{Caption}%
+    \label{fig:binGyro}%
+\end{figure}
 
@@
 
-%\begin{figure}[h]
-%	\includegraphics[width=0.45\textwidth]{../Plots/locked_kepler.pdf}
-%   \caption{Simulated distribution of P$_{rot}$ for tidally-locked (cross) and unlocked (circle) stellar binaries, colored by age for Model 2 (see $\S$~\ref{sec:results:kepler2}).
-Left: P$_{rot}$ as a function of stellar mass.
-Right: Marginalized P$_{rot}$ distribution for locked (blue) and unlocked (orange) binaries.
-Tidally-locked binaries rotate more rapidly than unlocked binaries, displaying an increased density in the P$_{rot}$ distribution for P$_{rot} {\lsim} 20$ d, and with a median P$_{rot} = 19.8$ d, compared to the unlocked median P$_{rot} = 27.5$ d.
-Unlocked binaries on average rotate more quickly than single stars (compare to single distribution in Fig.~\ref{fig:kepler2}) due to the influence of tidal torques.}
-%    \label{fig:locking}
-%\end{figure}
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
+   \caption{Caption}%
+    \label{fig:lockingCPL}%
+\end{figure}
+
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
+   \caption{Caption}%
+    \label{fig:lockingCTL}%
+\end{figure}
 

### 2018-11-09 Paper/sync.tex

@@
 
+- Consult Lurie's paper for CPL vs CTL (e.g. why not 3:2 population-> use CTL!) and for discussing sunsync population
+
 \end{abstract}
@@
 Ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that link the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, will likely be incorrect, however no previous study has quantified this effect.
-Moreover, no previous study has quantified how tidal-locking proceeds in low-mass stellar binaries using a realistic treatment of stellar evolution and magnetic braking.
-
-XXX mention main sequence lifetimes
+Moreover, no previous study has quantified how tidal-locking proceeds in stellar binaries using a realistic treatment of stellar evolution and magnetic braking over the full main sequence lifetimes of low-mass stars.
 
@@
 \begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
    \caption{Caption}%
@@
 \begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
    \caption{Caption}%

### 2018-11-12 Analysis/example.py
added full lurie dataset, updated figures

@@ -6,6 +6,26 @@
 This script plots the Prot evolution of 1-1 Msun binary undergoing stellar-tidal
 evolution with Porb = 50 d for various tidal Qs and taus.
 
+Interpretation:
+
+For all stars, early spin down is due to stellar radius contraction along the
+pre-main sequence.  All stars then begin to spin down over long timescales
+due to magnetic braking and tidal torques in binaries. For the CPL model,
+ALL binaries tidally lock while for the CTL model, only
+binaries with tau > 0.1 s tidally lock.  Binaries with tau < 0.1 s have tides
+that are weak and get overpowered by magnetic braking toques that spin down the
+star.  Note how when CPL models tidally lock, it's in synchronous rotation
+since e = 0.1  CTL models lock into slighty *supersynchronous* rotation since
+Peq = f(e, Porb) * Porb where f(e, Porb) < 1 for e > 0. Binaries with sufficiently
+large tau can also modify e owing to stronger tidal torques, working to
+circularize the orbit, to stars with large tau have larger Peq due to smaller e.
+For our choice of Q and tau, the CPL model has stronger torques than the CTL model,
+locking systems earlier.  Note that for long Porb and for systems of a given age,
+binaries will have longer Prot than single star systems owing to tidal torques
+spinning-down the star towards the tidally locked state.  Using gyrochronlogy
+models, this would incorrect making the systems appear old than they actually
+are.
+
 """
 
 import numpy as np

### 2018-11-12 Analysis/lurie.py

@@ -27,14 +27,9 @@ plotLurie = True
 cpl = pd.read_csv("../Data/mcCPLTorqueNov9.csv")
 ctl = pd.read_csv("../Data/mcCTLTorqueNov9.csv")
 
-lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
-                    names=["Porb", "Prot", "Ecc"])
-
-# Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
-cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
-cpl["Sec_LockTime"][cpl["Sec_LockTime"] < 0] = 7.0e9
-ctl["Pri_LockTime"][ctl["Pri_LockTime"] < 0] = 7.0e9
-ctl["Sec_LockTime"][ctl["Sec_LockTime"] < 0] = 7.0e9
+lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
+lurie["Prot"] = lurie["p_acf"].copy()
+lurie["Porb"] = lurie["p_orb"].copy()
 
 # 1st Fig modeled after Fig. 6, upper panel from Lurie+2017
 
@@ -52,7 +47,7 @@ im = ax0.scatter(cpl["Age_Porb"], cpl["Age_Porb"]/cpl["Pri_ProtAge"],
 if plotLurie:
     # Plot Lurie+2017 data
     ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=100, zorder=2,
-                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
+                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)", alpha=0.7)
 
     # Plot Lurie+2017 detection limit of 45 d
     ax0.plot([1.0, 100.0], [1.0/45.0, 100.0/45.0], ls="-", lw="5", color="k", zorder=3)
@@ -63,8 +58,8 @@ ax0.set_xlim(1,100)
 ax0.set_ylim(0.9, 30)
 ax0.set_xscale("log")
 ax0.set_yscale("log")
-ax0.set_xlabel("P$_{orb}$ [d]", fontsize=25)
-ax0.set_ylabel("P$_{orb}$ / P$_{rot}$", fontsize=25)
+ax0.set_xlabel("P$_{orb}$ [d]", fontsize=30)
[…]

### 2018-11-12 Analysis/porbProtPorbHist.py

@@ -28,11 +28,13 @@ cpl["Pri_dTidalQ"] = pd.Series(cpl["Pri_dTidaLQ"].values, index=cpl.index)
 ctl = pd.read_csv("../Data/mcCTLTorqueNov9.csv")
 
 # Read in lurie data
-lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
-                    names=["Porb", "Prot", "Ecc"])
+lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
+lurie["Prot"] = lurie["p_acf"].copy()
+lurie["Porb"] = lurie["p_orb"].copy()
 
 # Construct Porb, ecc bins based on assumed ranges
 num = 11
+bins = 20
 porbBinEdges = np.linspace(0, 100, num)
 porbProtBinEdges = np.linspace(0.4, 2.0, num)
 
@@ -107,14 +109,13 @@ for ii in range(len(porbBinEdges)-1):
 
 ax1.set_xlim(0.4, 2.0)
 ax1.set_ylim(0, 100)
-ax1.set_xlabel(r"P$_{orb}$/P$_{rot}$", fontsize=22)
-ax1.set_ylabel("Orbital Period [d]", fontsize=22)
-ax1.set_title("CPL")
+ax1.set_xlabel(r"P$_{orb}$/P$_{rot}$", fontsize=30)
+ax1.set_ylabel("Orbital Period [d]", fontsize=30)
 
 ### Colorbar ###
 cbaxes = fig.add_subplot(gs[2])
 cb = plt.colorbar(im, cax=cbaxes)
-cb.set_label(label="Median log$_{10}(Q)$", labelpad=30, rotation=270)
+cb.set_label(label="Median log$_{10}(Q)$", labelpad=30, fontsize=25)
 
 fig.savefig("../Plots/porbProtPorbQHist.pdf", bbox_inches="tight", dpi=600)
 
@@ -139,13 +140,12 @@ for ii in range(len(porbBinEdges)-1):
 
 ax2.set_xlim(0.4, 2.0)
 ax2.set_ylim(0, 100)
-ax2.set_xlabel(r"P$_{orb}$/P$_{rot}$", fontsize=22)
-ax2.set_title("CTL")
+ax2.set_xlabel(r"P$_{orb}$/P$_{rot}$", fontsize=30)
 
 ### Colorbar ###
 cbaxes = fig.add_subplot(gs[2])
[…]

### 2018-11-12 Analysis/protDist.py

@@ -27,7 +27,12 @@ single-star only models due to tidal torques.  Both tidal models predict a large
 population of rapid rotators (Prot < 7.5 d, < 20 d) than single star models fail
 to do, regardless of age, for stars with M > 0.5 Msun, aka GK stars which Kepler
 is most sensitive to. Therefore, fast rotators in Kepler field can be explained
-by tidally-interacting binaries.
+by tidally-interacting binaries. Furthermore, note that the binary Prot distribution
+does not always follow the expected result of single star gyrochronology that
+predicts that, in the long run, stars rotate more slowly with age.  Ages of
+binaries, especially short Prot tidally-influenced systems, do not always
+correlate with age as tides can fix Prot = Peq, or at least keep it near Peq,
+decoupling Prot from age, causing gyrochronlogy methods to fail in this instance. 
 
 """
 
@@ -107,7 +112,7 @@ ax2.set_rasterization_zorder(0)
 ax2.set_xlabel("Mass [M$_{\odot}$]")
 ax2.set_xlim(0.1, 1)
 ax2.set_ylim(0.0, 100)
-ax2.text(0.15, 90, "Single", ha="left", va="center", size=30, color="black",
+ax2.text(0.15, 90, "Single Star", ha="left", va="center", size=30, color="black",
          zorder=100)
 
 ### Colorbar for scatter plots ###

### 2018-11-12 Analysis/subsync.py

@@ -8,12 +8,18 @@ on the subsynchronous rotators.
 
 Script output:
 
-Median subsync tidal tau: 0.042416253599999995
-Median supersync tidal tau: 0.1245296974
+Median subsync tidal tau: 0.0790922272
+Median supersync tidal tau: 0.12508653609999998
 Median subsync ecc: 0.1288805
 Median supersync ecc: 0.151426
 Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6126373626373627
 Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.11401098901098901
+
+Interpretation:
+
+Although Lurie+2017 attributes differential rotation to the production of the
+subsynchronous population, we find that coupled stellar-tidal evolution naturally
+produces the population. XXX COMPARE FRACTIONS TO LURIE---IT's CLOSE
 
 """
 
@@ -40,12 +46,6 @@ ctl = pd.read_csv("../Data/mcCTLTorqueNov9.csv")
 lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
                     names=["Porb", "Prot", "Ecc"])
 
-# Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
-cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
-cpl["Sec_LockTime"][cpl["Sec_LockTime"] < 0] = 7.0e9
-ctl["Pri_LockTime"][ctl["Pri_LockTime"] < 0] = 7.0e9
-ctl["Sec_LockTime"][ctl["Sec_LockTime"] < 0] = 7.0e9
-
 ### 1st Fig modeled after Fig. 7 from Lurie+2017, sub zoomed in on subsync population ###
 
 fig = plt.figure(figsize=(9, 8))
@@ -68,9 +68,9 @@ if plotLurie:
 ax0.set_rasterization_zorder(0)
 ax0.set_xlim(0,10)
 ax0.set_ylim(0.4, 1.3)
-ax0.set_xlabel("P$_{orb}$ [d]", fontsize=25)
-ax0.set_ylabel("P$_{orb}$ / P$_{rot}$", fontsize=25)
-leg = ax0.legend(loc="lower left", framealpha=0.0, fontsize=18)
[…]

### 2018-11-12 Sims/tauQ/runQ.py

@@ -5,7 +5,7 @@ import os
 
 # Find all subdirectories
 dir_path = os.path.dirname(os.path.realpath(__file__))
-dirs = ["20_4", "20_5", "20_6", "20_7", "50_4", "50_5", "50_6", "50_7"]
+dirs = ["q4", "q5", "q6", "q7", "tau10", "tau1", "tau0_1", "tau0_01", "single"]
 
 # Run simulations
 for dir in dirs:

### 2018-11-12 Paper/sync.tex
Updates from ShareLaTeX

@@
 Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, numerical solutions for the CPL and CTL models can produce unphysical evolution.
-We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively, when P$_{rot}$ is within $1\%$ of P$_{eq}$.
+We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively, when two conditions are met: P$_{rot}$ is within $1\%$ of P$_{eq}$ and the P$_{rot}$ derivative points towards $P_{eq}$ on both sides of P$_{eq}$, i.e.
+when the gradient of P$_{rot}$ points towards the tidally-locked state when tides overpower magnetic braking.
 We find that this scheme produces physically and numerically accurate results.
@@
 
-\begin{figure}[h]
+\begin{figure}
 	\includegraphics[width=0.4\textwidth]{../Plots/tidalExample.pdf}
@@
 %% SECTION: Simulations + Initial conditions %%
-
 \section{Simulations} \label{sec:simulations}
@@
 
+\subsection{Coupled Stellar-Tidal Evolution}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
+   \caption{Caption.}%
+    \label{fig:coupledExample}%
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/eqPer.pdf}
+   \caption{Caption.}%
+    \label{fig:eqPer}%
+\end{figure}
+
 \subsection{Tidal-Locking Over $Q,\tau$ and P$_{orb}$}
@@
 \begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/qTauPorbLockHist.pdf}
+	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
    \caption{Caption}%
@@
 
-\subsection{Tidal-Locking as a Function of Eccentricity} \label{sec:results:ecc}
-
-\begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/eccPorbLockHist.pdf}
-   \caption{Caption.}%
-    \label{fig:eccLock}%
-\end{figure*}
-
-\begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/eccPorbRatioHist.pdf}
-   \caption{Caption.}%
-    \label{fig:eccRatio}%
-\end{figure*}
-
-XXX
-
-\subsubsection{Tidal-Locking as a Function of P$_{orb}$} \label{sec:results:porb}
-
-XXX
-
-\begin{figure}[h]
-	\includegraphics[width=0.4\textwidth]{../Plots/lockTimePorbHist.pdf}
-   \caption{Caption.}%
-    \label{fig:lockHistPorb}%
-\end{figure}
-
-\subsection{Comparison with \kepler} \label{sec:kepler}
-
-\begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/lurie.pdf}
-   \caption{Caption.}
-    \label{fig:lurie}
-\end{figure*}
-
 \subsubsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology}
-
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
-   \caption{Caption}%
-    \label{fig:binGyro}%
-\end{figure}
 
@@
 
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
+\begin{figure}[t]
+	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
    \caption{Caption}%
-    \label{fig:lockingCPL}%
+    \label{fig:protDist}%
 \end{figure}
 
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
+%% SECTION: COMPARISON TO KEPLER %%
+\section{Comparison to \kepler} \label{sec:kepler}
+
+\begin{figure}[t]
+	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
    \caption{Caption}%
-    \label{fig:lockingCTL}%
+    \label{fig:lurie7}%
+\end{figure}
+
+\subsection{The Subsynchronous Population} \label{sec:subsync}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/subsync.pdf}
+   \caption{Caption}%
+    \label{fig:subsync}%
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/subsyncTauHist.pdf}
+   \caption{Caption}%
+    \label{fig:subsyncTauHist}%
+\end{figure}
+
+\subsection{$Q$, $\tau$ Constraints from P$_{orb}$ and P$_{rot}$} \label{sec:qTau}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbQHist.pdf}
+   \caption{Caption}%
+    \label{fig:qmap}%
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbTauHist.pdf}
+   \caption{Caption}%
+    \label{fig:taumap}%
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/qLurie.pdf}
+   \caption{Caption}%
+    \label{fig:qLurie}%
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/tauLurie.pdf}
+   \caption{Caption}%
+    \label{fig:tauLurie}%
 \end{figure}
@@
 %% SECTION: DISCUSSION %%
-
 \section{Discussion} \label{sec:discussion}

### 2018-11-12 Analysis/porbProtPorbHist.py
tweaked plots again

@@ -140,6 +140,7 @@ for ii in range(len(porbBinEdges)-1):
 
 ax2.set_xlim(0.4, 2.0)
 ax2.set_ylim(0, 100)
+ax2.set_ylabel("Orbital Period [d]", fontsize=30)
 ax2.set_xlabel(r"P$_{orb}$/P$_{rot}$", fontsize=30)
 
 ### Colorbar ###
@@ -151,7 +152,7 @@ fig.savefig("../Plots/porbProtPorbTauHist.pdf", bbox_inches="tight", dpi=600)
 
 ################################################################################
 #
-#   Estimate tidal Q, tau for Porb < 10 d where Lurie+2017 sample is more complete
+#   Estimate tidal Q, tau for Lurie+2017 sample
 #
 ################################################################################
 

### 2018-11-12 Paper/sync.tex
Updates from ShareLaTeX

@@
 
--tidally locking before 5 Gyr occurs for most binaries below orbital periods of 60-70 days regardless of tidal parameter
-
--compute cumulative fraction of tidal-locking times in Porb bins plot so I can say x fraction are stars with Porb $< 60$d are tidally locked.
-
--make locking.py plot from previous paper with new data
-
 - 5 points with Porb < 10 days and Porb/Prot > 10 in lurie plot are likely young stars that have or continue to spin up via angular momentum conservation due to pre main sequence contraction.
@@
 
-- Lurie has some binaries with Porb/Prot = 0.9, subsynchronous rotators.  cite his explaination for them
-
 - Lurie finds tons of short Porb EBs because transit observational biases favor bodies with short orbital periods for a finite observing window, eg Kepler's lifetime
@@
 - Consult Lurie's paper for CPL vs CTL (e.g. why not 3:2 population-> use CTL!) and for discussing sunsync population
+
+- cite Matsen+2018 for "more than $50\%$ of thebinary period distribution will be imaged in about $85\%$ of the host stars" in paper
+about detecting unresolved binaries in TESS
 
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
@@
     \label{fig:lurie7}%
-\end{figure}
+\end{figure*}
 
@@
     \label{fig:subsync}%
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/subsyncTauHist.pdf}
-   \caption{Caption}%
-    \label{fig:subsyncTauHist}%
 \end{figure}

### 2018-11-12 Analysis/locking.py
updated locking script

@@ -11,12 +11,33 @@ in the Kepler field.
 Script Output:
 
 CPL locked Prot median: 27.049087
-CPL unlocked Prot median: 42.744551
+CPL unlocked Prot median: 41.0628235
+CPL interacting Prot median: 51.7563165
 CPL fraction of locked binaries: 0.2518
+CPL fraction of not locked but interacting binaries: 0.0956
 
 CTL locked Prot median: 4.683883
-CTL unlocked Prot median: 42.3556655
+CTL unlocked Prot median: 41.19235
+CTL interacting Prot median: 49.27336
 CTL fraction of locked binaries: 0.0202
+CTL fraction of not locked but interacting binaries: 0.1257
+
+CPL locked Age median: 4.511807432302286
+CPL unlocked Age median: 3.6993176339544354
+CPL interacting Age median: 4.518420938215314
+
+CTL locked Age median: 3.8579810329094926
+CTL unlocked Age median: 3.9204912513099366
+CTL interacting Age median: 4.604624291466229
+
+CPL locked Porb median: 27.947582500000003
+CPL unlocked Porb median: 60.4407475
+CPL interacting Porb median: 57.719126499999994
+
+CTL locked Porb median: 4.7715415
+CTL unlocked Porb median: 51.597271
+CTL interacting Porb median: 58.190884
+
 
 
 """
@@ -60,6 +81,19 @@ ctl["Sec_LockTime"][ctl["Sec_LockTime"] < 0] = 7.0e9
 cpl["Locked"] = pd.Series(cpl["Pri_LockTime"] < cpl["Age"], index=cpl.index)
 ctl["Locked"] = pd.Series(ctl["Pri_LockTime"] < ctl["Age"], index=ctl.index)
 
+# Make flag for binaries that are strongly tidally-influence, that is, stars with
+# Prot/Peq within [0.9, 1.1] but not locked
+mask = ((np.fabs(1.0 - (cpl["Pri_ProtAge"]/cpl["Age_Peq"])) <= 0.1).values) & (~cpl["Locked"].values)
+cpl["Interacting"] = pd.Series(mask, index=cpl.index)
+mask = ((np.fabs(1.0 - (ctl["Pri_ProtAge"]/ctl["Age_Peq"])) <= 0.1).values) & (~ctl["Locked"].values)
+ctl["Interacting"] = pd.Series(mask, index=ctl.index)
+
[…]

### 2018-11-12 Analysis/protDist.py

@@ -85,7 +85,7 @@ lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
 inds = np.random.choice(np.arange(len(cpl)), size=num, replace=False)
 
 # Plot!
-fig = plt.figure(figsize=(30, 8))
+fig = plt.figure(figsize=(30, 10))
 gs = GridSpec(3, 7,
               height_ratios=[0.05, 0.015, 1],
               width_ratios=[1, 0.001, 1, 0.001, 1, 0.001, 0.25],

### 2018-11-12 Paper/sync.tex
Updates from ShareLaTeX

@@
 
-\shorttitle{Tidal Locking in Stellar Binaries}
+\shorttitle{Tidal Torques in Stellar Binaries}
 \shortauthors{Fleming et al.}
@@
 
-\title{Tidal Locking in Low-Mass Stellar Binary Stars}
+\title{Rotation Periods in Low-Mass Binary Stars: The Impact of Tidal Torques}
 
@@
 A large sample of $816$ rotation period measurements derived using starspot modulations on primary stars of \kepler eclipsing binaries by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked, but also find tentative evidence for tidal-locking in binaries with P$_{orb}$ up to 45 d, their detection limit.
+Furthermore, \citet{Lurie2017} finds a subsynchrnous population of short-period binaries clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal-locking.
 Observations by the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will contain low-mass, short-period eclipsing binaries, permitting additional tests for models of stellar rotation evolution due to tidal torques.
@@
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
-We present our results in $\S$~\ref{sec:results} and discuss their implications in $\S$~\ref{sec:discussion}.
+We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{Stellar radius (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}).
+   \caption{Stellar radius (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} magnetic braking law.
 Each dot denotes the approximate time when each star reaches the main sequence.}%
@@
 
+Tidal torques always drive a body towards the tidally-locked state.
 When a body tidally locks, tidal torques fix P$_{rot}$ to the equilibrium P$_{rot}$, P$_{eq}$.
@@
 when P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
-Both the CPL and CTL model, however, predict that tidally-interacting bodies on sufficiently eccentric orbits undergo pseudosynchronous, or supersynchronous rotation, e.g.
+Although spin-orbit synchronization is an expected outcome of tidal evolution \citep{Counselman1973}, in general for tidally-locked bodies on non-circular orbits, both the CPL and CTL model predict pseudosynchronous, or supersynchronous rotation, e.g.
 Mercury's 3:2 spin-orbit resonance \citep[P$_{rot} = 2/3$ P$_{orb}$,][]{GoldreichPeale1966}.
@@
 
-Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, numerical solutions for the CPL and CTL models can produce unphysical evolution.
-We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively, when two conditions are met: P$_{rot}$ is within $1\%$ of P$_{eq}$ and the P$_{rot}$ derivative points towards $P_{eq}$ on both sides of P$_{eq}$, i.e.
-when the gradient of P$_{rot}$ points towards the tidally-locked state when tides overpower magnetic braking.
+Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, and due to the inherant discreteness of numerical integrations, numerical solutions for the CPL and CTL models can produce unphysical evolution.
+We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively when P$_{rot}$ is within $1\%$ of P$_{eq}$.
+To ensure that tidal torques dominate over torques due to magnetic braking and stellar evolution, we additionally require that the P$_{rot}$ derivative points towards $P_{eq}$ on both sides of P$_{eq}$, i.e.
+when the gradient of P$_{rot}$ points towards the tidally-locked state, before fixing P$_{rot} = $ P$_{eq}$.
 We find that this scheme produces physically and numerically accurate results.
@@
 Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2009}, however a wide range of values exist in the literature.
-Therefore, we consider a wide range of tidal parameters and sample stellar tidal Qs loguniformly over $[10^4,10^8]$ and $\tau$ loguniformly over $[10^{-3},1]$ s.
-
-We supplement our Monte Carlo study with simulations of interesting cases and simulations over grids of parameters, such as P$_{orb}$ and $Q$ or $\tau$, respectively, using the parameter ranges justified above.
-We describe those simulation setups below in $\S$~\ref{sec:results}.
+Therefore, we consider a wide range of tidal parameters and sample stellar tidal Qs loguniformly over $[10^4,10^7]$ and $\tau$ loguniformly over $[10^{-2},1]$ s.
 
@@
 
-\section{Results} \label{sec:results}
+\section{Role $Q$ and $\tau$}
 
 we can explain fast rotation observed in \citet{Meibom2007} to be due to tidal torques speeding up stellar Prots
-
-\subsection{Coupled Stellar-Tidal Evolution}
 
@@
 
+XXX
+
 \begin{figure}
@@
 
-\subsection{Tidal-Locking Over $Q,\tau$ and P$_{orb}$}
+XXX
 
@@
 
-\subsubsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology}
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbQHist.pdf}
+   \caption{Caption}%
+    \label{fig:qmap}%
+\end{figure}
 
-Here we examine the tidal locking in binaries using the simulations from XXX by plotting P$_{rot}$ as a function of mass and age for binaries in XXX.
-We find that $35.5\%$ of binaries tidally lock before the age of the system, typically for tidal $Q \lsim 10^7$.
+XXX
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbTauHist.pdf}
+   \caption{Caption}%
+    \label{fig:taumap}%
+\end{figure}
+
+XXX
+
+\section{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{../Plots/lockedCPL.pdf}
+   \caption{Caption.}%
+    \label{fig:lockedCPL}%
+\end{figure}
+
+XXX
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{../Plots/lockedCTL.pdf}
+   \caption{Caption.}%
+    \label{fig:lockedCTL}%
+\end{figure}
+
+XXX
+
+Here we compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking.
+We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar}.
+To generate the simulations, we sample from the same mass and initial P$_{rot}$ distributions we used for the binary simulations described $\S$~\ref{sec:methods}.
+In Fig.~\ref{fig:protDist}, we P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
+
 Tidally-locked binaries tend to rotate more rapidly than unlocked binaries, reflected in the increased density for P$_{rot} \lsim 20$ d in the marginal distribution for the locked population and in the distribution medians (locked median P$_{rot} = 19.8$ d compared to unlocked median P$_{rot} = 27.5$ d).
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
-   \caption{Caption}%
+   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations.
+The P$_{rot}$ distribution for each case, marginalized over stellar mass, is depicted in the right panel.
+For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
+Both tidal models predict a substantial population of fast rotators (P$_{rot} < 20$ d) not seen in the single star model at any age, except for young (age $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$).
+In many binary systems, both tidal models predict that age does not directly correlate with P$_{rot}$, especially in short P$_{orb}$ binaries where tidal torques are stronger, compared to the clear age-dependence of P$_{rot}$ seen in the single star models, the key assumption of gyrochronology methods.}%
     \label{fig:protDist}%
-\end{figure}
+\end{figure*}
+
+XXX
 
@@
 \section{Comparison to \kepler} \label{sec:kepler}
+
+XXX
 
@@
 
-\subsection{The Subsynchronous Population} \label{sec:subsync}
+XXX
+
+\subsection{P$_{orb} < 10$ d: The Subsynchronous Population} \label{sec:subsync}
+
+XXX
 
@@
 \subsection{$Q$, $\tau$ Constraints from P$_{orb}$ and P$_{rot}$} \label{sec:qTau}
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbQHist.pdf}
-   \caption{Caption}%
-    \label{fig:qmap}%
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbTauHist.pdf}
-   \caption{Caption}%
-    \label{fig:taumap}%
-\end{figure}
 

### 2018-11-12 Analysis/lurie.py
tweaked figure formatting

@@ -49,7 +49,7 @@ ax0.axhline(1, lw=3, color="black", ls="--", zorder=100)
 if plotLurie:
     # Plot Lurie+2017 data
     ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=100, zorder=2,
-                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)", alpha=0.7)
+                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
 # Format
 ax0.set_rasterization_zorder(0)
@@ -78,7 +78,7 @@ ax1.axhline(1, lw=3, color="black", ls="--", zorder=100)
 if plotLurie:
     # Plot Lurie+2017 data
     ax1.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=100, zorder=2,
-               marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)", alpha=0.7)
+               marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
 # Format
 ax1.set_rasterization_zorder(0)
@@ -114,7 +114,7 @@ ax0.axhline(1, lw=3, color="black", ls="--", zorder=100)
 if plotLurie:
     # Plot Lurie+2017 data
     ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=100, zorder=3,
-                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)", alpha=0.7)
+                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
 # Format
 ax0.set_rasterization_zorder(0)
@@ -141,7 +141,7 @@ ax1.axhline(1, lw=3, color="black", ls="--", zorder=100)
 if plotLurie:
     # Plot Lurie+2017 data
     ax1.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=100, zorder=3,
-               marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)", alpha=0.7)
+               marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
 # Format
 ax1.set_rasterization_zorder(0)

### 2018-11-12 Analysis/porbProtPorbHist.py

@@ -212,7 +212,7 @@ ax.axvline(down, ls="--", color="black", lw=2)
 ax.axvline(up, ls="--", color="black", lw=2)
 
 # Annotate
-ax.text(5, 0.5, (r"log$_{10}(Q) = %.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up-med, med-down)),
+ax.text(5.5, 0.425, (r"log$_{10}(Q) = %.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up-med, med-down)),
         ha="center", va="center", size=16, color="black", zorder=100,
         bbox=dict(boxstyle="square", fc="white", ec="white", alpha=0.9))
 

### 2018-11-12 Analysis/subsync.py

@@ -63,7 +63,7 @@ ax0.axhline(1, lw=2, color="black", ls="-", zorder=2)
 
 if plotLurie:
     # Plot Lurie+2017 data
-    ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], c=lurie["Ecc"], s=100, zorder=3,
+    ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], c=lurie["Ecc"], s=200, zorder=3,
                 marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
 # Format

### 2018-11-13 Analysis/lurie.py
tweaked plots to use better prots from Lurie+2017

@@ -28,7 +28,7 @@ cpl = pd.read_csv("../Data/mcCPLTorqueNov9.csv")
 ctl = pd.read_csv("../Data/mcCTLTorqueNov9.csv")
 
 lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
-lurie["Prot"] = lurie["p_acf"].copy()
+lurie["Prot"] = lurie["p_1_min"].copy() # As recommended by Lurie+2017 for equitorial Prot
 lurie["Porb"] = lurie["p_orb"].copy()
 
 # 1st Fig modeled after Fig. 6, upper panel from Lurie+2017

### 2018-11-13 Analysis/porbProtPorbHist.py

@@ -29,7 +29,7 @@ ctl = pd.read_csv("../Data/mcCTLTorqueNov9.csv")
 
 # Read in lurie data
 lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
-lurie["Prot"] = lurie["p_acf"].copy()
+lurie["Prot"] = lurie["p_1_min"].copy() # As recommended by Lurie+2017 for equitorial Prot
 lurie["Porb"] = lurie["p_orb"].copy()
 
 # Construct Porb, ecc bins based on assumed ranges
@@ -212,7 +212,7 @@ ax.axvline(down, ls="--", color="black", lw=2)
 ax.axvline(up, ls="--", color="black", lw=2)
 
 # Annotate
-ax.text(5.5, 0.425, (r"log$_{10}(Q) = %.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up-med, med-down)),
+ax.text(5.2, 0.425, (r"log$_{10}(Q) = %.2lf^{+%0.2lf}_{-%0.2lf}$" % (med, up-med, med-down)),
         ha="center", va="center", size=16, color="black", zorder=100,
         bbox=dict(boxstyle="square", fc="white", ec="white", alpha=0.9))
 

### 2018-11-13 Analysis/subsync.py

@@ -45,6 +45,7 @@ plotLurie = True
 cpl = pd.read_csv("../Data/mcCPLTorqueNov9.csv")
 ctl = pd.read_csv("../Data/mcCTLTorqueNov9.csv")
 
+# Load in cleaned sample of Prot and Ecc from Lurie+2017
 lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
                     names=["Porb", "Prot", "Ecc"])
 

### 2018-11-13 Paper/sync.tex
Updates from ShareLaTeX

@@
 
+- we use Lurie+2017 Prots that "Assuming solar-like differential rotation, P1,min will be closest to the equatorial rotation period.
+This provides a consistent reference point for the differential rotation discussion below."
+
 \end{abstract}
@@
 \section{Introduction} \label{sec:intro}
+
+talk about how most people have focused on circularization, sync is hard to measure, we focus on low-mass stars with convective envelopes where the eqtide model is applicable, not higher mass stars with radiative envelopes where dynamical tide effects likely dominate (e.g.
+stuff lurie talked about)
 
@@
 $k_2 \tau$ scaling in Eq.~(\ref{eqn:ctl:z}), so we instead examine how our results scale with $Q$ and $\tau$.
+Any constraints we derive for $Q$ and $\tau$ can trivially be scaled to other values of $k_2$.
 
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{../Plots/lockedCPL.pdf}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
    \caption{Caption.}%
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{../Plots/lockedCTL.pdf}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
    \caption{Caption.}%

### 2018-11-14 Analysis/eqPer.py
updated eqper plots

@@ -14,7 +14,7 @@ import matplotlib.pyplot as plt
 
 #Typical plot parameters that make for pretty plots
 mpl.rcParams['figure.figsize'] = (9,8)
-mpl.rcParams['font.size'] = 24.0
+mpl.rcParams['font.size'] = 20.0
 
 ## for Palatino and other serif fonts use:
 mpl.rc('font',**{'family':'serif'})
@@ -74,37 +74,33 @@ for ii, dir in enumerate(dirs):
         iLockIndCTL = np.argmin(np.fabs(ctl[:,0]-dLockTimeCTL))
 
     # Left: Prot/Peq
-    axes[0].plot(cpl[:,0], cpl[:,4]/cpl[:,5], lw=3, ls="-", color=colors[ii],
-                 label="P$_{orb}$ = %s d" % dir)
-    axes[0].plot(ctl[:,0], ctl[:,4]/ctl[:,5], lw=3, ls="--", color=colors[ii])
+    axes[0].plot(cpl[:,0], cpl[:,4]/cpl[:,5], lw=2.5, ls="-", color=colors[ii])
+    axes[0].plot(ctl[:,0], ctl[:,4]/ctl[:,5], lw=2.5, ls="--", color=colors[ii])
 
     # Right: dProt/dt
     cplDeriv = (cpl[:,6] + cpl[:,7])*YEARSEC/DAYSEC*1.0e9
     ctlDeriv = (ctl[:,6] + ctl[:,7])*YEARSEC/DAYSEC*1.0e9
-    axes[1].plot(cpl[:iLockIndCPL,0], cplDeriv[:iLockIndCPL], lw=3, ls="-",
-                 color=colors[ii])
-    axes[1].plot(ctl[:iLockIndCTL,0], ctlDeriv[:iLockIndCTL], lw=3, ls="--",
+    axes[1].plot(cpl[:iLockIndCPL,0], cplDeriv[:iLockIndCPL], lw=2, ls="-",
+                 color=colors[ii], label="P$_{orb}$ = %s d" % dir)
+    axes[1].plot(ctl[:iLockIndCTL,0], ctlDeriv[:iLockIndCTL], lw=2, ls="--",
                  color=colors[ii])
 
-    axes[1].set_ylabel("Orbital Period [d]")
+axes[1].plot([100], [100], lw=3, ls="-", color="grey", label="CPL")
[…]

### 2018-11-14 Paper/sync.tex

@@
 
-- Consult Lurie's paper for CPL vs CTL (e.g. why not 3:2 population-> use CTL!) and for discussing sunsync population
+- Consult Lurie's paper for CPL vs CTL (e.g.
+why not 3:2 population-> use CTL!) and for discussing sunsync population.
+stars prob not in 3:2 resonance since they likely dont have a fixed solid body shape as I assume, but perhaps in a decoupled core envelope model (bouvier?) envelope could get locked into resonance
 
-- cite Matsen+2018 for "more than $50\%$ of thebinary period distribution will be imaged in about $85\%$ of the host stars" in paper
+- cite Matsen+2018 for "more than $50\%$ of the binary period distribution will be imaged in about $85\%$ of the host stars" in paper
 about detecting unresolved binaries in TESS
@@
 This provides a consistent reference point for the differential rotation discussion below."
+
+- lower left cluster in ProtPorb space in Lurie plot are likely BD or exoplanets according to Lurie
+
+- compare results to Lurie section 4.4.2 discussion of dependence with eccentricity
+
+- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques
+
+- we can explain fast rotation observed in \citet{Meibom2007} to be due to tidal torques speeding up stellar Prots
 
@@
 
-\section{Role $Q$ and $\tau$}
-
-we can explain fast rotation observed in \citet{Meibom2007} to be due to tidal torques speeding up stellar Prots
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/example.pdf}
-   \caption{Caption.}%
-    \label{fig:coupledExample}%
-\end{figure}
+\section{The Rotational Equilibrium State} \label{sec:eq}
 
@@
 XXX
+
+\section{Role $Q$ and $\tau$}
 
@@
 \end{figure}
-
-XXX
-
-\section{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
-   \caption{Caption.}%
-    \label{fig:lockedCPL}%
-\end{figure}
-
-XXX
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
-   \caption{Caption.}%
-    \label{fig:lockedCTL}%
-\end{figure}
-
-XXX
-
-Here we compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking.
-We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar}.
-To generate the simulations, we sample from the same mass and initial P$_{rot}$ distributions we used for the binary simulations described $\S$~\ref{sec:methods}.
-In Fig.~\ref{fig:protDist}, we P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
-
-Tidally-locked binaries tend to rotate more rapidly than unlocked binaries, reflected in the increased density for P$_{rot} \lsim 20$ d in the marginal distribution for the locked population and in the distribution medians (locked median P$_{rot} = 19.8$ d compared to unlocked median P$_{rot} = 27.5$ d).
-Binaries readily tidally-lock at short P$_{orb}$ due to the tidal force's strong semi-major axis dependence, increasing the strength of tidal torques at shorter orbital separations.
-When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} = $P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.
-
-Tidal forces modify the spins of unlocked binaries as well, tending to spin-up stars against the effects of magnetic braking towards orbital synchronization, reflected in unlocked binaries' median P$_{rot} = 27.5$ d, compared to the single star median P$_{rot} = 30.6$ d.
-Tidal locking is not limited to P$_{orb} \lsim 10$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 70$ d, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars (compare to XXX).
-This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find that binaries can tidally lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-The observed age distribution of tidally-locked binaries has little dependence on P$_{rot}$, with both young and old systems tidally-locking at long and short P$_{rot}$, causing the predictions of gyrochronology models to fail in such systems.
-
-\begin{figure*}[t]
-	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
-   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations.
-The P$_{rot}$ distribution for each case, marginalized over stellar mass, is depicted in the right panel.
-For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
-Both tidal models predict a substantial population of fast rotators (P$_{rot} < 20$ d) not seen in the single star model at any age, except for young (age $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$).
-In many binary systems, both tidal models predict that age does not directly correlate with P$_{rot}$, especially in short P$_{orb}$ binaries where tidal torques are stronger, compared to the clear age-dependence of P$_{rot}$ seen in the single star models, the key assumption of gyrochronology methods.}%
-    \label{fig:protDist}%
-\end{figure*}
 
@@
 
+\section{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
+
+Here we compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking.
+We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar}.
+To generate the simulations, we sample from the same mass and initial P$_{rot}$ distributions we used for the binary simulations described $\S$~\ref{sec:methods}.
+In Fig.~\ref{fig:protDist}, we P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
+
+Tidally-locked binaries tend to rotate more rapidly than unlocked binaries, reflected in the increased density for P$_{rot} \lsim 20$ d in the marginal distribution for the locked population and in the distribution medians (locked median P$_{rot} = 19.8$ d compared to unlocked median P$_{rot} = 27.5$ d).
+Binaries readily tidally-lock at short P$_{orb}$ due to the tidal force's strong semi-major axis dependence, increasing the strength of tidal torques at shorter orbital separations.
+When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} = $P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.
+
+Tidal forces modify the spins of unlocked binaries as well, tending to spin-up stars against the effects of magnetic braking towards orbital synchronization, reflected in unlocked binaries' median P$_{rot} = 27.5$ d, compared to the single star median P$_{rot} = 30.6$ d.
+Tidal locking is not limited to P$_{orb} \lsim 10$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 70$ d, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars (compare to XXX).
+This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find that binaries can tidally lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
+The observed age distribution of tidally-locked binaries has little dependence on P$_{rot}$, with both young and old systems tidally-locking at long and short P$_{rot}$, causing the predictions of gyrochronology models to fail in such systems.
+
+\begin{figure*}[t]
+	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
+   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations.
+The P$_{rot}$ distribution for each case, marginalized over stellar mass, is depicted in the right panel.
+For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
+Both tidal models predict a substantial population of fast rotators (P$_{rot} < 20$ d) not seen in the single star model at any age, except for young (age $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$).
+In many binary systems, both tidal models predict that age does not directly correlate with P$_{rot}$, especially in short P$_{orb}$ binaries where tidal torques are stronger, compared to the clear age-dependence of P$_{rot}$ seen in the single star models, the key assumption of gyrochronology methods.}%
+    \label{fig:protDist}%
+\end{figure*}
+
+XXX
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
+   \caption{Caption.}%
+    \label{fig:lockedCPL}%
+\end{figure}
+
+XXX
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
+   \caption{Caption.}%
+    \label{fig:lockedCTL}%
+\end{figure}
+
+XXX
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedProtHist.pdf}
+   \caption{Caption.}%
+    \label{fig:lockedInteracting}%
+\end{figure}
+
+XXX
+
 %% SECTION: DISCUSSION %%

### 2018-11-18 Analysis/eqPer.py
added Porb eqPer plot, updated subsync figure

@@ -139,3 +139,32 @@ ax.set_ylim(0, 1.3)
 
 # Save!
 fig.savefig("../Plots/eqPer.pdf", bbox_inches="tight", dpi=600)
+
+### Short Porb Plot ###
+
+ctlDir = "../Sims/Peq"
+dirs = ["0.001", "0.01", "0.1"]
+labels = [r"$\tau = 10^{-3}$ s", r"$\tau = 10^{-2}$ s", r"$\tau = 10^{-1}$ s"]
+
+fig, ax = plt.subplots(figsize=(9,8))
+
+for ii, dir in enumerate(dirs):
+    # Load data
+    # saOutputOrder	Time -TotEn -TotAngMom -Radius -RotPer -EqRotPer
+    # DRotPerDtEqtide DRotPerDtStellar Ecce -OrbPer RadGyra -SurfEnFluxTotal
+    ctl = np.genfromtxt(os.path.join(ctlDir, dir, "bintides.secondary.forward"))
+
+    # Left: Prot/Peq
+    ax.plot(ctl[:,0], ctl[:,4]/ctl[:,5], lw=3, ls="-", label=labels[ii],
+            color="C%d" % ii)
+
+ax.legend(loc="best", framealpha=0, fontsize=18)
+ax.set_ylabel("P$_{rot}$/P$_{eq}$", fontsize=30)
+ax.set_xlabel("Time [yr]", fontsize=30)
+ax.set_xscale("log")
+ax.axhline(1, lw=2.5, ls=":", color="k")
+ax.set_xlim(1.0e6, cpl[-1,0])
+ax.set_ylim(0, 1.4)
+
+# Save!
+fig.savefig("../Plots/eqPerShortPorb.pdf", bbox_inches="tight", dpi=600)

### 2018-11-18 Analysis/subsync.py

@@ -64,8 +64,8 @@ ax0.axhline(1, lw=2, color="black", ls="-", zorder=2)
 
 if plotLurie:
     # Plot Lurie+2017 data
-    ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], c=lurie["Ecc"], s=200, zorder=3,
-                marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
+    ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], c=lurie["Ecc"], s=75, zorder=3,
+                marker="d", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
 # Format
 ax0.set_rasterization_zorder(0)

### 2018-11-18 Paper/sync.tex
Updates from ShareLaTeX

@@
 
-\title{Rotation Periods in Low-Mass Binary Stars: The Impact of Tidal Torques}
+\title{Rotation Periods in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking}
 
@@
 
-talk about how most people have focused on circularization, sync is hard to measure, we focus on low-mass stars with convective envelopes where the eqtide model is applicable, not higher mass stars with radiative envelopes where dynamical tide effects likely dominate (e.g.
-stuff lurie talked about)
-
 Angular momentum evolution in low-mass (M$\lsim 1$ M$_{\odot}$) short-period (P$_{orb} \lsim 10$ d) stellar binaries is dominated by tides.
 Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}.
-Orbital circularization in ubiquitous in short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
+Orbital circularization is ubiquitous in short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
 For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} finds that the orbit of solar twin binaries circularizes during the stellar pre-main sequence.
-Observations by \citet{Meibom2005} support this picture as they find short-period binaries in the ${\sim}150$ Myr old cluster, M35, tend to have circular orbits.
+Observations by \citet{Meibom2005} support this picture as they find short-period binaries in the ${\sim}150$ Myr old cluster M35 tend to have circular orbits.
 
 Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially so for the stellar spins.
-Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale since the stellar spins harbor much less angular momentum than the orbit \citep{Zahn1989,Witte2002,Mazeh2008}.
-Tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
-A large sample of $816$ rotation period measurements derived using starspot modulations on primary stars of \kepler eclipsing binaries by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked, but also find tentative evidence for tidal-locking in binaries with P$_{orb}$ up to 45 d, their detection limit.
-Furthermore, \citet{Lurie2017} finds a subsynchrnous population of short-period binaries clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal-locking.
-Observations by the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will contain low-mass, short-period eclipsing binaries, permitting additional tests for models of stellar rotation evolution due to tidal torques.
+Tidal torques drive stellar spins towards the tidally-locked state in which the stellar rotation periods (P$_{rot}$) are equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
+Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
 
-Rotation period distributions measured in clusters and field stars are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys due to low transit probabilities.
+A large, homogenous sample of $816$ rotation period measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d, their detection limit.
+Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking.
+These findings suggest that that synchronization for short P$_{orb}$ binaries is not the sole outcome of tidal interactions in low-mass stellar binaries, but rather that impact of tidal effects should be reconsidered for longer P$_{orb}$ and to examine if our current understanding of tidal torques is consistent with short P$_{orb}$ subsynchronous rotators.
+Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass, short-period eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
+
+Understanding angular momentum evolution in low-mass stellar binaries is paramount as rotation period distributions measured in clusters and field stars are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys due to low transit probabilities.
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P${rot}$ distributions.
-Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}.
-Ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that link the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, will likely be incorrect, however no previous study has quantified this effect.
-Moreover, no previous study has quantified how tidal-locking proceeds in stellar binaries using a realistic treatment of stellar evolution and magnetic braking over the full main sequence lifetimes of low-mass stars.
+Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, impacting a contaminating signal that current models do not account for.
+Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that link the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, will likely be incorrect owing to the influence of tidal torques.
+No previous study has quantified this effect.
 
-Here, we examine how tidal torques, coupled with a modern stellar evolution and magnetic braking models, shape the angular momentum evolution of low-mass binary stars.
-We investigate under what conditions tidal-locking occurs, and what the typical tidal-locking timescale is as a function of stellar mass and tidal dissipation parameters for two widely-used equilibrium tidal models.
-We show how tidal torques in binaries cause stellar rotation periods to not strongly correlate with age, causing the predictions of gyrochronlogy models to fail in such systems.
+Here, we examine angular momentum evolution over the full main sequence lifetimes of low-mass stellar binaries using a realistic treatment of stellar evolution, magnetic braking, and tidal torques.
+We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary orbital period and tidal dissipation parameters for two widely-used equilibrium tidal models.
+We show how tidal torques can impact stellar rotation rate in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
@@
 Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citet[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
+For stars, equilibrium tidal models assume that tidal forces dissipate energy in convective regions via viscous turbulence \citep[see][]{Zahn2008}, corresponding tidal dissipation in the outer convective layers of low-mass stars.
 Here, we consider two equilibrium tidal models to study the secular spin-orbital evolution of low-mass stellar binaries.
@@
 
+We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them.
 All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.} with accompanying descriptions and documentation.
@@
 
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/eqPer.pdf}
-   \caption{Caption.}%
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanel.pdf}
+   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively.
+Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d.
+The black dotted line indicates that tidally-locked state where P$_{rot} = $ P$_{eq}$.
+For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
+For all P$_{orb}$, all CPL models tidally-lock into synchronous rotation whereas the CTL model predicts subsynchronous (P$_{rot} > $ P$_{eq}$) rotation.
+In the CTL cases, the tidal torques are weaker than the magnetic braking torque, causing the stars to continue to spin down, preventing the systems from tidally-locking.
+Tidal torques strengthen for P$_{rot}$ far from P$_{eq}$, and magnetic braking torques weaken for long P$_{rot}$ \citep{Kawaler1988,Matt2015}, causing tidal torques to overpower magnetic braking, driving the stellar rotation back towards the tidally-locked state.
+Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.
+We truncate each curve when the respective system tidally-locks.
+Initially, all systems spin-up due to stellar contraction, except for the P$_{orb} = 5$ d case where tides overpower other torques, rapidly driving the stars towards synchronization.
+Once the stars reach the main sequence at ${\sim} 60$ Myr, tidal torques and magnetic braking combine to spin down the stars towards the tidally-locked state.
+For stars at long P$_{orb}$ that do not tidally-lock, after having spun-down to long P$_{rot}$, magnetic braking weakens, allowing tidal torques to sheppard P$_{rot}$ towards P$_{eq}$ as the P$_{rot}$ derivatives stay near 0, but with a slight negative value.}%
     \label{fig:eqPer}%
-\end{figure}
+\end{figure*}
 
@@
 	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
-   \caption{Caption}%
+   \caption{Median P$_{orb}/$P$_{rot}$ for the primary star at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
+The CPL model predicts that stellar binaries can tidally-lock out to P$_{orb} = 30$ d for most values of $Q$, with some locking all the way up to P$_{orb} = 100$ d for small $Q < 10^5$, substantially longer than what is expected.
+For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally-lock into synchronous rotation, regardless of $Q$.
+The CTL model, however, requires stars to have $\tau \geq 0.1$ s in order for stellar binaries with P$_{orb} < 10$ d to likely tidally-lock.
+Both models exhibit supersynchronous rotation (blue regions, P$_{rot} < $P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at these long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spin down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
+For large $Q$ ($Q > 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spining down stars past the tidally-locked state, producing a population of subsynchronous (red regions, P$_{rot} > $P$_{orb}$) rotators.}%
     \label{fig:qTauLock}%
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbQHist.pdf}
-   \caption{Caption}%
+   \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
+In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$.
+For all P$_{orb}$, synchronous and supersynchronous rotators have systematically low $Q$s as strong tidal torques are required to tidally-lock the systems.
+Note that for P$_{orb} < 10$ d, there are no rotators with $1.0 \leq$ P$_{orb}/$P$_{rot} \leq 1.5$ as strong tidal torques always tidally-lock the system, forcing P$_{rot} =$ P$_{eq}$, which can only take two values under the CPL model (see Eqn.~(\ref{eqn:cpl:eqPer})).
+Subsynchronous rotators have large $Q$s, and hence weak tidal torques, that are dominated by magnetic braking that spin-down the stars past the tidally-locked state.
+For P$_{orb} < 70$ d, no stars have P$_{orb}/$P$_{rot} > 1.5$ as in the CPL model, supersynchronous rotators can only tidally-lock into a 3:2 spin-orbit commensurability.
+At longer P$_{orb}$, supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.}%
     \label{fig:qmap}%
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbTauHist.pdf}
-   \caption{Caption}%
+   \caption{Same format as Fig.~\ref{fig:qmap}, but for the CTL model.
+As under the CPL model, stars tidally-lock when subject to strong tidal torques, typically for P$_{orb} < 50$ d and $\tau \gsim 0.25$ s.
+For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with low $tau < 0.1$ s as magnetic braking dominates over tidal torques, spinning down the star.
+For P$_{orb} > 50$ d, tides do not in general dominate the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s has upper and lower limits that encompass most of the range we considered for $\tau$.
+The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
+We do not observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).}%
     \label{fig:taumap}%
@@
 
-XXX
+Median subsync tidal tau: 0.0790922272
+Median supersync tidal tau: 0.12508653609999998
+Median subsync ecc: 0.1288805
+Median supersync ecc: 0.151426
+Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6126373626373627
+Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.11401098901098901
+
+Although Lurie+2017 attributes differential rotation to the production of the
+subsynchronous population, we find that coupled stellar-tidal evolution naturally
+produces the population. Compare with Lurie+2017 sample for 2 $<$ Porb $<$ 10 days:
+$72\%$ in [0.92,1.2] and $15\%$ in [0.84, 0.92].
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/subsync.pdf}
-   \caption{Caption}%
+   \caption{P$_{orb}/$P$_{rot}$ as a function for P$_{orb}$ simulated according to the CTL model (dots) compared with \kepler EBs from \citet{Lurie2017} (diamonds), colored by $e$.
+Our simulations are in excellent agreement with the observed orbital and rotational states of \kepler EBs.
+Our model predicts that all EBs with P$_{orb} < 3$ d are synchronized as is observed by \citet{Lurie2017}.
+We find that our model predicts a population of supersyncronous rotators on eccentric orbits and a population of subsyncronous rotators on nearly-circular orbits, both of which were observed by \citet{Lurie2017}.
+This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL formalism can accurately model the tidal evolution of low-mass stellar binaries.}%
     \label{fig:subsync}%

### 2018-11-19 Analysis/eqPer.py
added more sims, made plot

@@ -143,22 +143,30 @@ fig.savefig("../Plots/eqPer.pdf", bbox_inches="tight", dpi=600)
 ### Short Porb Plot ###
 
 ctlDir = "../Sims/Peq"
-dirs = ["0.001", "0.01", "0.1"]
-labels = [r"$\tau = 10^{-3}$ s", r"$\tau = 10^{-2}$ s", r"$\tau = 10^{-1}$ s"]
+cplDir = "../Sims/Peq"
+ctlDirs = ["0.1", "0.01", "0.001"]
+cplDirs = ["6", "7", "8"]
+labels = [r"$Q= 10^6$, $\tau = 10^{-1}$ s", r"$Q=10^7$, $\tau = 10^{-2}$ s",
+          r"$Q=10^8$, $\tau = 10^{-3}$ s"]
 
 fig, ax = plt.subplots(figsize=(9,8))
 
-for ii, dir in enumerate(dirs):
+for ii in range(len(ctlDirs)):
     # Load data
     # saOutputOrder	Time -TotEn -TotAngMom -Radius -RotPer -EqRotPer
     # DRotPerDtEqtide DRotPerDtStellar Ecce -OrbPer RadGyra -SurfEnFluxTotal
-    ctl = np.genfromtxt(os.path.join(ctlDir, dir, "bintides.secondary.forward"))
+    ctl = np.genfromtxt(os.path.join(ctlDir, ctlDirs[ii], "bintides.secondary.forward"))
+    cpl = np.genfromtxt(os.path.join(cplDir, cplDirs[ii], "bintides.secondary.forward"))
 
-    # Left: Prot/Peq
-    ax.plot(ctl[:,0], ctl[:,4]/ctl[:,5], lw=3, ls="-", label=labels[ii],
-            color="C%d" % ii)
+    # Prot/Peq
+    ax.plot(cpl[:,0], cpl[:,4]/cpl[:,5], lw=3, ls="-", color="C%d" % ii,
+            label=labels[ii])
+    ax.plot(ctl[:,0], ctl[:,4]/ctl[:,5], lw=3, ls="--", color="C%d" % ii)
 
-ax.legend(loc="best", framealpha=0, fontsize=18)
+ax.plot([100], [100], ls="-", lw=3, color="grey", label="CPL")
+ax.plot([100], [100], ls="--", lw=3, color="grey", label="CTL")
+
[…]

### 2018-11-19 Analysis/subsync.py

@@ -12,9 +12,10 @@ Median subsync tidal tau: 0.0790922272
 Median supersync tidal tau: 0.12508653609999998
 Median subsync ecc: 0.1288805
 Median supersync ecc: 0.151426
+Median subsync mass ratio: 0.6315316837433688
+Median supersync mass ratio: 0.6522318263105934
 Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6126373626373627
 Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.11401098901098901
-
 
 Interpretation:
 
@@ -127,9 +128,35 @@ ax.legend(loc="lower center", framealpha=0)
 
 fig.savefig("../Plots/subsyncEccHist.pdf", bbox_inches="tight", dpi=600)
 
-# Print median sub,supersync tidal taus
+# Print median sub,supersync tidal ecc
 print("Median subsync ecc:",np.median(ctl["Age_Ecc"][subMask].values))
 print("Median supersync ecc:",np.median(ctl["Age_Ecc"][~subMask].values))
+
+### Plot histogram of mass ratio above and below Prot=Peq=Porb=1 line
+
+subMask = (ctl["Age_Porb"]/ctl["Pri_ProtAge"] < 1)
+
+fig, ax = plt.subplots()
+
+mu = ctl["Sec_dMass"][subMask].values/ctl["Pri_dMass"][subMask].values
+ax.hist(mu, lw=3, histtype="step",
+        color="C0", label="Subsynchronous", bins="auto", density=True)
+
+mu = ctl["Sec_dMass"][~subMask].values/ctl["Pri_dMass"][~subMask].values
+ax.hist(mu, lw=3, histtype="step",
+        color="C1", label="Supersynchronous", bins="auto", density=True)
+
+ax.set_ylabel("Normalized Counts")
+ax.set_xlabel("Mass Ratio (M$_2$/M$_1$)")
+ax.legend(loc="lower center", framealpha=0)
+
+fig.savefig("../Plots/subsyncMuHist.pdf", bbox_inches="tight", dpi=600)
+
+# Print median sub,supersync tidal taus
+mu = ctl["Sec_dMass"][subMask].values/ctl["Pri_dMass"][subMask].values
+print("Median subsync mass ratio:",np.median(mu))
+mu = ctl["Sec_dMass"][~subMask].values/ctl["Pri_dMass"][~subMask].values
[…]

### 2018-11-19 Paper/sync.tex
Updates from ShareLaTeX

@@
 %%\author[0000-0001-9293-4043]{David P. Fleming}
-\author{David P. Fleming et al}
+\author{David P. Fleming}
 \affil{Astronomy Department, University of Washington \\
 Box 951580, Seattle, WA 98195}
+
+\author{Rory Barnes}
+\affiliation{Astronomy Department, University of Washington \\
+Box 951580, Seattle, WA 98195}
+
+\author{James R. A. Davenport}
+\affiliation{Astronomy Department, University of Washington \\
+Box 951580, Seattle, WA 98195}
+
+\author{Rodrigo Luger}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute \\
+New York, NY 10010}
 
@@
 
-- 5 points with Porb < 10 days and Porb/Prot > 10 in lurie plot are likely young stars that have or continue to spin up via angular momentum conservation due to pre main sequence contraction.
-evidence for young stars in the Kepler field, e.g.
-McQuillan+2014, Matt+2015, Davenport+2017,2018 support this theory
-
-- Lurie finds tons of short Porb EBs because transit observational biases favor bodies with short orbital periods for a finite observing window, eg Kepler's lifetime
-
-- Consult Lurie's paper for CPL vs CTL (e.g.
-why not 3:2 population-> use CTL!) and for discussing sunsync population.
-stars prob not in 3:2 resonance since they likely dont have a fixed solid body shape as I assume, but perhaps in a decoupled core envelope model (bouvier?) envelope could get locked into resonance
-
-- cite Matsen+2018 for "more than $50\%$ of the binary period distribution will be imaged in about $85\%$ of the host stars" in paper
-about detecting unresolved binaries in TESS
-
-- we use Lurie+2017 Prots that "Assuming solar-like differential rotation, P1,min will be closest to the equatorial rotation period.
-This provides a consistent reference point for the differential rotation discussion below."
-
-- lower left cluster in ProtPorb space in Lurie plot are likely BD or exoplanets according to Lurie
-
-- compare results to Lurie section 4.4.2 discussion of dependence with eccentricity
-
-- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques
-
-- we can explain fast rotation observed in \citet{Meibom2007} to be due to tidal torques speeding up stellar Prots
+Stellar rotation period (P$_{rot}$) studies have produced numerous insights into the angular momentum evolution of low-mass stars, but previous investigations have focused on single stars even though roughly half of Sun-like stars are in stellar binaries.
+Here, we examine how tidal torques, stellar evolution, and magnetic braking shape the P$_{rot}$ evolution of low-mass stellar binaries by performing Monte Carlo simulations of coupled stellar-tidal evolution.
+We find that tides can modify stellar rotations in binaries with orbital periods (P$_{orb}$) up to 100 days.
+At short P$_{orb}$, many stellar binaries tidally-lock into synchronous rotation, naturally producing a population of fast rotators (P$_{rot} < 10$ d) across all stellar masses as is observed in the \textit{Kepler} field, a feature that cannot be reproduced by single-star-only models of magnetic braking.
+Many binaries with longer P$_{orb}$ can tidally-lock, fixing the stellar P$_{rot}$ to the binary P$_{orb}$, causing P$_{rot}$ to not be a valid proxy for age in all cases, i.e.
+gyrochronology methods must be applied carefully.
+We compare our model predictions with \kepler eclipsing binary (EB) rotation periods and find that our predictions are in excellent agreement with observations, with the ``Constant Time Lag" equilibrium tidal best matching observations.
+We show that even for short P$_{orb}$ binaries, tidal torques can balance magnetic braking, naturally producing a population of subsynchronous rotators similar what \citet{Lurie2017} observed in their sample of \kepler EBs.
+By comparing observations of \kepler EBs with our simulations, we constrain the tidal time lag to $\tau = 0.19^{+0.43}_{-0.16}$ s.
 
@@
 Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially so for the stellar spins.
-Tidal torques drive stellar spins towards the tidally-locked state in which the stellar rotation periods (P$_{rot}$) are equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
+Tidal torques drive stellar spins towards the tidally-locked state in which the stellar P$_{rot}$ are equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
 Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
 
-A large, homogenous sample of $816$ rotation period measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d, their detection limit.
+The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d, their detection limit.
 Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking.
-These findings suggest that that synchronization for short P$_{orb}$ binaries is not the sole outcome of tidal interactions in low-mass stellar binaries, but rather that impact of tidal effects should be reconsidered for longer P$_{orb}$ and to examine if our current understanding of tidal torques is consistent with short P$_{orb}$ subsynchronous rotators.
+These findings suggest that that synchronization is not a given for tidally-interacting, short P$_{orb}$  stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$.
 Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass, short-period eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
 
-Understanding angular momentum evolution in low-mass stellar binaries is paramount as rotation period distributions measured in clusters and field stars are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys due to low transit probabilities.
+Understanding angular momentum evolution in low-mass stellar binaries is paramount as P$_{rot}$ distributions measured in clusters \citep[e.g.
+Praesepe, ][]{Douglas2017} and field stars \citep[e.g.
+\kepler, ][]{McQuillan2014,Simonian2018} are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys.
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P${rot}$ distributions.
-Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, impacting a contaminating signal that current models do not account for.
-Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that link the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, will likely be incorrect owing to the influence of tidal torques.
+Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, imparting a contaminating signal that is not currently accounted for by modern models.
+Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that links the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, will likely be incorrect owing to the influence of tidal torques.
 No previous study has quantified this effect.
@@
 We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary orbital period and tidal dissipation parameters for two widely-used equilibrium tidal models.
-We show how tidal torques can impact stellar rotation rate in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to fail in such systems.
+We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
@@
 
-We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
+We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, $R$, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
 We model the long-term stellar spin-down due to magnetic braking using the model derived in \citet{Matt2015} since it has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
@@
 \end{equation}
-where the moment of inertia $I = M r_g^2 R^2$, $\dot{J}_{mb}$ is the angular momentum loss due to magnetic braking, and the time derivatives of the stellar radius and $r_g$ are computed numerically using our interpolation of the \citet{Baraffe2015} stellar evolution grids.
+where the moment of inertia $I = M r_g^2 R^2$, $\dot{J}_{mb}$ is the angular momentum loss due to magnetic braking, and the time derivatives of the stellar $R$ and $r_g$ are computed numerically using our interpolation of the \citet{Baraffe2015} stellar evolution grids.
 
@@
 
-In Fig.~\ref{fig:stellarExample}, we plot the evolution of the radius, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar.
+In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar.
 We assume all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr.
 All stars' radii contract along the pre-main sequence, spinning the stars up (right panel).
-Once the stars reach the main sequence, their radii slowly grow and magnetic braking dominates the stellar angular momentum evolution, significantly spinning-down the stars over long timescales.
+Once the stars reach the main sequence, their structure changes slowly, allowing magnetic braking to dominate the stellar angular momentum evolution, significantly spinning-down the stars over long timescales.
 The $r_g$ evolution noticably differs between the stars as the M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
@@
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{Stellar radius (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} magnetic braking law.
+   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} magnetic braking law.
 Each dot denotes the approximate time when each star reaches the main sequence.}%
@@
 Under these assumptions, the tidal evolution is analogous to a driven, damped harmonic oscillator \citep{Greenberg2009}.
+For stars, equilibrium tidal models assume that tidal forces dissipate energy in convective regions via viscous turbulence \citep[see][]{Zahn2008}, corresponding to tidal dissipation in the outer convective layers of low-mass stars.
 We refer the reader to \citet{Barnes2017} for an in-depth discussion of the assumptions and limitations of equilibrium tidal models.
-Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citet[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
-For stars, equilibrium tidal models assume that tidal forces dissipate energy in convective regions via viscous turbulence \citep[see][]{Zahn2008}, corresponding tidal dissipation in the outer convective layers of low-mass stars.
+Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citep[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
 Here, we consider two equilibrium tidal models to study the secular spin-orbital evolution of low-mass stellar binaries.
@@
 \end{equation}
-Therefore, the CPL model predicts synchronous rotation for $e \lsim 0.23$, and a supersychronous 3:2 spin-orbit state otherwise.
+Therefore, the CPL model predicts synchronous rotation for $e \lsim 0.23$, and a supersychronous 3:2 spin-orbit state otherwise for tidally-locked rotators.
 
@@
 
-\subsection{Coupled Stellar-Tidal Evolution} \label{sec:coupled}
+\subsection{Coupled Stellar-Tidal Evolution For Tidally-Locked Systems} \label{sec:coupled}
 
-Following \citet{Fleming2018}, when one or both binary stars are tidally locked, we assume that tidal forces prevent magnetic braking from spinning down the tidally-locked star(s), but instead, any angular momentum lost is lost at the expense of the binary orbit, decreasing $a$ as a result \citep{Verbunt1981}.
+Following \citet{Fleming2018}, when one or both binary stars are tidally locked, tidal forces prevent magnetic braking from spinning down the tidally-locked star(s), and any angular momentum lost is lost at the expense of the binary orbit, decreasing $a$ as a result \citep{Verbunt1981}.
 Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify the $a$ decay equations due to stellar evolution and magnetic braking in tidally-locked binaries from \citet{Fleming2018}, their Eqs.
@@
 Given the inherent uncertainty in and complexity of the formation of binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$] and uniformly sampling the initial P$_{orb}$ over [$3,100$] d.
-Altough the CTL model is applicable for $e > 0.3$ while the CPL model is not, we restrict $e \leq 0.3$ to allow us to compare both models.
+Although the CTL model is applicable for $e > 0.3$, the CPL model is not, so we restrict $e \leq 0.3$ to allow us to compare both models.
+We do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binaries' dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.
 
-Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain, likely depend on complex viscous evolution within the stars \citep{Ogilvie2007}, differ for stars of the same spectral clase \citep{Barker2009}, and can vary as a function of stellar age \citep{Bolmont2016}.
+Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain, likely depend on complex viscous evolution within the stars \citep{Ogilvie2007}, can differ for stars of the same spectral class \citep{Barker2009}, and can vary as a function of stellar age \citep{Bolmont2016}, likely due to low-mass star's evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
 Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2009}, however a wide range of values exist in the literature.
-Therefore, we consider a wide range of tidal parameters and sample stellar tidal Qs loguniformly over $[10^4,10^7]$ and $\tau$ loguniformly over $[10^{-2},1]$ s.
+Therefore, we consider a wide range of tidal parameters by sampling stellar tidal Qs loguniformly over $[10^4,10^7]$ and $\tau$ loguniformly over $[10^{-2},1]$ s.
 
@@
 
-\section{The Rotational Equilibrium State} \label{sec:eq}
+\section{Results} \label{sec:results}
+
+\subsection{The Rotational Equilibrium State} \label{sec:eq}
+
+we actually see subsync rotation in Kepler-47, RUP whatever it's number (short period binaries!), and in the Kepler field!
+
+binaries WILL tidally-lock in the long term as P$_{eq}$ will decrease because P$_{orb}$ decreases in the long-run due to tidal energy dissipation (e.q.
+Eqn.~(\ref{eqn:tidal_locked_two}), adot from Fleming+2018 coupled stellar-tidal evolution, reference example tidal example above, Fleming+2018 paper).
+As P$_{orb}$ decreases, closer to P$_{rot}$ so easier locking, AND tidal torques become strong as distance decreases, allowing them to overpower magnetic braking.
+Can see this is slope of P$_{eq}/$P$_{rot}$ curve steepening.
+Dissipation occurs because weak tidal torques want to balance out magnetic braking, keeping Prot not equal to Peq, producing tidal dissipation.
+Tides plus magnetic braking still constantly spinning star down.
+Systems WILL inevitably tidally-lock, but at MUCH longer timescales as most of our Peq simulations do not lock by 7 Gyr even though Prot=Peq passings occured, but not locking because magnetic braking overpowered tidal torques at that time.
 
@@
 
-\section{Role $Q$ and $\tau$}
+\subsection{Role $Q$ and $\tau$}
 
@@
 %% SECTION: COMPARISON TO KEPLER %%
-\section{Comparison to \kepler} \label{sec:kepler}
+\subsection{Comparison to \kepler} \label{sec:kepler}
 
-XXX
+- Consult Lurie's paper for CPL vs CTL (e.g.
+why not 3:2 population-> use CTL!) and for discussing sunsync population.
+stars prob not in 3:2 resonance since they likely dont have a fixed solid body shape as I assume, but perhaps in a decoupled core envelope model (bouvier?) envelope could get locked into resonance
+
+- cite Matsen+2018 for "more than $50\%$ of the binary period distribution will be imaged in about $85\%$ of the host stars" in paper
+about detecting unresolved binaries in TESS
+
+- we use Lurie+2017 Prots that "Assuming solar-like differential rotation, P1,min will be closest to the equatorial rotation period.
+This provides a consistent reference point for the differential rotation discussion below."
+
+- compare results to Lurie section 4.4.2 discussion of dependence with eccentricity
+
+- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques
+
+- we can explain fast rotation observed in \citet{Meibom2007} to be due to tidal torques speeding up stellar Prots
 
@@
 	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
-   \caption{Caption}%
+   \caption{P$_{orb}/$P$_{rot}$ as a function for P$_{orb}$ simulated according to the CPL model (left) and the CTL model (right) colored by $e$, compared with \kepler EBs from \citet{Lurie2017} (red plus signs).
+Each simulation is integrated to system ages uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
+The \kepler EBs at low P$_{orb}$ and low P$_{orb}/$P$_{rot}$ are likely either brown dwarfs or exoplanets \citep{Lurie2017}, and hence not modeled by our simulations, so we do not consider them.
+TODO.}%
     \label{fig:lurie7}%
@@
 
-\subsection{P$_{orb} < 10$ d: The Subsynchronous Population} \label{sec:subsync}
+\subsubsection{P$_{orb} < 10$ d: The Subsynchronous Population} \label{sec:subsync}
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/subsync.pdf}
-   \caption{P$_{orb}/$P$_{rot}$ as a function for P$_{orb}$ simulated according to the CTL model (dots) compared with \kepler EBs from \citet{Lurie2017} (diamonds), colored by $e$.
+   \caption{Same format as Fig.~\ref{fig:lurie7}, with \citet{Lurie2017} data plotted as diamonds.
+All points are colored according to the binary orbital $e$.
 Our simulations are in excellent agreement with the observed orbital and rotational states of \kepler EBs.
-Our model predicts that all EBs with P$_{orb} < 3$ d are synchronized as is observed by \citet{Lurie2017}.
-We find that our model predicts a population of supersyncronous rotators on eccentric orbits and a population of subsyncronous rotators on nearly-circular orbits, both of which were observed by \citet{Lurie2017}.
+Our model predicts that nearly all EBs with P$_{orb} < 3$ d are synchronized as is observed by \citet{Lurie2017}.
+We find that our model predicts a population of supersyncronous rotators on eccentric orbits and a population of subsyncronous rotators with P$_{orb}/$P$_{rot} \approx 0.9$ on nearly-circular orbits, both of which were observed by \citet{Lurie2017}.
 This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL formalism can accurately model the tidal evolution of low-mass stellar binaries.}%
@@
 
-\subsection{$Q$, $\tau$ Constraints from P$_{orb}$ and P$_{rot}$} \label{sec:qTau}
+\subsubsection{$Q$, $\tau$ Constraints from P$_{orb}$ and P$_{rot}$} \label{sec:qTau}
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/qLurie.pdf}
-   \caption{Caption}%
+   \caption{Distribution of stellar tidal $Q$ produced by convolving \citet{Lurie2017} observations of EBs' P$_{rot}$ and P$_{orb}$ with our CPL simulation results shown in Fig.~\ref{fig:qmap}.
+The dashed black lines, from left to right, indicate the $16^{th}$ percentile, median, and $84^{th}$ percentile.
+Our $Q$ constraints are consistent with our flat prior distribution over log$_{10}(Q)$, suggesting that tidally-interacting stellar binaries can have a wide range of $Q$.}%
     \label{fig:qLurie}%
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/tauLurie.pdf}
-   \caption{Caption}%
+   \caption{Same as Fig.~\ref{fig:qLurie}, but using our CTL simulation results from Fig.~\ref{fig:taumap}.
+We find log$_{10}(\tau) = -0.73^{+0.52}_{-0.82}$ with an increase in density at larger $\tau$ deviating from our prior flat distribution, suggesting that time lags in low-mass stars are larger than the typical value of $\tau \approx 0.1$ s.
+A wide range of values for $\tau$ are still consistent with observations, however, indicating that tidally-interacting binaries can exhibit a wide range of tidal time lags, albeit with a preference for longer lags.}%
     \label{fig:tauLurie}%
@@
 
-\section{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
+\subsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
 
@@
 	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
-   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations.
+   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
 The P$_{rot}$ distribution for each case, marginalized over stellar mass, is depicted in the right panel.
@@
 Both tidal models predict a substantial population of fast rotators (P$_{rot} < 20$ d) not seen in the single star model at any age, except for young (age $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$).
-In many binary systems, both tidal models predict that age does not directly correlate with P$_{rot}$, especially in short P$_{orb}$ binaries where tidal torques are stronger, compared to the clear age-dependence of P$_{rot}$ seen in the single star models, the key assumption of gyrochronology methods.}%
+In many binary systems, both tidal models predict that age does not directly correlate with P$_{rot}$, especially in short P$_{orb}$ binaries where tidal torques are stronger, compared to the clear age-dependence of P$_{rot}$ seen in the single star models, the key assumption of gyrochronology methods.
+Any application of gyrochonlogy methods to age stars, especially those with P$_{rot} < 20$ d, should rule out or account for binarity.}%
     \label{fig:protDist}%
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
-   \caption{Caption.}%
+   \caption{Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
+Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
+Right: P$_{rot}$ distribution for each case, marginalized over stellar mass.
+Tidally-locked stars systematically rotate faster (median P$_{rot} = 27.0$ d) than interacting (median P$_{rot} = 51.8$ d) and not locked (median P$_{rot} = 41.1$ d) binaries as the locked population typically have shorter P$_{orb}$ where tidal torques are stronger, although binaries up to P$_{orb} = 100$ d tidally-lock.
+More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale with stellar mass, reflected by the enhanced density of locked systems at larger masses.
+The interacting population tends to rotate more slowly than the unlocked population as at short P$_{orb}$, and hence P$_{rot}$, the systems preferentially tidally-lock.
+At longer P$_{orb}$, magnetic braking spin down the stars while tidal torques sheppard P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
+$35\%$ of stars are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries.
+The P$_{rot}$ - mass distribution for not locked binaries resembles the single star sequence observed in the center right panel of Fig.~\ref{fig:protDist}.}%
     \label{fig:lockedCPL}%
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
-   \caption{Caption.}%
+   \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations.
+According to the CTL model, only $2\%$ of stellar binaries tidally-locked, mostly at short P$_{orb}$ (median P$_{orb} = 4.8$ d), where tidal torques are strong.
+The  $15\%$ of stars are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries.
+As in Fig.~\ref{fig:lockedCPL}, interacting binaries cluster at longer P$_{rot}$ as magnetic braking overpowers tidal torques to spin down stars, but tidal torques are strong enough to sheppard P$_{rot}$ near P$_{eq}$.}%
     \label{fig:lockedCTL}%
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/lockedProtHist.pdf}
-   \caption{Caption.}%
-    \label{fig:lockedInteracting}%
+   \caption{P$_{rot}$ distribution for tidally-locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.
+Strong tidal torques in the CPL model can tidally-lock binaries out to P$_{rot} = 100$ d, while weaker torques in the CTL model can only tidally-lock binary stars into a rapid rotation state with P$_{rot} < 20$ d.
+Under both models, binaries with a wide range of P$_{rot}$ can be strongly tidally-interacting.}%
+    \label{fig:lockedProtHist}%
 \end{figure}
+
+interacting systems tend to cluster near more massive stars for intermediate periods since their PMS lifetimes are shorter, so they have already spin down and tidal torques are stronger in those systems
 

### 2018-12-04 Paper/sync.tex

@@
 We compare our model predictions with \kepler eclipsing binary (EB) rotation periods and find that our predictions are in excellent agreement with observations, with the ``Constant Time Lag" equilibrium tidal best matching observations.
-We show that even for short P$_{orb}$ binaries, tidal torques can balance magnetic braking, naturally producing a population of subsynchronous rotators similar what \citet{Lurie2017} observed in their sample of \kepler EBs.
-By comparing observations of \kepler EBs with our simulations, we constrain the tidal time lag to $\tau = 0.19^{+0.43}_{-0.16}$ s.
+We show that even for short P$_{orb}$ binaries, tidal torques can balance magnetic braking, producing a population of subsynchronous rotators similar what \citet{Lurie2017} observed in their sample of \kepler EBs.
+By comparing observations of \kepler EBs with our simulations, we constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.
 
@@
 Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially so for the stellar spins.
-Tidal torques drive stellar spins towards the tidally-locked state in which the stellar P$_{rot}$ are equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
+Tidal torques drive stellar spins towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
 Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
 
-The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d, their detection limit.
+The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
 Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking.
 These findings suggest that that synchronization is not a given for tidally-interacting, short P$_{orb}$  stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$.
-Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass, short-period eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
+Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
 
@@
 Praesepe, ][]{Douglas2017} and field stars \citep[e.g.
-\kepler, ][]{McQuillan2014,Simonian2018} are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys.
+\kepler, ][]{McQuillan2014} are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys.
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P${rot}$ distributions.
@@
 Here, we examine angular momentum evolution over the full main sequence lifetimes of low-mass stellar binaries using a realistic treatment of stellar evolution, magnetic braking, and tidal torques.
-We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary orbital period and tidal dissipation parameters for two widely-used equilibrium tidal models.
+We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.
 We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to fail in such systems.
@@
 Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, and due to the inherant discreteness of numerical integrations, numerical solutions for the CPL and CTL models can produce unphysical evolution.
-We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively when P$_{rot}$ is within $1\%$ of P$_{eq}$.
+We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively, when P$_{rot}$ is within $1\%$ of P$_{eq}$.
 To ensure that tidal torques dominate over torques due to magnetic braking and stellar evolution, we additionally require that the P$_{rot}$ derivative points towards $P_{eq}$ on both sides of P$_{eq}$, i.e.
@@
 Therefore, we consider a wide range of tidal parameters by sampling stellar tidal Qs loguniformly over $[10^4,10^7]$ and $\tau$ loguniformly over $[10^{-2},1]$ s.
+All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can impart significant dynamical changes in the system \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.
 
@@
 
-\subsection{The Rotational Equilibrium State} \label{sec:eq}
+\subsection{The Weak Tidal Torque Regime} \label{sec:eq}
 
-we actually see subsync rotation in Kepler-47, RUP whatever it's number (short period binaries!), and in the Kepler field!
-
-binaries WILL tidally-lock in the long term as P$_{eq}$ will decrease because P$_{orb}$ decreases in the long-run due to tidal energy dissipation (e.q.
-Eqn.~(\ref{eqn:tidal_locked_two}), adot from Fleming+2018 coupled stellar-tidal evolution, reference example tidal example above, Fleming+2018 paper).
-As P$_{orb}$ decreases, closer to P$_{rot}$ so easier locking, AND tidal torques become strong as distance decreases, allowing them to overpower magnetic braking.
-Can see this is slope of P$_{eq}/$P$_{rot}$ curve steepening.
-Dissipation occurs because weak tidal torques want to balance out magnetic braking, keeping Prot not equal to Peq, producing tidal dissipation.
-Tides plus magnetic braking still constantly spinning star down.
-Systems WILL inevitably tidally-lock, but at MUCH longer timescales as most of our Peq simulations do not lock by 7 Gyr even though Prot=Peq passings occured, but not locking because magnetic braking overpowered tidal torques at that time.
-
-XXX
+Here we examine the P$_{rot}$ evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a focus on systems with weak tidal torques, i.e.
+long P$_{orb}$ and large $Q$ or small $\tau$, to examine the boundry between evolution dominated by tides or magnetic braking.
+In Fig.~\ref{fig:eqPer}, we plot the evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) formalisms.
+Both models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
+The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily to short P$_{orb}$ systems.
 
@@
 The black dotted line indicates that tidally-locked state where P$_{rot} = $ P$_{eq}$.
+Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.
+We truncate each curve when the binary tidally-locks.
 For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
-For all P$_{orb}$, all CPL models tidally-lock into synchronous rotation whereas the CTL model predicts subsynchronous (P$_{rot} > $ P$_{eq}$) rotation.
-In the CTL cases, the tidal torques are weaker than the magnetic braking torque, causing the stars to continue to spin down, preventing the systems from tidally-locking.
-Tidal torques strengthen for P$_{rot}$ far from P$_{eq}$, and magnetic braking torques weaken for long P$_{rot}$ \citep{Kawaler1988,Matt2015}, causing tidal torques to overpower magnetic braking, driving the stellar rotation back towards the tidally-locked state.
-Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.
-We truncate each curve when the respective system tidally-locks.
-Initially, all systems spin-up due to stellar contraction, except for the P$_{orb} = 5$ d case where tides overpower other torques, rapidly driving the stars towards synchronization.
-Once the stars reach the main sequence at ${\sim} 60$ Myr, tidal torques and magnetic braking combine to spin down the stars towards the tidally-locked state.
-For stars at long P$_{orb}$ that do not tidally-lock, after having spun-down to long P$_{rot}$, magnetic braking weakens, allowing tidal torques to sheppard P$_{rot}$ towards P$_{eq}$ as the P$_{rot}$ derivatives stay near 0, but with a slight negative value.}%
+For all P$_{orb}$, the CPL models tidally-lock into synchronous rotation whereas the CTL model predicts subsynchronous (P$_{rot} > $ P$_{eq}$) rotation that persists for Gyrs.}%
     \label{fig:eqPer}%
@@
 
-XXX
+% caption extra: 
+%In the CTL cases, magnetic braking overpowers tidal torques, spinning down the stars, preventing tidal-locking.
+Tidal torques eventually overpower magnetic braking for long P$_{rot}$ \citep{Kawaler1988,Matt2015}, slowly driving the stellar rotation back towards the tidally-locked state.
+% Initially, all systems spin-up due to stellar radius contraction.
+Once the stars reach the main sequence at ${\sim} 60$ Myr, tidal torques and magnetic braking combine to spin down the stars towards the tidally-locked state.
+For many CTL models, $\dot{\mathrm{P}}_{rot} > 0$ as magnetic braking overpowers tidal torques causing sustained subsynchronous rotation.
+After spinning down to long P$_{rot}$, magnetic braking weakens, allowing tidal torques to sheppard P$_{rot}$ towards P$_{eq}$ as the P$_{rot}$ derivatives stay near 0, but with a slight negative value.
 
-\subsection{Role $Q$ and $\tau$}
+The CTL model, however, predicts that for P$_{orb} \gsim 10$ d, magnetic braking overpowers weak tidal torques, spinning down the stars past the tidally-locked state.
+Magnetic braking pushes P$_{rot}/$P$_{eq} > 1$ with the maximum value set by the torque balance between tidal torques and magnetic braking.
+The maximum values grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
+Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution.
+After reaching the maximum, however, P$_{rot}$ decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
+First, as P$_{rot}$ continue to slow, magnetic braking weakens as its torque scales as $\omega$ or $\omega^3$ \citep{Kawaler1988}, depending on if the stars are in the saturated or unsaturated regime, respectively \citep{Matt2015}.
+Second, as P$_{rot}$ slows down further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example).
+Third, when P$_{rot} > $ P$_{eq}$, P$_{orb}$ decreases as tides transfer angular momentum from the orbit into stellar rotations, strengthening tidal torques that depend on the binary separation as $a^{-6}$.
+These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepparing them towards P$_{eq}$ in the long term.
+
+We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to stellar evolution, magnetic braking, and tidal torques.
+Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 100 Myr.
+After $R$ contraction, $\dot{\mathrm{P}}_{rot} > 0$ as tides and magnetic braking combine to spin down stars towards the tidally-locked state.
+For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$, but $\ddot{\mathrm{P}}_{rot} < 0$ as the three processes described above gradually strengthen tidal torques relative to magnetic braking.
+In the long-run, tidal torques overpower magnetic braking, with a slightly negative P$_{rot}$ derivative, slowly driving P$_{rot}$ towards P$_{eq}$ (see the P$_{orb} = 20, 40$ d cases in Fig.~\ref{fig:eqPer}).
+This slowly shifting torque balance produces a population of subsynchronous rotators with P$_{rot} \gsim $ P$_{eq}$ that can persist for Gyrs.
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorb.pdf}
+   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively.
+Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally-lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
+}%
+    \label{fig:eqPerShortPorb}%
+\end{figure}
+
+% caption extra:
+% In the long-term, magnetic braking weakens relative to tidal torques for longer P$_{rot}$, allowing tides to drive the stars back towards the tidally-locked state.
+
+Spin-orbit synchronization should not be assumed for short P$_{orb}$ binaries.
+In Fig.~\ref{fig:eqPerShortPorb}, we plot the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters and find that subsynchronous rotation occurs in the general case of weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
+Moreover, short P$_{orb}$ subsynchronous rotators can eventually tidally-lock via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
+Short P$_{orb}$ subsynchronous binaries exist in nature, such as some \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012} and a Ruprecht 147 eclipsing binary \citep{Torres2018}, suggesting that this theoretical observation is real.
+We explore this effect further in $\S$~\ref{sec:subsync} and describe how we can use it to constrain the tidal parameters $Q$ and $\tau$.
+
+\subsection{Influence of P$_{orb}$, $Q$ and $\tau$}
+
+We examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, respectfully, and connect them to observables.
+In Fig.~\ref{fig:qTauLock}, we bin our simulation results after the full 7 Gyr evolution by P$_{orb}$ and $Q$ or $\tau$ and compute the median P$_{orb}/$P$_{rot}$ in each bin, marginalizing over all other parameters.
+
+Spin-orbit synchronization is the typical outcome for tidally-interacting binaries according to the CPL model, with synchronization occurring for most binaries with P$_{orb} < 10$ d, regardless of $Q$.
+The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to P$_{orb} \approx 100 $ d for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
+According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, regardless of $\tau$.
+Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e.
+stars with large $Q$ or small $\tau$, that form via the mechanism described in $\S$~\ref{sec:eq}, with weaker the tidal torques allowing more subsynchronous rotation.
+The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) with P$_{orb} > 80$ d does not in general correspond to binaries tidally-locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally-locked state.
+Under both tidal models, there is a population of nearly synchronous rotators near P$_{orb} \approx 60$ d.
+This population corresponds to the evolution described in $\S$~\ref{sec:eq} in which magnetic braking initially spins down stars past the tidally-locked state, but in the long-term, tidal torques spin up the stars, shepparding them towards the tidally-locked state.
+This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ (see Figs.~\ref{fig:eqPer},\ref{fig:eqPerShortPorb}).
 
@@
 	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
-   \caption{Median P$_{orb}/$P$_{rot}$ for the primary star at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
-The CPL model predicts that stellar binaries can tidally-lock out to P$_{orb} = 30$ d for most values of $Q$, with some locking all the way up to P$_{orb} = 100$ d for small $Q < 10^5$, substantially longer than what is expected.
-For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally-lock into synchronous rotation, regardless of $Q$.
-The CTL model, however, requires stars to have $\tau \geq 0.1$ s in order for stellar binaries with P$_{orb} < 10$ d to likely tidally-lock.
-Both models exhibit supersynchronous rotation (blue regions, P$_{rot} < $P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at these long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spin down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
-For large $Q$ ($Q > 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spining down stars past the tidally-locked state, producing a population of subsynchronous (red regions, P$_{rot} > $P$_{orb}$) rotators.}%
+   \caption{Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
+For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally-lock into synchronous rotation, regardless of $Q$, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.
+Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spin down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
+For large $Q$ ($Q > 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous (red regions, P$_{rot} >$ P$_{orb}$) rotators.}%
     \label{fig:qTauLock}%
@@
 
-XXX
+% caption extra:
+% The CPL model predicts that stellar binaries can tidally-lock out to P$_{orb} = 30$ d for most values of $Q$, with some locking all the way up to P$_{orb} = 100$ d for small $Q < 10^5$.
+
+From Fig.~\ref{fig:qTauLock}, certain rotation states, e.g.
+synchronous rotation, typically occur in certain regions of parameter space, e.g.
+where tidal torques are dominant, small $Q$ or large $\tau$.
+Therefore by measuring P$_{rot}$ and P$_{orb}$ for the primary star in a binary system, one can in principle constrain the strength of the tidal interaction that led to that rotation state.
+We do so in Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap} by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$ and computing the median $Q$ and $\tau$, and the $16\%$, and $84\%$ percentiles, in each bin.
 
@@
    \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
-In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$.
-For all P$_{orb}$, synchronous and supersynchronous rotators have systematically low $Q$s as strong tidal torques are required to tidally-lock the systems.
-Note that for P$_{orb} < 10$ d, there are no rotators with $1.0 \leq$ P$_{orb}/$P$_{rot} \leq 1.5$ as strong tidal torques always tidally-lock the system, forcing P$_{rot} =$ P$_{eq}$, which can only take two values under the CPL model (see Eqn.~(\ref{eqn:cpl:eqPer})).
-Subsynchronous rotators have large $Q$s, and hence weak tidal torques, that are dominated by magnetic braking that spin-down the stars past the tidally-locked state.
-For P$_{orb} < 70$ d, no stars have P$_{orb}/$P$_{rot} > 1.5$ as in the CPL model, supersynchronous rotators can only tidally-lock into a 3:2 spin-orbit commensurability.
-At longer P$_{orb}$, supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.}%
+In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$ and display these values.
+Narrow constraints imply that certain tidal parameters are required to produce that rotation state, e.g.
+supersynchronous rotators near P$_{orb}/$P$_{rot} = 1.5$ require $Q \lsim 10^{5.5}$.}%
     \label{fig:qmap}%
@@
 
-XXX
+For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems, especially for P$_{orb} > 20$ d.
+For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ as strong tidal torques always tidally-lock the binaries.
+In this case, P$_{eq}$ can only take two discrete values under the CPL model, i.e.
+Eqn.~(\ref{eqn:cpl:eqPer}).
+No stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 70$ d as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability.
+
+Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking that spin-down the stars past the tidally-locked state.
+This effect can be seen for P$_{orb} < 50$ d where the median $Q$ gradually increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state that require low $Q$.
+This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.
+In this case, the more supersynchronous the rotation, the weaker the tidal torques must be, and hence the larger the $Q$ must be, as they fail to spin down stars towards P$_{orb}$.
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbTauHist.pdf}
-   \caption{Same format as Fig.~\ref{fig:qmap}, but for the CTL model.
-As under the CPL model, stars tidally-lock when subject to strong tidal torques, typically for P$_{orb} < 50$ d and $\tau \gsim 0.25$ s.
-For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with low $tau < 0.1$ s as magnetic braking dominates over tidal torques, spinning down the star.
-For P$_{orb} > 50$ d, tides do not in general dominate the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s has upper and lower limits that encompass most of the range we considered for $\tau$.
-The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
-We do not observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).}%
+   \caption{Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
     \label{fig:taumap}%
@@
 
-XXX
+As under the CPL model, stars tidally-lock when subject to strong tidal torques, typically for P$_{orb} < 50$ d and $\tau \gsim 0.25$ s.
+For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with $tau < 0.1$ s as magnetic braking dominates over tidal torques, spinning down the star.
+For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s has upper and lower limits that encompass most of the range we considered for $\tau$, suggesting that tides are not important in this region of parameter space.
+The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
+We do not observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).
 
@@
 \subsection{Comparison to \kepler} \label{sec:kepler}
+
+Here we compare our simulation results to P$_{rot}$ measurements of primary stars in stellar binaries by \citet{Lurie2017}.
+\citet{Lurie2017} measured 816 rotation periods for \kepler EBs with star spot modulations and visually inspected each light curve to ensure their accuracy.
+The \citet{Lurie2017} dataset is the largest homogenous set of P$_{rot}$ measurements available for low-mass stellar binaries and represents the state of the art benchmark for studies of tidal interactions in stellar binaries.
+We compare our results to the P$_{1,min}$ P$_{rot}$ values reported by \citet{Lurie2017} as the authors demonstrated that these values are likely to be close to the equatorial P$_{rot}$ that we track in our simulations.
+
+In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
 
@@
 
-- cite Matsen+2018 for "more than $50\%$ of the binary period distribution will be imaged in about $85\%$ of the host stars" in paper
-about detecting unresolved binaries in TESS
-
-- we use Lurie+2017 Prots that "Assuming solar-like differential rotation, P1,min will be closest to the equatorial rotation period.
-This provides a consistent reference point for the differential rotation discussion below."
-
-- compare results to Lurie section 4.4.2 discussion of dependence with eccentricity
+- compare results to Lurie section 4.4.2 discussion of dependence with eccentricity \citet{Burkart2014}
 
@@
 
-- we can explain fast rotation observed in \citet{Meibom2007} to be due to tidal torques speeding up stellar Prots
-
 \begin{figure*}[t]
 	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
-   \caption{P$_{orb}/$P$_{rot}$ as a function for P$_{orb}$ simulated according to the CPL model (left) and the CTL model (right) colored by $e$, compared with \kepler EBs from \citet{Lurie2017} (red plus signs).
-Each simulation is integrated to system ages uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
-The \kepler EBs at low P$_{orb}$ and low P$_{orb}/$P$_{rot}$ are likely either brown dwarfs or exoplanets \citep{Lurie2017}, and hence not modeled by our simulations, so we do not consider them.
-TODO.}%
+   \caption{P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ simulated according to the CPL model (left) and the CTL model (right) colored by $e$, compared with \kepler EBs from \citet{Lurie2017} (red plus signs).
+The \kepler EBs at low P$_{orb}$ and low P$_{orb}/$P$_{rot}$ are likely either brown dwarfs or exoplanets \citep{Lurie2017}, and hence not modeled by our simulations, so we do not consider them, but we display them for completeness.}%
     \label{fig:lurie7}%
@@
 Our simulations are in excellent agreement with the observed orbital and rotational states of \kepler EBs.
+}%
+    \label{fig:subsync}%
+\end{figure}
+
 Our model predicts that nearly all EBs with P$_{orb} < 3$ d are synchronized as is observed by \citet{Lurie2017}.
 We find that our model predicts a population of supersyncronous rotators on eccentric orbits and a population of subsyncronous rotators with P$_{orb}/$P$_{rot} \approx 0.9$ on nearly-circular orbits, both of which were observed by \citet{Lurie2017}.
-This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL formalism can accurately model the tidal evolution of low-mass stellar binaries.}%
-    \label{fig:subsync}%
-\end{figure}
+This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL formalism can accurately model the tidal evolution of low-mass stellar binaries.
 
@@
 Binaries readily tidally-lock at short P$_{orb}$ due to the tidal force's strong semi-major axis dependence, increasing the strength of tidal torques at shorter orbital separations.
-When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} = $P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.
+When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} =$ P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.
 
@@
 
-XXX
-
 %% SECTION: DISCUSSION %%
@@
 XXX
+
+- cite Matsen+2018 for "more than $50\%$ of the binary period distribution will be imaged in about $85\%$ of the host stars" in paper
+about detecting unresolved binaries in TESS
+
+- consider dynamical tidal effects (Zahn 1975)?
 

### 2018-12-07 Paper/sync.tex
Updates from ShareLaTeX

@@
 
-Stellar rotation period (P$_{rot}$) studies have produced numerous insights into the angular momentum evolution of low-mass stars, but previous investigations have focused on single stars even though roughly half of Sun-like stars are in stellar binaries.
-Here, we examine how tidal torques, stellar evolution, and magnetic braking shape the P$_{rot}$ evolution of low-mass stellar binaries by performing Monte Carlo simulations of coupled stellar-tidal evolution.
+ Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries, but few previous studies have examined tides in binaries out to intermediate orbital periods (P$_{orb}$).
+Here, we examine how tidal torques, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries with P$_{orb} \leq 100$ d by performing Monte Carlo simulations of coupled stellar-tidal evolution.
 We find that tides can modify stellar rotations in binaries with orbital periods (P$_{orb}$) up to 100 days.
-At short P$_{orb}$, many stellar binaries tidally-lock into synchronous rotation, naturally producing a population of fast rotators (P$_{rot} < 10$ d) across all stellar masses as is observed in the \textit{Kepler} field, a feature that cannot be reproduced by single-star-only models of magnetic braking.
+At short P$_{orb}$, many stellar binaries tidally-lock into synchronous rotation, naturally producing a population of fast rotators (P$_{rot} < 10$ d) as is observed in the \textit{Kepler} field, a feature that cannot be reproduced by single-star-only models of magnetic braking.
 Many binaries with longer P$_{orb}$ can tidally-lock, fixing the stellar P$_{rot}$ to the binary P$_{orb}$, causing P$_{rot}$ to not be a valid proxy for age in all cases, i.e.
 gyrochronology methods must be applied carefully.
-We compare our model predictions with \kepler eclipsing binary (EB) rotation periods and find that our predictions are in excellent agreement with observations, with the ``Constant Time Lag" equilibrium tidal best matching observations.
-We show that even for short P$_{orb}$ binaries, tidal torques can balance magnetic braking, producing a population of subsynchronous rotators similar what \citet{Lurie2017} observed in their sample of \kepler EBs.
-By comparing observations of \kepler EBs with our simulations, we constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.
+We compare our model predictions with \kepler eclipsing binary (EB) rotation periods and find that our predictions are in excellent agreement with observations, with the ``Constant Time Lag" equilibrium tidal model best matching observations.
+We recommend observational tests to discriminate between equilibrium tidal models for stars.
+We show that when tidal torques are weak, even in short P$_{orb}$ binaries, tidal torques can overpower magnetic braking, producing a long-lasting population of subsynchronous rotators similar what \citet{Lurie2017} observed in their sample of \kepler EBs.
+By comparing observations of \kepler EBs with our simulations, we loosely constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.
 
@@
 Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}.
-Orbital circularization is ubiquitous in short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
+Orbital circularization is ubiquitous in short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989,Claret1995} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
 For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} finds that the orbit of solar twin binaries circularizes during the stellar pre-main sequence.
@@
 Tidal torques drive stellar spins towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
-Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
+Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
 
@@
 
-We simulate coupled stellar-tidal evolution for stellar binaries using an improved version of the coupled stellar-tidal evolution model presented in \citet{Fleming2018}.
+We simulate coupled stellar-tidal evolution for low-mass binaries using an improved version of the coupled stellar-tidal evolution model presented in \citet{Fleming2018}.
 We implement our model in the open-source code VPLanet\footnote{VPLanet is publicly available
@@
 Although the CTL model is applicable for $e > 0.3$, the CPL model is not, so we restrict $e \leq 0.3$ to allow us to compare both models.
-We do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binaries' dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.
+We do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.
 
-Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain, likely depend on complex viscous evolution within the stars \citep{Ogilvie2007}, can differ for stars of the same spectral class \citep{Barker2009}, and can vary as a function of stellar age \citep{Bolmont2016}, likely due to low-mass star's evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
-Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2009}, however a wide range of values exist in the literature.
-Therefore, we consider a wide range of tidal parameters by sampling stellar tidal Qs loguniformly over $[10^4,10^7]$ and $\tau$ loguniformly over $[10^{-2},1]$ s.
-All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can impart significant dynamical changes in the system \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.
+Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain, likely depend on complex viscous evolution within the stars \citep{Ogilvie2007}, can differ for stars of the same spectral class \citep{Barker2009}, and can vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass star's evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
+Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a wide range of values exist in the literature.
+Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s loguniformly over $[10^4,10^7]$ and $\tau$ loguniformly over $[10^{-2},1]$ s.
+Note that in general, there is no expression to compute $Q$ as a function of $\tau$, or vice versa, except in some special cases where approximations exist, e.g.
+Eqn.
+(2) from \citet{Heller2011}.
+All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk, that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017}, would likely have dissipated \citep{Haisch2001}.
 
@@
 Moreover, short P$_{orb}$ subsynchronous rotators can eventually tidally-lock via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
-Short P$_{orb}$ subsynchronous binaries exist in nature, such as some \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012} and a Ruprecht 147 eclipsing binary \citep{Torres2018}, suggesting that this theoretical observation is real.
+Short P$_{orb}$ subsynchronous binaries exist in nature, such as some \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012} and EPIC 219394517 \citep{Torres2018}, suggesting that this theoretical observation is real.
 We explore this effect further in $\S$~\ref{sec:subsync} and describe how we can use it to constrain the tidal parameters $Q$ and $\tau$.
@@
 In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
-
-- Consult Lurie's paper for CPL vs CTL (e.g.
-why not 3:2 population-> use CTL!) and for discussing sunsync population.
-stars prob not in 3:2 resonance since they likely dont have a fixed solid body shape as I assume, but perhaps in a decoupled core envelope model (bouvier?) envelope could get locked into resonance
-
-- compare results to Lurie section 4.4.2 discussion of dependence with eccentricity \citet{Burkart2014}
-
-- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques
+Both models reproduce most trends in the \citet{Lurie2017} data: the large cluster of synchronized rotators with P$_{orb} \lsim 20$ d, a population of subsynchronous rotators with P$_{orb} \lsim 20$ d, and the sparse cluster of supersynchronous rotators at all P$_{orb}$ that correspond to young stars that have yet to significantly spin down due to magnetic braking or tides.
+The lower limit seen in both panels is a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down via magnetic braking in the weak tides regime over 7 Gyr, the longest age considered in our simulations.
 
@@
 
+The predictions of the CPL model fail to reproduce several key aspects of that data.
+First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} \lsim 1.5$, a prediction that is entirely inconsistent with the data as no obvious spin-orbit commensurablities, aside from 1:1 synchronization, is present in the data likely owing to the fact that stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
+The CPL model can produce the subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
+The CTL model, however, excels where the CPL models fails and naturally produces both populations, suggesting that the CTL model better describes tidal interactions in low-mass binaries.
+We examine this point further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}.
+
 XXX
 
-\subsubsection{P$_{orb} < 10$ d: The Subsynchronous Population} \label{sec:subsync}
+\citet{Claret1995} find that supersynchronous rotation their sample of eccentric binaries are in good agreement with the predictions of tidal models.
+
+presence of 3:2 cluster would be strong evidence in favor of CPL model.
+Can use presence/absence of this feature as a strong discriminators between tidal models.
+additionally, CTL predicts more subsync clusters that could be used for model descriptions.
+below, say how CPL model has locked binaries at long Porb, ctl doesn't above 20 d so that's another model descriimination.
+say each time that CTL passes, CPL doesn't according to lurie data.
+In general, talk about implications and use flowery language
+
+\subsubsection{P$_{orb} < 10$ d According to the CTL Model} \label{sec:subsync}
 
@@
 
+- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques.
+
 Although Lurie+2017 attributes differential rotation to the production of the
 subsynchronous population, we find that coupled stellar-tidal evolution naturally
-produces the population. Compare with Lurie+2017 sample for 2 $<$ Porb $<$ 10 days:
+produces the population. it's probably a combination of both effects. Compare with Lurie+2017 sample for 2 $<$ Porb $<$ 10 days:
 $72\%$ in [0.92,1.2] and $15\%$ in [0.84, 0.92].
@@
 \subsubsection{$Q$, $\tau$ Constraints from P$_{orb}$ and P$_{rot}$} \label{sec:qTau}
+
+XXX: say how our quoted $Q$ and $\tau$ ranges produced binaries consistent with what \kepler saw, so not only is our model a reasonable choice to model stellar-tidal evolution, but also our quoted tidal parameter ranges seem to bracket what stars actually have.
+We refine it in the below analysis
 
@@
 \subsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
+
+XXX:
+
+-hammer implications for gyrochronology, fast rotators in kepler field, compare to \citet{Simonian2018}
+
+- \citet{Abt2004} find evidence for tides modifying P$_{rot}$ in binaries with P$_{orb}$ up to 500 d.
 
@@
 
+- talk about how we need more binary observations with Prot, Porb, and ecc, especially at longer orbital periods, to more reliably discriminate between tidal models.
+At low eccentricity, it does matter (sync vs supersync in \kepler) as opposed to what \citet{Leconte2010} recommends
+
 - cite Matsen+2018 for "more than $50\%$ of the binary period distribution will be imaged in about $85\%$ of the host stars" in paper
 about detecting unresolved binaries in TESS
-
-- consider dynamical tidal effects (Zahn 1975)?
 

### 2018-12-10 Paper/sync.tex
added some text

@@
 The CPL model can produce the subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
-The CTL model, however, excels where the CPL models fails and naturally produces both populations, suggesting that the CTL model better describes tidal interactions in low-mass binaries.
-We examine this point further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}.
-
-XXX
-
-\citet{Claret1995} find that supersynchronous rotation their sample of eccentric binaries are in good agreement with the predictions of tidal models.
-
-presence of 3:2 cluster would be strong evidence in favor of CPL model.
-Can use presence/absence of this feature as a strong discriminators between tidal models.
-additionally, CTL predicts more subsync clusters that could be used for model descriptions.
-below, say how CPL model has locked binaries at long Porb, ctl doesn't above 20 d so that's another model descriimination.
-say each time that CTL passes, CPL doesn't according to lurie data.
-In general, talk about implications and use flowery language
+The CTL model, however, seems to excel where the CPL models fails by producing both populations.
+We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
 
@@
 
-Median subsync tidal tau: 0.0790922272
-Median supersync tidal tau: 0.12508653609999998
-Median subsync ecc: 0.1288805
-Median supersync ecc: 0.151426
-Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6126373626373627
-Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.11401098901098901
+Median supersync ecc: 0.15
 
-- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques.
-
-Although Lurie+2017 attributes differential rotation to the production of the
-subsynchronous population, we find that coupled stellar-tidal evolution naturally
-produces the population. it's probably a combination of both effects. Compare with Lurie+2017 sample for 2 $<$ Porb $<$ 10 days:
-$72\%$ in [0.92,1.2] and $15\%$ in [0.84, 0.92].
+Our model predicts that nearly all EBs with P$_{orb} < 3$ d are synchronized as is observed by \citet{Lurie2017}.
+We find that our model predicts a population of supersyncronous rotators on eccentric orbits and a population of subsyncronous rotators with P$_{orb}/$P$_{rot} \approx 0.9$ on nearly-circular orbits, both of which were observed by \citet{Lurie2017}.
+This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL formalism can accurately model the tidal evolution of low-mass stellar binaries.
 
@@
 
-Our model predicts that nearly all EBs with P$_{orb} < 3$ d are synchronized as is observed by \citet{Lurie2017}.
-We find that our model predicts a population of supersyncronous rotators on eccentric orbits and a population of subsyncronous rotators with P$_{orb}/$P$_{rot} \approx 0.9$ on nearly-circular orbits, both of which were observed by \citet{Lurie2017}.
-This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL formalism can accurately model the tidal evolution of low-mass stellar binaries.
+Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking naturally
+produces the population.
+Our mechanism for producing long-lasting subsynchronous rotation described in $\S$~(XXX sec ref) does produce the population seen in Fig.~\ref{fig:subsync} as the sunsynchronous rotators systematically have weaker tidal torques, with a median $\tau = 0.13$ s, compared to the median $\tau = 0.08$ s for the supersynchronous population in which strong tides tidally-lock the binaries.
+
+Our CTL model successfully reproduces the \citet{Lurie2017} data not just by qualitatively matching the spin-orbital populations observed in \kepler, but also in specific populations.
+\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $<$ Porb $<$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $11\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
+Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $61\%$ of our CTL population.
+Both are close matches, especially given that we did not hand-tune our simulation initial conditions, suggesting that our coupled stellar-tidal model using the CTL formalism is an accurate description of the physics of tidally-interacting binaries.
+These findings suggest that the CTL model describes tidal interactions in low-mass binaries better than the CPL model, but in the absence of reliable uncertanties on the \citet{Lurie2017} data, we cannot definitively assert that in general, the CTL model is superior to the CPL model for tidal torques in low-mass binaries.
+We discuss means to discriminate between the two models in $\S$~\ref{sec:whichModel}.
 
@@
 
+\subsection{CPL or CTL?} \label{sec:whichModel}
+
+Here we offer observational tests to discriminate between the two models in short P$_{orb}$ binaries where tidal effects are more likely to be significant.
+First, one could obtain a series of time-resolved spectra of the short P$_{orb}$ supersynchronous binaries studied by \citet{Lurie2017} to derive more robust $e$ constraints than derived using the approximations in \citet{Lurie2017}.
+For binaries with P$_{orb} < 10$ and $e < \sqrt{1/19}$, the CPL model does not permit supersynchronous rotation except in extremely young systems that are still contracting along the pre-main sequence, or that have just reached the main sequence, while the CTL model does allow for it.
+If supersynchronous rotation is observed in such systems, it is evidence in favor of the CTL model.
+
+Second, for tidally-locked binaries with $e > \sqrt{1/19}$ for short P$_{orb}$, the CPL model predicts supersynchronous rotation in the form of a 3:2 spin-orbit comensurability, e.g.
+the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permited.
+The CTL model, however, permits a continuum of supersynchronous states in eccentric tidally-locked rotators, e.g.
+right panel of Fig.~\ref{fig:lurie7}, that seems to provide a much better match to the \citet{Lurie2017} data, providing strong evidence in favor of the CLT model.
+The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making this distinction conclusive in this work.
+Again, more robust $e$ constrains via time-dependent spectra would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a gaussian process model as in (XXX cite GP rasumussen, and correct ruth angus paper).
+Accurate determinations of the spin-orbital state of short P$_{orb}$ binaries are critical to determine which tidal model best describes the evolution of low-mass stellar binaries.
+
 %% SECTION: DISCUSSION %%
@@
 XXX
+
+presence of 3:2 cluster would be strong evidence in favor of CPL model.
+Can use presence/absence of this feature as a strong discriminators between tidal models.
+additionally, CTL predicts more subsync clusters that could be used for model descriptions.
+below, say how CPL model has locked binaries at long Porb, ctl doesn't above 20 d so that's another model descriimination.
+say each time that CTL passes, CPL doesn't according to lurie data.
+In general, talk about implications and use flowery language
 

### 2018-12-12 Paper/sync.tex
1st draft

--- 2018-12-07 Paper/sync.tex
@@
 
-\title{Rotation Periods in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking}
+\title{Rotation Period Evolution in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking}
 
@@
 
- Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries, but few previous studies have examined tides in binaries out to intermediate orbital periods (P$_{orb}$).
-Here, we examine how tidal torques, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries with P$_{orb} \leq 100$ d by performing Monte Carlo simulations of coupled stellar-tidal evolution.
-We find that tides can modify stellar rotations in binaries with orbital periods (P$_{orb}$) up to 100 days.
-At short P$_{orb}$, many stellar binaries tidally-lock into synchronous rotation, naturally producing a population of fast rotators (P$_{rot} < 10$ d) as is observed in the \textit{Kepler} field, a feature that cannot be reproduced by single-star-only models of magnetic braking.
-Many binaries with longer P$_{orb}$ can tidally-lock, fixing the stellar P$_{rot}$ to the binary P$_{orb}$, causing P$_{rot}$ to not be a valid proxy for age in all cases, i.e.
+ Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries.
+Here, we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d and across a wide range tidal dissipation parameters.
+In many cases, we find that magnetic braking can overpower tides, producing a long-lasting population of subsynchronous rotators, even in short P$_{orb}$ binaries, reproducing the population observed by \citet{Lurie2017} in their sample of \kepler eclipsing binaries (EBs).
+At short P$_{orb}$, many stellar binaries tidally-lock, naturally producing a population of fast rotators as is observed in the \textit{Kepler} field, a feature that cannot be reproduced by single-star-only models of magnetic braking.
+Many binaries can tidally-lock even out to P$_{orb} \approx 100$ d, fixing the stellar P$_{rot}$ to the binary P$_{orb}$, causing P$_{rot}$ to not be a valid proxy for age in all cases, i.e.
 gyrochronology methods must be applied carefully.
-We compare our model predictions with \kepler eclipsing binary (EB) rotation periods and find that our predictions are in excellent agreement with observations, with the ``Constant Time Lag" equilibrium tidal model best matching observations.
-We recommend observational tests to discriminate between equilibrium tidal models for stars.
-We show that when tidal torques are weak, even in short P$_{orb}$ binaries, tidal torques can overpower magnetic braking, producing a long-lasting population of subsynchronous rotators similar what \citet{Lurie2017} observed in their sample of \kepler EBs.
-By comparing observations of \kepler EBs with our simulations, we loosely constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.
+We compare our model predictions with \kepler EB rotation periods and find that our predictions are in excellent agreement with observations.
+We compare the predictions of two common equilibrium tidal models and show that the ``Constant Time Lag" equilibrium tidal model provides a better match to observations.
+We recommend observational tests to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
+By comparing observations of \kepler EBs with our simulations, we constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.
 
@@
 Praesepe, ][]{Douglas2017} and field stars \citep[e.g.
-\kepler, ][]{McQuillan2014} are likely contaminated by unresolved binaries, given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys.
-In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P${rot}$ distributions.
+\kepler, ][]{McQuillan2014} are likely contaminated by unresolved binaries given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys.
+In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.
 Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, imparting a contaminating signal that is not currently accounted for by modern models.
-Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that links the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, will likely be incorrect owing to the influence of tidal torques.
+Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that links the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, could be incorrect owing to the influence of tidal torques.
 No previous study has quantified this effect.
@@
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
-We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.
+We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:protDist} and $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.
 
@@
 Both models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
-The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily to short P$_{orb}$ systems.
+The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.
 
@@
 Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution.
-After reaching the maximum, however, P$_{rot}$ decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
-First, as P$_{rot}$ continue to slow, magnetic braking weakens as its torque scales as $\omega$ or $\omega^3$ \citep{Kawaler1988}, depending on if the stars are in the saturated or unsaturated regime, respectively \citep{Matt2015}.
+
+P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
+First, as P$_{rot}$ continue to slow due to magnetic braking, magnetic braking itself weakens as its torque scales as $\omega$ or $\omega^3$ \citep{Kawaler1988}, depending on if the stars are in the saturated or unsaturated regime, respectively \citep{Matt2015}.
 Second, as P$_{rot}$ slows down further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example).
-Third, when P$_{rot} > $ P$_{eq}$, P$_{orb}$ decreases as tides transfer angular momentum from the orbit into stellar rotations, strengthening tidal torques that depend on the binary separation as $a^{-6}$.
-These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepparing them towards P$_{eq}$ in the long term.
+Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$ and strengthening tidal torques that strongly depend on the binary separation as $a^{-6}$.
+These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long term.
 
 We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to stellar evolution, magnetic braking, and tidal torques.
-Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 100 Myr.
+Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 60 Myr when the stars reach the main sequence.
 After $R$ contraction, $\dot{\mathrm{P}}_{rot} > 0$ as tides and magnetic braking combine to spin down stars towards the tidally-locked state.
@@
 
-Spin-orbit synchronization should not be assumed for short P$_{orb}$ binaries.
-In Fig.~\ref{fig:eqPerShortPorb}, we plot the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters and find that subsynchronous rotation occurs in the general case of weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
-Moreover, short P$_{orb}$ subsynchronous rotators can eventually tidally-lock via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
-Short P$_{orb}$ subsynchronous binaries exist in nature, such as some \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012} and EPIC 219394517 \citep{Torres2018}, suggesting that this theoretical observation is real.
-We explore this effect further in $\S$~\ref{sec:subsync} and describe how we can use it to constrain the tidal parameters $Q$ and $\tau$.
+In Fig.~\ref{fig:eqPerShortPorb}, we examine this effect in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters.
+Subsynchronous rotation occurs in the general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
+Previous theoretical studies have also predicted subsynchronous rotation in short P$_{orb}$ binaries arising from the balance between tidal torques and magnetic braking \citep[e.g.][]{Habets1989,Zahn1994,Keppens1997} suggesting that this behavior is not an artifact of our choice of tidal or magnetic braking models, but rather a general outcome of the competition between magnetic braking and tidal evolution in low-mass binaries.
+Short P$_{orb}$ subsynchronous rotators can eventually tidally-lock via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
 
-\subsection{Influence of P$_{orb}$, $Q$ and $\tau$}
+Short P$_{orb}$ subsynchronous binaries exist in nature, such as many \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012}, EPIC 219394517 \citep{Torres2018}, and in ``Binary 6211" observed by \citet{Meibom2006}, suggesting that this theoretical observation is real and borne out in nature.
+Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries.
+We explore this effect further and test it against observations of \kepler EBs in $\S$~\ref{sec:subsync}.
 
-We examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, respectfully, and connect them to observables.
+\subsection{Influence of P$_{orb}$, $Q$ and $\tau$} \label{sec:qTauMaps}
+
+We examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, respectfully.
 In Fig.~\ref{fig:qTauLock}, we bin our simulation results after the full 7 Gyr evolution by P$_{orb}$ and $Q$ or $\tau$ and compute the median P$_{orb}/$P$_{rot}$ in each bin, marginalizing over all other parameters.
@@
 Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e.
-stars with large $Q$ or small $\tau$, that form via the mechanism described in $\S$~\ref{sec:eq}, with weaker the tidal torques allowing more subsynchronous rotation.
+stars with large $Q$ or small $\tau$, that form via the mechanism described in $\S$~\ref{sec:eq}, with weaker tidal torques allowing more subsynchronous rotation.
 The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) with P$_{orb} > 80$ d does not in general correspond to binaries tidally-locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally-locked state.
-Under both tidal models, there is a population of nearly synchronous rotators near P$_{orb} \approx 60$ d.
-This population corresponds to the evolution described in $\S$~\ref{sec:eq} in which magnetic braking initially spins down stars past the tidally-locked state, but in the long-term, tidal torques spin up the stars, shepparding them towards the tidally-locked state.
+
+Both tidal models predict a population of nearly synchronous rotators near P$_{orb} \approx 60$ d.
+This population corresponds to the evolution described in $\S$~\ref{sec:eq} in which magnetic braking initially spins down stars past the tidally-locked state, but in the long-term, tidal torques spin up the stars, shepherding them towards the tidally-locked state.
 This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ (see Figs.~\ref{fig:eqPer},\ref{fig:eqPerShortPorb}).
@@
 Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spin down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
-For large $Q$ ($Q > 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous (red regions, P$_{rot} >$ P$_{orb}$) rotators.}%
+For large $Q$ ($Q > 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
     \label{fig:qTauLock}%
@@
 
-From Fig.~\ref{fig:qTauLock}, certain rotation states, e.g.
-synchronous rotation, typically occur in certain regions of parameter space, e.g.
-where tidal torques are dominant, small $Q$ or large $\tau$.
-Therefore by measuring P$_{rot}$ and P$_{orb}$ for the primary star in a binary system, one can in principle constrain the strength of the tidal interaction that led to that rotation state.
-We do so in Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap} by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$ and computing the median $Q$ and $\tau$, and the $16\%$, and $84\%$ percentiles, in each bin.
+We isolate the impact of $Q$ and $\tau$ have in shaping the spin-orbital state of a stellar binary for a given P$_{orb}$, e.g.
+whether or not a binary is synchronized, by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
+In each bin, we compute the median $Q$ and $\tau$, and the $16\%$, and $84\%$ percentiles to quantify the distribution of tidal strength parameters that are consistent with each spin-orbital state.
 
@@
 Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking that spin-down the stars past the tidally-locked state.
-This effect can be seen for P$_{orb} < 50$ d where the median $Q$ gradually increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state that require low $Q$.
+This effect can be seen for P$_{orb} < 50$ d where the median $Q$ gradually increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state.
 This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.
@@
 
-As under the CPL model, stars tidally-lock when subject to strong tidal torques, typically for P$_{orb} < 50$ d and $\tau \gsim 0.25$ s.
-For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with $tau < 0.1$ s as magnetic braking dominates over tidal torques, spinning down the star.
+According to the CTL model, many binaries tidally-lock for P$_{orb} \lsim 20$ d.
+Tidally-locking can occur for P$_{orb} < 50$ d and $\tau \gsim 0.25$ s.
+For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with $tau < 0.1$ s as magnetic braking dominates over tidal torques.
 For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s has upper and lower limits that encompass most of the range we considered for $\tau$, suggesting that tides are not important in this region of parameter space.
@@
 
+\subsection{P$_{rot}$ Distribution of Stellar Binaries} \label{sec:protDist}
+
+Here we examine how the competition between tidal torques and magnetic braking shape the P$_{rot}$ distribution of low-mass stellar binaries.
+We consider two cases where tidal torques dominate: ``Locked" where P$_{rot} = $ P$_{eq}$ and ``Interacting" where P$_{rot}$ is within $10\%$ of P$_{eq}$ as in this regime, tides are likely shepherding P$_{rot}$ towards the tidally-locked state keeping P$_{rot}$ close to P$_{eq}$, typically within about $10\%$ as we demonstrated in $\S$~\ref{sec:eq}, e.g.
+Fig.~\ref{fig:eqPer}.
+We refer to the remaining binaries as ``Unlocked" as magnetic braking and stellar evolution likely dominate their angular momentum evolution.
+In Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, we plot P$_{rot}$ as a function of mass for the primary stars in stellar binaries for both the CPL and CTL model, respectively, integrated to system ages uniformly sampled over $1-7$ Gyr.
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
+   \caption{Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
+Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
+Right: Marginalized P$_{rot}$ distribution for each case.}%
+    \label{fig:lockedCPL}%
+\end{figure}
+
+Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.0$ d and $4.68$ d) than unlocked (median CPL, CTL P$_{rot} = 41.1$ d and $41.2$ d) binaries.
+More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale as $R^5$, and $R$ increases with stellar mass.
+This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model.
+
+Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
+This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
+Moreover, \citet{Abt2004} find evidence for tides modifying P$_{rot}$ in binaries with P$_{orb}$ up to 500 d, implying tidal torques matter out long P$_{orb}$.
+Under the CTL model, however, binaries can only tidally-lock out to P$_{orb} \approx 20$ d, with only $2\%$ of binaries becoming tidally-locked, in stark contrast to the CPL model's predictions.
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
+   \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
+    \label{fig:lockedCTL}%
+\end{figure}
+
+The interacting population tends to rotate more slowly than the unlocked population as at short P$_{orb}$, and hence P$_{rot}$, binaries preferentially tidally-lock due to stronger tidal torques.
+At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthing enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
+The CPL and CTL models predict that $35\%$ and $15\%$ of stars, respectfully, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
+The P$_{rot}$ - mass distribution for unlocked binaries resembles the single star sequence observed in the center right panel of Fig.~\ref{fig:protDist} indicating that magnetic braking and stellar evolution dictate their angular momentum evolution as expected.
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/lockedProtHist.pdf}
+   \caption{P$_{rot}$ distribution for tidally-locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.}%
+    \label{fig:lockedProtHist}%
+\end{figure}
+
+As shown in Fig.~\ref{fig:lockedProtHist}, both models predict that tidal torques can influence the P$_{rot}$ evolution out to P$_{orb} \approx 80$ d, causing a significant departure from the long-term expected spin down of single stars due to magnetic braking.
+The CTL model predicts that binaries can tidally-lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL P$_{orb} = 4.8$ d.
+The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states, from P$_{rot} < 20$ d out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
+The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
+We discuss this point further in $\S$~\ref{sec:whichModel}.
+
+\subsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
+
+In the absence of external forces, the long-term P$_{rot}$ evolution of low-mass stars is governed by magnetic braking.
+In stellar binaries, however, tidal torques compete against magnetic braking to drive P$_{rot}$ towards P$_{eq}$, e.g.
+Fig.~\ref{fig:eqPer}, causing their P$_{rot}$ evolution to differ from that of a single star.
+Here we compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking.
+We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar} with initial conditions sampled from the same mass and P$_{rot}$ distributions used for the binary simulations described $\S$~\ref{sec:methods}.
+In Fig.~\ref{fig:protDist}, we display P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
+
+\begin{figure*}[t]
+	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
+   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
+For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
+Right: The P$_{rot}$ distribution for each case, marginalized over stellar mass.}%
+    \label{fig:protDist}%
+\end{figure*}
+
+Both tidal models predict a substantial population of fast rotators (P$_{rot} < 20$ d) not seen in the single star model at any age, except for young (age $\lsim 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$).
+Under the CPL and CTL models, $19\%$ and $17\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.
+At the long P$_{rot}$ end, the CPL predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models.
+This feature is not due to long-term spin-down via magnetic braking as this population lies above the old edge of single star sequence.
+Instead tidal-locking in long P$_{orb}$ binaries produces these slow rotators as shown in Fig.~\ref{fig:lockedCPL}.
+
+There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
+In stark contrast for many binaries, especially for P$_{rot} < 20 $ d, both tidal models predict that age does not strongly correlate with P$_{rot}$.
+We find that the median ages and $68\%$ interval for the CPL and CTL models fast rotators (P$_{rot} \lsim 20$ d) are $3.1^{+2.8}_{-1.7}$ Gyr and $2.8^{+2.8}_{-1.5}$ Gyr, respectfully, compared to much younger stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr for the single star model.
+
+The only rapidly rotating single stars are young stars with ages $< 2$ Gyr that have not yet experienced substantial angular momentum loss due to magnetic braking.
+In binaries, however, tidal torques can lock stellar P$_{rot}$ to P$_{eq}$, especially at short P$_{orb}$ where tidal torques are stronger, creating a long-lasting population of fast rotators with a wide age range.
+If one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict that ages $\lsim 2$ Gyr, e.g.
+the right panel of Fig.~\ref{fig:protDist}.
+If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling its P$_{rot}$ from age, causing the predictions of gyrochronology models to fail.
+This effect poses a significant problem for studies of P$_{rot}$ distributions and their ability to constrain stellar ages.
+This problem likely exists in the \kepler field as \citet{Simonian2018} finds that many \kepler fast rotators are likely unresolved photometric binaries.
+
+At the long P$_{rot}$ end of the distribution, the CPL model predicts a population of tidally-locked extremely slow rotators that single star models fail to produce.
+Gyrochronology methods would incorrectly infer old ages $> 7$ Gyr for this slowly rotating population, whereas most of these binaries have ages $\lsim 6$ Gyr, with some being $\sim 1-4$ Gyr old.
+We caution that any application of gyrochonlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d or P$_{rot} \gsim 70$, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
+
 %% SECTION: COMPARISON TO KEPLER %%
@@
 
-Here we compare our simulation results to P$_{rot}$ measurements of primary stars in stellar binaries by \citet{Lurie2017}.
-\citet{Lurie2017} measured 816 rotation periods for \kepler EBs with star spot modulations and visually inspected each light curve to ensure their accuracy.
-The \citet{Lurie2017} dataset is the largest homogenous set of P$_{rot}$ measurements available for low-mass stellar binaries and represents the state of the art benchmark for studies of tidal interactions in stellar binaries.
+We compare our simulation results to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing stellar binaries by \citet{Lurie2017}.
+\citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with starspot modulations and visually inspected each light curve to ensure their accuracy.
+The \citet{Lurie2017} dataset is the largest homogenous set of P$_{rot}$ measurements available for low-mass stellar binaries and represents the state of the art benchmark for studies of the influence of tides on P$_{rot}$ in stellar binaries.
 We compare our results to the P$_{1,min}$ P$_{rot}$ values reported by \citet{Lurie2017} as the authors demonstrated that these values are likely to be close to the equatorial P$_{rot}$ that we track in our simulations.
@@
 In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
-Both models reproduce most trends in the \citet{Lurie2017} data: the large cluster of synchronized rotators with P$_{orb} \lsim 20$ d, a population of subsynchronous rotators with P$_{orb} \lsim 20$ d, and the sparse cluster of supersynchronous rotators at all P$_{orb}$ that correspond to young stars that have yet to significantly spin down due to magnetic braking or tides.
-The lower limit seen in both panels is a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down via magnetic braking in the weak tides regime over 7 Gyr, the longest age considered in our simulations.
+Both models reproduce most trends in the \citet{Lurie2017} data, namely the large cluster of synchronized rotators and the population of subsynchronous rotators, both for P$_{orb} \lsim 20$ d.
+Magnetic braking and age imparts two noticable features in the data.
+First, the sparse cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} > 1.5$ at all P$_{orb}$ corresponds to young stars that have yet to significantly spin down due to magnetic braking or tides.
+Second, the lower limit seen in both panels is a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down via magnetic braking in the weak tides regime over 7 Gyr, the longest age considered in our simulations.
 
@@
 
-The predictions of the CPL model fail to reproduce several key aspects of that data.
-First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} \lsim 1.5$, a prediction that is entirely inconsistent with the data as no obvious spin-orbit commensurablities, aside from 1:1 synchronization, is present in the data likely owing to the fact that stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
-The CPL model can produce the subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
-The CTL model, however, excels where the CPL models fails and naturally produces both populations, suggesting that the CTL model better describes tidal interactions in low-mass binaries.
-We examine this point further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}.
-
-XXX
-
-\citet{Claret1995} find that supersynchronous rotation their sample of eccentric binaries are in good agreement with the predictions of tidal models.
-
-presence of 3:2 cluster would be strong evidence in favor of CPL model.
-Can use presence/absence of this feature as a strong discriminators between tidal models.
-additionally, CTL predicts more subsync clusters that could be used for model descriptions.
-below, say how CPL model has locked binaries at long Porb, ctl doesn't above 20 d so that's another model descriimination.
-say each time that CTL passes, CPL doesn't according to lurie data.
-In general, talk about implications and use flowery language
+Both models can reproduce the major trends in the data, but the predictions of the CPL model fail to reproduce several key aspects.
+First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all tidally-locked sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.
+This prediction is inconsistent with the data as no obvious spin-orbit commensurablities, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
+The CPL model can produce the population of subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
+The CTL model, however, seems to excel where the CPL models fails by producing both populations.
+We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
 
@@
 
-Median subsync tidal tau: 0.0790922272
-Median supersync tidal tau: 0.12508653609999998
-Median subsync ecc: 0.1288805
-Median supersync ecc: 0.151426
-Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6126373626373627
-Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.11401098901098901
-
-- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques.
-
-Although Lurie+2017 attributes differential rotation to the production of the
-subsynchronous population, we find that coupled stellar-tidal evolution naturally
-produces the population. it's probably a combination of both effects. Compare with Lurie+2017 sample for 2 $<$ Porb $<$ 10 days:
-$72\%$ in [0.92,1.2] and $15\%$ in [0.84, 0.92].
+Our coupled stellar-tidal model accurately reproduces the spin and orbital states of low-mass \kepler EBs when using the CTL formalism.
+In Fig.~\ref{fig:subsync}, we demonstrate this finding by comparing our simulated systems (dots) with the observed \citet{Lurie2017} EB data (diamonds), where both points are colored by $e$.
+Nearly all simulated binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in excellent agreement with the \citet{Lurie2017} observations.
+At very short P$_{orb}$ in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries.
+Beyond P$_{orb} \approx 4$ d, tidally-locked, eccentric binaries populate the supersynchronous population, again in excellent agreement with the \citet{Lurie2017} \kepler data.
+These systems have tidally-locked but their orbits have not yet circularized, even though they likely experience strong tidal torques at short separations, because the tidal locking time scale is 2-3 orders of magnitude less than the circularization timescale \citep{Mazeh2008}.
 
@@
 
-Our model predicts that nearly all EBs with P$_{orb} < 3$ d are synchronized as is observed by \citet{Lurie2017}.
-We find that our model predicts a population of supersyncronous rotators on eccentric orbits and a population of subsyncronous rotators with P$_{orb}/$P$_{rot} \approx 0.9$ on nearly-circular orbits, both of which were observed by \citet{Lurie2017}.
-This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL formalism can accurately model the tidal evolution of low-mass stellar binaries.
+For P$_{orb} \gsim 4$ d, our model produces a population of subsynchronous rotators, including a cluster of subsynchronously rotating stellar binaries with nearly circular orbits, again in excellent agreement with the cluster of binaries with nearly circular orbits near P$_{orb}/$P$_{rot} = 0.9$ observed by \citet{Lurie2017}.
+Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces this population.
+Our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.13$ s, compared to the median $\tau = 0.08$ s for the tidally-locked supersynchronous population.
+
+\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $<$ Porb $<$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $11\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
+Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $61\%$ of our CTL population.
+This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL equilibrium tide formalism can accurately model the tidal evolution of low-mass stellar binaries.
+The comparison between theory and observations, however, is limited in that the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and because \citet{Lurie2017} uses transit durations and ingress/egress times to approximate the EB orbital $e$.
+In the absence of more precise and reliable constraints in the \citet{Lurie2017} \kepler data, however, we cannot definitively assert that the CTL model is superior to the CPL formalism for modeling tidal torques in low-mass binaries.
+We discuss observational tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
 
@@
 
-XXX: say how our quoted $Q$ and $\tau$ ranges produced binaries consistent with what \kepler saw, so not only is our model a reasonable choice to model stellar-tidal evolution, but also our quoted tidal parameter ranges seem to bracket what stars actually have.
-We refine it in the below analysis
+In $\S$~\ref{sec:qTauMaps}, we demonstrated that certain rotation states typically occur for a limited subset of tidal parameters.
+For example for a low-mass short P$_{orb}$ stellar binary, if we observe subsynchronous rotation, we can infer that tidal torques must be weak since magnetic braking dominates the P$_{rot}$ evolution, implying that those stars have large $Q$ or small $\tau$, as demonstrated in Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
+By measuring P$_{rot}$ for the primary star in a stellar binary and P$_{orb}$, one can therefore constrain the strength of the tidal interaction that led to that rotation state.
+Here, we compare \kepler EB spin-orbital states measured by \citet{Lurie2017} with our simulation results to derive approximate constraints on the tidal model parameters, $Q$ and $\tau$.
+We assign each binary in the \citet{Lurie2017} dataset to its corresponding P$_{orb}$, P$_{orb}/$P$_{rot}$ bin in in Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap} and draw 100 samples of $Q$ and $\tau$, with replacement, from our simulated CPL and CTL model data, respectively, in the same bin to build approximate $Q$ and $\tau$ distribution for each system.
+We combine the $Q$ and $\tau$ distributions, respectively, for all binaries in the \citet{Lurie2017} data to produce approximate distributions for $Q$ and $\tau$, marginalized over all other parameters, and display them in Fig.~\ref{fig:qLurie} and Fig.~\ref{fig:tauLurie}.
 
@@
    \caption{Distribution of stellar tidal $Q$ produced by convolving \citet{Lurie2017} observations of EBs' P$_{rot}$ and P$_{orb}$ with our CPL simulation results shown in Fig.~\ref{fig:qmap}.
-The dashed black lines, from left to right, indicate the $16^{th}$ percentile, median, and $84^{th}$ percentile.
-Our $Q$ constraints are consistent with our flat prior distribution over log$_{10}(Q)$, suggesting that tidally-interacting stellar binaries can have a wide range of $Q$.}%
+The dashed black lines, from left to right, indicate the $16^{th}$ percentile, median, and $84^{th}$ percentile.}%
     \label{fig:qLurie}%
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/tauLurie.pdf}
-   \caption{Same as Fig.~\ref{fig:qLurie}, but using our CTL simulation results from Fig.~\ref{fig:taumap}.
-We find log$_{10}(\tau) = -0.73^{+0.52}_{-0.82}$ with an increase in density at larger $\tau$ deviating from our prior flat distribution, suggesting that time lags in low-mass stars are larger than the typical value of $\tau \approx 0.1$ s.
-A wide range of values for $\tau$ are still consistent with observations, however, indicating that tidally-interacting binaries can exhibit a wide range of tidal time lags, albeit with a preference for longer lags.}%
+   \caption{Same as Fig.~\ref{fig:qLurie}, but using our CTL simulation results from Fig.~\ref{fig:taumap}.}%
     \label{fig:tauLurie}%
@@
 
-\subsection{Deviations From Single Star P$_{rot}$ Evolution: Implications for Gyrochronology} \label{sec:gyro}
+We derive a wide $Q$ constraint, log$_{10}(Q) = 5.55^{+0.98}_{-1.06}$, that is consistent with our flat prior distribution, suggesting that tidally-interacting stellar binaries can have a wide range of $Q$.
+Our $\tau$ constraint, log$_{10}(\tau) = -0.73^{+0.52}_{-0.82}$, however, displays a noticeable increase in density at larger $\tau$ near 1 s.
+This is a significant deviation from our flat prior distribution suggesting that time lags in low-mass stars are larger than the typical value of $\tau \approx 0.1$ s.
+Still, a wide range of values for both $Q$ and $\tau$ are consistent with observations indicating that tidally-interacting binaries can exhibit a wide range of tidal parameters, albeit with a slightl preference for larger time lags under the CTL formalism.
 
-XXX:
+A better method to constrain $Q$ or $\tau$ would be to model the stellar-tidal evolution of an individual binary system in an Markov chain Monte Carlo (MCMC) framework, conditioning on observed parameters, like P$_{orb}$ and $e$, to derive full posterior probability distributions for $Q$ or $\tau$.
+We lack the precise constraints on the stellar and orbital parameters, and their uncertainties, for the binaries in the \citet{Lurie2017} dataset required for an MCMC analysis, so we leave that for future work.
 
--hammer implications for gyrochronology, fast rotators in kepler field, compare to \citet{Simonian2018}
+\subsection{CPL or CTL?} \label{sec:whichModel}
 
-- \citet{Abt2004} find evidence for tides modifying P$_{rot}$ in binaries with P$_{orb}$ up to 500 d.
+Accurate measurements of P$_{rot}$ and $e$, especially out to long P$_{orb}$, can discriminate between which equilibrium tidal model best describes tidal interactions in low-mass stellar binaries.
+Here, we outline three observational tests that can discriminate between the two models.
+The first two require obtaining a series of time-resolved spectra over full orbital phase of short P$_{orb}$ supersynchronous binaries, i.e.
+those identified by \citet{Lurie2017}, to compute radial velocities over the binary's orbit.
+Combined with \kepler photometry, one can derive much more robust $e$ constraints than those computed using the transit approximations in \citet{Lurie2017}.
 
-Here we compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking.
-We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar}.
-To generate the simulations, we sample from the same mass and initial P$_{rot}$ distributions we used for the binary simulations described $\S$~\ref{sec:methods}.
-In Fig.~\ref{fig:protDist}, we P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
+The first test considers binaries with P$_{orb} < 10$ d that are likely tidally-locked on eccentric orbits, but with $e < 0.23$.
+In this $e$ regime, the CPL model predicts that the majority of systems are tidally-locked into synchronous rotation and does not permit a supersynchronous rotation state, e.g.
+Eqn.~(\ref{eqn:cpl:eqPer}).
+The CTL model, however, predicts a continuum of supersynchronous rotators on eccentric orbits in this regime, e.g.
+Eqn.~(\ref{eqn:ctl:eqPer}).
+Supersynchronous rotation that is not due to tidal interactions can occur in extremely young, rapidly rotating systems that are still contracting along the pre-main sequence, or that have recently reached the main sequence.
+These young, supersynchronous rotators are unlikely to be tidally-locked and usually have P$_{orb}/$P$_{rot} > 1.5$, distinguishing them from tidally-locked binaries (see Fig.~\ref{fig:lurie7}).
+If supersynchronous rotation is observed in binaries with P$_{orb} < 10$ d, P$_{orb}/$P$_{rot} < 1.5$, and $0 < e \lsim 0.23$, it is evidence in favor of the CTL model.
 
-Tidally-locked binaries tend to rotate more rapidly than unlocked binaries, reflected in the increased density for P$_{rot} \lsim 20$ d in the marginal distribution for the locked population and in the distribution medians (locked median P$_{rot} = 19.8$ d compared to unlocked median P$_{rot} = 27.5$ d).
-Binaries readily tidally-lock at short P$_{orb}$ due to the tidal force's strong semi-major axis dependence, increasing the strength of tidal torques at shorter orbital separations.
-When stellar binaries approach the tidally-locked state, tidal torques overpower the torque due to magnetic braking, fixing P$_{rot} =$ P$_{orb}$, resulting in long-lasting rapid rotation in short P$_{orb}$ binaries.
+Second, for tidally-locked binaries with $e > 0.23$, the CPL model predicts supersynchronous rotation in the form of a 3:2 spin-orbit comensurability, e.g.
+the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permitted.
+The CTL model, however, permits a continuum of supersynchronous rotation states in eccentric tidally-locked rotators.
+If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model.
 
-Tidal forces modify the spins of unlocked binaries as well, tending to spin-up stars against the effects of magnetic braking towards orbital synchronization, reflected in unlocked binaries' median P$_{rot} = 27.5$ d, compared to the single star median P$_{rot} = 30.6$ d.
-Tidal locking is not limited to P$_{orb} \lsim 10$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 70$ d, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars (compare to XXX).
-This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find that binaries can tidally lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-The observed age distribution of tidally-locked binaries has little dependence on P$_{rot}$, with both young and old systems tidally-locking at long and short P$_{rot}$, causing the predictions of gyrochronology models to fail in such systems.
+The continuous supersynchronous rotation predicted by the CTL model seen in right panel of Fig.~\ref{fig:lurie7} seems to provide a much better match to the \citet{Lurie2017} data for both tests described above, providing evidence in favor of the CLT model.
+The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making this distinction conclusive.
+More robust phase-resolved spectroscopic $e$ constrains would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a Gaussian process model as was done by \citet{Angus2018}.
 
-\begin{figure*}[t]
-	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
-   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
-The P$_{rot}$ distribution for each case, marginalized over stellar mass, is depicted in the right panel.
-For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
-Both tidal models predict a substantial population of fast rotators (P$_{rot} < 20$ d) not seen in the single star model at any age, except for young (age $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$).
-In many binary systems, both tidal models predict that age does not directly correlate with P$_{rot}$, especially in short P$_{orb}$ binaries where tidal torques are stronger, compared to the clear age-dependence of P$_{rot}$ seen in the single star models, the key assumption of gyrochronology methods.
-Any application of gyrochonlogy methods to age stars, especially those with P$_{rot} < 20$ d, should rule out or account for binarity.}%
-    \label{fig:protDist}%
-\end{figure*}
-
-XXX
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
-   \caption{Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
-Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
-Right: P$_{rot}$ distribution for each case, marginalized over stellar mass.
-Tidally-locked stars systematically rotate faster (median P$_{rot} = 27.0$ d) than interacting (median P$_{rot} = 51.8$ d) and not locked (median P$_{rot} = 41.1$ d) binaries as the locked population typically have shorter P$_{orb}$ where tidal torques are stronger, although binaries up to P$_{orb} = 100$ d tidally-lock.
-More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale with stellar mass, reflected by the enhanced density of locked systems at larger masses.
-The interacting population tends to rotate more slowly than the unlocked population as at short P$_{orb}$, and hence P$_{rot}$, the systems preferentially tidally-lock.
-At longer P$_{orb}$, magnetic braking spin down the stars while tidal torques sheppard P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
-$35\%$ of stars are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries.
-The P$_{rot}$ - mass distribution for not locked binaries resembles the single star sequence observed in the center right panel of Fig.~\ref{fig:protDist}.}%
-    \label{fig:lockedCPL}%
-\end{figure}
-
-XXX
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
-   \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations.
-According to the CTL model, only $2\%$ of stellar binaries tidally-locked, mostly at short P$_{orb}$ (median P$_{orb} = 4.8$ d), where tidal torques are strong.
-The  $15\%$ of stars are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries.
-As in Fig.~\ref{fig:lockedCPL}, interacting binaries cluster at longer P$_{rot}$ as magnetic braking overpowers tidal torques to spin down stars, but tidal torques are strong enough to sheppard P$_{rot}$ near P$_{eq}$.}%
-    \label{fig:lockedCTL}%
-\end{figure}
-
-XXX
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedProtHist.pdf}
-   \caption{P$_{rot}$ distribution for tidally-locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.
-Strong tidal torques in the CPL model can tidally-lock binaries out to P$_{rot} = 100$ d, while weaker torques in the CTL model can only tidally-lock binary stars into a rapid rotation state with P$_{rot} < 20$ d.
-Under both models, binaries with a wide range of P$_{rot}$ can be strongly tidally-interacting.}%
-    \label{fig:lockedProtHist}%
-\end{figure}
-
-interacting systems tend to cluster near more massive stars for intermediate periods since their PMS lifetimes are shorter, so they have already spin down and tidal torques are stronger in those systems
+Third, as shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 70$ d would provide strong evidence in favor of the CPL model as the CTL model cannot tidally-lock binaries past P$_{orb} \approx 20$ d, e.g.
+Fig.~\ref{fig:lockedCTL}, whereas the CPL model can tidally-lock binaries out to P$_{orb} \approx 100$ d.
+We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small starspot modulation amplitudes \citep{Lurie2017,Reinhold2018}.
+Long term spectroscopic monitoring may be warranted in such cases.
 
@@
 
-XXX
+In this work, we probed the long-term angular momentum evolution of low-mass stellar binaries, with a focus P$_{rot}$ in intermediate P$_{orb}$ binaries.
+We considered the impact of two common equilibrium tidal models, magnetic braking, and stellar evolution.
+We performed a large suite of simulations for binaries with physically-motivated initial conditions out to P$_{orb} = 100$ and across a wide range of tidal dissipation parameters to examine the competition between tidal torques and magnetic braking for controlling the stellar P$_{rot}$ evolution.
+In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have synchronized spins and circularized orbits.
+We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking.
+Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact out to P$_{orb} \approx 80$ d, causing their P$_{rot}$ evolution to differ from the long-term spin down due to magnetic braking experience by single stars.
+Many binaries with P$_{orb} \lsim 20$ d tidally-lock, but the CPL model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
+When binaries tidally-lock, tides fix P$_{rot} = $ P$_{eq}$, preventing further spin-down via magnetic braking, decoupling P$_{rot}$ from age.
+In such systems, we caution that gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails.
+We caution any application of gyrochronlogy methods to stars with P$_{rot} \lsim 20$ d or P$_{rot} \gsim 70$ should account for the possibility of stellar binarity to prevent deriving incorrect ages.
 
-- talk about how we need more binary observations with Prot, Porb, and ecc, especially at longer orbital periods, to more reliably discriminate between tidal models.
-At low eccentricity, it does matter (sync vs supersync in \kepler) as opposed to what \citet{Leconte2010} recommends
+We compare the predictions of both the CPL and CTL models with observations of P$_{rot}$ and P$_{orb}$ of \kepler EBs by \citet{Lurie2017} and find that both can reproduce the data, validating our approach and suggesting that equilibrium tidal models can accurately reproduce stellar-tidal evolution in low-mass stellar binaries.
+We show that the CTL model best reproduces the data, but the lack of uncertainties on P$_{rot}$ and the approximate orbital eccentricities derived by \citet{Lurie2017} prevent us from conclusively ruling in favor of the CTL model.
+We offer three observational tests that can distinguish between which equilibrium tidal model best describes tidal interactions in low-mass stellar binaries.
+To perform these tests, we that recommend observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d and in short P$_{orb}$ subsynchronous rotators.
 
-- cite Matsen+2018 for "more than $50\%$ of the binary period distribution will be imaged in about $85\%$ of the host stars" in paper
-about detecting unresolved binaries in TESS
+The excellent agreement we find between our theory and \kepler observations outlines a critical point: one cannot simply observe a short P$_{orb}$ binary on a circular orbit and assume synchronization, nor can one observe a binary with P$_{orb} \gsim 20$ d and assume that tides have not shaped that system's angular momentum evolution.
+For example, stellar-tidal interactions can produce synchronous and subsynchronous rotation for short P$_{orb}$ binaries on circular orbits, e.g.
+Fig.~\ref{fig:subsync}, depending on the age of the system, e.g.
+Fig.~\ref{fig:eqPerShortPorb}, and the strength of tidal dissipation, e.g.
+Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
+Understanding the long-term angular momentum evolution of stellar binaries out to P$_{orb} = 100$ d requires detailed modeling of its coupled-stellar tidal evolution, and characterizing tidal dissipation parameters, e.g.
+how $Q$ and $\tau$ vary from star-to-star.
+Many new eclipsing stellar binaries will be discovered by TESS \citep[e.g.][]{Sullivan2015,Matson2018} and in analysis of K2 data.
+Obtaining precise orbital and rotational constraints for stellar binaries will permit detailed characterization of tidal interactions between low-mass stars and shed light into the long-term angular momentum evolution in stellar binaries.
 

### 2018-12-20 Paper/sync.tex
Updates from ShareLaTeX

@@
  Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries.
-Here, we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d and across a wide range tidal dissipation parameters.
-In many cases, we find that magnetic braking can overpower tides, producing a long-lasting population of subsynchronous rotators, even in short P$_{orb}$ binaries, reproducing the population observed by \citet{Lurie2017} in their sample of \kepler eclipsing binaries (EBs).
-At short P$_{orb}$, many stellar binaries tidally-lock, naturally producing a population of fast rotators as is observed in the \textit{Kepler} field, a feature that cannot be reproduced by single-star-only models of magnetic braking.
-Many binaries can tidally-lock even out to P$_{orb} \approx 100$ d, fixing the stellar P$_{rot}$ to the binary P$_{orb}$, causing P$_{rot}$ to not be a valid proxy for age in all cases, i.e.
+Here we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters.
+In many cases, we find that magnetic braking can overpower tides, producing a long-lasting population of subsynchronous rotators, even in short P$_{orb}$ binaries, reproducing the observed \kepler eclipsing binaries (EBs) population.
+Furthermore, at short P$_{orb}$, tidal locking naturally producing a population of fast rotators that cannot be reproduced by single-star-only models of magnetic braking.
+Many binaries can tidally-lock even out to P$_{orb} \approx 100$ d, fixing P$_{rot}$ to P$_{orb}$, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
 gyrochronology methods must be applied carefully.
-We compare our model predictions with \kepler EB rotation periods and find that our predictions are in excellent agreement with observations.
 We compare the predictions of two common equilibrium tidal models and show that the ``Constant Time Lag" equilibrium tidal model provides a better match to observations.
 We recommend observational tests to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
-By comparing observations of \kepler EBs with our simulations, we constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.
+Finally, by comparing observations of \kepler EBs with our simulations, we constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.
 
@@
 
-Angular momentum evolution in low-mass (M$\lsim 1$ M$_{\odot}$) short-period (P$_{orb} \lsim 10$ d) stellar binaries is dominated by tides.
+Angular momentum evolution in low-mass (M$\lsim 1$ M$_{\odot}$) short-period (P$_{orb} \lsim 10$ d) stellar binaries is strongly influenced by tides.
 Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}.
-Orbital circularization is ubiquitous in short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989,Claret1995} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
+Orbital circularization is ubiquitous for short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989,Claret1995} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
 For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} finds that the orbit of solar twin binaries circularizes during the stellar pre-main sequence.
@@
 
-Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially so for the stellar spins.
-Tidal torques drive stellar spins towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
+Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially for stellar spins.
+Tidal torques drive them towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
 Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
@@
 Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking.
-These findings suggest that that synchronization is not a given for tidally-interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$.
-Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
+These findings suggest that synchronization is not a given for tidally-interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$.
+XXX Not sure why suggests consider longer Porb XXX Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
 
-Understanding angular momentum evolution in low-mass stellar binaries is paramount as P$_{rot}$ distributions measured in clusters \citep[e.g.
+Understanding how tidal torques affect angular momentum evolution in low-mass stellar binaries is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g.
 Praesepe, ][]{Douglas2017} and field stars \citep[e.g.
@@
 We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:protDist} and $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.
+
+XXX Why not discuss tidal braking more?
+I expected an intro that described tidal affects on binaries, magnetic braking, and then stating that no study had properly linked the two.
+The tidal review is good, but more on MB would be helpful.
+XXX
+
+tides impact binary ecc, tides can brake spins, tidally-lock systems, keeping them fast rotators etc.
+Talk about magnetic braking as both magnetic braking and tides act in binaries.
+Talk about observations of Prot, like  \citet{Abt2004} citation and similar things like Lurie
 
@@
 
-\subsubsection{Example Stellar Evolution} \label{sec:methods:stellarExample}
-
 In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar.
@@
 
+XXX Paragraph, equation and plot for Matt15? XXX
+
 \begin{figure*}[ht]
@@
    \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} magnetic braking law.
-Each dot denotes the approximate time when each star reaches the main sequence.}%
+Each dot denotes the approximate time when each star reaches the main sequence.
+XXX Reiners & Mohanty??
+XXX}%
     \label{fig:stellarExample}%
@@
 
- Equilibrium tidal models, first introduced by \citep{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega$, and obliquities $\psi$, of the both gravitating bodies due to tidal torques.
-Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the both bodies' centers of mass due to friction within each body.
-The tidal bulges cause torques that drive the exchange of angular momentum between the orbit and both bodies' spins.
+ Equilibrium tidal models, first introduced by \citep{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega$, and obliquities $\psi$, of both gravitating bodies due to tidal torques.
+Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the bodies' centers of mass due to friction within each.
+The tidal bulges cause torques that permit the exchange of angular momentum between the orbit and both bodies' spins.
 Equilibrium tidal models are linear since they assume that the tidal waves that comprise the tidal bulge raised on a body are uncoupled.
@@
 
+XXX discuss Rory's concern of how the above discussion only applies in the truncated series CPL formalism, whereas Murray and Dermott derived a continuous Peq for the CPL model as well
+
 Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, and due to the inherant discreteness of numerical integrations, numerical solutions for the CPL and CTL models can produce unphysical evolution.
@@
 We plot the tidal evolution for $a$, $e$, and $\omega$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively, in Fig.~\ref{fig:tidalExample}.
-Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increasing as tides force the spins toward the tidally-locked state, transfering rotational angular momentum into the orbit in the process.
+Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally-locked state, transfering rotational angular momentum into the orbit in the process.
 At late times, both the CPL and CTL drive the binaries towards orbital circularization, with tidal dissipation decreasing P$_{orb}$.
 The predictions of the CPL and CTL model, differ, however, when the binaries tidally-lock.
-Under the CPL model, the binary tidally-locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
+Under the CPL model, the binary tidally-locks into a synchronous orbit when $e < \sqrt{1/19}$, e.g.
 Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence, e.g.
@@
    \caption{Tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary's $e$ (top), P$_{orb}$ (middle) and P$_{rot}$ (bottom) for the CPL (blue) and CTL (orange) model.
-The blue (CPL) and orange (CTL) vertical dashed lines denote when the stellar binary tidally-locks.
+The blue (CPL) and orange (CTL) vertical dashed lines denote when the stellar binary tidally locks.
 Both the CPL and CTL model predict the same qualitative evolution.
@@
 
-Following \citet{Fleming2018}, when one or both binary stars are tidally locked, tidal forces prevent magnetic braking from spinning down the tidally-locked star(s), and any angular momentum lost is lost at the expense of the binary orbit, decreasing $a$ as a result \citep{Verbunt1981}.
+Following \citet{Fleming2018}, when one or both binary stars are tidally locked, tidal forces prevent magnetic braking from spinning down the tidally-locked star(s), and any angular momentum lost comes at the expense of the binary orbit, decreasing $a$ as a result \citep{Verbunt1981}.
 Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify the $a$ decay equations due to stellar evolution and magnetic braking in tidally-locked binaries from \citet{Fleming2018}, their Eqs.
@@
 For each simulation, we sample the primary's mass uniformly over $[0.1, 1]$ M$_{\odot}$.
-Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of young stars in the ${\sim}2$ Myr old ONC.
+Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of young stars in the ${\sim}2$ Myr old Orion Nebula Cluster \citep{Matt2015}.
 We compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010,Moe2018}.
@@
 
-Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain, likely depend on complex viscous evolution within the stars \citep{Ogilvie2007}, can differ for stars of the same spectral class \citep{Barker2009}, and can vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass star's evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
+Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain due to complex viscous evolution within the stars \citep{Ogilvie2007}, and can differ for stars of the same spectral class \citep{Barker2009}.
+These parameters can also vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass stars' evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
 Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a wide range of values exist in the literature.
-Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s loguniformly over $[10^4,10^7]$ and $\tau$ loguniformly over $[10^{-2},1]$ s.
+Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s log-uniformly over $[10^4,10^7]$ and $\tau$ log-uniformly over $[10^{-2},1]$ s.
 Note that in general, there is no expression to compute $Q$ as a function of $\tau$, or vice versa, except in some special cases where approximations exist, e.g.
@@
 Here we examine the P$_{rot}$ evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a focus on systems with weak tidal torques, i.e.
-long P$_{orb}$ and large $Q$ or small $\tau$, to examine the boundry between evolution dominated by tides or magnetic braking.
+long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking.
 In Fig.~\ref{fig:eqPer}, we plot the evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) formalisms.
@@
 Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d.
-The black dotted line indicates that tidally-locked state where P$_{rot} = $ P$_{eq}$.
+The black dotted line indicates the tidally-locked state where P$_{rot} = $ P$_{eq}$.
 Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.
-We truncate each curve when the binary tidally-locks.
+We truncate each curve when the binary tidally locks.
 For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
@@
 Second, as P$_{rot}$ slows down further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example).
-Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$ and strengthening tidal torques that strongly depend on the binary separation as $a^{-6}$.
+Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$ and strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.
 These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long term.
@@
    \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively.
-Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally-lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
+Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
 }%
@@
    \caption{Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
-For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally-lock into synchronous rotation, regardless of $Q$, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.
-Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spin down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
-For large $Q$ ($Q > 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
+For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, regardless of $Q$, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.
+Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spun down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
+For large $Q$ ($> 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
     \label{fig:qTauLock}%
@@
 Narrow constraints imply that certain tidal parameters are required to produce that rotation state, e.g.
-supersynchronous rotators near P$_{orb}/$P$_{rot} = 1.5$ require $Q \lsim 10^{5.5}$.}%
+supersynchronous rotators near P$_{orb}/$P$_{rot} = 1.5$ require $Q \lsim 10^{5.5}$.
+XXX This fig and the next are real eye charts.
+Can you blow them up to be full page?
+XXX}%
     \label{fig:qmap}%
@@
 Tidally-locking can occur for P$_{orb} < 50$ d and $\tau \gsim 0.25$ s.
-For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with $tau < 0.1$ s as magnetic braking dominates over tidal torques.
+For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with $\tau < 0.1$ s as magnetic braking dominates over tidal torques.
 For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s has upper and lower limits that encompass most of the range we considered for $\tau$, suggesting that tides are not important in this region of parameter space.
@@
 
-Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we observe stellar binaries to tidally lock out up to P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
+Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we observe stellar binaries to tidally lock up to P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
 This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
@@
 At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthing enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
-The CPL and CTL models predict that $35\%$ and $15\%$ of stars, respectfully, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
+The CPL and CTL models predict that $35\%$ and $15\%$ of stars, respectivey, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
 The P$_{rot}$ - mass distribution for unlocked binaries resembles the single star sequence observed in the center right panel of Fig.~\ref{fig:protDist} indicating that magnetic braking and stellar evolution dictate their angular momentum evolution as expected.
@@
 In Fig.~\ref{fig:protDist}, we display P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
+XXX IS this paragraph necessary?
 
@@
 
+XXX This subsection is long and meandering. And obvious. I think you can just spend a couple paragraphs pointing out the top-level statistics and then go for the gyrochronology jugular. XXX
+
+be direct, scare gyro people
+
 %% SECTION: COMPARISON TO KEPLER %%
@@
 The CTL model, however, seems to excel where the CPL models fails by producing both populations.
-We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
+XXX This is a big point that I'm not sure is obvious.
+Does a simple 1-D histogram show it?
+We really struggled to understand this population with John, so if this is really the explanation, it's a big deal.
+XXX We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
 
@@
 
-\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $<$ Porb $<$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $11\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
+\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $11\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
 Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $61\%$ of our CTL population.
 This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL equilibrium tide formalism can accurately model the tidal evolution of low-mass stellar binaries.
-The comparison between theory and observations, however, is limited in that the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and because \citet{Lurie2017} uses transit durations and ingress/egress times to approximate the EB orbital $e$.
+The comparison between theory and observations, however, is limited in that the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and because \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$.
 In the absence of more precise and reliable constraints in the \citet{Lurie2017} \kepler data, however, we cannot definitively assert that the CTL model is superior to the CPL formalism for modeling tidal torques in low-mass binaries.
 We discuss observational tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
+XXX Almost seems like this should be a separate paper...
+XXX
 
-\subsubsection{$Q$, $\tau$ Constraints from P$_{orb}$ and P$_{rot}$} \label{sec:qTau}
+%\subsubsection{$Q$, $\tau$ Constraints from P$_{orb}$ and P$_{rot}$} \label{sec:qTau}
 
-In $\S$~\ref{sec:qTauMaps}, we demonstrated that certain rotation states typically occur for a limited subset of tidal parameters.
+%In $\S$~\ref{sec:qTauMaps}, we demonstrated that certain rotation states typically occur for a limited subset of tidal parameters.
 For example for a low-mass short P$_{orb}$ stellar binary, if we observe subsynchronous rotation, we can infer that tidal torques must be weak since magnetic braking dominates the P$_{rot}$ evolution, implying that those stars have large $Q$ or small $\tau$, as demonstrated in Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
@@
 Here, we compare \kepler EB spin-orbital states measured by \citet{Lurie2017} with our simulation results to derive approximate constraints on the tidal model parameters, $Q$ and $\tau$.
-We assign each binary in the \citet{Lurie2017} dataset to its corresponding P$_{orb}$, P$_{orb}/$P$_{rot}$ bin in in Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap} and draw 100 samples of $Q$ and $\tau$, with replacement, from our simulated CPL and CTL model data, respectively, in the same bin to build approximate $Q$ and $\tau$ distribution for each system.
+%We assign each binary in the \citet{Lurie2017} dataset to its corresponding P$_{orb}$, P$_{orb}/$P$_{rot}$ bin in in Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap} and draw 100 samples of $Q$ and $\tau$, with replacement, from our simulated CPL and CTL model data, respectively, in the same bin to build approximate $Q$ and $\tau$ distribution for each system.
 We combine the $Q$ and $\tau$ distributions, respectively, for all binaries in the \citet{Lurie2017} data to produce approximate distributions for $Q$ and $\tau$, marginalized over all other parameters, and display them in Fig.~\ref{fig:qLurie} and Fig.~\ref{fig:tauLurie}.
 
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/qLurie.pdf}
-   \caption{Distribution of stellar tidal $Q$ produced by convolving \citet{Lurie2017} observations of EBs' P$_{rot}$ and P$_{orb}$ with our CPL simulation results shown in Fig.~\ref{fig:qmap}.
-The dashed black lines, from left to right, indicate the $16^{th}$ percentile, median, and $84^{th}$ percentile.}%
-    \label{fig:qLurie}%
-\end{figure}
+%\begin{figure}
+%	\includegraphics[width=0.45\textwidth]{../Plots/qLurie.pdf}
+%   \caption{Distribution of stellar tidal $Q$ produced by convolving \citet{Lurie2017} observations of EBs' P$_{rot}$ and P$_{orb}$ with our CPL simulation results shown in Fig.~\ref{fig:qmap}.
+The dashed black lines, from left to right, indicate the $16^{th}$ percentile, median, and $84^{th}$ percentile.
+XXX Looks like your priors need to be wider :( XXX}%
+%    \label{fig:qLurie}%
+%\end{figure}
 
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/tauLurie.pdf}
-   \caption{Same as Fig.~\ref{fig:qLurie}, but using our CTL simulation results from Fig.~\ref{fig:taumap}.}%
+%\begin{figure}
+%	\includegraphics[width=0.45\textwidth]{../Plots/tauLurie.pdf}
+%   \caption{Same as Fig.~\ref{fig:qLurie}, but using our CTL simulation results from Fig.~\ref{fig:taumap}.}%
     \label{fig:tauLurie}%
-\end{figure}
+%\end{figure}
 
-We derive a wide $Q$ constraint, log$_{10}(Q) = 5.55^{+0.98}_{-1.06}$, that is consistent with our flat prior distribution, suggesting that tidally-interacting stellar binaries can have a wide range of $Q$.
+%We derive a wide $Q$ constraint, log$_{10}(Q) = 5.55^{+0.98}_{-1.06}$, that is consistent with our flat prior distribution, suggesting that tidally-interacting stellar binaries can have a wide range of $Q$.
 Our $\tau$ constraint, log$_{10}(\tau) = -0.73^{+0.52}_{-0.82}$, however, displays a noticeable increase in density at larger $\tau$ near 1 s.
@@
 
-A better method to constrain $Q$ or $\tau$ would be to model the stellar-tidal evolution of an individual binary system in an Markov chain Monte Carlo (MCMC) framework, conditioning on observed parameters, like P$_{orb}$ and $e$, to derive full posterior probability distributions for $Q$ or $\tau$.
+%A better method to constrain $Q$ or $\tau$ would be to model the stellar-tidal evolution of an individual binary system in an Markov chain Monte Carlo (MCMC) framework, conditioning on observed parameters, like P$_{orb}$ and $e$, to derive full posterior probability distributions for $Q$ or $\tau$.
 We lack the precise constraints on the stellar and orbital parameters, and their uncertainties, for the binaries in the \citet{Lurie2017} dataset required for an MCMC analysis, so we leave that for future work.
@@
 
+XXX The abov discussion is a bit tricky.
+Remember that CPL and CTL are truncated infinite series, so locking into 1:1 vs 3:2 might not be the only possible outcomes.
+In Murray & Dermott they show how one can use a CPL model to produce the "pseudo-equlibrium" period that is a continuous function of $e$.
+XXX
+
 The continuous supersynchronous rotation predicted by the CTL model seen in right panel of Fig.~\ref{fig:lurie7} seems to provide a much better match to the \citet{Lurie2017} data for both tests described above, providing evidence in favor of the CLT model.
@@
 
-In this work, we probed the long-term angular momentum evolution of low-mass stellar binaries, with a focus P$_{rot}$ in intermediate P$_{orb}$ binaries.
+In this work, we probed the long-term angular momentum evolution of low-mass stellar binaries, with a focus on P$_{rot}$ in intermediate P$_{orb}$ binaries.
 We considered the impact of two common equilibrium tidal models, magnetic braking, and stellar evolution.
@@
 We caution any application of gyrochronlogy methods to stars with P$_{rot} \lsim 20$ d or P$_{rot} \gsim 70$ should account for the possibility of stellar binarity to prevent deriving incorrect ages.
+XXX Whew, there's a lot here!
+Maybe a bulleted list is in order?
+And make sure a sentence for each is in the abstract.
+XXX
 
@@
 DPF and RB acknowledge that this work was supported by the NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement number NNA13AA93A.
+XXX We should probably put VPL in our affiliations XXX
 

### 2018-12-20 Paper/sync.tex
added new Reiners model sims, plots, text in paper

@@
 We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, $R$, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
-We model the long-term stellar spin-down due to magnetic braking using the model derived in \citet{Matt2015} since it has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
+We consider two separate models for the long-term stellar spin-down due to magnetic braking to gauge how sensitive the interaction between tides and magnetic braking is in modifying P$_{rot}$ evolution in stellar binaries.
+
+The magnetic braking model presented in \citet{Matt2015} has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field and is strongly dependent on the Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
+The \citet{Matt2015} model predicts that below a certain $Ro$ for rapidly rotating stars, stellar magnetic activity saturates at a constant value, producing a magnetic braking torque that is directly proportional to the stellar rotation rate.
+For more slowly-rotating unsaturated stars,
+\begin{equation} \label{eqn:mattUnSat}
+\frac{dJ}{dt} = -\frac{dJ}{dt}\Bigg|_0 \left( \frac{\tau_{cz}}{\tau_{cz \odot}} \right) \left( \frac{\omega}{\omega_{\odot}} \right)^3 
+\end{equation}
+while for rapidly rotating saturated stars,
+\begin{equation} \label{eqn:mattSat}
+\frac{dJ}{dt} = -\frac{dJ}{dt}\Bigg|_0 \chi^2 \left( \frac{\omega}{\omega_{\odot}} \right) 
+\end{equation}
+where
+\begin{equation} \label{eqn:matt0}
+\frac{dJ}{dt}\Bigg|_0 = 9.5 \times 10^{30} \mathrm{erg} \left( \frac{R}{R_{\odot}} \right)^{3.1} \left( \frac{M}{M_{\odot}} \right)^{0.5}.
+\end{equation}
+Saturated magnetic braking occurs for $Ro \leq Ro_{\odot}/\chi$ for $\chi = 10$ where \citet{Matt2015} defines $\chi = Ro_{\odot}/Ro_{sat}$.
+We adopt all model parameters given in Table 1 from \citet{Matt2015} and compute $\tau_{cz}$ using Eqn.
+(36) from \citet{Cranmer2011}.
+
+The second model we consider was developed by \citet{Reiners2012}, who modeled stellar spin-down as a function stellar magnetic field strength.
+Similar to the \citet{Matt2015} model, \citet{Reiners2012} suggests that magnetic braking occurs in a unsaturated ($\omega < \omega_{crit}$) and unsaturated ($\omega \geq \omega_{crit}$) regime as follows:
+\begin{equation} \label{eqn:reiners}
+\begin{split}
+\frac{dJ_{\star}}{dt} & = -C \left[ \omega \left(\frac{R^{16}}{m^2} \right)^{1/3} \right] \text{for $\omega \geq \omega_{crit}$} \\
+\frac{dJ_{\star}}{dt} & = -C \left[ \left( \frac{\omega}{\omega_{crit}} \right)^4 \omega \left(\frac{R^{16}}{m^2} \right)^{1/3} \right] \text{for $\omega < \omega_{crit}$}.
+\end{split}
+\end{equation}
+\citet{Reiners2012} tuned their model to match the Sun's P$_{rot}$ and the distribution of P$_{rot}$ of Gyr-old low-mass field stars and find $C = 2.66 \times 10^3 \text{ (gm$^5$ cm$^{-10}$ s$^3$)$^{1/3}$}$, $\omega_{crit} = 8.56 \times 10^{-6}\text{ s$^{-1}$}$ for $M > 0.35$ M$_{\odot}$, and for late M dwarfs with $M \leq 0.35 M_{\odot}$, $\omega_{crit} = 1.82 \times 10^{-6} \text{ s$^{-1}$}$.
+
 We model the net change in the stellar rotation rate due to stellar evolution and magnetic braking via the following equation 
@@
 Once the stars reach the main sequence, their structure changes slowly, allowing magnetic braking to dominate the stellar angular momentum evolution, significantly spinning-down the stars over long timescales.
-The $r_g$ evolution noticably differs between the stars as the M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
-
-XXX Paragraph, equation and plot for Matt15? XXX
+The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} magnetic braking law.
-Each dot denotes the approximate time when each star reaches the main sequence.
-XXX Reiners & Mohanty??
-XXX}%
+   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} and \citet{Reiners2012} magnetic braking equations.
+Each dot denotes the approximate time when each star reaches the main sequence.}%
     \label{fig:stellarExample}%
@@
 
+XXX Discuss how at long Porb, the CTL dissipation becomes super weak.
+Cite Leconte 2010 equations to justify how if we converted tau to Q, we'd get super large Q, and hence weak tides.
+This is a generic result.
+
 Due to the discontinuities in the equilibrium tidal model equations, for example in Eq.~(\ref{eqn:cpl:eps}) when $\omega \approx n$, and due to the inherant discreteness of numerical integrations, numerical solutions for the CPL and CTL models can produce unphysical evolution.
@@
 
-\subsection{The Weak Tidal Torque Regime} \label{sec:eq}
+\subsection{Interaction Between Magnetic and Tidal Braking} \label{sec:eq}
+
+XXX discuss \citet{Matt2015} and \citet{Reiners2012} interactions
 
@@
 supersynchronous rotators near P$_{orb}/$P$_{rot} = 1.5$ require $Q \lsim 10^{5.5}$.
-XXX This fig and the next are real eye charts.
-Can you blow them up to be full page?
-XXX}%
+}%
     \label{fig:qmap}%
@@
 
-In the absence of external forces, the long-term P$_{rot}$ evolution of low-mass stars is governed by magnetic braking.
+We compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking to gauge the impact of tidal torques on shaping P$_{rot}$ distributions.
+We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar} with initial conditions sampled from the same mass and P$_{rot}$ distributions used for the binary simulations described $\S$~\ref{sec:methods}.
+In Fig.~\ref{fig:protDist}, we display P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
+
+% In the absence of external forces, the long-term P$_{rot}$ evolution of low-mass stars is governed by magnetic braking.
 In stellar binaries, however, tidal torques compete against magnetic braking to drive P$_{rot}$ towards P$_{eq}$, e.g.
 Fig.~\ref{fig:eqPer}, causing their P$_{rot}$ evolution to differ from that of a single star.
-Here we compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking.
-We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar} with initial conditions sampled from the same mass and P$_{rot}$ distributions used for the binary simulations described $\S$~\ref{sec:methods}.
-In Fig.~\ref{fig:protDist}, we display P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
-XXX IS this paragraph necessary?
+Here w
 
@@
 
-Both tidal models predict a substantial population of fast rotators (P$_{rot} < 20$ d) not seen in the single star model at any age, except for young (age $\lsim 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$).
+There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
+The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) that are nearly at or have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.
+In stark contrast, both tidal models predict a substantial population of tidally-locked fast rotators (P$_{rot} < 20$ d).
 Under the CPL and CTL models, $19\%$ and $17\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.
-At the long P$_{rot}$ end, the CPL predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models.
-This feature is not due to long-term spin-down via magnetic braking as this population lies above the old edge of single star sequence.
-Instead tidal-locking in long P$_{orb}$ binaries produces these slow rotators as shown in Fig.~\ref{fig:lockedCPL}.
+At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models.
+Tidal-locking in long P$_{orb}$ binaries with low tidal $Qs$, e.g.
+Fig.~\ref{fig:lockedCPL}, produces these slow rotators.
 
-There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
-In stark contrast for many binaries, especially for P$_{rot} < 20 $ d, both tidal models predict that age does not strongly correlate with P$_{rot}$.
-We find that the median ages and $68\%$ interval for the CPL and CTL models fast rotators (P$_{rot} \lsim 20$ d) are $3.1^{+2.8}_{-1.7}$ Gyr and $2.8^{+2.8}_{-1.5}$ Gyr, respectfully, compared to much younger stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr for the single star model.
+Both tidal models predict that age does not strongly correlate with P$_{rot}$, especially for P$_{rot} < 20 $ d, as many of these systems tidally-lock at short P$_{orb}$.
+The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $3.1^{+2.8}_{-1.7}$ Gyr and $2.8^{+2.8}_{-1.5}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
 
-The only rapidly rotating single stars are young stars with ages $< 2$ Gyr that have not yet experienced substantial angular momentum loss due to magnetic braking.
-In binaries, however, tidal torques can lock stellar P$_{rot}$ to P$_{eq}$, especially at short P$_{orb}$ where tidal torques are stronger, creating a long-lasting population of fast rotators with a wide age range.
+Tidal torques pose a fundamental problem for inferring ages of rapidly-rotating stars, especially in the case where stellar bin
+
 If one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict that ages $\lsim 2$ Gyr, e.g.
@@
 
-At the long P$_{rot}$ end of the distribution, the CPL model predicts a population of tidally-locked extremely slow rotators that single star models fail to produce.
-Gyrochronology methods would incorrectly infer old ages $> 7$ Gyr for this slowly rotating population, whereas most of these binaries have ages $\lsim 6$ Gyr, with some being $\sim 1-4$ Gyr old.
 We caution that any application of gyrochonlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d or P$_{rot} \gsim 70$, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
 
-XXX This subsection is long and meandering. And obvious. I think you can just spend a couple paragraphs pointing out the top-level statistics and then go for the gyrochronology jugular. XXX
+XXX This subsection is long and meandering.
+And obvious.
+I think you can just spend a couple paragraphs pointing out the top-level statistics and then go for the gyrochronology jugular.
+be direct, scare gyro people
 
-be direct, scare gyro people
+% In binaries, however, tidal torques can lock stellar P$_{rot}$ to P$_{eq}$, especially at short P$_{orb}$ where tidal torques are stronger, creating a long-lasting population of fast rotators with a wide age range.
+% This feature is not due to long-term spin-down via magnetic braking as this population lies above the old edge of single star sequence. Instead
 
@@
 We discuss observational tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
-XXX Almost seems like this should be a separate paper...
-XXX
 
@@
 Remember that CPL and CTL are truncated infinite series, so locking into 1:1 vs 3:2 might not be the only possible outcomes.
-In Murray & Dermott they show how one can use a CPL model to produce the "pseudo-equlibrium" period that is a continuous function of $e$.
+In Murray and Dermott they show how one can use a CPL model to produce the "pseudo-equlibrium" period that is a continuous function of $e$.
 XXX

### 2018-12-20 Analysis/eqPerMatt.py

@@ -1,0 +1,178 @@
+"""
+
+@author: David P. Fleming, University of Washington, Oct 2018
+@email: dflemin3 (at) uw (dot) edu
+
+This script plots stellar-tidal evolution
+
+"""
+
+import numpy as np
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 20.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+# Path to sim results
+cplDir = "../Sims/CPLTestMatt"
+ctlDir = "../Sims/CTLTestMatt"
+logfile_name = "bintides.log"
+dirs = ["5", "10", "20", "30", "40", "50", "60"]
+colors = ["C0", "C1", "C2", "C3", "C4", "C5", "C6"]
+YEARSEC = 3.154e+7 # Seconds per year
+DAYSEC = 24.0*3600.0 # Seconds per day
+
+### Two Panel Plot ###
+
+fig, axes = plt.subplots(ncols=2, figsize=(14, 6))
+
+for ii, dir in enumerate(dirs):
+    # Load data
+    # saOutputOrder	Time -TotEn -TotAngMom -Radius -RotPer -EqRotPer DRotPerDtEqtide DRotPerDtStellar Ecce -OrbPer RadGyra -SurfEnFluxTotal
+    cpl = np.genfromtxt(os.path.join(cplDir, dir, "bintides.secondary.forward"))
+    ctl = np.genfromtxt(os.path.join(ctlDir, dir, "bintides.secondary.forward"))
+
+    # Pull LockTimes out of logfile
+    with open(os.path.join(os.path.join(cplDir, dir, logfile_name)), 'r') as f:
+        logfile_in = f.read()
+
+    matches = []
+    for line in logfile_in.split("\n"):
+        if line.startswith("(LockTime)"):
+            matches.append((line.split()[-1]))
+
+    # Convert lock times to years
+    dLockTimeCPL = float(matches[3])/YEARSEC
+
[…]

### 2018-12-20 Analysis/eqPerReiners.py

@@ -21,8 +21,8 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Path to sim results
-cplDir = "../Sims/CPLTest"
-ctlDir = "../Sims/CTLTest"
+cplDir = "../Sims/CPLTestReiners"
+ctlDir = "../Sims/CTLTestReiners"
 logfile_name = "bintides.log"
 dirs = ["5", "10", "20", "30", "40", "50", "60"]
 colors = ["C0", "C1", "C2", "C3", "C4", "C5", "C6"]
@@ -107,7 +107,7 @@ axes[1].axhline(0, lw=2.5, ls=":", color="k")
 
 # Save!
 fig.tight_layout()
-fig.savefig("../Plots/eqPerTwoPanel.pdf",bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/eqPerTwoPanelReiners.pdf",bbox_inches="tight", dpi=600)
 
 ### One Panel Plot ###
 
@@ -138,12 +138,12 @@ ax.set_xlim(1.0e6, cpl[-1,0])
 ax.set_ylim(0, 1.3)
 
 # Save!
-fig.savefig("../Plots/eqPer.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/eqPerReiners.pdf", bbox_inches="tight", dpi=600)
 
 ### Short Porb Plot ###
 
-ctlDir = "../Sims/Peq"
-cplDir = "../Sims/Peq"
+ctlDir = "../Sims/PeqReiners"
+cplDir = "../Sims/PeqReiners"
 ctlDirs = ["0.1", "0.01", "0.001"]
 cplDirs = ["6", "7", "8"]
 labels = [r"$Q= 10^6$, $\tau = 10^{-1}$ s", r"$Q=10^7$, $\tau = 10^{-2}$ s",
@@ -172,7 +172,7 @@ ax.set_xlabel("Time [yr]", fontsize=30)
 ax.set_xscale("log")
 ax.axhline(1, lw=2.5, ls=":", color="k")
 ax.set_xlim(1.0e6, cpl[-1,0])
-ax.set_ylim(0, 1.4)
+ax.set_ylim(0, 2.5)
 
 # Save!
-fig.savefig("../Plots/eqPerShortPorb.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/eqPerShortPorbReiners.pdf", bbox_inches="tight", dpi=600)

### 2018-12-20 Analysis/locking.py

@@ -10,18 +10,22 @@ in the Kepler field.
 
 Script Output:
 
+Prot Stats:
+CPL Prot median: 39.7426215
 CPL locked Prot median: 27.049087
 CPL unlocked Prot median: 41.0628235
 CPL interacting Prot median: 51.7563165
 CPL fraction of locked binaries: 0.2518
 CPL fraction of not locked but interacting binaries: 0.0956
 
+CTL Prot median: 41.87753
 CTL locked Prot median: 4.683883
 CTL unlocked Prot median: 41.19235
 CTL interacting Prot median: 49.27336
 CTL fraction of locked binaries: 0.0202
 CTL fraction of not locked but interacting binaries: 0.1257
 
+Age Stats:
 CPL locked Age median: 4.511807432302286
 CPL unlocked Age median: 3.6993176339544354
 CPL interacting Age median: 4.518420938215314
@@ -30,6 +34,7 @@ CTL locked Age median: 3.8579810329094926
 CTL unlocked Age median: 3.9204912513099366
 CTL interacting Age median: 4.604624291466229
 
+Porb Stats:
 CPL locked Porb median: 27.947582500000003
 CPL unlocked Porb median: 60.4407475
 CPL interacting Porb median: 57.719126499999994
@@ -37,8 +42,6 @@ CPL interacting Porb median: 57.719126499999994
 CTL locked Porb median: 4.7715415
 CTL unlocked Porb median: 51.597271
 CTL interacting Porb median: 58.190884
-
-
 
 """
 
@@ -201,12 +204,16 @@ fig.savefig("../Plots/lockedCTL.pdf", bbox_inches="tight", dpi=600)
 
 # Prots in d
 print("Prot Stats:")
+print("Single Prot median:",np.median(single["Pri_ProtAge"]))
+print()
+print("CPL Prot median:",np.median(cpl["Pri_ProtAge"]))
 print("CPL locked Prot median:",np.median(cpl["Pri_ProtAge"][cpl["Locked"]]))
 print("CPL unlocked Prot median:",np.median(cpl["Pri_ProtAge"][cpl["Free"]]))
 print("CPL interacting Prot median:",np.median(cpl["Pri_ProtAge"][cpl["Interacting"]]))
 print("CPL fraction of locked binaries:",np.mean(cpl["Locked"]))
 print("CPL fraction of not locked but interacting binaries:",np.mean(cpl["Interacting"]))
[…]

### 2018-12-20 Analysis/subsync.py

@@ -41,10 +41,11 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 plotLurie = True
+bins = "auto"
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLTorqueNov9.csv")
-ctl = pd.read_csv("../Data/mcCTLTorqueNov9.csv")
+cpl = pd.read_csv("../Data/mcCPLTorqueUniformDec19.csv")
+ctl = pd.read_csv("../Data/mcCTLTorqueLargeTauUniformDec19.csv")
 
 # Load in cleaned sample of Prot and Ecc from Lurie+2017
 lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
@@ -71,7 +72,7 @@ if plotLurie:
 # Format
 ax0.set_rasterization_zorder(0)
 ax0.set_xlim(0,10)
-ax0.set_ylim(0.4, 1.3)
+ax0.set_ylim(0.6, 1.25)
 ax0.set_xlabel("P$_{orb}$ [d]", fontsize=30)
 ax0.set_ylabel("P$_{orb}$ / P$_{rot}$", fontsize=30)
 leg = ax0.legend(loc="lower left", framealpha=0.0, fontsize=22)
@@ -87,6 +88,117 @@ cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Eccentricity", fontsize=25)
 
 fig.savefig("../Plots/subsync.pdf", bbox_inches="tight", dpi=600)
+
+### Make version of above plot with marginal Porb/Prot histogram with CTL ###
+
+fig = plt.figure(figsize=(9,9))
+
+gs = GridSpec(4,4)
+
+ax = fig.add_subplot(gs[1:4,0:3])
+ax_marg_y = fig.add_subplot(gs[1:4,3])
+
+im = ax.scatter(ctl["Age_Porb"], ctl["Age_Porb"]/ctl["Pri_ProtAge"],
+                c=ctl["Age_Ecc"].values, cmap="viridis", zorder=1,
+                s=40, marker="o", vmin=0, vmax=0.3, label="CTL")
+
+# Line at synchronization
+ax.axhline(1, lw=2, color="black", ls="-", zorder=2)
+
+if plotLurie:
+    # Plot Lurie+2017 data
[…]

### 2018-12-20 Sims/PeqMatt/run.py

@@ -1,0 +1,15 @@
+# Run all vplanet simulations
+import subprocess
+import sys
+import os
+
+# Find all subdirectories
+dir_path = os.path.dirname(os.path.realpath(__file__))
+dirs = ["0.001", "0.01", "0.1", "6", "7", "8", "9"]
+
+# Run simulations
+for dir in dirs:
+    os.chdir(os.path.join(dir_path,dir))
+
+    # Run simulation
+    subprocess.call(['vplanet', 'vpl.in'])

### 2018-12-27 Analysis/protDistReiners.py
adding reiners simulation data

@@ -1,0 +1,177 @@
+"""
+
+@author: David P. Fleming, University of Washington, Nov 2018
+@email: dflemin3 (at) uw (dot) edu
+
+Prot distributions as a function of stellar mass and age for the CPL, CTL, and
+single star models uniform over ages [1,7] Gyr, roughly field star ages.
+
+Script Output:
+
+CPL median Prot: 29.687965000000002
+CTL median Prot: 29.1996125
+Single median Prot: 31.452574500000004
+
+CPL Fraction of Stars with Prot < 7.5 d: 0.0692
+CTL Fraction of Stars with Prot < 7.5 d: 0.0673
+Single Fraction of Stars with Prot < 7.5 d: 0.0269
+
+CPL Fraction of Stars with Prot < 20 d: 0.228
+CTL Fraction of Stars with Prot < 20 d: 0.233
+Single Fraction of Stars with Prot < 20 d: 0.1082
+
+CPL Median Age of Stars with Prot < 20 d: 2.6490630945051183
+CTL Median Age of Stars with Prot < 20 d: 2.7346587999659704
+Single Mediage Age of Stars with Prot < 20 d: 1.6656389380019303
+
+CPL + Age of Stars with Prot < 20 d: 5.434789920095749
+CTL + Age of Stars with Prot < 20 d: 5.474016962602814
+Single + Age of Stars with Prot < 20 d: 2.605113602559185
+
+CPL - Age of Stars with Prot < 20 d: 1.384067339193213
+CTL - Age of Stars with Prot < 20 d: 1.4189953162291897
+Single - Age of Stars with Prot < 20 d: 1.189225831520236
+
+
+
+Interpretation:
+
+Binary models, regardless of tidal model, systematic rotate more quickly than
+single-star only models due to tidal torques.  Both tidal models predict a large
+population of rapid rotators (Prot < 7.5 d, < 20 d) than single star models fail
+to do, regardless of age, for stars with M > 0.5 Msun, aka GK stars which Kepler
+is most sensitive to. Therefore, fast rotators in Kepler field can be explained
+by tidally-interacting binaries. Furthermore, note that the binary Prot distribution
[…]

### 2018-12-27 Analysis/subsync.py

@@ -8,6 +8,8 @@ on the subsynchronous rotators.
 
 Script output:
 
+CTL:
+
 Median subsync tidal tau: 0.17078373819999998
 Median supersync tidal tau: 0.466656378
 Median subsync ecc: 0.107152
@@ -16,6 +18,11 @@ Median subsync mass ratio: 0.6416154898583578
 Median supersync mass ratio: 0.6614121797592005
 Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.7182910547396528
 Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.09078771695594126
+
+CPL:
+
+Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6887052341597796
+Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.01928374655647383
 
 Interpretation:
 
@@ -88,6 +95,22 @@ cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Eccentricity", fontsize=25)
 
 fig.savefig("../Plots/subsync.pdf", bbox_inches="tight", dpi=600)
+
+# Compare with Lurie+2017 sample for 2 < Porb < 10 days
+# Lurie+2017: 72% in [0.92,1.2]
+# Lurie+2017: 15% in [0.84, 0.92]
+
+porbMask = (ctl["Age_Porb"] < 10) & (ctl["Age_Porb"] > 2)
+mask = (ctl["Age_Porb"]/ctl["Pri_ProtAge"] >= 0.92)
+mask = mask & (ctl["Age_Porb"]/ctl["Pri_ProtAge"] < 1.2)
+mask = mask & porbMask
+print("Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d:",np.sum(mask)/np.sum(porbMask))
+
+mask = (ctl["Age_Porb"]/ctl["Pri_ProtAge"] >= 0.84)
+mask = mask & (ctl["Age_Porb"]/ctl["Pri_ProtAge"] < 0.92)
+mask = mask & porbMask
+print("Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d:",np.sum(mask)/np.sum(porbMask))
+
 
 ### Make version of above plot with marginal Porb/Prot histogram with CTL ###
 
[…]

### 2018-12-27 Paper/sync.tex

@@
 Box 951580, Seattle, WA 98195}
+\affil{NASA Astrobiology Institute - Virtual Planetary Laboratory Lead Team, USA}
 
@@
 New York, NY 10010}
+\affil{NASA Astrobiology Institute - Virtual Planetary Laboratory Lead Team, USA}
 
@@
 Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries.
+Here we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models.
+We find that many binaries with P$_{orb} \lsim 20$ d tidally-lock, and most with $P_{orb} \lsim 4$ d tidally-lock into synchronous rotation on circularized orbits.
+At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.
+In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, reproducing the subsynchronous primary stars in eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}.
+Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the CPL equilibrium tidal model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
+Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
+gyrochronology methods can fail unless one accounts for binarity.
+We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to disciminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
+
+ %Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries.
 Here we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters.
@@
 These findings suggest that synchronization is not a given for tidally-interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$.
-XXX Not sure why suggests consider longer Porb XXX Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
+Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
 
@@
 The \citet{Matt2015} model predicts that below a certain $Ro$ for rapidly rotating stars, stellar magnetic activity saturates at a constant value, producing a magnetic braking torque that is directly proportional to the stellar rotation rate.
-For more slowly-rotating unsaturated stars,
+The angular momentum loss for rapidly-rotating saturated stars is given by
 \begin{equation} \label{eqn:mattUnSat}
@@
 \end{equation}
-while for rapidly rotating saturated stars,
+while for more slowly-rotating unsaturated stars,
 \begin{equation} \label{eqn:mattSat}
@@
 The second model we consider was developed by \citet{Reiners2012}, who modeled stellar spin-down as a function stellar magnetic field strength.
-Similar to the \citet{Matt2015} model, \citet{Reiners2012} suggests that magnetic braking occurs in a unsaturated ($\omega < \omega_{crit}$) and unsaturated ($\omega \geq \omega_{crit}$) regime as follows:
+Similar to the \citet{Matt2015} model, \citet{Reiners2012} posits that magnetic braking can be divided into a unsaturated ($\omega < \omega_{crit}$) and unsaturated ($\omega \geq \omega_{crit}$) regime as follows:
 \begin{equation} \label{eqn:reiners}
 \begin{split}
-\frac{dJ_{\star}}{dt} & = -C \left[ \omega \left(\frac{R^{16}}{m^2} \right)^{1/3} \right] \text{for $\omega \geq \omega_{crit}$} \\
-\frac{dJ_{\star}}{dt} & = -C \left[ \left( \frac{\omega}{\omega_{crit}} \right)^4 \omega \left(\frac{R^{16}}{m^2} \right)^{1/3} \right] \text{for $\omega < \omega_{crit}$}.
+\frac{dJ_{\star}}{dt} & = -C \left[ \omega \left(\frac{R^{16}}{M^2} \right)^{1/3} \right] \text{for $\omega \geq \omega_{crit}$} \\
+\frac{dJ_{\star}}{dt} & = -C \left[ \left( \frac{\omega}{\omega_{crit}} \right)^4 \omega \left(\frac{R^{16}}{M^2} \right)^{1/3} \right] \text{for $\omega < \omega_{crit}$}.
 \end{split}
@@
 
-In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar.
+In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar, using both magnetic braking models.
 We assume all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr.
@@
 Once the stars reach the main sequence, their structure changes slowly, allowing magnetic braking to dominate the stellar angular momentum evolution, significantly spinning-down the stars over long timescales.
+Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution.
 The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
@@
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} and \citet{Reiners2012} magnetic braking equations.
+   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} (solid) and \citet{Reiners2012} (dashed) magnetic braking equations.
 Each dot denotes the approximate time when each star reaches the main sequence.}%
@@
  Equilibrium tidal models, first introduced by \citep{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega$, and obliquities $\psi$, of both gravitating bodies due to tidal torques.
-Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the bodies' centers of mass due to friction within each.
+Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the bodies' centers of mass due to friction within each body.
+This assumption is typically referred to as the ``weak friction approximation" \citep{Zahn2008}.
 The tidal bulges cause torques that permit the exchange of angular momentum between the orbit and both bodies' spins.
@@
 Under these assumptions, the tidal evolution is analogous to a driven, damped harmonic oscillator \citep{Greenberg2009}.
-For stars, equilibrium tidal models assume that tidal forces dissipate energy in convective regions via viscous turbulence \citep[see][]{Zahn2008}, corresponding to tidal dissipation in the outer convective layers of low-mass stars.
+For low-mass stars, equilibrium tidal models assume that tidal forces dissipate energy in the outer-convective regions via viscous turbulence \citep[see][]{Zahn2008}.
+Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citep[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
 We refer the reader to \citet{Barnes2017} for an in-depth discussion of the assumptions and limitations of equilibrium tidal models.
-Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citep[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
-Here, we consider two equilibrium tidal models to study the secular spin-orbital evolution of low-mass stellar binaries.
+Here, we consider two common equilibrium tidal models to study the secular spin-orbital evolution of low-mass stellar binaries.
 
@@
 
-The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass.
+The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass and is $2^{nd}$ order in $e$.
 We use the \eqtide implementation of the CPL model in VPLanet following the derivation of \citet{FerrazMello2008}.
@@
 The ``Constant Time Lag" (CTL) \citep[][]{Hut1981,Leconte2010} equilibrium tidal model assumes a constant time interval between the body's tidal bulge and the passage of the tidally-interacting companion.
-In this formalism, unlike the CPL model, the CTL model is continuous over a range of tidal wave frequencies.
+In this formalism, unlike the CPL model, the CTL model is continuous over a range of tidal wave frequencies and applicable for large $e$.
 However if the assumption of linearity is relaxed, i.e.
-frequencies associate with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}, hence we do not make this assumption and hold the tidal time lag fixed.
+frequencies associated with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}, hence we do not make this assumption and hold the tidal time lag fixed.
 We use the \eqtide implementation of the CTL model in VPLanet following the derivation of \citet{Leconte2010}.
@@
 
-The CTL model, however, is continuous over a range of tidal frequencies and therefore predicts a P$_{eq}$ that is a continuous function of both $e$ and $\psi$.
+We note that two discrete rotation states are not the only permitted ones for tidally-locked systems under the CPL formalism.
+For example, An alternate derivation of P$_{eq}$ for orbiters with rotation axes perpendicular to the orbital plane under the CPL model predicts
+\begin{equation} \label{eqn:cpl:eqPerCont}
+P_{eq} = \frac{P_{orb}}{1 + 9.5e^2},
+\end{equation}
+a continous function of $e$ \citep{Goldreich1966b,Murray1999}.
+Here, we follow the suggestions of both \citet{Barnes2013} and \citet{Barnes2017} and use the discrete P$_{eq}$ version of the CPL model for self-consistency.
+
+The CTL model is continuous over a range of tidal frequencies and therefore predicts a P$_{eq}$ that is a continuous function of both $e$ and $\psi$.
 Following \citet{Barnes2017}, we define the CTL P$_{eq}$ by
@@
 
-XXX discuss Rory's concern of how the above discussion only applies in the truncated series CPL formalism, whereas Murray and Dermott derived a continuous Peq for the CPL model as well
+In general, a continuous P$_{eq}$ and the discrete 1:1 and 3:2 spin-orbit commensurabilities are not the only equilibrium rotation states for tidally-locked rotators predicted by equilibrium tidal models.
+For example, \citet{Rodriguez2012} show that tidally-interacting bodies can get captured into many spin-orbit resonances states, e.g.
+2:1, 5:2, 4:3, etc.
+Although our models do not resolve capture into such states, we search for evidence of them in data of the spin-orbital states \kepler EBs.
 
-XXX Discuss how at long Porb, the CTL dissipation becomes super weak.
-Cite Leconte 2010 equations to justify how if we converted tau to Q, we'd get super large Q, and hence weak tides.
-This is a generic result.
+\subsubsection{Numerical Details of Tidal Locking}
 
@@
 We plot the tidal evolution for $a$, $e$, and $\omega$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively, in Fig.~\ref{fig:tidalExample}.
-Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally-locked state, transfering rotational angular momentum into the orbit in the process.
+Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally-locked state, transferring rotational angular momentum into the orbit in the process.
 At late times, both the CPL and CTL drive the binaries towards orbital circularization, with tidal dissipation decreasing P$_{orb}$.
@@
 For each simulation, we sample the primary's mass uniformly over $[0.1, 1]$ M$_{\odot}$.
-Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of young stars in the ${\sim}2$ Myr old Orion Nebula Cluster \citep{Matt2015}.
+Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of young stars in the ${\sim}2$ Myr old Orion Nebula Cluster \citep{Stassun1999,Herbst2001,Herbst2002,Rodriguez-Ledesma2009}.
 We compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010,Moe2018}.
-Given the inherent uncertainty in and complexity of the formation of binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$] and uniformly sampling the initial P$_{orb}$ over [$3,100$] d.
+Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$] and uniformly sampling the initial P$_{orb}$ over [$3,100$] d.
 Although the CTL model is applicable for $e > 0.3$, the CPL model is not, so we restrict $e \leq 0.3$ to allow us to compare both models.
@@
 These parameters can also vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass stars' evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
-Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a wide range of values exist in the literature.
-Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s log-uniformly over $[10^4,10^7]$ and $\tau$ log-uniformly over $[10^{-2},1]$ s.
-Note that in general, there is no expression to compute $Q$ as a function of $\tau$, or vice versa, except in some special cases where approximations exist, e.g.
+Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a range of values exist in the literature.
+Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s log-uniformly over $[10^4,10^8]$ and $\tau$ log-uniformly over $[10^{-2},10]$ s.
+There is no general expression to compute $Q$ as a function of $\tau$, or vice versa, except in some special cases where approximations exist, e.g.
 Eqn.
 (2) from \citet{Heller2011}.
-All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk, that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017}, would likely have dissipated \citep{Haisch2001}.
+
+We explore the impact of both the \citet{Matt2015} and \citet{Reiners2012} magnetic braking models, but default to using the \citet{Matt2015} unless noted otherwise.
+All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.
 
@@
 \subsection{Interaction Between Magnetic and Tidal Braking} \label{sec:eq}
-
-XXX discuss \citet{Matt2015} and \citet{Reiners2012} interactions
 
@@
 
+\subsubsection{\citet{Reiners2012}}
+
+The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but rather is a generic outcome of the coupling of magnetic braking and weak tidal torques.
+XXX
+
 \begin{figure}
@@
 
-Spin-orbit synchronization is the typical outcome for tidally-interacting binaries according to the CPL model, with synchronization occurring for most binaries with P$_{orb} < 10$ d, regardless of $Q$.
+Spin-orbit synchronization is the typical outcome for binaries with P$_{orb} < 10$ d according to the CPL model for most values of $Q$.
 The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to P$_{orb} \approx 100 $ d for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
-According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, regardless of $\tau$.
+According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \approx 10$ s.
 Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e.
@@
 This population corresponds to the evolution described in $\S$~\ref{sec:eq} in which magnetic braking initially spins down stars past the tidally-locked state, but in the long-term, tidal torques spin up the stars, shepherding them towards the tidally-locked state.
-This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ (see Figs.~\ref{fig:eqPer},\ref{fig:eqPerShortPorb}).
+This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ and $Q$ or $\tau$ (see Figs.~\ref{fig:eqPer},\ref{fig:eqPerShortPorb}).
 
@@
 For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems, especially for P$_{orb} > 20$ d.
-For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ as strong tidal torques always tidally-lock the binaries.
-In this case, P$_{eq}$ can only take two discrete values under the CPL model, i.e.
+For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ as strong tidal torques always tidally-lock the binaries into 1:1 or 3:2 spin-orbit commesurability.
+No stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 70$ d as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
 Eqn.~(\ref{eqn:cpl:eqPer}).
-No stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 70$ d as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability.
 
@@
 
-Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.0$ d and $4.68$ d) than unlocked (median CPL, CTL P$_{rot} = 41.1$ d and $41.2$ d) binaries.
+Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.6$ d and $7.4$ d) than unlocked (median CPL, CTL P$_{rot} = 41.3$ d and $40.8$ d) binaries.
 More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale as $R^5$, and $R$ increases with stellar mass.
@@
 
-Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we observe stellar binaries to tidally lock up to P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
+Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we find stellar binaries can tidally-lock over a wide range of P$_{rot}$ up to P$_{rot} =$ P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
 This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
 Moreover, \citet{Abt2004} find evidence for tides modifying P$_{rot}$ in binaries with P$_{orb}$ up to 500 d, implying tidal torques matter out long P$_{orb}$.
-Under the CTL model, however, binaries can only tidally-lock out to P$_{orb} \approx 20$ d, with only $2\%$ of binaries becoming tidally-locked, in stark contrast to the CPL model's predictions.
+Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d, in stark contrast to the CPL model's predictions.
+Some binaries can tidally-lock or tidally-interact out to $P_{rot} \approx 80$ d under the CTL model, but not nearly as many as the CPL model, nor can the CTL model tidally-lock binaries out to P$_{orb} = 100$ d, regardless of $\tau$.
+Only $7.5\%$ of binaries tidally-lock according to the CTL model, compared with $25\%$ of CPL model binaries, as tidal torques are generally stronger in the CPL formalism.
 
@@
 The interacting population tends to rotate more slowly than the unlocked population as at short P$_{orb}$, and hence P$_{rot}$, binaries preferentially tidally-lock due to stronger tidal torques.
-At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthing enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
-The CPL and CTL models predict that $35\%$ and $15\%$ of stars, respectivey, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
+At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthening enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
+The CPL and CTL models predict that $34\%$ and $25\%$ of stars, respectively, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
 The P$_{rot}$ - mass distribution for unlocked binaries resembles the single star sequence observed in the center right panel of Fig.~\ref{fig:protDist} indicating that magnetic braking and stellar evolution dictate their angular momentum evolution as expected.
@@
 As shown in Fig.~\ref{fig:lockedProtHist}, both models predict that tidal torques can influence the P$_{rot}$ evolution out to P$_{orb} \approx 80$ d, causing a significant departure from the long-term expected spin down of single stars due to magnetic braking.
-The CTL model predicts that binaries can tidally-lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL P$_{orb} = 4.8$ d.
-The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states, from P$_{rot} < 20$ d out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
+The CTL model predicts that the vast majority tidally-locked binaries, $86\%$, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL P$_{orb} = 7.6$ d.
+The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states with only $39\%$ locking into P$_{rot} < 20$ d, and many locking out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
 The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
@@
 In Fig.~\ref{fig:protDist}, we display P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
+We perform this procedure for stars losing angular momentum via magnetic braking according to both the \citet{Matt2015} and \citet{Reiners2012} magnetic braking models.
+
+\subsubsection{\citet{Matt2015} Magnetic Braking}
 
@@
 In stark contrast, both tidal models predict a substantial population of tidally-locked fast rotators (P$_{rot} < 20$ d).
-Under the CPL and CTL models, $19\%$ and $17\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.
+Under the CPL and CTL models, $18\%$ and $19\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.
 At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models.
@@
 
-Both tidal models predict that age does not strongly correlate with P$_{rot}$, especially for P$_{rot} < 20 $ d, as many of these systems tidally-lock at short P$_{orb}$.
-The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $3.1^{+2.8}_{-1.7}$ Gyr and $2.8^{+2.8}_{-1.5}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
+Both tidal models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{rot} < 20 $ d, as many of these systems tidally-lock at short P$_{orb}$.
+The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
 
-Tidal torques pose a fundamental problem for inferring ages of rapidly-rotating stars, especially in the case where stellar bin
+\subsubsection{\citet{Reiners2012} Magnetic Braking}
 
-If one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict that ages $\lsim 2$ Gyr, e.g.
+XXX
+
+\subsubsection{Impact of Binarity on Gyrochronology}
+
+Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronlogy.
+Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$, decoupling P$_{rot}$ from age.
+For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr, e.g.
 the right panel of Fig.~\ref{fig:protDist}.
-If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling its P$_{rot}$ from age, causing the predictions of gyrochronology models to fail.
-This effect poses a significant problem for studies of P$_{rot}$ distributions and their ability to constrain stellar ages.
-This problem likely exists in the \kepler field as \citet{Simonian2018} finds that many \kepler fast rotators are likely unresolved photometric binaries.
+If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail.
+This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal.
 
-We caution that any application of gyrochonlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d or P$_{rot} \gsim 70$, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
+In general, itt is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another, exacerbating this issue.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a significant contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are subject to large systematic errors.
+Moreover, this problem could be more significant that the aforementioned estimate implies as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
+We caution that any application of gyrochonlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
 
-XXX This subsection is long and meandering.
-And obvious.
-I think you can just spend a couple paragraphs pointing out the top-level statistics and then go for the gyrochronology jugular.
-be direct, scare gyro people
+Furthermore, any study of the angular momentum evolution of a stellar population will be fundamentally flawed unless one robustly accounts for binarity.
+Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
+Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
+The application of or the calibration of magnetic braking models to a sample of stellar rotation periods will be inherently flawed unless one controls for binarity.
 
@@
 
-We compare our simulation results to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing stellar binaries by \citet{Lurie2017}.
+We compare our simulation results using both tidal models and the \citet{Matt2015} magnetic braking model to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing stellar binaries by \citet{Lurie2017}.
 \citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with starspot modulations and visually inspected each light curve to ensure their accuracy.
@@
 Both models reproduce most trends in the \citet{Lurie2017} data, namely the large cluster of synchronized rotators and the population of subsynchronous rotators, both for P$_{orb} \lsim 20$ d.
-Magnetic braking and age imparts two noticable features in the data.
-First, the sparse cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} > 1.5$ at all P$_{orb}$ corresponds to young stars that have yet to significantly spin down due to magnetic braking or tides.
+Magnetic braking and age imparts two noticeable features in the data.
+First, supersynchronous rotators with P$_{orb}/$P$_{rot} > 1.5$, especially at long P$_{orb}$, corresponds to young stars that have yet to significantly spin down due to magnetic braking or tides.
 Second, the lower limit seen in both panels is a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down via magnetic braking in the weak tides regime over 7 Gyr, the longest age considered in our simulations.
@@
 The CTL model, however, seems to excel where the CPL models fails by producing both populations.
-XXX This is a big point that I'm not sure is obvious.
-Does a simple 1-D histogram show it?
-We really struggled to understand this population with John, so if this is really the explanation, it's a big deal.
-XXX We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
+We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
 
@@
 Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces this population.
-Our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.13$ s, compared to the median $\tau = 0.08$ s for the tidally-locked supersynchronous population.
+Our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
 
-\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $11\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
-Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $61\%$ of our CTL population.
-This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL equilibrium tide formalism can accurately model the tidal evolution of low-mass stellar binaries.
+\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, the cluster of stars those authors identified near $P_{orb}/$P$_{rot} = 0.9$, compared with $9\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
+Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$, compared with $72\%$ of our CTL population.
+In comparison, the CPL model (not pictured) predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, reproducing the synchronized \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
+
+This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL equilibrium tide formalism can accurately model the tidal evolution of low-mass stellar binaries, especially at short P$_{orb}$.
 The comparison between theory and observations, however, is limited in that the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and because \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$.
@@
 %   \caption{Distribution of stellar tidal $Q$ produced by convolving \citet{Lurie2017} observations of EBs' P$_{rot}$ and P$_{orb}$ with our CPL simulation results shown in Fig.~\ref{fig:qmap}.
-The dashed black lines, from left to right, indicate the $16^{th}$ percentile, median, and $84^{th}$ percentile.
-XXX Looks like your priors need to be wider :( XXX}%
+The dashed black lines, from left to right, indicate the $16^{th}$ percentile, median, and $84^{th}$ percentile.}%
 %    \label{fig:qLurie}%
@@
 %   \caption{Same as Fig.~\ref{fig:qLurie}, but using our CTL simulation results from Fig.~\ref{fig:taumap}.}%
-    \label{fig:tauLurie}%
+%    \label{fig:tauLurie}%
 %\end{figure}
@@
 
-Third, as shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 70$ d would provide strong evidence in favor of the CPL model as the CTL model cannot tidally-lock binaries past P$_{orb} \approx 20$ d, e.g.
-Fig.~\ref{fig:lockedCTL}, whereas the CPL model can tidally-lock binaries out to P$_{orb} \approx 100$ d.
+Third, as shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 70$ d would provide strong evidence in favor of the CPL model as the CTL model is unlikely to tidally-lock binaries past P$_{orb} \approx 20$ d, e.g.
+Fig.~\ref{fig:lockedCTL}, whereas the CPL model readily tidally-locks binaries out to P$_{orb} \approx 100$ d.
 We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small starspot modulation amplitudes \citep{Lurie2017,Reinhold2018}.
@@
 
-In this work, we probed the long-term angular momentum evolution of low-mass stellar binaries, with a focus on P$_{rot}$ in intermediate P$_{orb}$ binaries.
-We considered the impact of two common equilibrium tidal models, magnetic braking, and stellar evolution.
+In this work, we probed the long-term angular momentum evolution of low-mass stellar binaries, with a focus on P$_{rot}$ in short and intermediate P$_{orb}$ binaries.
+We considered the impact of two common equilibrium tidal models, two magnetic braking models, and stellar evolution.
 We performed a large suite of simulations for binaries with physically-motivated initial conditions out to P$_{orb} = 100$ and across a wide range of tidal dissipation parameters to examine the competition between tidal torques and magnetic braking for controlling the stellar P$_{rot}$ evolution.
-In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have synchronized spins and circularized orbits.
+In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have synchronized spins and circularized orbits, in good agreement with observations of \kepler EBs.
 We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking.
 Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact out to P$_{orb} \approx 80$ d, causing their P$_{rot}$ evolution to differ from the long-term spin down due to magnetic braking experience by single stars.
-Many binaries with P$_{orb} \lsim 20$ d tidally-lock, but the CPL model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
+Many binaries with P$_{orb} \lsim 20$ d tidally-lock, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.
 When binaries tidally-lock, tides fix P$_{rot} = $ P$_{eq}$, preventing further spin-down via magnetic braking, decoupling P$_{rot}$ from age.
@@
 We caution any application of gyrochronlogy methods to stars with P$_{rot} \lsim 20$ d or P$_{rot} \gsim 70$ should account for the possibility of stellar binarity to prevent deriving incorrect ages.
-XXX Whew, there's a lot here!
-Maybe a bulleted list is in order?
-And make sure a sentence for each is in the abstract.
-XXX
 
@@
 DPF and RB acknowledge that this work was supported by the NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement number NNA13AA93A.
-XXX We should probably put VPL in our affiliations XXX
 

### 2018-12-27 Paper/sync.tex
almost 2nd draft

@@
 
-Angular momentum evolution in low-mass (M$\lsim 1$ M$_{\odot}$) short-period (P$_{orb} \lsim 10$ d) stellar binaries is strongly influenced by tides.
+The long-term angular momentum evolution of low-mass (M$\lsim 1$ M$_{\odot}$) stars is controlled by magnetic braking, the torque exerted on stars due to the coupling of stellar winds to the surface magnetic field \citep{Dunn1961,Mestel1968}.
+Early on in stellar lifetimes, stars spin-up as they contract along the premain sequence.
+Once stars reach the main sequence, stellar radii remain mostly constant and magnetic braking saps angular momentum from low-mass stars, gradually spinning them down over time \citep{Skumanich1972}.
+Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \citep[e.g.
+Praesepe, ][]{Matt2015,Douglas2017} and field stars \citep[e.g.
+\kepler, ][]{Reiners2012,Matt2015,vanSaders2018}.
+Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as gyrochronlogy \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.
+
+In contrast, the angular momentum evolution in low-mass short-period (P$_{orb} \lsim 10$ d) stellar binaries is dominated by tides.
 Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}.
@@
 Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially for stellar spins.
-Tidal torques drive them towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
-Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
+Tidal torques drive binaries towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
+Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is less angular momentum in stellar spins than the binary orbit.
+As a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
 
-The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
+In low-mass binaries, both magnetic braking and tidal torques compete to shape the stellar P$_{rot}$ evolution.
+When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$.
+In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2017}.
+Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[e.g.
+P$_{rot} >$ P$_{orb}$, ][]{Habets1989,Zahn1994,Keppens1997}.
+This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations.
+
+The influence of tides is not necessarily limited to short P$_{orb}$ systems either, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered.
+The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.
+
+%The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
 Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking.
@@
 
-Understanding how tidal torques affect angular momentum evolution in low-mass stellar binaries is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g.
+Understanding how tidal torques and magnetic braking impact angular momentum evolution in low-mass stellar binaries is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g.
 Praesepe, ][]{Douglas2017} and field stars \citep[e.g.
@@
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.
-Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, imparting a contaminating signal that is not currently accounted for by modern models.
-Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that links the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, could be incorrect owing to the influence of tidal torques.
+Tidal torques can modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, or compete against magnetic braking to form subsychronous rotators, imparting a contaminating signal that is not currently accounted for by models.
+Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology could be incorrect owing to the influence of tidal torques.
 No previous study has quantified this effect.
 
-Here, we examine angular momentum evolution over the full main sequence lifetimes of low-mass stellar binaries using a realistic treatment of stellar evolution, magnetic braking, and tidal torques.
+There is currently a large number of \kepler binaries with known P$_{rot}$ and P$_{orb}$ \citep[e.g.][]{Lurie2017}.
+Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} are expected to detect additional low-mass eclipsing binaries, with Gaia parallaxes \citep{Gaia2016} poised to help refine these stellar parameters, potentially creating a rich dataset of the angular momentum budgets of low-mass binaries.
+In today's data-rich era of astronomy, it is critical to develop a theoretical framework to understand the angular momentum evolution of low-mass binaries due to stellar evolution, magnetic braking, and tidal torques.
+Such a framework can enable the characterization of the nature of tidal torques in binaries and allow for the long-term modeling of the evolutionary histories of individual systems.
+
+Here, we present a model for the angular momentum evolution of low-mass stellar binaries over their full premain and main sequence lifetimes using a realistic treatment of stellar evolution, magnetic braking, and tidal torques.
 We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.
@@
 We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:protDist} and $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.
-
-XXX Why not discuss tidal braking more?
-I expected an intro that described tidal affects on binaries, magnetic braking, and then stating that no study had properly linked the two.
-The tidal review is good, but more on MB would be helpful.
-XXX
-
-tides impact binary ecc, tides can brake spins, tidally-lock systems, keeping them fast rotators etc.
-Talk about magnetic braking as both magnetic braking and tides act in binaries.
-Talk about observations of Prot, like  \citet{Abt2004} citation and similar things like Lurie
 
@@
 Once the stars reach the main sequence, their structure changes slowly, allowing magnetic braking to dominate the stellar angular momentum evolution, significantly spinning-down the stars over long timescales.
-Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution.
+Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution, with the \citet{Reiners2012} predicting marginally stronger torques.
 The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
@@
 long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking.
+
+\subsubsection{Using the \citet{Matt2015} Magnetic Braking Model}
+
 In Fig.~\ref{fig:eqPer}, we plot the evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) formalisms.
@@
 
-\subsubsection{\citet{Reiners2012}}
+\subsubsection{Using the \citet{Reiners2012} Magnetic Braking Model}
 
 The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but rather is a generic outcome of the coupling of magnetic braking and weak tidal torques.
-XXX
+As demonstrated above, strong tidal torques under the CPL model tidally-lock each binary out to P$_{orb} = 60$ d.
+The CTL model, however, can only tidally-lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the other binaries into subsynchronous rotation, similar to the \citet{Matt2015} simulations.
+Curiously, all CTL subsynchronous rotators seem to converge to P$_{rot}/$P$_{eq} \approx 1.1$.
+In this state, tidal torques nearly balance magnetic braking, producing a long-lasting population of subsynchronous rotators, similar to what was observed above.
+In the right panel of Fig.~\ref{fig:eqPerReiners}, we see that the net P$_{rot}$ derivatives for these binaries oscillate about 0, but are slightly negative due to the mechanism described above.
+This slight negative derivative is effectively negligible, however, as the torque balance will likely hold the binaries in subsynchronous rotation for 10s of Gyrs.
+We note that in other exploratory simulations with larger $\tau$, we find that the 3 part mechanism described above holds, and tides win out over magnetic braking in the long-run, driving P$_{rot}$ back towards P$_{eq}$.
+
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelReiners.pdf}
+   \caption{Same format as Fig.~\ref{fig:eqPer}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
+    \label{fig:eqPerReiners}%
+\end{figure*}
+
+\subsubsection{Subsynchronous Rotation at Short P$_{orb}$}
+
+In Fig.~\ref{fig:eqPerShortPorb}, we examine this effect in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters.
+Subsynchronous rotation occurs in the general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
+Previous theoretical studies have also predicted subsynchronous rotation in short P$_{orb}$ binaries arising from the balance between tidal torques and magnetic braking \citep[e.g.][]{Habets1989,Zahn1994,Keppens1997} suggesting that this behavior is not an artifact of our choice of tidal or magnetic braking models, but rather a general outcome of the competition between magnetic braking and tidal evolution in low-mass binaries.
+Short P$_{orb}$ subsynchronous rotators can eventually tidally-lock via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
 
@@
 % In the long-term, magnetic braking weakens relative to tidal torques for longer P$_{rot}$, allowing tides to drive the stars back towards the tidally-locked state.
-
-In Fig.~\ref{fig:eqPerShortPorb}, we examine this effect in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters.
-Subsynchronous rotation occurs in the general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
-Previous theoretical studies have also predicted subsynchronous rotation in short P$_{orb}$ binaries arising from the balance between tidal torques and magnetic braking \citep[e.g.][]{Habets1989,Zahn1994,Keppens1997} suggesting that this behavior is not an artifact of our choice of tidal or magnetic braking models, but rather a general outcome of the competition between magnetic braking and tidal evolution in low-mass binaries.
-Short P$_{orb}$ subsynchronous rotators can eventually tidally-lock via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
-   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
+   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model.
 For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
@@
 
-XXX
+Similar to the \citet{Matt2015} simulations discussed above, the single star simulations display a clear age-dependence of P$_{rot}$.
+The \citet{Reiners2012} model's magnetic braking transition at $M = 0.35$ M$_{\odot}$ produces a sharp cutoff in the P$_{rot} - M$ distribution between late M dwarfs and more massive stars.
+More massive stars ($M \geq 0.35$ M$_{\odot}$) follow a smooth P$_{rot} -$ age sequence, while the late M dwarfs can remain rapidly rotating if they are young and in the saturated regime or extremely slowly rotating if the are older than about 4 Gyr.
+As in the \citet{Matt2015} model, the only rapid rotators (P$_{rot} \lsim 20$ d) are very young stars with ages $\lsim 2$ Gyr.
+The presence of extremely slowly rotating late M dwarfs, absent in the predictions of the \citet{Matt2015} model but seemingly present in nature \citep[e.g.][]{Newton2018}, could be used as an observational discriminant between the \citet{Matt2015} and \citet{Reiners2012} model, but a full investigation of which magnetic braking model accurately describes angular momentum loss in low mass stars is beyond the scope of this work.
+
+\begin{figure*}[t]
+	\includegraphics[width=\textwidth]{../Plots/protDistReiners.pdf}
+   \caption{Same format as Fig.~\ref{fig:protDist}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
+    \label{fig:protDistReiners}%
+\end{figure*}
+
+The general trends found in the coupled stellar-tidal simulations using the \citet{Matt2015} model persist in the simulations evolved using the \citet{Reiners2012} model, demonstrating that these findings are not magnetic braking model-dependant, but rather a generic result of the interaction between magnetic braking and tidal torques in low-mass binaries.
+Both the CPL and CTL simulations display a clear single-star magnetic braking-dominated sequence, but as before, tidally-locked binaries appear across a wide range of P$_{rot}$ and primary star masses.
+Both the CPL and CTL model tidally-lock many short P$_{orb}$ stars into rapid rotation across all masses, with both models predicting $23\%$ of stars have P$_{rot} < 20$ d, compared to $11\%$ of single-stars, mostly for late M dwarfs in the latter case.
+As shown above, the CTL model predicts that binaries can strongly tidally-interact out to long P$_{orb}$, seen out to P$_{rot} = 80$ d, while the CPL model can tidally-lock stars out to P$_{orb} \approx 100$ d.
+These strong tidal interactions decouple P$_{rot}$ from age in binaries, mirroring the behavior discussed above, especially for rapid rotators.
+The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.6^{+2.8}_{-1.3}$ Gyr and $2.7^{+2.7}_{-1.4}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.7^{+0.9}_{-0.5}$ Gyr, similar to the age distributions for rapid rotators evolved using the \citet{Matt2015} model.
 
@@
 Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronlogy.
-Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$, decoupling P$_{rot}$ from age.
+Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
 For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr, e.g.
-the right panel of Fig.~\ref{fig:protDist}.
+the right panel of Fig.~\ref{fig:protDist} or Fig.~\ref{fig:protDistReiners}.
 If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail.
@@
 
-In general, itt is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another, exacerbating this issue.
-Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a significant contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are subject to large systematic errors.
-Moreover, this problem could be more significant that the aforementioned estimate implies as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
-We caution that any application of gyrochonlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
+In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another, exacerbating this issue.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a significant contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are potentially subject to large systematic errors.
+Moreover, this problem could be more significant than the aforementioned estimate implies as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
+We caution that any application of gyrochronlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
 
@@
 First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all tidally-locked sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.
-This prediction is inconsistent with the data as no obvious spin-orbit commensurablities, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
+This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
 The CPL model can produce the population of subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
@@
 
-XXX The abov discussion is a bit tricky.
-Remember that CPL and CTL are truncated infinite series, so locking into 1:1 vs 3:2 might not be the only possible outcomes.
-In Murray and Dermott they show how one can use a CPL model to produce the "pseudo-equlibrium" period that is a continuous function of $e$.
-XXX
-
 The continuous supersynchronous rotation predicted by the CTL model seen in right panel of Fig.~\ref{fig:lurie7} seems to provide a much better match to the \citet{Lurie2017} data for both tests described above, providing evidence in favor of the CLT model.
-The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making this distinction conclusive.
+Moreover, it is unclear if low-mass binary stars can even lock into spin-orbit resonances, such as those discussed in \citet{Rodriguez2012}, as stellar convective envelopes lack a fixed shape \citep{Burkart2014,Lurie2017}.
+There is no obvious clustering of stars near any spin-orbit resonance, aside from 1:1 spin-orbit synchronization in the \citet{Lurie2017} \kepler EB data.
+The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making any distinction between models conclusive.
 More robust phase-resolved spectroscopic $e$ constrains would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a Gaussian process model as was done by \citet{Angus2018}.
 
-Third, as shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 70$ d would provide strong evidence in favor of the CPL model as the CTL model is unlikely to tidally-lock binaries past P$_{orb} \approx 20$ d, e.g.
-Fig.~\ref{fig:lockedCTL}, whereas the CPL model readily tidally-locks binaries out to P$_{orb} \approx 100$ d.
+These two tests can fail to discriminate between the CPL and CTL model, however, if the CPL model P$_{eq}$ is a continuous function of $e$, e.g.
+Eqn.~(\ref{eqn:cpl:eqPerCont}), as was argued by \citet{Goldreich1966b} and derived by \citet{Murray1999}.
+In such a case, one would need a large number of accurate and precise measurements P$_{orb}$ and $e$, with robust uncertainties, for tidally-interacting binaries to discriminate between the CPL and CTL continuous P$_{eq}$, e.g.
+Eqn.~(\ref{eqn:cpl:eqPerCont}) versus Eqn.~(\ref{eqn:ctl:eqPer}).
+In practice, this is likely too difficult as it requires extensive photometric and spectroscopic observations of many binaries.
+Since there is no obvious clustering of \kepler EBs near any spin-orbit resonance, the above two tests are likely insufficient.
+We offer a third observational test that provides the best method to discriminate between models.
+
+As shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 80$ d would provide strong evidence in favor of the CPL model, as the CTL model is unlikely to tidally-lock binaries past P$_{orb} \approx 20$ d and cannot tidally-lock stars beyond P$_{orb} \approx 80$ d, even with extreme $\tau$, e.g.
+Fig.~\ref{fig:lockedCTL}.
+The CPL model, however, readily tidally-locks binaries out to P$_{orb} \approx 100$ d.
 We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small starspot modulation amplitudes \citep{Lurie2017,Reinhold2018}.
@@
 Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact out to P$_{orb} \approx 80$ d, causing their P$_{rot}$ evolution to differ from the long-term spin down due to magnetic braking experience by single stars.
+We show that our predictions are not dependant on the choice of magnetic braking model, but rather are generic outcomes of the interaction between magnetic braking and tidal torques.
+
 Many binaries with P$_{orb} \lsim 20$ d tidally-lock, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.

### 2018-12-28 Analysis/eqPerMatt.py
added some annotations

@@ -96,6 +96,9 @@ axes[0].axhline(1, lw=2.5, ls=":", color="k")
 axes[0].set_xlim(1.0e6, cpl[-1,0])
 axes[0].set_ylim(0, 1.2)
 
+axes[0].text(1.2e6, 1.12, "Matt et al. (2015)",
+         ha="left", va="center", size=15, color="black", zorder=100)
+
 # Format right axis
 axes[1].set_ylabel("$d$P$_{rot}$/$dt$ [d/Gyr]", fontsize=25)
 axes[1].set_xlabel("Time [yr]", fontsize=25)

### 2018-12-28 Analysis/eqPerReiners.py

@@ -96,6 +96,9 @@ axes[0].axhline(1, lw=2.5, ls=":", color="k")
 axes[0].set_xlim(1.0e6, cpl[-1,0])
 axes[0].set_ylim(0, 1.2)
 
+axes[0].text(1.2e6, 1.12, "Reiners \& Mohanty (2012)",
+         ha="left", va="center", size=15, color="black", zorder=100)
+
 # Format right axis
 axes[1].set_ylabel("$d$P$_{rot}$/$dt$ [d/Gyr]", fontsize=25)
 axes[1].set_xlabel("Time [yr]", fontsize=25)

### 2018-12-28 Analysis/locking.py

@@ -158,9 +158,10 @@ color = (cpl.iloc[inds]["Pri_ProtAge"][cpl.iloc[inds]["Locked"]]/cpl.iloc[inds][
 im = ax.scatter(x, y, color="C0", marker="x", s=50, zorder=3, label="Locked")
 
 # Format plot
-ax.set_xlabel("Mass [M$_{\odot}$]", fontsize=28)
+ax.set_xlabel("Mass [M$_{\odot}$]", fontsize=30)
 ax.set_xlim(np.min(cpl["Pri_dMass"]), np.max(cpl["Pri_dMass"]))
-ax.set_ylabel("P$_{rot}$ [d]", fontsize=28)
+ax.set_ylabel("P$_{rot}$ [d]", fontsize=30)
+ax.set_title("CPL", fontsize=28)
 ax.set_ylim(ylims)
 ax.set_rasterization_zorder(0)
 leg = ax.legend(loc="upper left", framealpha=0.5, fontsize=15)
@@ -209,9 +210,10 @@ y = ctl.iloc[inds]["Pri_ProtAge"][ctl.iloc[inds]["Locked"]]
 im = ax.scatter(x, y, color="C0", marker="x", s=50, zorder=3, label="Locked")
 
 # Format plot
-ax.set_xlabel("Mass [M$_{\odot}$]", fontsize=28)
+ax.set_xlabel("Mass [M$_{\odot}$]", fontsize=30)
 ax.set_xlim(np.min(ctl["Pri_dMass"]), np.max(ctl["Pri_dMass"]))
-ax.set_ylabel("P$_{rot}$ [d]", fontsize=28)
+ax.set_ylabel("P$_{rot}$ [d]", fontsize=30)
+ax.set_title("CTL", fontsize=28)
 ax.set_ylim(ylims)
 ax.set_rasterization_zorder(0)
 leg = ax.legend(loc="upper left", framealpha=0.0, fontsize=15)

### 2018-12-28 Analysis/protDist.py

@@ -104,6 +104,9 @@ ax0.set_ylim(0.0, 100)
 ax0.text(0.15, 90, "CPL", ha="left", va="center", size=30, color="black",
          zorder=100)
 
+ax0.text(0.13, 112, "Matt et al. (2015)", ha="left", va="center", size=30,
+         color="black", zorder=100)
+
 ### CTL Plot ###
 ax1 = fig.add_subplot(gs[16])
 

### 2018-12-28 Analysis/protDistReiners.py

@@ -105,6 +105,9 @@ ax0.set_ylim(0.0, 100)
 ax0.text(0.4, 90, "CPL", ha="left", va="center", size=30, color="black",
          zorder=100)
 
+ax0.text(0.13, 112, "Reiners \& Mohanty (2012)", ha="left", va="center", size=30,
+         color="black", zorder=100)
+
 ### CTL Plot ###
 ax1 = fig.add_subplot(gs[16])
 

### 2018-12-28 Analysis/stellar.py

@@ -58,9 +58,9 @@ ax[0].scatter(time[ind_k], k[ind_k,1], s=75, color="C1")
 ax[0].scatter(time[ind_m], m[ind_m,1], s=75, color="C2")
 
 # Format
-ax[0].legend(loc="best", framealpha=0, fontsize=25)
-ax[0].set_ylabel("Radius [R$_{\odot}$]")
-ax[0].set_xlabel("$t - t_0$ [yr]")
+ax[0].legend(loc="best", framealpha=0, fontsize=26)
+ax[0].set_ylabel("Radius [R$_{\odot}$]", fontsize=30)
+ax[0].set_xlabel("Time [yr]", fontsize=30)
 
 ax[0].set_xlim(1.0e6,time[-1])
 ax[0].set_xscale("log")
@@ -76,8 +76,8 @@ ax[1].scatter(time[ind_k], k[ind_k,3], s=75, color="C1")
 ax[1].scatter(time[ind_m], m[ind_m,3], s=75, color="C2")
 
 # Format
-ax[1].set_ylabel("Radius of Gyration")
-ax[1].set_xlabel("$t - t_0$ [yr]")
+ax[1].set_ylabel("Radius of Gyration", fontsize=30)
+ax[1].set_xlabel("Time [yr]", fontsize=30)
 ax[1].set_xlim(1.0e6,time[-1])
 ax[1].set_xscale("log")
 
@@ -104,13 +104,13 @@ ax[2].plot([100], [100], lw=2.5, ls="-", color="grey", label="Matt et al. (2015)
 ax[2].plot([100], [100], lw=2.5, ls="--", color="grey", label="Reiners \& Mohanty (2012)")
 
 # Format
-ax[2].set_ylabel("Rotation Period [d]")
-ax[2].set_xlabel("$t - t_0$ [yr]")
+ax[2].set_ylabel("Rotation Period [d]", fontsize=30)
+ax[2].set_xlabel("Time [yr]", fontsize=30)
 ax[2].set_xlim(1.0e6,time[-1])
 ax[2].set_ylim(0.5e-1,1.0e2)
 ax[2].set_xscale("log")
 ax[2].set_yscale("log")
-ax[2].legend(loc="upper left", framealpha=0, fontsize=22)
+ax[2].legend(loc="upper left", framealpha=0, fontsize=24)
 
 # Save!
 fig.tight_layout()

### 2018-12-28 Analysis/subsync.py

@@ -82,12 +82,15 @@ ax0.set_xlim(0,10)
 ax0.set_ylim(0.6, 1.25)
 ax0.set_xlabel("P$_{orb}$ [d]", fontsize=30)
 ax0.set_ylabel("P$_{orb}$ / P$_{rot}$", fontsize=30)
-leg = ax0.legend(loc="lower left", framealpha=0.0, fontsize=22)
+leg = ax0.legend(loc="lower left", framealpha=0.5, fontsize=22)
 leg.legendHandles[0]._sizes = [100]
 leg.legendHandles[0].set_color('k')
 if plotLurie:
     leg.legendHandles[1].set_color('k')
     leg.legendHandles[1]._sizes = [100]
+
+ax0.text(0.5, 1.2, "CTL", ha="left", va="center", size=28,
+         color="black", zorder=100)
 
 ### Colorbar ###
 cbaxes = fig.add_subplot(gs[2])

### 2018-12-28 Paper/sync.tex

@@
 At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.
-In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, reproducing the subsynchronous primary stars in eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}.
+In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}.
 Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the CPL equilibrium tidal model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
@@
 Praesepe, ][]{Matt2015,Douglas2017} and field stars \citep[e.g.
-\kepler, ][]{Reiners2012,Matt2015,vanSaders2018}.
+the \kepler field, ][]{Reiners2012,Matt2015,vanSaders2018}.
 Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as gyrochronlogy \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.
@@
 In today's data-rich era of astronomy, it is critical to develop a theoretical framework to understand the angular momentum evolution of low-mass binaries due to stellar evolution, magnetic braking, and tidal torques.
-Such a framework can enable the characterization of the nature of tidal torques in binaries and allow for the long-term modeling of the evolutionary histories of individual systems.
+Such a framework can enable the characterization of the nature of tidal torques in binaries by conditioning on the large datasets of the spin and orbital states of stellar binaries, and also allowing the long-term modeling of the evolutionary histories of individual systems.
 
@@
 We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.
-We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to fail in such systems.
+We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, causing the predictions of gyrochronlogy models to fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
-We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:protDist} and $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.
+We discuss our results in $\S$~\ref{sec:results}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.
 
@@
 We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, $R$, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
-We consider two separate models for the long-term stellar spin-down due to magnetic braking to gauge how sensitive the interaction between tides and magnetic braking is in modifying P$_{rot}$ evolution in stellar binaries.
+We consider two separate magnetic braking models to gauge how sensitive the interaction between tides and magnetic braking is in modifying P$_{rot}$ evolution in stellar binaries.
 
-The magnetic braking model presented in \citet{Matt2015} has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field and is strongly dependent on the Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
+The first magnetic braking model we use was derived by \citet{Matt2015} and has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
+This model is strongly dependent on the stellar Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
 The \citet{Matt2015} model predicts that below a certain $Ro$ for rapidly rotating stars, stellar magnetic activity saturates at a constant value, producing a magnetic braking torque that is directly proportional to the stellar rotation rate.
 The angular momentum loss for rapidly-rotating saturated stars is given by
+\begin{equation} \label{eqn:mattSat}
+\frac{dJ}{dt} = -\frac{dJ}{dt}\Bigg|_0 \chi^2 \left( \frac{\omega}{\omega_{\odot}} \right) 
+\end{equation}
+while for more slowly-rotating unsaturated stars,
 \begin{equation} \label{eqn:mattUnSat}
@@
 \end{equation}
-while for more slowly-rotating unsaturated stars,
-\begin{equation} \label{eqn:mattSat}
-\frac{dJ}{dt} = -\frac{dJ}{dt}\Bigg|_0 \chi^2 \left( \frac{\omega}{\omega_{\odot}} \right) 
-\end{equation}
 where
 \begin{equation} \label{eqn:matt0}
-\frac{dJ}{dt}\Bigg|_0 = 9.5 \times 10^{30} \mathrm{erg} \left( \frac{R}{R_{\odot}} \right)^{3.1} \left( \frac{M}{M_{\odot}} \right)^{0.5}.
+\frac{dJ}{dt}\Bigg|_0 = 9.5 \times 10^{30} \ \mathrm{erg} \ \left( \frac{R}{R_{\odot}} \right)^{3.1} \left( \frac{M}{M_{\odot}} \right)^{0.5}.
 \end{equation}
@@
 The second model we consider was developed by \citet{Reiners2012}, who modeled stellar spin-down as a function stellar magnetic field strength.
-Similar to the \citet{Matt2015} model, \citet{Reiners2012} posits that magnetic braking can be divided into a unsaturated ($\omega < \omega_{crit}$) and unsaturated ($\omega \geq \omega_{crit}$) regime as follows:
+Similar to the \citet{Matt2015} model, \citet{Reiners2012} posit that magnetic braking can be divided into a unsaturated ($\omega < \omega_{crit}$) and unsaturated ($\omega \geq \omega_{crit}$) regime as follows:
 \begin{equation} \label{eqn:reiners}
@@
 
-The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass and is $2^{nd}$ order in $e$.
+The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass.
 We use the \eqtide implementation of the CPL model in VPLanet following the derivation of \citet{FerrazMello2008}.
@@
 
+\small
 \begin{equation}\label{eqn:ctl:a}
@@
 \end{equation}
+\normalsize
 where the intermediate variables are given by 
@@
 We note that two discrete rotation states are not the only permitted ones for tidally-locked systems under the CPL formalism.
-For example, An alternate derivation of P$_{eq}$ for orbiters with rotation axes perpendicular to the orbital plane under the CPL model predicts
+For example, an alternate derivation of P$_{eq}$ for orbiters with rotation axes perpendicular to the orbital plane under the CPL model predicts
 \begin{equation} \label{eqn:cpl:eqPerCont}
@@
 \dot{a}_{coupled}^{(2)} = \frac{-\dot{J}_{mb} - 2 \omega \left( \sum_{i=1}^{2} m_i r_{g,i}^2 R_i \dot{R_i} + m_i r_{g,i} \dot{r}_{g,i} R_i^2 \right)}
-{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} \left( m_1 r_{g,1}^2 R_1^2 + m_2 r_{g,2}^2 R_2^2 \right)}.
+{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} \left( m_1 r_{g,1}^2 R_1^2 + m_2 r_{g,2}^2 R_2^2 \right)}
 \end{split}
@@
 
-We examine stellar angular momentum evolution in low-mass binaries by performing a Monte Carlo study of two sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism.
+We examine stellar angular momentum evolution in low-mass binaries by performing a Monte Carlo study of four sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism, using both magnetic braking models.
 We simulate both stars' spin evolution but mainly consider the P$_{rot}$ evolution for the primary, i.e.
@@
 In the long-run, tidal torques overpower magnetic braking, with a slightly negative P$_{rot}$ derivative, slowly driving P$_{rot}$ towards P$_{eq}$ (see the P$_{orb} = 20, 40$ d cases in Fig.~\ref{fig:eqPer}).
-This slowly shifting torque balance produces a population of subsynchronous rotators with P$_{rot} \gsim $ P$_{eq}$ that can persist for Gyrs.
+This slowly shifting torque balance produces a population of subsynchronous rotators with P$_{rot} > $ P$_{eq}$ that can persist for Gyrs.
 
@@
 The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but rather is a generic outcome of the coupling of magnetic braking and weak tidal torques.
-As demonstrated above, strong tidal torques under the CPL model tidally-lock each binary out to P$_{orb} = 60$ d.
-The CTL model, however, can only tidally-lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the other binaries into subsynchronous rotation, similar to the \citet{Matt2015} simulations.
+The CTL model can only tidally-lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the longer P$_{orb}$ binaries into subsynchronous rotation.
 Curiously, all CTL subsynchronous rotators seem to converge to P$_{rot}/$P$_{eq} \approx 1.1$.
 In this state, tidal torques nearly balance magnetic braking, producing a long-lasting population of subsynchronous rotators, similar to what was observed above.
-In the right panel of Fig.~\ref{fig:eqPerReiners}, we see that the net P$_{rot}$ derivatives for these binaries oscillate about 0, but are slightly negative due to the mechanism described above.
+In the right panel of Fig.~\ref{fig:eqPerReiners}, we see that the net P$_{rot}$ derivatives for these binaries oscillate about 0, but are slightly negative due to the 3 simultaneous physical effects described above.
 This slight negative derivative is effectively negligible, however, as the torque balance will likely hold the binaries in subsynchronous rotation for 10s of Gyrs.
-We note that in other exploratory simulations with larger $\tau$, we find that the 3 part mechanism described above holds, and tides win out over magnetic braking in the long-run, driving P$_{rot}$ back towards P$_{eq}$.
+We note that in other exploratory simulations with larger $\tau$, we find that the tides do in fact evnetually win out over magnetic braking in the long-run, driving P$_{rot}$ back towards P$_{eq}$.
 
@@
 
-In Fig.~\ref{fig:eqPerShortPorb}, we examine this effect in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters.
-Subsynchronous rotation occurs in the general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
+In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters.
+Subsynchronous rotation occurs in general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
 Previous theoretical studies have also predicted subsynchronous rotation in short P$_{orb}$ binaries arising from the balance between tidal torques and magnetic braking \citep[e.g.][]{Habets1989,Zahn1994,Keppens1997} suggesting that this behavior is not an artifact of our choice of tidal or magnetic braking models, but rather a general outcome of the competition between magnetic braking and tidal evolution in low-mass binaries.
@@
 The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to P$_{orb} \approx 100 $ d for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
-According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \approx 10$ s.
+According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \approx 3$ s.
 Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e.
@@
 
-We isolate the impact of $Q$ and $\tau$ have in shaping the spin-orbital state of a stellar binary for a given P$_{orb}$, e.g.
-whether or not a binary is synchronized, by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
+We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
 In each bin, we compute the median $Q$ and $\tau$, and the $16\%$, and $84\%$ percentiles to quantify the distribution of tidal strength parameters that are consistent with each spin-orbital state.
@@
 For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems, especially for P$_{orb} > 20$ d.
-For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ as strong tidal torques always tidally-lock the binaries into 1:1 or 3:2 spin-orbit commesurability.
-No stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 70$ d as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
+For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 70$ d as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
 Eqn.~(\ref{eqn:cpl:eqPer}).
@@
 This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.
-In this case, the more supersynchronous the rotation, the weaker the tidal torques must be, and hence the larger the $Q$ must be, as they fail to spin down stars towards P$_{orb}$.
+In this case, the more supersynchronous the rotation, the weaker the tidal torques must be, and hence the larger the $Q$ must be.
 
@@
 Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we find stellar binaries can tidally-lock over a wide range of P$_{rot}$ up to P$_{rot} =$ P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
-This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-Moreover, \citet{Abt2004} find evidence for tides modifying P$_{rot}$ in binaries with P$_{orb}$ up to 500 d, implying tidal torques matter out long P$_{orb}$.
+This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d and with the observations of \citet{Abt2004}.
 Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d, in stark contrast to the CPL model's predictions.
-Some binaries can tidally-lock or tidally-interact out to $P_{rot} \approx 80$ d under the CTL model, but not nearly as many as the CPL model, nor can the CTL model tidally-lock binaries out to P$_{orb} = 100$ d, regardless of $\tau$.
+Some binaries can tidally-lock or tidally-interact out to $P_{rot} \approx 80$ d, but not nearly as many as the CPL model, nor can the CTL model tidally-lock binaries out to P$_{orb} = 100$ d, regardless of $\tau$.
 Only $7.5\%$ of binaries tidally-lock according to the CTL model, compared with $25\%$ of CPL model binaries, as tidal torques are generally stronger in the CPL formalism.
@@
 The CTL model predicts that the vast majority tidally-locked binaries, $86\%$, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL P$_{orb} = 7.6$ d.
-The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states with only $39\%$ locking into P$_{rot} < 20$ d, and many locking out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
+The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states with only $39\%$ of locking binaries have P$_{rot} < 20$ d, while many lock out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
 The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
@@
 
-We compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking to gauge the impact of tidal torques on shaping P$_{rot}$ distributions.
+We compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking to gauge the impact of tidal torques on driving P$_{rot}$ distributions away from that of single stars.
 We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar} with initial conditions sampled from the same mass and P$_{rot}$ distributions used for the binary simulations described $\S$~\ref{sec:methods}.
@@
 In Fig.~\ref{fig:subsync}, we demonstrate this finding by comparing our simulated systems (dots) with the observed \citet{Lurie2017} EB data (diamonds), where both points are colored by $e$.
+\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, the cluster of stars those authors identified near $P_{orb}/$P$_{rot} = 0.9$, compared with $9\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
+Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$, compared with $72\%$ of our CTL population, in perfect agreement.
+In comparison, the CPL model (not pictured) predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, reproducing the synchronized \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
+
 Nearly all simulated binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in excellent agreement with the \citet{Lurie2017} observations.
@@
 
-For P$_{orb} \gsim 4$ d, our model produces a population of subsynchronous rotators, including a cluster of subsynchronously rotating stellar binaries with nearly circular orbits, again in excellent agreement with the cluster of binaries with nearly circular orbits near P$_{orb}/$P$_{rot} = 0.9$ observed by \citet{Lurie2017}.
+For P$_{orb} \gsim 4$ d, our model produces a population of subsynchronous rotators, including a cluster of subsynchronously rotating stellar binaries with nearly circular orbits.
 Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces this population.
 Our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
-
-\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, the cluster of stars those authors identified near $P_{orb}/$P$_{rot} = 0.9$, compared with $9\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
-Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$, compared with $72\%$ of our CTL population.
-In comparison, the CPL model (not pictured) predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, reproducing the synchronized \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
 
@@
 In the absence of more precise and reliable constraints in the \citet{Lurie2017} \kepler data, however, we cannot definitively assert that the CTL model is superior to the CPL formalism for modeling tidal torques in low-mass binaries.
-We discuss observational tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
 
@@
 We performed a large suite of simulations for binaries with physically-motivated initial conditions out to P$_{orb} = 100$ and across a wide range of tidal dissipation parameters to examine the competition between tidal torques and magnetic braking for controlling the stellar P$_{rot}$ evolution.
-In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have synchronized spins and circularized orbits, in good agreement with observations of \kepler EBs.
+
+In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have tidally-synchronized spins and circularized orbits, in good agreement with observations of \kepler EBs.
 We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking.
@@
 In such systems, we caution that gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails.
-We caution any application of gyrochronlogy methods to stars with P$_{rot} \lsim 20$ d or P$_{rot} \gsim 70$ should account for the possibility of stellar binarity to prevent deriving incorrect ages.
+We caution any application of gyrochronlogy methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
 

### 2018-12-28 Paper/sync.tex

@@
 Early on in stellar lifetimes, stars spin-up as they contract along the premain sequence.
-Once stars reach the main sequence, stellar radii remain mostly constant and magnetic braking saps angular momentum from low-mass stars, gradually spinning them down over time \citep{Skumanich1972}.
+Once stars reach the main sequence, stellar radii remain mostly constant while magnetic braking removes angular momentum from the stars, gradually spinning them down over time \citep{Skumanich1972}.
 Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \citep[e.g.
@@
 Tidal torques drive binaries towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
-Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is less angular momentum in stellar spins than the binary orbit.
+Tidal-locking occurs much earlier than orbital circularization with the tidal-locking timescale estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is less angular momentum in stellar spins than the binary orbit.
 As a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
@@
 
-We simulate coupled stellar-tidal evolution for low-mass binaries using an improved version of the coupled stellar-tidal evolution model presented in \citet{Fleming2018}.
+We simulate coupled stellar-tidal evolution for low-mass binaries using an improved version of the model presented in \citet{Fleming2018}.
 We implement our model in the open-source code VPLanet\footnote{VPLanet is publicly available
@@
 	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanel.pdf}
-   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively.
+   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model.
 Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d.
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorb.pdf}
-   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively.
+   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
 Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbQHist.pdf}
-   \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
+   \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.
 In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$ and display these values.

### 2018-12-29 Analysis/stellar.py
tweaking stellar fig fontsizes

@@ -17,7 +17,7 @@ import matplotlib.pyplot as plt
 
 #Typical plot parameters that make for pretty plots
 mpl.rcParams['figure.figsize'] = (9,8)
-mpl.rcParams['font.size'] = 30.0
+mpl.rcParams['font.size'] = 23.0
 
 ## for Palatino and other serif fonts use:
 mpl.rc('font',**{'family':'serif'})
@@ -44,8 +44,7 @@ ind_k = np.argmin(np.fabs(time-1.0e8))
 ind_m = np.argmin(np.fabs(time-3.0e8))
 
 # Plot!
-ncols = 3
-fig, ax = plt.subplots(ncols=ncols, figsize=(9*ncols,8))
+fig, ax = plt.subplots(ncols=3, figsize=(20,6))
 
 # Left panel: Stellar radius evolution
 ax[0].plot(time, g[:,1], lw=2.5, color="C0", label=r"$1$ M$_{\odot}$")
@@ -58,9 +57,9 @@ ax[0].scatter(time[ind_k], k[ind_k,1], s=75, color="C1")
 ax[0].scatter(time[ind_m], m[ind_m,1], s=75, color="C2")
 
 # Format
-ax[0].legend(loc="best", framealpha=0, fontsize=26)
-ax[0].set_ylabel("Radius [R$_{\odot}$]", fontsize=30)
-ax[0].set_xlabel("Time [yr]", fontsize=30)
+ax[0].legend(loc="best", framealpha=0, fontsize=20)
+ax[0].set_ylabel("Radius [R$_{\odot}$]", fontsize=25)
+ax[0].set_xlabel("Time [yr]", fontsize=25)
 
 ax[0].set_xlim(1.0e6,time[-1])
 ax[0].set_xscale("log")
@@ -76,8 +75,8 @@ ax[1].scatter(time[ind_k], k[ind_k,3], s=75, color="C1")
 ax[1].scatter(time[ind_m], m[ind_m,3], s=75, color="C2")
 
 # Format
-ax[1].set_ylabel("Radius of Gyration", fontsize=30)
-ax[1].set_xlabel("Time [yr]", fontsize=30)
+ax[1].set_ylabel("Radius of Gyration", fontsize=25)
+ax[1].set_xlabel("Time [yr]", fontsize=25)
 ax[1].set_xlim(1.0e6,time[-1])
 ax[1].set_xscale("log")
 
[…]

### 2018-12-29 Paper/sync.tex
Updates from ShareLaTeX

@@
 
-The influence of tides is not necessarily limited to short P$_{orb}$ systems either, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered.
+The influence of tides is not necessarily limited to short P$_{orb}$ systems, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered.
 The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.
@@
 
-Understanding how tidal torques and magnetic braking impact angular momentum evolution in low-mass stellar binaries is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g.
-Praesepe, ][]{Douglas2017} and field stars \citep[e.g.
-\kepler, ][]{McQuillan2014} are likely contaminated by unresolved binaries given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys.
+Understanding the interaction between tidal torques and magnetic braking is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g.
+Praesepe, ][]{Agueros2011,Douglas2017} and field stars \citep[e.g.
+\kepler, ][]{Reinhold2013,McQuillan2014} are likely contaminated by unresolved binaries given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys.
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.
@@
 Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} are expected to detect additional low-mass eclipsing binaries, with Gaia parallaxes \citep{Gaia2016} poised to help refine these stellar parameters, potentially creating a rich dataset of the angular momentum budgets of low-mass binaries.
-In today's data-rich era of astronomy, it is critical to develop a theoretical framework to understand the angular momentum evolution of low-mass binaries due to stellar evolution, magnetic braking, and tidal torques.
-Such a framework can enable the characterization of the nature of tidal torques in binaries by conditioning on the large datasets of the spin and orbital states of stellar binaries, and also allowing the long-term modeling of the evolutionary histories of individual systems.
+Developing a framework for the angular momentum evolution of low-mass binaries can enable the characterization of the nature of tidal torques in binaries by conditioning on datasets of the spin and orbital states of stellar binaries.
 
 Here, we present a model for the angular momentum evolution of low-mass stellar binaries over their full premain and main sequence lifetimes using a realistic treatment of stellar evolution, magnetic braking, and tidal torques.
-We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.
-We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, causing the predictions of gyrochronlogy models to fail in such systems.
+We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models and two magnetic braking models.
+We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, causing the predictions of gyrochronlogy models to potentially fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanel.pdf}
+	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
    \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model.
@@
 
-In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have tidally-synchronized spins and circularized orbits, in good agreement with observations of \kepler EBs.
+In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have tidally-synchronized spins and circularized orbits, in good agreement with observations of \kepler EBs and binaries in the field.
 We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking.
-Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact out to P$_{orb} \approx 80$ d, causing their P$_{rot}$ evolution to differ from the long-term spin down due to magnetic braking experience by single stars.
-We show that our predictions are not dependant on the choice of magnetic braking model, but rather are generic outcomes of the interaction between magnetic braking and tidal torques.
-
-Many binaries with P$_{orb} \lsim 20$ d tidally-lock, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.
-When binaries tidally-lock, tides fix P$_{rot} = $ P$_{eq}$, preventing further spin-down via magnetic braking, decoupling P$_{rot}$ from age.
-In such systems, we caution that gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails.
-We caution any application of gyrochronlogy methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
+Our predictions are not strongly dependant on the choice of magnetic braking model, but rather are generic outcomes of the interaction between magnetic braking and tidal torques.
+Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact out to P$_{orb} \approx 80$ d.
+Many binaries with P$_{orb} \lsim 20$ d tidally-lock according to both models, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.
+Tidal interactions can cause P$_{rot}$ evolution in stellar binaries to differ from the long-term spin down due to magnetic braking experienced by single stars, decoupling P$_{rot}$ from age.
+In tidally-interacting binaries, gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails.
+We caution that any application of gyrochronlogy methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
 
@@
 We offer three observational tests that can distinguish between which equilibrium tidal model best describes tidal interactions in low-mass stellar binaries.
-To perform these tests, we that recommend observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d and in short P$_{orb}$ subsynchronous rotators.
+To perform these tests, we that recommend observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d.
 

### 2019-02-04 Analysis/protDist.py
fixed plot name conflict

@@ -151,7 +151,7 @@ ax3.legend(loc="best", framealpha=0.0, fontsize=17)
 plt.setp(ax3.get_yticklabels(), visible=False);
 ax3.set_ylim(0.0, 100)
 
-fig.savefig("protDist.pdf", dpi=600, bbox_inches="tight")
+fig.savefig("../Plots/protDist.pdf", dpi=600, bbox_inches="tight")
 
 ### Age histogram of rapid rotators (Prot <= 20 d) ###
 fig, ax = plt.subplots()
@@ -194,5 +194,3 @@ print()
 print("CPL - Age of Stars with Prot < 20 d:",np.percentile(cpl["Age"][cpl["Pri_ProtAge"] < 20]/1.0e9, 16))
 print("CTL - Age of Stars with Prot < 20 d:",np.percentile(ctl["Age"][ctl["Pri_ProtAge"] < 20]/1.0e9, 16))
 print("Single - Age of Stars with Prot < 20 d:",np.percentile(single["Age"][single["Pri_ProtAge"] < 20]/1.0e9, 16))
-
-fig.savefig("../Plots/protDist.pdf", bbox_inches="tight", dpi=600)

### 2019-02-04 Paper/sync.tex
Updates from ShareLaTeX

@@
 gyrochronology methods can fail unless one accounts for binarity.
-We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to disciminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
+We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
 
@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorb.pdf}
+	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
    \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.

### 2019-02-04 Analysis/protDistAgeHist.py
tweaking plot font

@@ -53,21 +53,21 @@ ax.hist(single["Age"][single["Pri_ProtAge"] <= 20]/1.0e9, bins="auto",
 # Annotate young stars
 ax.annotate(s="", xy=(2.5, 430), xytext=(0.8, 430),
             arrowprops={"arrowstyle" : "<->", "color" : "black"})
-ax.text(1.65, 450, "Young Systems", ha="center", va="center", size=18, color="k")
+ax.text(1.65, 460, "Young \n Rapid Rotators", ha="center", va="center", size=21, color="k")
 
 # Annotate locked population
 ax.annotate(s="", xy=(7, 180), xytext=(2.75, 180),
             arrowprops={"arrowstyle" : "<->", "color" : "black"})
-ax.text(4.85, 200, "Tidally-Locked Binaries", ha="center", va="center", size=18,
+ax.text(4.85, 200, "Tidally-Locked Binaries", ha="center", va="center", size=21,
         color="k")
 
 ax.set_xlabel("Age [Gyr]", fontsize=30)
 ax.set_ylabel("Counts", fontsize=30)
 ax.set_xlim(0.8,7.2)
-ax.set_ylim(0, 475)
-ax.legend(loc="best", framealpha=0.0, fontsize=20)
+ax.set_ylim(0, 490)
+ax.legend(loc="best", framealpha=0.0, fontsize=22)
 
-fig.savefig("../Plots/protAgeHist.png", dpi=200, bbox_inches="tight")
+fig.savefig("../Plots/protAgeHist.pdf", dpi=400, bbox_inches="tight")
 
 ### Age histogram of rapid rotators (Prot <= 20 d) highlighting single stars ###
 fig, ax = plt.subplots()

### 2019-02-04 Paper/sync.tex
Updates from ShareLaTeX

@@
 Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
-For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr, e.g.
-the right panel of Fig.~\ref{fig:protDist} or Fig.~\ref{fig:protDistReiners}.
-If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail.
-This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal.
+For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr.
+We highlight this point in Fig.~\ref{fig:ageHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
+If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs.
+This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), e.g.
+Fig.~\ref{fig:protAgeHist}, but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, e.g.
+Fig.~\ref{fig:protDist} and Fig.~\ref{fig:protDistReiners}.
+
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
+   \caption{Histogram of system ages for stars with P$_{rot} < 20$ d evolved using the \citet{Matt2015} magnetic braking model.}%
+    \label{fig:protAgeHist}%
+\end{figure}
 

### 2019-02-06 Analysis/analyticTorqueBalance.py
adding analytic torque balance calculation

@@ -1,0 +1,91 @@
+"""
+
+@author: David P. Fleming, University of Washington, Feb. 2019
+@email: dflemin3 (at) uw (dot) edu
+
+Analytic torque balance for CTL dw/dt and Matt+2015 magnetic braking.
+
+"""
+
+import numpy as np
+import os
+import pandas as pd
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from scipy.optimize import fsolve
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['font.size'] = 22.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+DAYSEC = 86400.0 # Number of seconds in a day
+
+def ctlMatt2015Obj(Prot, Porb, k2Tau):
+    """
+    CTL, Matt+2015 torque balance
+    """
+
+    return (Prot**3 / Porb**5) * (1.0 - Porb/Prot) - 6.17e-16/k2Tau
+# end function
+
+
+def ctlMatt2015(Porb, logK2Tau):
+    """
+    Solve for the Prot such that dw/dt_tides + dw/dt_mg = 0
+
+    Parameters
+    ----------
+    Porb : float
+        Orbital period [d]
+    logK2Tau : float
+        log10(tidal time lag in s multipled by the Love number of degree 2, k2)
+
+    Returns
+    -------
+    Prot : float
+        rotation period [d] such that
+    """
+
+    # Convert Porb from [d] to [s]
+    Pb = Porb*DAYSEC
+
+    # Unlog k2Tau
+    k2Tau = 10**logK2Tau
+
+    # Solve assuming Porb as the initial guess
+    Pr = fsolve(ctlMatt2015Obj, Pb, args=(Pb, k2Tau))[0]
+
+    # Return Prot in [d]
+    return Pr/DAYSEC
+# end function
+
+nbins = 10
+logK2Taus = np.log10(0.5 * np.logspace(-1.8, 0.3, nbins))[::-1]
+Porbs = np.linspace(20, 80, nbins)
+res = np.zeros((nbins, nbins)) # = Prot/Peq
+
+for ii in range(nbins):
+    for jj in range(nbins):
+        # Since e = 0, psi = 0, Peq = Porb in CTL model
[…]

### 2019-02-06 Analysis/lurie.py

@@ -94,7 +94,7 @@ cbaxes = fig.add_subplot(gs[4])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Eccentricity")
 
-fig.savefig("../Plots/lurieFig6.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/lurieFig6.pdf", bbox_inches="tight", dpi=200)
 
 ### 2nd Fig modeled after Fig. 7 from Lurie+2017 ###
 
@@ -106,20 +106,20 @@ gs = GridSpec(1, 5, width_ratios=[1, 0.05, 1, 0.01, 0.075], wspace=0.05)
 # Plot simulated data
 ax0 = fig.add_subplot(gs[0])
 im = ax0.scatter(cpl["Age_Porb"], cpl["Age_Porb"]/cpl["Pri_ProtAge"],
-                 c=cpl["Age_Ecc"].values, cmap="viridis", zorder=1,
+                 c=cpl["Age_Ecc"].values, cmap="viridis", zorder=3, alpha=0.5,
                  s=40, marker="o", vmin=0, vmax=0.3, label="Simulated")
 
 ax0.axhline(1, lw=3, color="black", ls="--", zorder=100)
 
 if plotLurie:
     # Plot Lurie+2017 data
-    ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=100, zorder=3,
+    ax0.scatter(lurie["Porb"], lurie["Porb"]/lurie["Prot"], color="red", s=150, zorder=1,
                 marker="+", vmin=0, vmax=0.3, label="Lurie et al. (2017)")
 
 # Format
 ax0.set_rasterization_zorder(0)
-ax0.set_xlim(0,50)
-ax0.set_ylim(0.0, 3)
+ax0.set_xlim(0, 30)
+ax0.set_ylim(0.25, 2)
 ax0.set_xlabel("P$_{orb}$ [d]", fontsize=30)
 ax0.set_ylabel("P$_{orb}$ / P$_{rot}$", fontsize=30)
 ax0.set_title("CPL")
@@ -133,20 +133,20 @@ if plotLurie:
 # Plot simulated data
 ax1 = fig.add_subplot(gs[2])
 im = ax1.scatter(ctl["Age_Porb"], ctl["Age_Porb"]/ctl["Pri_ProtAge"],
[…]

### 2019-02-06 Paper/sync.tex

@@
 
-We examine stellar angular momentum evolution in low-mass binaries by performing a Monte Carlo study of four sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism, using both magnetic braking models.
+We examine stellar angular momentum evolution in low-mass binaries by simulating four sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism, using both magnetic braking models.
 We simulate both stars' spin evolution but mainly consider the P$_{rot}$ evolution for the primary, i.e.
@@
 We compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010,Moe2018}.
-Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$] and uniformly sampling the initial P$_{orb}$ over [$3,100$] d.
+Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$], consistent with eccentricities of field binaries that likely have not been tidally-processed \citep{Raghavan2010}.
 Although the CTL model is applicable for $e > 0.3$, the CPL model is not, so we restrict $e \leq 0.3$ to allow us to compare both models.
-We do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.
+We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.
 
@@
 P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
-First, as P$_{rot}$ continue to slow due to magnetic braking, magnetic braking itself weakens as its torque scales as $\omega$ or $\omega^3$ \citep{Kawaler1988}, depending on if the stars are in the saturated or unsaturated regime, respectively \citep{Matt2015}.
-Second, as P$_{rot}$ slows down further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example).
+First, as P$_{rot}$ continue to increase due to magnetic braking, magnetic braking itself weakens as its torque scales as $\omega$ or $\omega^3$ \citep{Kawaler1988}, depending on if the stars are in the saturated or unsaturated regime, respectively \citep{Matt2015}.
+Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example).
 Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$ and strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.
@@
 For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr.
-We highlight this point in Fig.~\ref{fig:ageHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
+We highlight this point in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
 If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs.
@@
 
+\subsection{Inferring Binary Tidal Properties from \kepler} \label{sec:inference}
+
+XXX
+
 %% SECTION: DISCUSSION %%

### 2019-02-06 Paper/sync.tex

@@
 At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.
-In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}.
+In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, qualitatively reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}.
 Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the CPL equilibrium tidal model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
@@
 Tidal torques drive binaries towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
-Tidal-locking occurs much earlier than orbital circularization with the tidal-locking timescale estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is less angular momentum in stellar spins than the binary orbit.
+Tidal-locking occurs much earlier than orbital circularization with the tidal-locking timescale estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is typically much less angular momentum in stellar spins than the binary orbit.
 As a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.
@@
 In low-mass binaries, both magnetic braking and tidal torques compete to shape the stellar P$_{rot}$ evolution.
-When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$.
+When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$, or more generally P$_{rot} =$ P$_{eq}$ for eccentric orbits.
 In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2017}.
 Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[e.g.
-P$_{rot} >$ P$_{orb}$, ][]{Habets1989,Zahn1994,Keppens1997}.
+P$_{rot} >$ P$_{eq}$, ][]{Habets1989,Zahn1994,Keppens1997}.
 This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations.
@@
 Under these assumptions, the tidal evolution is analogous to a driven, damped harmonic oscillator \citep{Greenberg2009}.
-For low-mass stars, equilibrium tidal models assume that tidal forces dissipate energy in the outer-convective regions via viscous turbulence \citep[see][]{Zahn2008}.
+For low-mass stars, equilibrium tidal models assume that tidal forces primarily dissipate energy in the outer-convective regions via viscous turbulence \citep[see][]{Zahn2008}.
 Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citep[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
@@
 $k_2 \tau$ scaling in Eq.~(\ref{eqn:ctl:z}), so we instead examine how our results scale with $Q$ and $\tau$.
-Any constraints we derive for $Q$ and $\tau$ can trivially be scaled to other values of $k_2$.
+Any constraints we derive as a function $Q$ or $\tau$ can trivially be scaled to other values of $k_2$.
 
@@
 
-Tidal torques always drive a body towards the tidally-locked state.
+Tidal torques drive a body's rotation rate towards the tidally-locked state.
 When a body tidally locks, tidal torques fix P$_{rot}$ to the equilibrium P$_{rot}$, P$_{eq}$.
@@
 We follow \citet{Barnes2013} and \citet{Fleming2018} and fix P$_{rot} = $ P$_{eq}$ according to Eq.~(\ref{eqn:cpl:eqPer}) or Eq.~(\ref{eqn:ctl:eqPer}) for the CPL and CTL models, respectively, when P$_{rot}$ is within $1\%$ of P$_{eq}$.
-To ensure that tidal torques dominate over torques due to magnetic braking and stellar evolution, we additionally require that the P$_{rot}$ derivative points towards $P_{eq}$ on both sides of P$_{eq}$, i.e.
+To ensure that tidal torques dominate over torques due to magnetic braking and stellar evolution when forcing tidal-locking, we additionally require that the P$_{rot}$ derivative points towards $P_{eq}$ on both sides of P$_{eq}$, i.e.
 when the gradient of P$_{rot}$ points towards the tidally-locked state, before fixing P$_{rot} = $ P$_{eq}$.
@@
 Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$], consistent with eccentricities of field binaries that likely have not been tidally-processed \citep{Raghavan2010}.
-Although the CTL model is applicable for $e > 0.3$, the CPL model is not, so we restrict $e \leq 0.3$ to allow us to compare both models.
+Although the CTL model is applicable for $e \gsim 0.3$, the CPL model is not, so we restrict $e \leq 0.3$ to allow us to compare both models.
 We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.
@@
 
-\subsubsection{Using the \citet{Matt2015} Magnetic Braking Model}
+\subsubsection{Analytic Torque Balance}
+
+In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation.
+Since magnetic braking scales as P$_{rot}^{-3}$ for unsaturated rotators, e.g.
+Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as stellar rotation slows, so at some P$_{rot}$, tidal torques that try to spin-up sunsynchronous rotators back towards P$_{eq}$ will balance the spin-down torque of magnetic braking, producing a long-lasting state of subsynchronous rotation.
+Given an initial binary system configuration, we can analytically determine at which P$_{rot}$ torques due to magnetic braking and tides balance.
+
+We derive where this balance occurs by considering tidal torques under the CTL formalism and magnetic braking under the \citet{Matt2015} model to match the simulations performed in $\S$~\ref{sec:mattBalance}.
+We assume torque balance by setting the sum of Eqn.~(\ref{eqn:ctl:omega}), noting that $d\omega/dt = 1/I dJ\dt$, and Eqn.~(\ref{eqn:mattUnSat}) equal to 0.
+We use the unsaturated \citet{Matt2015} magnetic braking torque since in the weak tides regime, the tidal and magnetic braking torques will only balance for slow rotators, placing the stars in the unsatured regime \citep{Matt2015}.
+For simplicity, we assume both stars are solar-mass with 0 obliquity, the binary orbit is circular, and that the torque balance occurs on the main sequence, allowing us to set the $R = 1 R_{\odot}$.
+Under these assumptions, by specifying a P$_{orb}$ and $k_2 \tau$, we can compute the P$_{rot}$ at which tidal and magnetic braking torques balance.
+
+\subsubsection{Using the \citet{Matt2015} Magnetic Braking Model} \label{sec:mattBalance}
 
@@
 This slight negative derivative is effectively negligible, however, as the torque balance will likely hold the binaries in subsynchronous rotation for 10s of Gyrs.
-We note that in other exploratory simulations with larger $\tau$, we find that the tides do in fact evnetually win out over magnetic braking in the long-run, driving P$_{rot}$ back towards P$_{eq}$.
+We note that in other exploratory simulations with larger $\tau$, we find that the tides do in fact eventually win out over magnetic braking in the long-run, driving P$_{rot}$ back towards P$_{eq}$.
 
@@
 The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to P$_{orb} \approx 100 $ d for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
-According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \approx 3$ s.
+According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \gsim 3$ s.
 Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e.
@@
 As shown in Fig.~\ref{fig:lockedProtHist}, both models predict that tidal torques can influence the P$_{rot}$ evolution out to P$_{orb} \approx 80$ d, causing a significant departure from the long-term expected spin down of single stars due to magnetic braking.
-The CTL model predicts that the vast majority tidally-locked binaries, $86\%$, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL P$_{orb} = 7.6$ d.
+The CTL model predicts that the majority tidally-locked binaries, $86\%$, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL P$_{orb} = 7.6$ d.
 The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states with only $39\%$ of locking binaries have P$_{rot} < 20$ d, while many lock out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
@@
 There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
-The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) that are nearly at or have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.
+The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.
 In stark contrast, both tidal models predict a substantial population of tidally-locked fast rotators (P$_{rot} < 20$ d).
 Under the CPL and CTL models, $18\%$ and $19\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.
-At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models.
-Tidal-locking in long P$_{orb}$ binaries with low tidal $Qs$, e.g.
-Fig.~\ref{fig:lockedCPL}, produces these slow rotators.
+At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models, and is produces by tidal-locking in long P$_{orb}$ binaries with low tidal $Qs$, e.g.
+Fig.~\ref{fig:lockedCPL}.
 
@@
 The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
+We highlight this point in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
+
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
+   \caption{Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
+Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
+    \label{fig:protAgeHist}%
+\end{figure}
 
@@
 For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr.
-We highlight this point in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
-If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs.
-This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), e.g.
-Fig.~\ref{fig:protAgeHist}, but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, e.g.
+If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs, e.g.
+Fig.~\ref{fig:protAgeHist}.
+This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, e.g.
 Fig.~\ref{fig:protDist} and Fig.~\ref{fig:protDistReiners}.
 
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
-   \caption{Histogram of system ages for stars with P$_{rot} < 20$ d evolved using the \citet{Matt2015} magnetic braking model.}%
-    \label{fig:protAgeHist}%
-\end{figure}
-
-In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another, exacerbating this issue.
-Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a significant contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are potentially subject to large systematic errors.
+In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are potentially subject to large systematic errors.
 Moreover, this problem could be more significant than the aforementioned estimate implies as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
@@
 
-Furthermore, any study of the angular momentum evolution of a stellar population will be fundamentally flawed unless one robustly accounts for binarity.
+Furthermore, any study of the angular momentum evolution of a stellar population will be flawed unless one robustly accounts for binarity.
 Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
 Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
-The application of or the calibration of magnetic braking models to a sample of stellar rotation periods will be inherently flawed unless one controls for binarity.
+We recommend that the application of or the calibration of magnetic braking models to a sample of stellar rotation periods control for binarity.
 
@@
 In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
-Both models reproduce most trends in the \citet{Lurie2017} data, namely the large cluster of synchronized rotators and the population of subsynchronous rotators, both for P$_{orb} \lsim 20$ d.
+Both models qualitatively reproduce most trends in the \citet{Lurie2017} data, namely the large cluster of synchronized rotators and the population of subsynchronous rotators, both for P$_{orb} \lsim 20$ d.
 Magnetic braking and age imparts two noticeable features in the data.
@@
 First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all tidally-locked sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.
-This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
+This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless coupling occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
 The CPL model can produce the population of subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
-The CTL model, however, seems to excel where the CPL models fails by producing both populations.
+The CTL model, however, seems to excel where the CPL models fails by qualitatively reproducing both populations.
 We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
@@
 
-Our coupled stellar-tidal model accurately reproduces the spin and orbital states of low-mass \kepler EBs when using the CTL formalism.
-In Fig.~\ref{fig:subsync}, we demonstrate this finding by comparing our simulated systems (dots) with the observed \citet{Lurie2017} EB data (diamonds), where both points are colored by $e$.
+Our coupled stellar-tidal model reproduces many of the the spin and orbital features seen in low-mass \kepler EBs when using the CTL formalism.
+In Fig.~\ref{fig:subsync}, we compare our simulated systems (dots) with the observed \citet{Lurie2017} EB data (crosses), where both points are colored by $e$.
 \citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, the cluster of stars those authors identified near $P_{orb}/$P$_{rot} = 0.9$, compared with $9\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
-Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$, compared with $72\%$ of our CTL population, in perfect agreement.
-In comparison, the CPL model (not pictured) predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, reproducing the synchronized \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
+Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$, compared with $72\%$ of our CTL population, in good agreement.
+In comparison, the CPL model (not pictured) predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, producing a population of synchronized binaries similar to what is seen in the \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/subsync.pdf}
-   \caption{Same format as Fig.~\ref{fig:lurie7}, with \citet{Lurie2017} data plotted as diamonds.
+   \caption{Same format as Fig.~\ref{fig:lurie7}, with \citet{Lurie2017} data plotted as crosses.
 All points are colored according to the binary orbital $e$.
-Our simulations are in excellent agreement with the observed orbital and rotational states of \kepler EBs.
-}%
+Our simulations reproduce many of the observed orbital and rotational features seen in the \kepler EB population.}%
     \label{fig:subsync}%
@@
 For P$_{orb} \gsim 4$ d, our model produces a population of subsynchronous rotators, including a cluster of subsynchronously rotating stellar binaries with nearly circular orbits.
-Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces this population.
+Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces subsynchronous rotators.
 Our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
-
-This excellent agreement between theory and observations validates our model and provides strong evidence that the CTL equilibrium tide formalism can accurately model the tidal evolution of low-mass stellar binaries, especially at short P$_{orb}$.
-The comparison between theory and observations, however, is limited in that the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and because \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$.
-In the absence of more precise and reliable constraints in the \citet{Lurie2017} \kepler data, however, we cannot definitively assert that the CTL model is superior to the CPL formalism for modeling tidal torques in low-mass binaries.
+Our simulations can qualitatively reproduce many features seen in the \citet{Lurie2017} \kepler data, indicating that the equilibrium tide formalism can successfully model the tidal evolution of low-mass stellar binaries, especially at short P$_{orb}$.
+The comparison between theory and observations, however, is limited to more qualitative comparisons for two reasons.
+First, the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$, leading to inaccurate $e$ determinations.
+In the absence of more accurate constraints in the \citet{Lurie2017} \kepler data, we cannot robustly compare out simulation results with observations to discriminate between which equilibrium tidal model better describes tidal torques in low-mass.
+Second, without reliable and accurate prior distributions for unconstrained quantities like $Q$, $\tau$, and the initial P$_{orb}$, it is difficult to perform a true Monte Carlo study that would allow us to directly compare our simulated distributions to the observations.
+The lack of good priors and accurate data inhibits us from conditioning our model on the observed data to infer properties such as $Q$ and $\tau$, perhaps as a function of mass, but still permits us to qualitatively compare our results with observations and demonstrate how our model can reproduce most features seen in the data.
+We discuss this point further in $\S$~\ref{sec:inference}.
 
@@
 If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model.
-
-The continuous supersynchronous rotation predicted by the CTL model seen in right panel of Fig.~\ref{fig:lurie7} seems to provide a much better match to the \citet{Lurie2017} data for both tests described above, providing evidence in favor of the CLT model.
-Moreover, it is unclear if low-mass binary stars can even lock into spin-orbit resonances, such as those discussed in \citet{Rodriguez2012}, as stellar convective envelopes lack a fixed shape \citep{Burkart2014,Lurie2017}.
+It is unclear, however, if low-mass binary stars can even lock into spin-orbit resonances, such as those discussed in \citet{Rodriguez2012}, as stellar convective envelopes lack a fixed shape \citep{Burkart2014,Lurie2017}.
 There is no obvious clustering of stars near any spin-orbit resonance, aside from 1:1 spin-orbit synchronization in the \citet{Lurie2017} \kepler EB data.
@@
 
+% The continuous supersynchronous rotation predicted by the CTL model seen in right panel of Fig.~\ref{fig:lurie7} seems to provide a much better match to the \citet{Lurie2017} data for both tests described above, providing evidence in favor of the CLT model.
+Moreover, i
+
 \subsection{Inferring Binary Tidal Properties from \kepler} \label{sec:inference}
@@
 XXX
+
+The limitations/biases in the Lurie data set are not well known, besides the obvious lack of longer Porb binaries, so any direct comparison of my simulations with the data is not robust.
+Instead, I think it would be better to focus on how my simulations can produce the features seen in the Lurie data, like the subsynchronous rotators, to show how the interaction between tidal torques and magnetic braking can produce rich dynamical interactions with varied outcomes.
 
@@
 
-We compare the predictions of both the CPL and CTL models with observations of P$_{rot}$ and P$_{orb}$ of \kepler EBs by \citet{Lurie2017} and find that both can reproduce the data, validating our approach and suggesting that equilibrium tidal models can accurately reproduce stellar-tidal evolution in low-mass stellar binaries.
-We show that the CTL model best reproduces the data, but the lack of uncertainties on P$_{rot}$ and the approximate orbital eccentricities derived by \citet{Lurie2017} prevent us from conclusively ruling in favor of the CTL model.
-We offer three observational tests that can distinguish between which equilibrium tidal model best describes tidal interactions in low-mass stellar binaries.
-To perform these tests, we that recommend observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d.
+We compare the predictions of both the CPL and CTL models with observations of P$_{rot}$ and P$_{orb}$ of \kepler EBs by \citet{Lurie2017} and find that both can qualitatively reproduce many features seen in the data, validating our approach and suggesting that equilibrium tidal models can accurately reproduce stellar-tidal evolution in low-mass stellar binaries.
+The lack of uncertainties on P$_{rot}$ and the approximate orbital eccentricities derived by \citet{Lurie2017} prevent us from discriminating between which tidal model best describes tidal torques in low-mass binaries.
+We offer three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
+To perform these tests, we that recommend observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d, although we note that these are non-trivial observations.
 
-The excellent agreement we find between our theory and \kepler observations outlines a critical point: one cannot simply observe a short P$_{orb}$ binary on a circular orbit and assume synchronization, nor can one observe a binary with P$_{orb} \gsim 20$ d and assume that tides have not shaped that system's angular momentum evolution.
-For example, stellar-tidal interactions can produce synchronous and subsynchronous rotation for short P$_{orb}$ binaries on circular orbits, e.g.
+Our theoretical predictions outline a critical point: one cannot simply observe a short P$_{orb}$ binary on a circular orbit and assume synchronization, nor can one observe a binary with P$_{orb} \gsim 20$ d and assume that tides have not shaped that system's angular momentum evolution.
+Stellar-tidal interactions can produce synchronous and subsynchronous rotation for short P$_{orb}$ binaries on circular orbits, e.g.
 Fig.~\ref{fig:subsync}, depending on the age of the system, e.g.

### 2019-02-06 Analysis/analyticTorqueBalance.py
updated analytic, lurie figures

@@ -62,7 +62,7 @@ def ctlMatt2015(Porb, logK2Tau):
     return Pr/DAYSEC
 # end function
 
-nbins = 10
+nbins = 100
 logK2Taus = np.log10(0.5 * np.logspace(-1.8, 0.3, nbins))[::-1]
 Porbs = np.linspace(20, 80, nbins)
 res = np.zeros((nbins, nbins)) # = Prot/Peq

### 2019-02-06 Analysis/lurie.py

@@ -27,9 +27,13 @@ plotLurie = True
 cpl = pd.read_csv("../Data/mcCPLDec22.csv")
 ctl = pd.read_csv("../Data/mcCTLDec22.csv")
 
-lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
-lurie["Prot"] = lurie["p_1_min"].copy() # As recommended by Lurie+2017 for equitorial Prot
-lurie["Porb"] = lurie["p_orb"].copy()
+#lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
+#lurie["Prot"] = lurie["p_1_min"].copy() # As recommended by Lurie+2017 for equitorial Prot
+#lurie["Porb"] = lurie["p_orb"].copy()
+
+# Load in cleaned sample of Prot and Ecc from Lurie+2017
+lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
+                    names=["Porb", "Prot", "Ecc"])
 
 # 1st Fig modeled after Fig. 6, upper panel from Lurie+2017
 
@@ -98,60 +102,58 @@ fig.savefig("../Plots/lurieFig6.pdf", bbox_inches="tight", dpi=200)
 
 ### 2nd Fig modeled after Fig. 7 from Lurie+2017 ###
 
-fig = plt.figure(figsize=(19, 8))
-gs = GridSpec(1, 5, width_ratios=[1, 0.05, 1, 0.01, 0.075], wspace=0.05)
+fig = plt.figure(figsize=(24, 7))
+gs = GridSpec(1, 7, width_ratios=[1, 0.05, 1, 0.05, 1, 0.01, 0.075], wspace=0.05)
 
 ### CPL Plot ###
 
 # Plot simulated data
 ax0 = fig.add_subplot(gs[0])
 im = ax0.scatter(cpl["Age_Porb"], cpl["Age_Porb"]/cpl["Pri_ProtAge"],
-                 c=cpl["Age_Ecc"].values, cmap="viridis", zorder=3, alpha=0.5,
-                 s=40, marker="o", vmin=0, vmax=0.3, label="Simulated")
-
+                 c=cpl["Age_Ecc"].values, cmap="viridis", zorder=3,
+                 s=40, marker="o", vmin=0, vmax=0.3)
 ax0.axhline(1, lw=3, color="black", ls="--", zorder=100)
-
[…]

### 2019-02-06 Paper/sync.tex

@@
 
-\subsection{Interaction Between Magnetic and Tidal Braking} \label{sec:eq}
+\subsection{Interaction Between Magnetic and Tidal Braking: Subsynchronous Rotation} \label{sec:eq}
 
-Here we examine the P$_{rot}$ evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a focus on systems with weak tidal torques, i.e.
-long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking.
+Here we focus on binaries in weak tides regime, i.e.
+long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking via analytic calculations and simulations.
 
-\subsubsection{Analytic Torque Balance}
+\subsubsection{Analytic Torque Balance} \label{sec:analytic}
 
 In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation.
+In this case when P$_{rot} >$ P$_{eq}$ for long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated magnetic braking regime \citep{Matt2015}.
 Since magnetic braking scales as P$_{rot}^{-3}$ for unsaturated rotators, e.g.
-Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as stellar rotation slows, so at some P$_{rot}$, tidal torques that try to spin-up sunsynchronous rotators back towards P$_{eq}$ will balance the spin-down torque of magnetic braking, producing a long-lasting state of subsynchronous rotation.
-Given an initial binary system configuration, we can analytically determine at which P$_{rot}$ torques due to magnetic braking and tides balance.
+Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as stellar rotation slows, so at some P$_{rot}$, tidal torques that try to spin-up subsynchronous rotators back towards P$_{eq}$ will balance the spin-down torque of magnetic braking, producing a long-lasting equilibrium state of subsynchronous rotation.
 
-We derive where this balance occurs by considering tidal torques under the CTL formalism and magnetic braking under the \citet{Matt2015} model to match the simulations performed in $\S$~\ref{sec:mattBalance}.
-We assume torque balance by setting the sum of Eqn.~(\ref{eqn:ctl:omega}), noting that $d\omega/dt = 1/I dJ\dt$, and Eqn.~(\ref{eqn:mattUnSat}) equal to 0.
-We use the unsaturated \citet{Matt2015} magnetic braking torque since in the weak tides regime, the tidal and magnetic braking torques will only balance for slow rotators, placing the stars in the unsatured regime \citep{Matt2015}.
-For simplicity, we assume both stars are solar-mass with 0 obliquity, the binary orbit is circular, and that the torque balance occurs on the main sequence, allowing us to set the $R = 1 R_{\odot}$.
-Under these assumptions, by specifying a P$_{orb}$ and $k_2 \tau$, we can compute the P$_{rot}$ at which tidal and magnetic braking torques balance.
+We compute the P$_{rot}$ at which this balance occurs by considering tidal torques under the CTL formalism and magnetic braking under the \citet{Matt2015} model.
+We assume torque balance by setting the sum of Eqn.~(\ref{eqn:ctl:omega}) and Eqn.~(\ref{eqn:mattUnSat}) equal to 0, noting that $d\omega/dt = (1/\mathrm{I}) dJ/dt$ for fixed moment of inertia, I.
+For simplicity, we assume both stars are solar-mass with 0 obliquity, the binary orbit is circular, and that the torque balance occurs on the main sequence where stellar properties change slowly, allowing us to set $R = 1 R_{\odot}$ and assume constant moments of inertia.
+Given these assumptions, we can specify P$_{orb}$, $k_2$, and $\tau$ to compute the equilibrium P$_{rot}$ at which tidal and magnetic braking torques balance.
+We display the result of this calculation in Fig.~\ref{fig:analyticBalance} and normalize P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$ (see Eqn.~(\ref{eqn:ctl:eqPer})).
 
-\subsubsection{Using the \citet{Matt2015} Magnetic Braking Model} \label{sec:mattBalance}
+\begin{figure}[h]
+	\includegraphics[width=0.5\textwidth]{../Plots/analyticTorque.pdf}
+   \caption{The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit computed according to the analytic calculation described in $\S$~\ref{sec:analytic}.
+The white dashed line indicates our fiducial values for $k_2$ and $\tau$, 0.5 and 0.1 s, respectively, that we adopt in the simulations in $\S$~\ref{sec:mattBalance}.
+}%
+    \label{fig:analyticBalance}%
+\end{figure}
 
-In Fig.~\ref{fig:eqPer}, we plot the evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) formalisms.
+In general as tides weaken, i.e.
+increasing P$_{orb}$ or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where magnetic braking torques are weaker, or equivalently at larger P$_{rot}/$P$_{eq}$.
+Conversely, moderate to strong tidal torques, $k_2 \tau \lsim -1$ or P$_{orb} \lsim 20$ d, overpower magnetic braking for most orbital periods, tidally-locking the binaries into synchronous rotation.
+Given this calculation we expect that for many orbital periods and tidal time lags, stellar binaries will reach a state of subsynchronous rotation.
+For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), we predict that solar-twin binaries on low eccentricity orbits will reach the subsynchronous state with P$_{rot}/$P$_{eq} \approx 1.1-1.4$.
+This simple analytic calculation, however, does not account for stellar evolution or secular orbital evolution due to tides, e.g.
+tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we cannot estimate when the binaries reach the subsynchronous state nor how long it will last.
+
+Below, we verify the results of our analytic calculations by examining how stellar binaries evolve towards subsynchronous rotation, and under what conditions it occurs in the weak tides regime.
+We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a combination of both tidal models and both magnetic braking models.
+
+\subsubsection{Torque Balance Using the \citet{Matt2015} Magnetic Braking Model} \label{sec:mattBalance}
+
+In Fig.~\ref{fig:eqPer}, we plot the simulated evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models.
 Both models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
@@
 The CTL model, however, predicts that for P$_{orb} \gsim 10$ d, magnetic braking overpowers weak tidal torques, spinning down the stars past the tidally-locked state.
-Magnetic braking pushes P$_{rot}/$P$_{eq} > 1$ with the maximum value set by the torque balance between tidal torques and magnetic braking.
-The maximum values grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
-Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution.
+Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.1$ with the maximum value set by the torque balance between tidal torques and magnetic braking, in good agreement with the analytic calculations in $\S$~\ref{sec:analytic}.
+As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
+Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution, driving the binaries towards subsynchronous rotation.
 
-P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
-First, as P$_{rot}$ continue to increase due to magnetic braking, magnetic braking itself weakens as its torque scales as $\omega$ or $\omega^3$ \citep{Kawaler1988}, depending on if the stars are in the saturated or unsaturated regime, respectively \citep{Matt2015}.
+Subsynchronous rotation does not persist indefinitely, however, as P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
+First, as P$_{rot}$ continue to increase due to magnetic braking, magnetic braking itself weakens as its torque scales as P$_{rot}^{-3}$ for unsaturated rotators \citep{Matt2015}.
 Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example).
@@
 
-We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to stellar evolution, magnetic braking, and tidal torques.
+We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to stellar evolution, magnetic braking, and tidal torques as a function of time in our simulations.
 Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 60 Myr when the stars reach the main sequence.
 After $R$ contraction, $\dot{\mathrm{P}}_{rot} > 0$ as tides and magnetic braking combine to spin down stars towards the tidally-locked state.
-For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$, but $\ddot{\mathrm{P}}_{rot} < 0$ as the three processes described above gradually strengthen tidal torques relative to magnetic braking.
-In the long-run, tidal torques overpower magnetic braking, with a slightly negative P$_{rot}$ derivative, slowly driving P$_{rot}$ towards P$_{eq}$ (see the P$_{orb} = 20, 40$ d cases in Fig.~\ref{fig:eqPer}).
+For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$ driving the stars towards subsynchronous rotation.
+$\ddot{\mathrm{P}}_{rot} < 0$, however, as the three processes described above gradually strengthen tidal torques relative to magnetic braking.
+In the long-run, tidal torques overpower magnetic braking, with a slightly negative P$_{rot}$ derivative, slowly driving P$_{rot}$ back towards P$_{eq}$ (see the P$_{orb} = 20, 40$ d cases in Fig.~\ref{fig:eqPer}).
 This slowly shifting torque balance produces a population of subsynchronous rotators with P$_{rot} > $ P$_{eq}$ that can persist for Gyrs.
 
-\subsubsection{Using the \citet{Reiners2012} Magnetic Braking Model}
+\subsubsection{Torque Balance Using the \citet{Reiners2012} Magnetic Braking Model}
 
@@
 Short P$_{orb}$ subsynchronous binaries exist in nature, such as many \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012}, EPIC 219394517 \citep{Torres2018}, and in ``Binary 6211" observed by \citet{Meibom2006}, suggesting that this theoretical observation is real and borne out in nature.
-Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries.
-We explore this effect further and test it against observations of \kepler EBs in $\S$~\ref{sec:subsync}.
+Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries and sunsynchronous rotation should be expected in many tidally-interacting binaries.
+We explore these effect further and compare our theory to observations of \kepler EBs in $\S$~\ref{sec:subsync}.
 
@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbQHist.pdf}
+	\includegraphics[width=0.5\textwidth]{../Plots/porbProtPorbQHist.pdf}
    \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.
@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/porbProtPorbTauHist.pdf}
+	\includegraphics[width=0.5\textwidth]{../Plots/porbProtPorbTauHist.pdf}
    \caption{Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
@@
 
-\subsection{P$_{rot}$ Distribution of Stellar Binaries} \label{sec:protDist}
+\subsection{P$_{rot}$ Distribution of a Synthetic Population of Stellar Binaries} \label{sec:protDist}
 
@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCPL.pdf}
+	\includegraphics[width=0.5\textwidth]{../Plots/lockedCPL.pdf}
    \caption{Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedCTL.pdf}
+	\includegraphics[width=0.5\textwidth]{../Plots/lockedCTL.pdf}
    \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/lockedProtHist.pdf}
+	\includegraphics[width=0.5\textwidth]{../Plots/lockedProtHist.pdf}
    \caption{P$_{rot}$ distribution for tidally-locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.}%

### 2019-02-06 Paper/sync.tex

@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/subsync.pdf}
+	\includegraphics[width=0.5\textwidth]{../Plots/subsync.pdf}
    \caption{Same format as Fig.~\ref{fig:lurie7}, with \citet{Lurie2017} data plotted as crosses.
@@
 
-XXX
+In principle, we could have conditioned our model on the \citet{Lurie2017} data to infer population-level properties of stellar binaries, such as tidal $Q$ and $\tau$ values for stars, perhaps as a function of stellar mass, or the initial binary P$_{orb}$ distribution, using hierarchical probabilistic modeling instead of qualitatively comparing of our simulations with the data.
+Such an analysis is beyond the scope of this work because the biases in the \citet{Lurie2017} data are not well known.
+For example, the bias of the obvious lack of longer P$_{orb}$ binaries in the \citet{Lurie2017} data arises from the fact that it is difficult to detect binaries with longer P$_{orb}$ in a finite time window via the transit method as longer P$_{orb}$ systems transit fewer times.
+What is not known, however, is how this bias impacts the completeness of binary detections as a function of P$_{orb}$, nor is the completeness known as a function of stellar P$_{rot}$.
+These biases, combined with the lack of P$_{rot}$ uncertainties, inaccurate binary eccentricities, and lack of binary masses and ages makes it difficult to directly compare our simulations with the data to robustly infer properties of low-mass binaries.
 
-The limitations/biases in the Lurie data set are not well known, besides the obvious lack of longer Porb binaries, so any direct comparison of my simulations with the data is not robust.
-Instead, I think it would be better to focus on how my simulations can produce the features seen in the Lurie data, like the subsynchronous rotators, to show how the interaction between tidal torques and magnetic braking can produce rich dynamical interactions with varied outcomes.
+Instead, one could use our model to infer tidal properties for individual systems with well-constrained properties.
+For example, Kepler-47 \citep{Orosz2012} is a low-mass binary whose primary star rotates subsynchronously.
+This system has been extensively monitored via photometry and spectroscopy because it harbors multiple transiting circumbinary planets, and as a result has well-constrained properties.
+One could infer the tidal properties for this binary, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties.
+This analysis, however, is beyond the scope of this work and we leave it for future endeavors.
 

### 2019-02-27 Analysis/analyticTorqueBalance.py
tweaked a bunch of plots

@@ -62,7 +62,7 @@ def ctlMatt2015(Porb, logK2Tau):
     return Pr/DAYSEC
 # end function
 
-nbins = 100
+nbins = 20
 logK2Taus = np.log10(0.5 * np.logspace(-1.8, 0.3, nbins))[::-1]
 Porbs = np.linspace(20, 80, nbins)
 res = np.zeros((nbins, nbins)) # = Prot/Peq

### 2019-02-27 Analysis/eqPerMatt.py

@@ -96,7 +96,7 @@ axes[0].axhline(1, lw=2.5, ls=":", color="k")
 axes[0].set_xlim(1.0e6, cpl[-1,0])
 axes[0].set_ylim(0, 1.2)
 
-axes[0].text(1.2e6, 1.12, "Matt et al. (2015)",
+axes[0].text(1.2e6, 1.12, "Matt et al. (2015) + tides",
          ha="left", va="center", size=18, color="black", zorder=100)
 
 # Format right axis

### 2019-02-27 Analysis/eqPerReiners.py

@@ -96,7 +96,7 @@ axes[0].axhline(1, lw=2.5, ls=":", color="k")
 axes[0].set_xlim(1.0e6, cpl[-1,0])
 axes[0].set_ylim(0, 1.2)
 
-axes[0].text(1.05e6, 1.14, "Reiners \& Mohanty (2012)",
+axes[0].text(1.05e6, 1.12, "Reiners \& Mohanty (2012) \n \t + tides",
          ha="left", va="center", size=17, color="black", zorder=100)
 
 # Format right axis

### 2019-02-27 Analysis/example.py

@@ -38,7 +38,7 @@ mpl.rcParams['figure.figsize'] = (9,8)
 mpl.rcParams['font.size'] = 22.0
 
 ## for Palatino and other serif fonts use:
-mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
+mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Where simulation results live

### 2019-02-27 Paper/sync.tex
Updates from Overleaf

@@
 
-Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries.
-Here we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models.
+We examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries up to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models.
 We find that many binaries with P$_{orb} \lsim 20$ d tidally-lock, and most with $P_{orb} \lsim 4$ d tidally-lock into synchronous rotation on circularized orbits.
@@
 The long-term angular momentum evolution of low-mass (M$\lsim 1$ M$_{\odot}$) stars is controlled by magnetic braking, the torque exerted on stars due to the coupling of stellar winds to the surface magnetic field \citep{Dunn1961,Mestel1968}.
-Early on in stellar lifetimes, stars spin-up as they contract along the premain sequence.
+Early in stellar lifetimes, stars spin-up as they contract along the pre-main sequence.
 Once stars reach the main sequence, stellar radii remain mostly constant while magnetic braking removes angular momentum from the stars, gradually spinning them down over time \citep{Skumanich1972}.
@@
 Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}.
-Orbital circularization is ubiquitous for short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989,Claret1995} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
-For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} finds that the orbit of solar twin binaries circularizes during the stellar pre-main sequence.
+Orbital circularization is ubiquitous for short-period binaries, owing to the tidal torque's strong radius and semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989,Claret1995} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized.
+For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} found that the orbit of solar twin binaries circularize during the stellar pre-main sequence.
 Observations by \citet{Meibom2005} support this picture as they find short-period binaries in the ${\sim}150$ Myr old cluster M35 tend to have circular orbits.
@@
 
-%The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
+%The large sample of $816$ P$_{rot}$ measurements derived using star spot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
 Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking.
@@
 We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models and two magnetic braking models.
-We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, causing the predictions of gyrochronlogy models to potentially fail in such systems.
+We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to potentially fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
@@
 Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution, with the \citet{Reiners2012} predicting marginally stronger torques.
-The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
+The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains fully convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
 
@@
 
- Equilibrium tidal models, first introduced by \citep{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega$, and obliquities $\psi$, of both gravitating bodies due to tidal torques.
+ Equilibrium tidal models, first introduced by \citet{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega_i$, and obliquities $\psi_i$, of both gravitating bodies due to tidal torques.
 Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the bodies' centers of mass due to friction within each body.
@@
 For low-mass stars, equilibrium tidal models assume that tidal forces primarily dissipate energy in the outer-convective regions via viscous turbulence \citep[see][]{Zahn2008}.
-Although simple, equilibrium tidal models have been used to robustly model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citep[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
+Although simple, equilibrium tidal models have been used to model the secular orbital and rotation evolution of both Solar System bodies and exoplanets \citep[e.g.][]{Goldreich1966,Jackson2009,Leconte2010,Heller2011,Barnes2013,Barnes2017} and stellar binaries \citep[e.g.][]{Zahn1989,Zahn2008,Khaliullin2011,Repetto2014,Fleming2018}.
 We refer the reader to \citet{Barnes2017} for an in-depth discussion of the assumptions and limitations of equilibrium tidal models.
@@
 
-The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with their own associated frequency, that maintain a fixed phase with respect to the line connecting the two stars' centers of mass.
+The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with its own associated frequency, that maintain a fixed phase offset with respect to the line connecting the two stars' centers of mass.
 We use the \eqtide implementation of the CPL model in VPLanet following the derivation of \citet{FerrazMello2008}.
@@
 However if the assumption of linearity is relaxed, i.e.
-frequencies associated with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}, hence we do not make this assumption and hold the tidal time lag fixed.
+frequencies associated with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}.
 We use the \eqtide implementation of the CTL model in VPLanet following the derivation of \citet{Leconte2010}.
@@
 Since magnetic braking scales as P$_{rot}^{-3}$ for unsaturated rotators, e.g.
-Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as stellar rotation slows, so at some P$_{rot}$, tidal torques that try to spin-up subsynchronous rotators back towards P$_{eq}$ will balance the spin-down torque of magnetic braking, producing a long-lasting equilibrium state of subsynchronous rotation.
+Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as the stellar rotation slows down, so at some P$_{rot}$, tidal torques that try to spin-up subsynchronous rotators back towards P$_{eq}$ will balance the spin-down torque of magnetic braking, producing a long-lasting state of subsynchronous rotation.
 
-We compute the P$_{rot}$ at which this balance occurs by considering tidal torques under the CTL formalism and magnetic braking under the \citet{Matt2015} model.
-We assume torque balance by setting the sum of Eqn.~(\ref{eqn:ctl:omega}) and Eqn.~(\ref{eqn:mattUnSat}) equal to 0, noting that $d\omega/dt = (1/\mathrm{I}) dJ/dt$ for fixed moment of inertia, I.
-For simplicity, we assume both stars are solar-mass with 0 obliquity, the binary orbit is circular, and that the torque balance occurs on the main sequence where stellar properties change slowly, allowing us to set $R = 1 R_{\odot}$ and assume constant moments of inertia.
+We compute the P$_{rot}$ at which this balance occurs by by setting the sum of Eqn.~(\ref{eqn:ctl:omega}) and Eqn.~(\ref{eqn:mattUnSat}) equal to 0, noting that $d\omega/dt = (1/\mathrm{I}) dJ/dt$ for fixed moment of inertia, I.
+In this calculation, we consider tidal torques under the CTL formalism and magnetic braking under the \citet{Matt2015} model.
+For simplicity, we assume both stars are solar-mass with 0 obliquity, a circular binary orbit, and that the torque balance occurs while the stars are on the main sequence where stellar properties change slowly, allowing us to set $R = 1 R_{\odot}$ and assume constant moments of inertia.
+We find below in simulations that these assumptions are well-justified.
 Given these assumptions, we can specify P$_{orb}$, $k_2$, and $\tau$ to compute the equilibrium P$_{rot}$ at which tidal and magnetic braking torques balance.
-We display the result of this calculation in Fig.~\ref{fig:analyticBalance} and normalize P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$ (see Eqn.~(\ref{eqn:ctl:eqPer})).
+We derive this balance, given by Eqn.~\ref{eqn:appendixBalance}, in $\S$~\ref{sec:appendix:balance} and visualize it in Fig.~\ref{fig:analyticBalance}.
+We normalize P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$ (see Eqn.~(\ref{eqn:ctl:eqPer})).
 
@@
 	\includegraphics[width=0.5\textwidth]{../Plots/analyticTorque.pdf}
-   \caption{The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit computed according to the analytic calculation described in $\S$~\ref{sec:analytic}.
-The white dashed line indicates our fiducial values for $k_2$ and $\tau$, 0.5 and 0.1 s, respectively, that we adopt in the simulations in $\S$~\ref{sec:mattBalance}.
-}%
+   \caption{The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}.
+The white dashed line indicates our fiducial values for $k_2$ and $\tau$, 0.5 and 0.1 s, respectively, that we adopt in the simulations in $\S$~\ref{sec:mattBalance}.}%
     \label{fig:analyticBalance}%
@@
 In general as tides weaken, i.e.
-increasing P$_{orb}$ or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where magnetic braking torques are weaker, or equivalently at larger P$_{rot}/$P$_{eq}$.
-Conversely, moderate to strong tidal torques, $k_2 \tau \lsim -1$ or P$_{orb} \lsim 20$ d, overpower magnetic braking for most orbital periods, tidally-locking the binaries into synchronous rotation.
-Given this calculation we expect that for many orbital periods and tidal time lags, stellar binaries will reach a state of subsynchronous rotation.
-For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), we predict that solar-twin binaries on low eccentricity orbits will reach the subsynchronous state with P$_{rot}/$P$_{eq} \approx 1.1-1.4$.
+increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where magnetic braking torques are weaker, or equivalently at larger P$_{rot}/$P$_{eq}$.
+Conversely, in the strong tides regime, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for most orbital periods, tidally-locking binaries into synchronous rotation.
+Given this calculation, we expect that for many orbital periods and tidal time lags, stellar binaries will reach a state of subsynchronous rotation.
+For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries on low eccentricity orbits reach the subsynchronous state with P$_{rot}/$P$_{eq} \approx 1.1-1.4$, with this value increasing with P$_{orb}$.
 This simple analytic calculation, however, does not account for stellar evolution or secular orbital evolution due to tides, e.g.
-tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we cannot estimate when the binaries reach the subsynchronous state nor how long it will last.
+tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we cannot precisely determine when the binaries reach the subsynchronous state nor how long it will last.
+For that, we turn to simulations.
 
-Below, we verify the results of our analytic calculations by examining how stellar binaries evolve towards subsynchronous rotation, and under what conditions it occurs in the weak tides regime.
+Below, we verify the results of our analytic calculations by examining how stellar binaries evolve towards subsynchronous rotation, under what conditions, and when it occurs in the weak tides regime.
 We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a combination of both tidal models and both magnetic braking models.
+In Fig.~\ref{fig:eqPer}, we plot the simulated evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models.
 
@@
 
-In Fig.~\ref{fig:eqPer}, we plot the simulated evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models.
-Both models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
+Both tidal models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
 The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.
+The CTL model, however, predicts that for P$_{orb} \gsim 10$ d, magnetic braking overpowers weak tidal torques, spinning down the stars past the tidally-locked state.
+Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.1$ with the maximum value set by the torque balance between tidal torques and magnetic braking, in good agreement with the analytic calculations in $\S$~\ref{sec:analytic}.
+As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
+Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution, driving the binaries towards subsynchronous rotation.
 
@@
 After spinning down to long P$_{rot}$, magnetic braking weakens, allowing tidal torques to sheppard P$_{rot}$ towards P$_{eq}$ as the P$_{rot}$ derivatives stay near 0, but with a slight negative value.
-
-The CTL model, however, predicts that for P$_{orb} \gsim 10$ d, magnetic braking overpowers weak tidal torques, spinning down the stars past the tidally-locked state.
-Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.1$ with the maximum value set by the torque balance between tidal torques and magnetic braking, in good agreement with the analytic calculations in $\S$~\ref{sec:analytic}.
-As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
-Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution, driving the binaries towards subsynchronous rotation.
 
@@
 Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries and sunsynchronous rotation should be expected in many tidally-interacting binaries.
-We explore these effect further and compare our theory to observations of \kepler EBs in $\S$~\ref{sec:subsync}.
+We explore these effect further and compare our theory to observations of \kepler EBs in $\S$~\ref{sec:kepler}.
 
@@
 
-\subsubsection{\citet{Matt2015} Magnetic Braking}
+\subsubsection{\citet{Matt2015} Magnetic Braking} \label{sec:mattDist}
 
@@
 Under the CPL and CTL models, $18\%$ and $19\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.
-At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models, and is produces by tidal-locking in long P$_{orb}$ binaries with low tidal $Qs$, e.g.
+At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models, and is produced by tidal-locking in long P$_{orb}$ binaries with low tidal $Qs$, e.g.
 Fig.~\ref{fig:lockedCPL}.
 
-Both tidal models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{rot} < 20 $ d, as many of these systems tidally-lock at short P$_{orb}$.
-The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
-We highlight this point in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
-
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
-   \caption{Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
-Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
-    \label{fig:protAgeHist}%
-\end{figure}
-
-\subsubsection{\citet{Reiners2012} Magnetic Braking}
+\subsubsection{\citet{Reiners2012} Magnetic Braking} \label{sec:reinersDist}
 
@@
 As shown above, the CTL model predicts that binaries can strongly tidally-interact out to long P$_{orb}$, seen out to P$_{rot} = 80$ d, while the CPL model can tidally-lock stars out to P$_{orb} \approx 100$ d.
-These strong tidal interactions decouple P$_{rot}$ from age in binaries, mirroring the behavior discussed above, especially for rapid rotators.
-The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.6^{+2.8}_{-1.3}$ Gyr and $2.7^{+2.7}_{-1.4}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.7^{+0.9}_{-0.5}$ Gyr, similar to the age distributions for rapid rotators evolved using the \citet{Matt2015} model.
+These strong tidal interactions seen both here and in $\S$~\ref{sec:mattDist} decouple P$_{rot}$ from age in binaries, especially for rapid rotators.
 
 \subsubsection{Impact of Binarity on Gyrochronology}
+
+Our models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{rot} < 20 $ d, as many of these systems tidally-lock at short P$_{orb}$.
+Under the \citet{Matt2015} model discussed in $\S$~\ref{sec:mattDist}, the median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
+Similarly for the models evolved with the \citet{Reiners2012} model discussed in $\S$~\ref{sec:reinersDist}, the median ages and $68\%$ interval for fast rotators are $2.6^{+2.8}_{-1.3}$ Gyr and $2.7^{+2.7}_{-1.4}$ Gyr according to the CPL and CTL models, respectfully, compared to single stars with ages of $1.7^{+0.9}_{-0.5}$ Gyr, We highlight this stark dichotomy between the age distribution of single and double stars in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
+
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
+   \caption{Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
+Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
+    \label{fig:protAgeHist}%
+\end{figure}
 
@@
 We compare our simulation results using both tidal models and the \citet{Matt2015} magnetic braking model to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing stellar binaries by \citet{Lurie2017}.
-\citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with starspot modulations and visually inspected each light curve to ensure their accuracy.
+The purpose of this comparison is to see if our model predictions, which by design populate a wide, but physically-plausible, region of parameter space, can reproduce features observed in the data.
+\citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with star spot modulations and visually inspected each light curve to ensure their accuracy.
 The \citet{Lurie2017} dataset is the largest homogenous set of P$_{rot}$ measurements available for low-mass stellar binaries and represents the state of the art benchmark for studies of the influence of tides on P$_{rot}$ in stellar binaries.
 We compare our results to the P$_{1,min}$ P$_{rot}$ values reported by \citet{Lurie2017} as the authors demonstrated that these values are likely to be close to the equatorial P$_{rot}$ that we track in our simulations.
+In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}, and the \citet{Lurie2017} data.
 
-In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
-Both models qualitatively reproduce most trends in the \citet{Lurie2017} data, namely the large cluster of synchronized rotators and the population of subsynchronous rotators, both for P$_{orb} \lsim 20$ d.
-Magnetic braking and age imparts two noticeable features in the data.
-First, supersynchronous rotators with P$_{orb}/$P$_{rot} > 1.5$, especially at long P$_{orb}$, corresponds to young stars that have yet to significantly spin down due to magnetic braking or tides.
-Second, the lower limit seen in both panels is a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down via magnetic braking in the weak tides regime over 7 Gyr, the longest age considered in our simulations.
-
+Both tidal models qualitatively reproduce most features seen in the \citet{Lurie2017} data, especially at short P$_{orb}$.
+Nearly all simulated binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in agreement with the \citet{Lurie2017} observations.
+At very short P$_{orb}$, in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries, in good agreement with theoretical predictions \citep{Counselman1973}.
+For P$_{orb} \gsim 4$ d, our models produce large populations of subsynchronous rotators.
+Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces subsynchronous rotators.
+For example under the CTL model, our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
 \begin{figure*}[t]
 	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
-   \caption{P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ simulated according to the CPL model (left) and the CTL model (right) colored by $e$, compared with \kepler EBs from \citet{Lurie2017} (red plus signs).
-The \kepler EBs at low P$_{orb}$ and low P$_{orb}/$P$_{rot}$ are likely either brown dwarfs or exoplanets \citep{Lurie2017}, and hence not modeled by our simulations, so we do not consider them, but we display them for completeness.}%
+   \caption{P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ according to the CPL model (left) and the CTL model (middle), and \citet{Lurie2017} \kepler EB observations (right).
+All points are colored by $e$.
+In the right panel, the \kepler EBs at low P$_{orb}$ and low P$_{orb}/$P$_{rot}$ are likely either brown dwarfs or exoplanets \citep{Lurie2017}, and hence are not modeled by our simulations, so we do not consider them, but we display them for completeness.}%
     \label{fig:lurie7}%
@@
 
-Both models can reproduce the major trends in the data, but the predictions of the CPL model fail to reproduce several key aspects.
-First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all tidally-locked sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.
-This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless coupling occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
-The CPL model can produce the population of subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
-The CTL model, however, seems to excel where the CPL models fails by qualitatively reproducing both populations.
-We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.
+The CPL model can produce subsynchronous rotators similar to the observed population with $10 \lsim$ P$_{orb} \lsim 20$ d, but struggles to populate the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
+The CTL model, on the other hand, populates a much wider swath of parameter space than the CPL model can, qualitatively reproducing many clusters seen in the \citet{Lurie2017} data.
+\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, the cluster of stars those authors identified near $P_{orb}/$P$_{rot} = 0.9$, compared with $9\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
+Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$, compared with $72\%$ of our CTL population, in good agreement.
+In comparison, the CPL model predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, producing a population of synchronized binaries similar to what is seen in the \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
 
-\subsubsection{P$_{orb} < 10$ d According to the CTL Model} \label{sec:subsync}
+The CPL model cannot produce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d present in the \citet{Lurie2017} data.
+Instead, owing to the its discrete P$_{eq}$, the CPL model predicts that all tidally-locked supersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.
+This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the \citet{Lurie2017} data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless coupling occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
+On the other hand, the CTL model produces a large number of supersynchronous binaries with eccentric orbits.
+These systems have tidally-locked but their orbits have not yet circularized, even though they likely experience strong tidal torques at short separations, because the tidal locking time scale is 2-3 orders of magnitude less than the circularization timescale \citep{Mazeh2008}.
 
-Our coupled stellar-tidal model reproduces many of the the spin and orbital features seen in low-mass \kepler EBs when using the CTL formalism.
+Both models predict a large number of extremely subsynchronous rotators with P$_{orb}/$P$_{rot} < 0.7$ across all P$_{orb}$ that is not seen in the \citet{Lurie2017} data.
+Magnetic braking and age combine to produce the lower limit of this subsynchronous population.
+The limit is a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down via magnetic braking in the weak tides regime over 7 Gyr, the longest age considered in our simulations.
+Our choice of prior distributions for both $Q$ and $\tau$ permit very weak tidal interactions, e.g.
+$Q \gsim 10^7$ and $\tau < 0.1$ s, that likely gives rise to this extremely subsynchronous population and suggests that our prior does not reflect the underlying distribution of stellar tidal parameters in nature.
+Alternatively, the data could be incomplete where our models predict slowly-rotating subsynchronous rotators as the photometric amplitude of star spot modulations tends to decrease with increasing P$_{rot}$, making detecting reliable rotation periods difficult \citep{McQuillan2014,Lurie2017,Reinhold2018}.
+
+Both tidal models reproduce many features observed in the \kepler EB distribution, suggesting that our model is a fair representation of the dynamical interactions of tidally-evolving, low-mass stellar binaries.
+Our comparison between theory and observations is limited, however, not just because our study focuses on examining the P$_{rot}$ evolution of binary stars subject to stellar evolution, magnetic braking, and tides throughout parameter space.
+The \citet{Lurie2017} P$_{rot}$ data lack uncertainties and \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$, leading to inaccurate $e$ determinations.
+Furthermore, biases in the data, e.g.
+the lack of long P$_{orb}$ binaries, inhibit our ability to robustly compare our predictions with the data.
+We discuss this point further in $\S$~\ref{sec:inference} and offer observational tests to discriminate between which of the two models better describes tidal interactions in low-mass binaries in $\S$~\ref{sec:whichModel}.
+
+% extra stuff
+% Second, In principle, we could correct for such a bias in the \citet{Lurie2017} data and condition our model predictions on the data to infer tidal properties, but that analysis is beyond the scope of this work for reasons we discuss in $\S$~\ref{sec:inference}.
+%  Our simulations can qualitatively reproduce many features seen in the \citet{Lurie2017} \kepler data, indicating that the equilibrium tide formalism can successfully model the tidal evolution of low-mass stellar binaries, especially at short P$_{orb}$.
+%, without reliable and accurate prior distributions for unconstrained quantities like $Q$, $\tau$ it is difficult to perform a true Monte Carlo study that would allow us to directly compare our simulated distributions to the observations.
+The lack of good priors and accurate data inhibits us from conditioning our model on the observed data to infer properties such as $Q$ and $\tau$, perhaps as a function of mass, but still permits us to qualitatively compare our results with observations and demonstrate how our model can reproduce most features seen in the data.
+%We examine the predictions of the CTL model further by comparing them with the \citet{Lurie2017} data where the data is likely most complete, P$_{orb} < 10$ d.
 In Fig.~\ref{fig:subsync}, we compare our simulated systems (dots) with the observed \citet{Lurie2017} EB data (crosses), where both points are colored by $e$.
@@
 In comparison, the CPL model (not pictured) predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, producing a population of synchronized binaries similar to what is seen in the \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
-
-Nearly all simulated binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in excellent agreement with the \citet{Lurie2017} observations.
-At very short P$_{orb}$ in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries.
-Beyond P$_{orb} \approx 4$ d, tidally-locked, eccentric binaries populate the supersynchronous population, again in excellent agreement with the \citet{Lurie2017} \kepler data.
-These systems have tidally-locked but their orbits have not yet circularized, even though they likely experience strong tidal torques at short separations, because the tidal locking time scale is 2-3 orders of magnitude less than the circularization timescale \citep{Mazeh2008}.
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{../Plots/subsync.pdf}
-   \caption{Same format as Fig.~\ref{fig:lurie7}, with \citet{Lurie2017} data plotted as crosses.
-All points are colored according to the binary orbital $e$.
-Our simulations reproduce many of the observed orbital and rotational features seen in the \kepler EB population.}%
-    \label{fig:subsync}%
-\end{figure}
-
-For P$_{orb} \gsim 4$ d, our model produces a population of subsynchronous rotators, including a cluster of subsynchronously rotating stellar binaries with nearly circular orbits.
-Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces subsynchronous rotators.
-Our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
-Our simulations can qualitatively reproduce many features seen in the \citet{Lurie2017} \kepler data, indicating that the equilibrium tide formalism can successfully model the tidal evolution of low-mass stellar binaries, especially at short P$_{orb}$.
-The comparison between theory and observations, however, is limited to more qualitative comparisons for two reasons.
-First, the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$, leading to inaccurate $e$ determinations.
-In the absence of more accurate constraints in the \citet{Lurie2017} \kepler data, we cannot robustly compare out simulation results with observations to discriminate between which equilibrium tidal model better describes tidal torques in low-mass.
-Second, without reliable and accurate prior distributions for unconstrained quantities like $Q$, $\tau$, and the initial P$_{orb}$, it is difficult to perform a true Monte Carlo study that would allow us to directly compare our simulated distributions to the observations.
-The lack of good priors and accurate data inhibits us from conditioning our model on the observed data to infer properties such as $Q$ and $\tau$, perhaps as a function of mass, but still permits us to qualitatively compare our results with observations and demonstrate how our model can reproduce most features seen in the data.
-We discuss this point further in $\S$~\ref{sec:inference}.
 
@@
 
-Accurate measurements of P$_{rot}$ and $e$, especially out to long P$_{orb}$, can discriminate between which equilibrium tidal model best describes tidal interactions in low-mass stellar binaries.
+Accurate measurements of P$_{rot}$ and $e$, especially out to long P$_{orb}$, can potentially discriminate between which equilibrium tidal model best describes tidal interactions in low-mass stellar binaries.
 Here, we outline three observational tests that can discriminate between the two models.
@@
 Eqn.~(\ref{eqn:cpl:eqPer}).
-The CTL model, however, predicts a continuum of supersynchronous rotators on eccentric orbits in this regime, e.g.
+The CTL model, however, predicts a continuum of supersynchronous rotators on eccentric orbits, e.g.
 Eqn.~(\ref{eqn:ctl:eqPer}).
 Supersynchronous rotation that is not due to tidal interactions can occur in extremely young, rapidly rotating systems that are still contracting along the pre-main sequence, or that have recently reached the main sequence.
-These young, supersynchronous rotators are unlikely to be tidally-locked and usually have P$_{orb}/$P$_{rot} > 1.5$, distinguishing them from tidally-locked binaries (see Fig.~\ref{fig:lurie7}).
-If supersynchronous rotation is observed in binaries with P$_{orb} < 10$ d, P$_{orb}/$P$_{rot} < 1.5$, and $0 < e \lsim 0.23$, it is evidence in favor of the CTL model.
+These young, supersynchronous rotators are unlikely to be tidally-locked, usually have P$_{orb}/$P$_{rot} > 1.5$, and do not stay supersynchronous for long given that solar mass pre-main sequence lifetimes are $\lsim 100$ Myr, distinguishing them from tidally-locked binaries (see Fig.~\ref{fig:lurie7}).
+If supersynchronous rotation is observed in binaries with P$_{orb} < 10$ d, P$_{orb}/$P$_{rot} < 1.5$, and $0 < e \lsim 0.23$, it is evidence in favor of the CTL model over our discrete CPL model.
 
 Second, for tidally-locked binaries with $e > 0.23$, the CPL model predicts supersynchronous rotation in the form of a 3:2 spin-orbit comensurability, e.g.
-the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permitted.
-The CTL model, however, permits a continuum of supersynchronous rotation states in eccentric tidally-locked rotators.
+the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permitted, compared to the continuum of supersynchronous rotation states in eccentric tidally-locked rotators predicted by the CTL model.
 If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model.
@@
 Eqn.~(\ref{eqn:cpl:eqPerCont}) versus Eqn.~(\ref{eqn:ctl:eqPer}).
-In practice, this is likely too difficult as it requires extensive photometric and spectroscopic observations of many binaries.
+In practice, this is extremely observationally expensive as it requires extensive photometric and spectroscopic observations of many binaries.
 Since there is no obvious clustering of \kepler EBs near any spin-orbit resonance, the above two tests are likely insufficient.
@@
 The CPL model, however, readily tidally-locks binaries out to P$_{orb} \approx 100$ d.
-We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small starspot modulation amplitudes \citep{Lurie2017,Reinhold2018}.
+We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small star spot modulation amplitudes \citep{McQuillan2014,Lurie2017,Reinhold2018}.
 Long term spectroscopic monitoring may be warranted in such cases.
@@
 
-\subsection{Inferring Binary Tidal Properties from \kepler} \label{sec:inference}
+\subsection{Inferring Stellar and Binary Properties from \kepler} \label{sec:inference}
 
 In principle, we could have conditioned our model on the \citet{Lurie2017} data to infer population-level properties of stellar binaries, such as tidal $Q$ and $\tau$ values for stars, perhaps as a function of stellar mass, or the initial binary P$_{orb}$ distribution, using hierarchical probabilistic modeling instead of qualitatively comparing of our simulations with the data.
-Such an analysis is beyond the scope of this work because the biases in the \citet{Lurie2017} data are not well known.
+Such an analysis is beyond the scope of this work, however.
+Furthermore, the biases in the \citet{Lurie2017} data are not well known.
 For example, the bias of the obvious lack of longer P$_{orb}$ binaries in the \citet{Lurie2017} data arises from the fact that it is difficult to detect binaries with longer P$_{orb}$ in a finite time window via the transit method as longer P$_{orb}$ systems transit fewer times.
 What is not known, however, is how this bias impacts the completeness of binary detections as a function of P$_{orb}$, nor is the completeness known as a function of stellar P$_{rot}$.
-These biases, combined with the lack of P$_{rot}$ uncertainties, inaccurate binary eccentricities, and lack of binary masses and ages makes it difficult to directly compare our simulations with the data to robustly infer properties of low-mass binaries.
+Unconstrained observational biases, combined with the lack of P$_{rot}$ uncertainties, potentially inaccurate binary eccentricities, and lack of binary masses and ages makes it difficult to directly compare our simulations with the data to robustly infer properties of low-mass binaries.
 
@@
 This system has been extensively monitored via photometry and spectroscopy because it harbors multiple transiting circumbinary planets, and as a result has well-constrained properties.
-One could infer the tidal properties for this binary, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties.
-This analysis, however, is beyond the scope of this work and we leave it for future endeavors.
+One could infer the tidal properties of this binary, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties and reasonable prior probability distributions for parameters like the initial binary $e$.
+This analysis, however, is also beyond the scope of this work and we leave it for future endeavors.
 
@@
 In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have tidally-synchronized spins and circularized orbits, in good agreement with observations of \kepler EBs and binaries in the field.
-We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking.
+We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking, or supersynchronously if they tidally-lock on eccentric orbits.
 Our predictions are not strongly dependant on the choice of magnetic braking model, but rather are generic outcomes of the interaction between magnetic braking and tidal torques.
-Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact out to P$_{orb} \approx 80$ d.
-Many binaries with P$_{orb} \lsim 20$ d tidally-lock according to both models, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.
+Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact at longer P$_{orb}$ than have previously been considered, out to P$_{orb} \approx 80$ d.
+Many binaries with P$_{orb} \lsim 20$ d tidally-lock according to both models, in good agreement with previous results, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.
 Tidal interactions can cause P$_{rot}$ evolution in stellar binaries to differ from the long-term spin down due to magnetic braking experienced by single stars, decoupling P$_{rot}$ from age.
@@
 
-We compare the predictions of both the CPL and CTL models with observations of P$_{rot}$ and P$_{orb}$ of \kepler EBs by \citet{Lurie2017} and find that both can qualitatively reproduce many features seen in the data, validating our approach and suggesting that equilibrium tidal models can accurately reproduce stellar-tidal evolution in low-mass stellar binaries.
-The lack of uncertainties on P$_{rot}$ and the approximate orbital eccentricities derived by \citet{Lurie2017} prevent us from discriminating between which tidal model best describes tidal torques in low-mass binaries.
+We compare the predictions of both the CPL and CTL models with observations of P$_{rot}$ and P$_{orb}$ of \kepler EBs by \citet{Lurie2017} and find that both can qualitatively reproduce many features seen in the data, validating our approach and suggesting that equilibrium tidal models can accurately model stellar-tidal evolution in low-mass stellar binaries.
+The lack of uncertainties on P$_{rot}$, the approximate orbital eccentricities derived by \citet{Lurie2017}, and unconstrained completeness estimates prevent us from discriminating between which tidal model best describes tidal torques in low-mass binaries and from inferring tidal properties of low-mass stars given the \kepler EB data.
 We offer three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
@@
 
-Our theoretical predictions outline a critical point: one cannot simply observe a short P$_{orb}$ binary on a circular orbit and assume synchronization, nor can one observe a binary with P$_{orb} \gsim 20$ d and assume that tides have not shaped that system's angular momentum evolution.
+Our theoretical predictions outline a critical point: one cannot simply observe a short P$_{orb}$ binary on a circular orbit and assume synchronization, nor can one observe a binary with P$_{orb} \gsim 20$ d and assume that tides have not impacted that system's angular momentum evolution.
 Stellar-tidal interactions can produce synchronous and subsynchronous rotation for short P$_{orb}$ binaries on circular orbits, e.g.
@@
 
+\appendix
+
+\section{Analytic Torque Balance} \label{sec:appendix:balance}
+
+Here we derive the equation for the stellar P$_{rot}$ at which tidal torques balance magnetic braking discussed in $\S$~\ref{sec:analytic}.
+As in $\S$~\ref{sec:analytic}, we assume that both stars have M $= 1$M$_{\odot}$, 0 obliquity, and we assume a circular binary orbit.
+We assume that the torque balance occurs while the stars are on the main sequence, where stellar properties change slowly, so the angular momentum evolution is controlled by the balance between tidal torques and magnetic braking, not stellar radius contraction.
+Under this assumption, we can set $R = 1 R_{\odot}$ and assume constant moments of inertia.
+For simplicity, we assume that magnetic braking proceeds under the \citet{Matt2015} model and the CTL model describes tidal torques.
+
+As discussed in $\S$~\ref{sec:analytic}, both stars are in the unsaturated rotation regime, so the torque due to magnetic braking is given by Eqn.~(\ref{eqn:mattUnSat}), which under the aforementioned assumptions, reduces to
+\begin{equation}
+    \frac{dJ}{dt}\Bigg|_{MB} = -C_{MB} \left( \frac{P_{rot,\odot}}{P_{rot}} \right)^3
+\end{equation}
+where P$_{rot} = 2 \pi / \omega$ and $C_{MB} = 9.5 \times 10^{30}$ ergs \citep{Matt2015}.
+
+Under the CTL model and our assumptions, the change in rotation rate due to tidal torques, Eqn.~\ref{eqn:ctl:omega}, reduces to 
+\begin{equation}
+    \frac{d\omega}{dt}\Bigg|_{tides} = \frac{\mathrm{P}_{orb}\mathrm{Z}_{CTL}}{2 \pi M r_g^2 R^2}\left( 1 - \frac{\mathrm{P}_{orb}}{\mathrm{P}_{rot}} \right)
+\end{equation}
+where P$_{orb} = 2 \pi/n$.  For fixed moment of inertia, $dJ/dt =I d\omega/dt$, and after inserting Eqn.~\ref{eqn:ctl:z} for Z$_{CTL}$, the tidal torque on the stellar rotations becomes
+\begin{equation}
+    \frac{dJ}{dt}\Bigg|_{tides} = \frac{C_{tides} k_2 \tau}{\mathrm{P}_{orb}^5}\left( 1 - \frac{\mathrm{P}_{orb}}{\mathrm{P}_{rot}} \right).
+\end{equation}
+where $C_{tides} = 24 \pi^5 R_{\odot}^5 / G$.
+
+The torques due to tides and magnetic braking balance when $ \frac{dJ}{dt}|_{tides} +  \frac{dJ}{dt}|_{MB} = 0$, 
+\begin{equation} \label{eqn:appendixBalance}
+    \frac{C_{tides} k_2 \tau}{\mathrm{P}_{orb}^5}\left( 1 - \frac{\mathrm{P}_{orb}}{\mathrm{P}_{rot}} \right) -C_{MB} \left( \frac{P_{rot,\odot}}{P_{rot}} \right)^3 = 0.
+\end{equation}
+By specifying P$_{orb}$ and the product, $k_2 \tau$, we can numerically solve Eqn.~(\ref{eqn:appendixBalance}) for the P$_{rot}$ at which torques due to magnetic braking and tides balance, often producing subsynchronous rotation as seen in Fig.~\ref{fig:analyticBalance} and our simulations in $\S$~\ref{sec:eq}.
+
 %% BIBLIOGRAPHY %%

### 2019-02-27 Paper/sync.tex
Updates from Overleaf

@@
 
-The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but rather is a generic outcome of the coupling of magnetic braking and weak tidal torques.
+The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but is a generic outcome of the coupling of magnetic braking and weak tidal torques.
 The CTL model can only tidally-lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the longer P$_{orb}$ binaries into subsynchronous rotation.
@@
 
-We examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, respectfully.
+We next examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, respectfully.
 In Fig.~\ref{fig:qTauLock}, we bin our simulation results after the full 7 Gyr evolution by P$_{orb}$ and $Q$ or $\tau$ and compute the median P$_{orb}/$P$_{rot}$ in each bin, marginalizing over all other parameters.
@@
 We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
-In each bin, we compute the median $Q$ and $\tau$, and the $16\%$, and $84\%$ percentiles to quantify the distribution of tidal strength parameters that are consistent with each spin-orbital state.
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{../Plots/porbProtPorbQHist.pdf}
+In each bin, we compute the median $Q$ and $\tau$, and the $16\%$, and $84\%$ percentiles, a measure of the spread of the distribution, to quantify the distribution of tidal strength parameters that are consistent with each spin-orbital state.
+
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../Plots/porbProtPorbQHist.pdf}
    \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.
@@
     \label{fig:qmap}%
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{../Plots/porbProtPorbTauHist.pdf}
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../Plots/porbProtPorbTauHist.pdf}
    \caption{Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
     \label{fig:taumap}%
-\end{figure}
+\end{figure*}
 
@@
 There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
-The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.
+The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) that have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.
 In stark contrast, both tidal models predict a substantial population of tidally-locked fast rotators (P$_{rot} < 20$ d).
@@
 The \citet{Reiners2012} model's magnetic braking transition at $M = 0.35$ M$_{\odot}$ produces a sharp cutoff in the P$_{rot} - M$ distribution between late M dwarfs and more massive stars.
-More massive stars ($M \geq 0.35$ M$_{\odot}$) follow a smooth P$_{rot} -$ age sequence, while the late M dwarfs can remain rapidly rotating if they are young and in the saturated regime or extremely slowly rotating if the are older than about 4 Gyr.
+More massive stars ($M \geq 0.35$ M$_{\odot}$) follow a smooth P$_{rot} -$ age sequence, while the late M dwarfs can remain rapidly rotating if they are young and in the saturated regime or extremely slowly rotating if they are older than about 4 Gyr.
 As in the \citet{Matt2015} model, the only rapid rotators (P$_{rot} \lsim 20$ d) are very young stars with ages $\lsim 2$ Gyr.
@@
 
-Our models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{rot} < 20 $ d, as many of these systems tidally-lock at short P$_{orb}$.
+Our models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{orb} < 20 $ d.
 Under the \citet{Matt2015} model discussed in $\S$~\ref{sec:mattDist}, the median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
@@
 In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another.
-Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are potentially subject to large systematic errors.
-Moreover, this problem could be more significant than the aforementioned estimate implies as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are potentially subject to systematic errors.
+Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
 We caution that any application of gyrochronlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
@@
 Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
-We recommend that the application of or the calibration of magnetic braking models to a sample of stellar rotation periods control for binarity.
+We recommend that the application, or calibration, magnetic braking models to a sample of stellar rotation periods control for binarity.
 
@@
 $Q \gsim 10^7$ and $\tau < 0.1$ s, that likely gives rise to this extremely subsynchronous population and suggests that our prior does not reflect the underlying distribution of stellar tidal parameters in nature.
-Alternatively, the data could be incomplete where our models predict slowly-rotating subsynchronous rotators as the photometric amplitude of star spot modulations tends to decrease with increasing P$_{rot}$, making detecting reliable rotation periods difficult \citep{McQuillan2014,Lurie2017,Reinhold2018}.
+Alternatively, the data could be incomplete where our models predict slowly-rotating subsynchronous rotators as the photometric amplitude of star spot modulations tends to decrease with increasing P$_{rot}$, making reliable rotation periods difficult to detect \citep{McQuillan2014,Lurie2017,Reinhold2018}.
 
 Both tidal models reproduce many features observed in the \kepler EB distribution, suggesting that our model is a fair representation of the dynamical interactions of tidally-evolving, low-mass stellar binaries.
-Our comparison between theory and observations is limited, however, not just because our study focuses on examining the P$_{rot}$ evolution of binary stars subject to stellar evolution, magnetic braking, and tides throughout parameter space.
+Our comparison between theory and observations is limited, however.
 The \citet{Lurie2017} P$_{rot}$ data lack uncertainties and \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$, leading to inaccurate $e$ determinations.
@@
 The lack of uncertainties on P$_{rot}$, the approximate orbital eccentricities derived by \citet{Lurie2017}, and unconstrained completeness estimates prevent us from discriminating between which tidal model best describes tidal torques in low-mass binaries and from inferring tidal properties of low-mass stars given the \kepler EB data.
-We offer three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
-To perform these tests, we that recommend observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d, although we note that these are non-trivial observations.
+We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
+To perform these tests, we recommend that observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d, although we note that these are non-trivial observations.
 
@@
 \acknowledgments
-This work was facilitated though the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system and funded by the STF at the University of Washington.
-This work was supported by NASA Headquarters under the NASA Earth and Space Science Fellowship Program - Grant 80NSSC17K0482.
-DPF and RB acknowledge that this work was supported by the NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement number NNA13AA93A.
+This work was facilitated though the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system and funded by the Student Technology Fund at the University of Washington.
+DPF was supported by NASA Headquarters under the NASA Earth and Space Science Fellowship Program - Grant 80NSSC17K0482.
+RB acknowledges support from the NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement number NNA13AA93A.
 

### 2019-03-05 Paper/sync.tex

@@
 We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them.
-All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.} with accompanying descriptions and documentation.
+All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}.
 
@@
 Stellar-tidal interactions can produce synchronous and subsynchronous rotation for short P$_{orb}$ binaries on circular orbits, e.g.
-Fig.~\ref{fig:subsync}, depending on the age of the system, e.g.
+Fig.~\ref{fig:eqPer}, depending on the age of the system, e.g.
 Fig.~\ref{fig:eqPerShortPorb}, and the strength of tidal dissipation, e.g.
@@
 RB acknowledges support from the NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement number NNA13AA93A.
+JRAD acknowledges support from the DIRAC Institute in the Department of Astronomy at the University of Washington.
+The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
 

### 2019-03-27 Paper/sync.tex
Updates from Overleaf

@@
 
-\keywords{binaries: close}
+\keywords{binaries: close, stars: evolution, stars: kinematics and dynamics, stars: rotation}
 
@@
 \begin{equation} \label{eqn:matt0}
-\frac{dJ}{dt}\Bigg|_0 = 9.5 \times 10^{30} \ \mathrm{erg} \ \left( \frac{R}{R_{\odot}} \right)^{3.1} \left( \frac{M}{M_{\odot}} \right)^{0.5}.
+\frac{dJ}{dt}\Bigg|_0 = 6.3 \times 10^{30} \ \mathrm{erg} \ \left( \frac{R}{R_{\odot}} \right)^{3.1} \left( \frac{M}{M_{\odot}} \right)^{0.5}.
 \end{equation}
 Saturated magnetic braking occurs for $Ro \leq Ro_{\odot}/\chi$ for $\chi = 10$ where \citet{Matt2015} defines $\chi = Ro_{\odot}/Ro_{sat}$.
-We adopt all model parameters given in Table 1 from \citet{Matt2015} and compute $\tau_{cz}$ using Eqn.
+We adopt all model parameters given in Table 1 from \citet{Matt2015}, with the correction from \citet{Matt2019}, and compute $\tau_{cz}$ using Eqn.
 (36) from \citet{Cranmer2011}.
@@
 \end{equation}
-where P$_{rot} = 2 \pi / \omega$ and $C_{MB} = 9.5 \times 10^{30}$ ergs \citep{Matt2015}.
+where P$_{rot} = 2 \pi / \omega$ and $C_{MB} = 6.3 \times 10^{30}$ ergs \citep{Matt2015,Matt2019}.
 

### 2019-03-27 Analysis/lockTimeDist.py
updated plots with corrected Matt+2015 model

@@ -8,10 +8,10 @@ period bins.
 
 Script output:
 
-CPL Primary Locked Fraction: 0.3119
-CTL Primary Locked Fraction: 0.0768
-CPL Secondary Locked Fraction: 0.2929
-CTL Secondary Locked Fraction: 0.0881
+CPL Primary Locked Fraction: 0.3084
+CTL Primary Locked Fraction: 0.0856
+CPL Secondary Locked Fraction: 0.3017
+CTL Secondary Locked Fraction: 0.0962
 
 """
 
@@ -30,8 +30,8 @@ mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
 cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9

### 2019-03-27 Analysis/locking.py

@@ -10,73 +10,39 @@ in the Kepler field.
 
 Script Output:
 
-CPL Prot median: 40.0140705
-CPL locked Prot median: 27.605545999999997
-CPL unlocked Prot median: 41.29818
-CPL interacting Prot median: 51.323143
-CPL fraction of locked binaries: 0.247
-CPL fraction of not locked but interacting binaries: 0.0971
-CPL fraction of locked binaries with Prot < 20 d: 0.097
+CPL Prot median: 31.550073499999996
+CPL locked Prot median: 22.636840499999998
+CPL unlocked Prot median: 32.429345
+CPL interacting Prot median: 43.517854
+CPL fraction of locked binaries: 0.2354
+CPL fraction of not locked but interacting binaries: 0.071
+CPL fraction of locked binaries with Prot < 20 d: 0.46219201359388273
 
-CTL Prot median: 39.7350345
-CTL locked Prot median: 7.400932000000001
-CTL unlocked Prot median: 40.804916999999996
-CTL interacting Prot median: 45.6187995
-CTL fraction of locked binaries: 0.0747
-CTL fraction of not locked but interacting binaries: 0.1756
-CTL fraction of locked binaries with Prot < 20 d: 0.0643
-
-Age Stats:
-CPL locked Age median: 4.395778067659794
-CPL unlocked Age median: 3.6588904832188893
-CPL interacting Age median: 4.436232401852751
-
-CTL locked Age median: 3.955133657998956
-CTL unlocked Age median: 3.8779651074930905
-CTL interacting Age median: 4.550487764517989
-
-Porb Stats:
-CPL locked Porb median: 28.608129499999997
-CPL unlocked Porb median: 60.951834999999996
-CPL interacting Porb median: 58.36535500000001
-
-CTL locked Porb median: 7.559052
-CTL unlocked Porb median: 55.369326
-CTL interacting Porb median: 52.700736
-[dflemin3 Analysis]$ python locking.py
-CPL Prot median: 40.0140705
-CPL locked Prot median: 27.605545999999997
-CPL unlocked Prot median: 41.29818
-CPL interacting Prot median: 51.323143
-CPL fraction of locked binaries: 0.247
-CPL fraction of not locked but interacting binaries: 0.0971
-CPL fraction of locked binaries with Prot < 20 d: 0.39271255060728744
-
-CTL Prot median: 39.7350345
-CTL locked Prot median: 7.400932000000001
[…]

### 2019-03-27 Analysis/lockingHistograms.py

@@ -28,8 +28,8 @@ seed = 42
 np.random.seed(seed)
 
 # Load in data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Create indices for a random sample of num points to make scatterplot legible
 inds = np.random.choice(np.arange(len(cpl)), size=num, replace=False)

### 2019-03-27 Analysis/lurie.py

@@ -24,8 +24,8 @@ mpl.rc('text', usetex=True)
 plotLurie = True
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 #lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
 #lurie["Prot"] = lurie["p_1_min"].copy() # As recommended by Lurie+2017 for equitorial Prot

### 2019-03-27 Analysis/pEqDist.py

@@ -25,8 +25,8 @@ mpl.rc('text', usetex=True)
 bins = 20
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 ### Total marginal histogram of Peq/Prot ###
 

### 2019-03-27 Analysis/porbProtPorbHist.py

@@ -23,9 +23,9 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
 cpl["Pri_dTidalQ"] = pd.Series(cpl["Pri_dTidaLQ"].values, index=cpl.index)
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Read in lurie data
 lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
@@ -118,7 +118,7 @@ cbaxes = fig.add_subplot(gs[2])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Median log$_{10}(Q)$", labelpad=30, fontsize=30)
 
-fig.savefig("../Plots/porbProtPorbQHist.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/porbProtPorbQHist.pdf", bbox_inches="tight", dpi=200)
 
 ### CTL Plot ###
 fig = plt.figure(figsize=(9, 8))
@@ -149,7 +149,7 @@ cbaxes = fig.add_subplot(gs[2])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label=r"Median log$_{10}(\tau[\mathrm{s}])$", labelpad=30, fontsize=30)
 
-fig.savefig("../Plots/porbProtPorbTauHist.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/porbProtPorbTauHist.pdf", bbox_inches="tight", dpi=200)
 
 ################################################################################
 #

### 2019-03-27 Analysis/porbProtPorbHistAge.py

@@ -23,9 +23,9 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
 cpl["Pri_dTidalQ"] = pd.Series(cpl["Pri_dTidaLQ"].values, index=cpl.index)
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Read in lurie data
 lurie = pd.read_csv("../Data/Lurie2017Full.csv", header=0)
@@ -118,7 +118,7 @@ cbaxes = fig.add_subplot(gs[2])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Median log$_{10}(Q)$", labelpad=30, fontsize=25)
 
-fig.savefig("../Plots/porbProtPorbQHistAge.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/porbProtPorbQHistAge.pdf", bbox_inches="tight", dpi=200)
 
 ### CTL Plot ###
 fig = plt.figure(figsize=(9, 8))
@@ -149,7 +149,7 @@ cbaxes = fig.add_subplot(gs[2])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label=r"Median log$_{10}(\tau[\mathrm{s}])$", labelpad=30, fontsize=25)
 
-fig.savefig("../Plots/porbProtPorbTauHistAge.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/porbProtPorbTauHistAge.pdf", bbox_inches="tight", dpi=200)
 
 ################################################################################
 #
@@ -219,7 +219,7 @@ if estQTau:
             ha="center", va="center", size=16, color="black", zorder=100,
             bbox=dict(boxstyle="square", fc="white", ec="white", alpha=0.9))
 
-    fig.savefig("../Plots/qLurieAge.pdf", bbox_inches="tight", dpi=600)
+    fig.savefig("../Plots/qLurieAge.pdf", bbox_inches="tight", dpi=200)
 
     # Tidal tau figure
     fig, ax = plt.subplots()
@@ -241,4 +241,4 @@ if estQTau:
[…]

### 2019-03-27 Analysis/protDist.py

@@ -8,28 +8,28 @@ single star models uniform over ages [1,7] Gyr, roughly field star ages.
 
 Script Output:
 
-CPL median Prot: 40.0140705
-CTL median Prot: 39.7350345
+CPL median Prot: 31.550073499999996
+CTL median Prot: 31.293733500000002
 Single median Prot: 46.4415785
 
-CPL Fraction of Stars with Prot < 7.5 d: 0.0566
-CTL Fraction of Stars with Prot < 7.5 d: 0.0625
+CPL Fraction of Stars with Prot < 7.5 d: 0.0968
+CTL Fraction of Stars with Prot < 7.5 d: 0.0845
 Single Fraction of Stars with Prot < 7.5 d: 0.0172
 
-CPL Fraction of Stars with Prot < 20 d: 0.1766
-CTL Fraction of Stars with Prot < 20 d: 0.1885
+CPL Fraction of Stars with Prot < 20 d: 0.2665
+CTL Fraction of Stars with Prot < 20 d: 0.268
 Single Fraction of Stars with Prot < 20 d: 0.0523
 
-CPL Median Age of Stars with Prot < 20 d: 2.7893700005017488
-CTL Median Age of Stars with Prot < 20 d: 3.108449437935221
+CPL Median Age of Stars with Prot < 20 d: 2.256742668821268
+CTL Median Age of Stars with Prot < 20 d: 2.386626069632524
 Single Mediage Age of Stars with Prot < 20 d: 1.346166303555715
 
-CPL + Age of Stars with Prot < 20 d: 5.705097268321147
-CTL + Age of Stars with Prot < 20 d: 5.695241667822796
+CPL + Age of Stars with Prot < 20 d: 5.13638640139413
+CTL + Age of Stars with Prot < 20 d: 5.410552087675718
 Single + Age of Stars with Prot < 20 d: 1.8214255111813729
 
-CPL - Age of Stars with Prot < 20 d: 1.315608590645681
-CTL - Age of Stars with Prot < 20 d: 1.3464971366685323
+CPL - Age of Stars with Prot < 20 d: 1.2826632777047906
+CTL - Age of Stars with Prot < 20 d: 1.3182898896016302
 Single - Age of Stars with Prot < 20 d: 1.0843035833816197
 
 
@@ -75,8 +75,8 @@ num = 2500
 np.random.seed(seed)
 
 # Load data
[…]

### 2019-03-27 Analysis/protDistAgeHist.py

@@ -34,8 +34,8 @@ num = 2500
 np.random.seed(seed)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 single = pd.read_csv("../Data/mcSingle.csv")
 lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
                     names=["Porb", "Prot", "Ecc"])
@@ -51,20 +51,20 @@ ax.hist(single["Age"][single["Pri_ProtAge"] <= 20]/1.0e9, bins="auto",
         histtype="step", lw=3, label="Single", color="C2", density=False)
 
 # Annotate young stars
-ax.annotate(s="", xy=(2.5, 430), xytext=(0.8, 430),
+ax.annotate(s="", xy=(2.5, 560), xytext=(0.8, 560),
             arrowprops={"arrowstyle" : "<->", "color" : "black"})
-ax.text(1.65, 460, "Young \n Rapid Rotators", ha="center", va="center", size=21, color="k")
+ax.text(1.65, 600, "Young \n Rapid Rotators", ha="center", va="center", size=21, color="k")
 
 # Annotate locked population
-ax.annotate(s="", xy=(7, 180), xytext=(2.75, 180),
+ax.annotate(s="", xy=(7, 200), xytext=(2.75, 200),
             arrowprops={"arrowstyle" : "<->", "color" : "black"})
-ax.text(4.85, 200, "Tidally-Locked Binaries", ha="center", va="center", size=21,
+ax.text(4.85, 220, "Tidally-Locked Binaries", ha="center", va="center", size=21,
         color="k")
 
 ax.set_xlabel("Age [Gyr]", fontsize=30)
 ax.set_ylabel("Counts", fontsize=30)
 ax.set_xlim(0.8,7.2)
-ax.set_ylim(0, 490)
+ax.set_ylim(0, 650)
 ax.legend(loc="best", framealpha=0.0, fontsize=22)
 
 fig.savefig("../Plots/protAgeHist.pdf", dpi=200, bbox_inches="tight")

### 2019-03-27 Analysis/qTauPorbHist.py

@@ -27,8 +27,8 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
 cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9

### 2019-03-27 Analysis/qTauPorbHistAge.py

@@ -27,8 +27,8 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
 cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
@@ -109,7 +109,7 @@ cbaxes = fig.add_subplot(gs[4])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Median P$_{eq}$/P$_{rot}$", labelpad=30, rotation=270)
 
-fig.savefig("../Plots/qTauPorbPeqAge.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/qTauPorbPeqAge.pdf", bbox_inches="tight", dpi=200)
 # Done!
 
 ################################################################################
@@ -146,5 +146,5 @@ cbaxes = fig.add_subplot(gs[4])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Median P$_{orb}$/P$_{rot}$", labelpad=30, rotation=270, fontsize=25)
 
-fig.savefig("../Plots/qTauPorbRatioHistAge.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/qTauPorbRatioHistAge.pdf", bbox_inches="tight", dpi=200)
 # Done!

### 2019-03-27 Analysis/ratioPorbHist.py

@@ -24,8 +24,8 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
 cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
@@ -105,7 +105,7 @@ cbaxes = fig.add_subplot(gs[4])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Median P$_{eq}$/P$_{rot}$", labelpad=30, rotation=270)
 
-fig.savefig("../Plots/ratioPorbPeq.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/ratioPorbPeq.pdf", bbox_inches="tight", dpi=200)
 # Done!
 
 ################################################################################
@@ -141,5 +141,5 @@ cbaxes = fig.add_subplot(gs[4])
 cb = plt.colorbar(im, cax=cbaxes)
 cb.set_label(label="Median P$_{orb}$/P$_{rot}$", labelpad=30, rotation=270)
 
-fig.savefig("../Plots/ratioPorbRatioHist.pdf", bbox_inches="tight", dpi=600)
+fig.savefig("../Plots/ratioPorbRatioHist.pdf", bbox_inches="tight", dpi=200)
 # Done!

### 2019-03-27 Analysis/subsync.py

@@ -10,19 +10,19 @@ Script output:
 
 CTL:
 
-Median subsync tidal tau: 0.17078373819999998
-Median supersync tidal tau: 0.466656378
-Median subsync ecc: 0.107152
-Median supersync ecc: 0.15391400000000002
-Median subsync mass ratio: 0.6416154898583578
-Median supersync mass ratio: 0.6614121797592005
-Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.7182910547396528
-Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.09078771695594126
+Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.7367688022284122
+Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.07520891364902507
+Median subsync tidal tau: 0.13994108759999999
+Median supersync tidal tau: 0.413785876
+Median subsync ecc: 0.098839
+Median supersync ecc: 0.15375899999999998
+Median subsync mass ratio: 0.625039651747849
+Median supersync mass ratio: 0.6587939375776
 
 CPL:
 
-Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6887052341597796
-Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.01928374655647383
+Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.7332457293035479
+Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.01576872536136662
 
 Interpretation:
 
@@ -51,8 +51,8 @@ plotLurie = True
 bins = "auto"
 
 # Load data
-cpl = pd.read_csv("../Data/mcCPLDec22.csv")
-ctl = pd.read_csv("../Data/mcCTLDec22.csv")
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
 
 # Load in cleaned sample of Prot and Ecc from Lurie+2017
 lurie = pd.read_csv("../Data/Lurie2017.csv", comment="#", header=None,
@@ -166,7 +166,7 @@ ax_marg_y.legend(loc="best", framealpha=0.0, fontsize=15)
 ax_marg_y.set_ylim(0.6, 1.25)
[…]

### 2019-03-27 Paper/sync.tex

@@
 The CTL model, however, predicts that for P$_{orb} \gsim 10$ d, magnetic braking overpowers weak tidal torques, spinning down the stars past the tidally-locked state.
-Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.1$ with the maximum value set by the torque balance between tidal torques and magnetic braking, in good agreement with the analytic calculations in $\S$~\ref{sec:analytic}.
+Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.05$ with the maximum value set by the torque balance between tidal torques and magnetic braking, in good agreement with the analytic calculations in $\S$~\ref{sec:analytic}.
 As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.

### 2019-04-08 Paper/sync.tex

@@
 
-The long-term angular momentum evolution of low-mass (M$\lsim 1$ M$_{\odot}$) stars is controlled by magnetic braking, the torque exerted on stars due to the coupling of stellar winds to the surface magnetic field \citep{Dunn1961,Mestel1968}.
+The long-term angular momentum evolution of low-mass (M$\lsim 1$ M$_{\odot}$) stars is controlled by magnetic braking, the torque exerted on stars due to the coupling of stellar winds to the surface magnetic field \citep{Mestel1968}.
 Early in stellar lifetimes, stars spin-up as they contract along the pre-main sequence.

### 2019-04-09 Analysis/cmap.py
adding gyro plots, updating figures per ref's request

@@ -4,57 +4,21 @@ https://stackoverflow.com/questions/7404116/defining-the-midpoint-of-a-colormap-
 """
 
 import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-from mpl_toolkits.axes_grid1 import AxesGrid
+from matplotlib import colors
 
-def shiftedColorMap(cmap, start=0, midpoint=0.5, stop=1.0, name='shiftedcmap'):
-    '''
-    Function to offset the "center" of a colormap. Useful for
-    data with a negative min and positive max and you want the
-    middle of the colormap's dynamic range to be at zero.
+# set the colormap and centre the colorbar
+class MidpointNormalize(colors.Normalize):
+	"""
+	Normalise the colorbar so that diverging bars work there way either side from a prescribed midpoint value)
 
-    Input
-    -----
-      cmap : The matplotlib colormap to be altered
-      start : Offset from lowest point in the colormap's range.
-          Defaults to 0.0 (no lower offset). Should be between
-          0.0 and `midpoint`.
-      midpoint : The new center of the colormap. Defaults to
-          0.5 (no shift). Should be between 0.0 and 1.0. In
-          general, this should be  1 - vmax / (vmax + abs(vmin))
-          For example if your data range from -15.0 to +5.0 and
-          you want the center of the colormap at 0.0, `midpoint`
-          should be set to  1 - 5/(5 + 15)) or 0.75
-      stop : Offset from highest point in the colormap's range.
-          Defaults to 1.0 (no upper offset). Should be between
-          `midpoint` and 1.0.
-    '''
-    cdict = {
-        'red': [],
-        'green': [],
-        'blue': [],
-        'alpha': []
-    }
+	e.g. im=ax1.imshow(array, norm=MidpointNormalize(midpoint=0.,vmin=-100, vmax=100))
+	"""
+	def __init__(self, vmin=None, vmax=None, midpoint=None, clip=False):
+		self.midpoint = midpoint
+		colors.Normalize.__init__(self, vmin, vmax, clip)
 
-    # regular index to compute the colors
-    reg_index = np.linspace(start, stop, 257)
-
-    # shifted index to match the data
-    shift_index = np.hstack([
-        np.linspace(0.0, midpoint, 128, endpoint=False),
-        np.linspace(midpoint, 1.0, 129, endpoint=True)
-    ])
-
[…]

### 2019-04-09 Analysis/gyro.py

@@ -1,0 +1,142 @@
+"""
+
+@author: David P Fleming, University of Washington, April 2019
+@email dflemin3 (at) uw (dot) edu
+
+Determine the age error introduced by assuming that tidally-interacting and
+tidally-locked binaries follow the same angular momentum evolution as single
+stars, i.e. by applying gyrochronology to binaries.
+
+"""
+
+import numpy as np
+import os
+import pandas as pd
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from cmap import MidpointNormalize
+from matplotlib.gridspec import GridSpec
+
+#Typical plot parameters that make for pretty plots
+mpl.rcParams['figure.figsize'] = (9,8)
+mpl.rcParams['font.size'] = 20.0
+
+## for Palatino and other serif fonts use:
+mpl.rc('font',**{'family':'serif'})
+mpl.rc('text', usetex=True)
+
+# Plot constants
+num = 16 # 1 more than the number of bins
+cmap = mpl.cm.RdBu
+cmap.set_bad('white')
+vmin = -300
+vmax = 25
+norm = MidpointNormalize(vmin=vmin, vmax=vmax, midpoint=0)
+
+# Load in data
+cpl = pd.read_csv("../Data/mcCPLMarch27.csv")
+ctl = pd.read_csv("../Data/mcCTLMarch27.csv")
+single = pd.read_csv("../Data/mcSingleMarch27.csv")
+
+# Lock times < 0 -> Not locked, set them to 7e9 (last simulation output time)
+cpl["Pri_LockTime"][cpl["Pri_LockTime"] < 0] = 7.0e9
+cpl["Sec_LockTime"][cpl["Sec_LockTime"] < 0] = 7.0e9
+ctl["Pri_LockTime"][ctl["Pri_LockTime"] < 0] = 7.0e9
+ctl["Sec_LockTime"][ctl["Sec_LockTime"] < 0] = 7.0e9
+
+# Make flag for binaries who are locked by the age of the system
+cpl["Locked"] = pd.Series(cpl["Pri_LockTime"] < cpl["Age"], index=cpl.index)
+ctl["Locked"] = pd.Series(ctl["Pri_LockTime"] < ctl["Age"], index=ctl.index)
+
+# Make flag for binaries that are strongly tidally-influence, that is, stars with
[…]

### 2019-04-09 Analysis/locking.py

@@ -124,7 +124,7 @@ color = (cpl.iloc[inds]["Pri_ProtAge"][cpl.iloc[inds]["Locked"]]/cpl.iloc[inds][
 im = ax.scatter(x, y, color="C0", marker="x", s=50, zorder=3, label="Locked")
 
 # Format plot
-ax.set_xlabel("Mass [M$_{\odot}$]", fontsize=30)
+ax.set_xlabel(r"Mass [M$_{\odot}$]", fontsize=30)
 ax.set_xlim(np.min(cpl["Pri_dMass"]), np.max(cpl["Pri_dMass"]))
 ax.set_ylabel("P$_{rot}$ [d]", fontsize=30)
 ax.set_title("CPL", fontsize=30)
@@ -176,7 +176,7 @@ y = ctl.iloc[inds]["Pri_ProtAge"][ctl.iloc[inds]["Locked"]]
 im = ax.scatter(x, y, color="C0", marker="x", s=50, zorder=3, label="Locked")
 
 # Format plot
-ax.set_xlabel("Mass [M$_{\odot}$]", fontsize=30)
+ax.set_xlabel(r"Mass [M$_{\odot}$]", fontsize=30)
 ax.set_xlim(np.min(ctl["Pri_dMass"]), np.max(ctl["Pri_dMass"]))
 ax.set_ylabel("P$_{rot}$ [d]", fontsize=30)
 ax.set_title("CTL", fontsize=30)

### 2019-04-09 Analysis/porbProtPorbHist.py

@@ -105,8 +105,8 @@ for ii in range(len(porbBinEdges)-1):
     for jj in range(len(porbProtBinEdges)-1):
         x = 0.5*(porbProtBinEdges[jj] + porbProtBinEdges[jj+1])
         if not np.isnan(cplTidalQ[ii,jj]):
-            ax1.text(x, y, ("$%.1lf^{+%0.1lf}_{-%0.1lf}$" % (cplTidalQ[ii,jj], cplTidalQUp[ii,jj], cplTidalQDown[ii,jj])),
-                     ha="center", va="center", size=14, color="white", zorder=100)
+            ax1.text(x, y, ("$%.1lf$" % cplTidalQ[ii,jj]),
+                     ha="center", va="center", size=18, color="white", zorder=100)
 
 ax1.set_xlim(0.4, 2.0)
 ax1.set_ylim(0, 100)
@@ -136,8 +136,9 @@ for ii in range(len(porbBinEdges)-1):
     for jj in range(len(porbProtBinEdges)-1):
         x = 0.5*(porbProtBinEdges[jj] + porbProtBinEdges[jj+1])
         if not np.isnan(ctlTidalTau[ii,jj]):
-            ax2.text(x, y, ("$%.1lf^{+%0.1lf}_{-%0.1lf}$" % (ctlTidalTau[ii,jj], ctlTidalTauUp[ii,jj], ctlTidalTauDown[ii,jj])),
-                     ha="center", va="center", size=12, color="white", zorder=100)
+            # Pick best font color
+            ax2.text(x, y, ("$%.1lf$" % ctlTidalTau[ii,jj]),
+                     ha="center", va="center", size=18, color="white", zorder=100)
 
 ax2.set_xlim(0.4, 2.0)
 ax2.set_ylim(0, 100)

### 2019-04-09 Analysis/protDist.py

@@ -104,9 +104,6 @@ ax0.set_ylim(0.0, 100)
 ax0.text(0.15, 90, "CPL", ha="left", va="center", size=30, color="black",
          zorder=100)
 
-ax0.text(0.13, 112, "Matt et al. (2015)", ha="left", va="center", size=30,
-         color="black", zorder=100)
-
 ### CTL Plot ###
 ax1 = fig.add_subplot(gs[16])
 

### 2019-04-09 Paper/sync.tex
Updates from Overleaf

@@
 
+\newcommand{\xxx}[1]{{\textbf{#1}}}
 \newcommand{\vplanet}[0]{\texttt{VPLanet}\xspace}
@@
 In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, qualitatively reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}.
-Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the CPL equilibrium tidal model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
+Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the \xxx{Constant Phase Lag tidal} model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
 Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
@@
 Once stars reach the main sequence, stellar radii remain mostly constant while magnetic braking removes angular momentum from the stars, gradually spinning them down over time \citep{Skumanich1972}.
-Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \citep[e.g.
-Praesepe, ][]{Matt2015,Douglas2017} and field stars \citep[e.g.
-the \kepler field, ][]{Reiners2012,Matt2015,vanSaders2018}.
-Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as gyrochronlogy \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.
+Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \textbf{\citep[e.g.
+Praesepe, ][]{Reiners2012,Matt2015,Douglas2017}} and field stars \citep[e.g.
+the \kepler field, ][]{Matt2015,vanSaders2018}.
+Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as \xxx{gyrochronology} \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.
 
@@
 We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models and two magnetic braking models.
-We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to potentially fail in such systems.
+We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of \xxx{gyrochronology} models to potentially fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
@@
 
-The first magnetic braking model we use was derived by \citet{Matt2015} and has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
+\xxx{We simulate magnetic braking using the model derived by \citet{Matt2015} as this formalism} has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
 This model is strongly dependent on the stellar Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
@@
 \end{equation}
-while for more slowly-rotating unsaturated stars,
+\xxx{while for more slowly-rotating unsaturated stars,}
 \begin{equation} \label{eqn:mattUnSat}
-\frac{dJ}{dt} = -\frac{dJ}{dt}\Bigg|_0 \left( \frac{\tau_{cz}}{\tau_{cz \odot}} \right) \left( \frac{\omega}{\omega_{\odot}} \right)^3 
+\frac{dJ}{dt} = -\frac{dJ}{dt}\Bigg|_0 \left( \frac{\tau_{cz}}{\tau_{cz \odot}} \right)^2 \left( \frac{\omega}{\omega_{\odot}} \right)^3
 \end{equation}
@@
 
-The second model we consider was developed by \citet{Reiners2012}, who modeled stellar spin-down as a function stellar magnetic field strength.
+%The second model we consider was developed by \citet{Reiners2012}, who modeled stellar spin-down as a function stellar magnetic field strength.
 Similar to the \citet{Matt2015} model, \citet{Reiners2012} posit that magnetic braking can be divided into a unsaturated ($\omega < \omega_{crit}$) and unsaturated ($\omega \geq \omega_{crit}$) regime as follows:
-\begin{equation} \label{eqn:reiners}
-\begin{split}
-\frac{dJ_{\star}}{dt} & = -C \left[ \omega \left(\frac{R^{16}}{M^2} \right)^{1/3} \right] \text{for $\omega \geq \omega_{crit}$} \\
-\frac{dJ_{\star}}{dt} & = -C \left[ \left( \frac{\omega}{\omega_{crit}} \right)^4 \omega \left(\frac{R^{16}}{M^2} \right)^{1/3} \right] \text{for $\omega < \omega_{crit}$}.
-\end{split}
-\end{equation}
-\citet{Reiners2012} tuned their model to match the Sun's P$_{rot}$ and the distribution of P$_{rot}$ of Gyr-old low-mass field stars and find $C = 2.66 \times 10^3 \text{ (gm$^5$ cm$^{-10}$ s$^3$)$^{1/3}$}$, $\omega_{crit} = 8.56 \times 10^{-6}\text{ s$^{-1}$}$ for $M > 0.35$ M$_{\odot}$, and for late M dwarfs with $M \leq 0.35 M_{\odot}$, $\omega_{crit} = 1.82 \times 10^{-6} \text{ s$^{-1}$}$.
+%\begin{equation} \label{eqn:reiners}
+%\begin{split}
+%\frac{dJ_{\star}}{dt} & = -C \left[ \omega \left(\frac{R^{16}}{M^2} \right)^{1/3} \right] \text{for $\omega \geq \omega_{crit}$} \\
+%\frac{dJ_{\star}}{dt} & = -C \left[ \left( \frac{\omega}{\omega_{crit}} \right)^4 \omega \left(\frac{R^{16}}{M^2} \right)^{1/3} \right] %\text{for $\omega < \omega_{crit}$}.
+%\end{split}
+%\end{equation}
+%\citet{Reiners2012} tuned their model to match the Sun's P$_{rot}$ and the distribution of P$_{rot}$ of Gyr-old low-mass field stars and find $C = 2.66 \times 10^3 \text{ (gm$^5$ cm$^{-10}$ s$^3$)$^{1/3}$}$, $\omega_{crit} = 8.56 \times 10^{-6}\text{ s$^{-1}$}$ for $M > 0.35$ M$_{\odot}$, and for late M dwarfs with $M \leq 0.35 M_{\odot}$, $\omega_{crit} = 1.82 \times 10^{-6} \text{ s$^{-1}$}$.
 
@@
 Once the stars reach the main sequence, their structure changes slowly, allowing magnetic braking to dominate the stellar angular momentum evolution, significantly spinning-down the stars over long timescales.
-Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution, with the \citet{Reiners2012} predicting marginally stronger torques.
 The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains fully convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.
+
+% Cut:
+% Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution, with the \citet{Reiners2012} predicting marginally stronger torques.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} (solid) and \citet{Reiners2012} (dashed) magnetic braking equations.
+   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \xxx{\citet{Matt2015} magnetic braking model}.
 Each dot denotes the approximate time when each star reaches the main sequence.}%
@@
 These parameters can also vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass stars' evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
-Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectfully \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a range of values exist in the literature.
+Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, \xxx{respectively} \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a range of values exist in the literature.
 Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s log-uniformly over $[10^4,10^8]$ and $\tau$ log-uniformly over $[10^{-2},10]$ s.
@@
 (2) from \citet{Heller2011}.
-
-We explore the impact of both the \citet{Matt2015} and \citet{Reiners2012} magnetic braking models, but default to using the \citet{Matt2015} unless noted otherwise.
-All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.
+\xxx{All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.}
 
@@
 All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}.
+
+% cut:
+% We explore the impact of both the \citet{Matt2015} and \citet{Reiners2012} magnetic braking models, but default to using the \citet{Matt2015} unless noted otherwise.
 
@@
 
-\subsubsection{Torque Balance Using the \citet{Matt2015} Magnetic Braking Model} \label{sec:mattBalance}
+\subsubsection{\xxx{Torque Balance}} \label{sec:mattBalance}
 
@@
 
-\subsubsection{Torque Balance Using the \citet{Reiners2012} Magnetic Braking Model}
-
-The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but is a generic outcome of the coupling of magnetic braking and weak tidal torques.
+% Cut as per ref's suggestion
+%\subsubsection{Torque Balance Using the \citet{Reiners2012} Magnetic Braking Model}
+
+%The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but is a generic outcome of the coupling of magnetic braking and weak tidal torques.
 The CTL model can only tidally-lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the longer P$_{orb}$ binaries into subsynchronous rotation.
@@
 
-\begin{figure*}
-	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelReiners.pdf}
-   \caption{Same format as Fig.~\ref{fig:eqPer}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
-    \label{fig:eqPerReiners}%
-\end{figure*}
+%\begin{figure*}
+%	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelReiners.pdf}
+%   \caption{Same format as Fig.~\ref{fig:eqPer}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
+%    \label{fig:eqPerReiners}%
+%\end{figure*}
 
@@
 
-We next examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, respectfully.
+We next examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, \xxx{respectively}.
 In Fig.~\ref{fig:qTauLock}, we bin our simulation results after the full 7 Gyr evolution by P$_{orb}$ and $Q$ or $\tau$ and compute the median P$_{orb}/$P$_{rot}$ in each bin, marginalizing over all other parameters.
@@
 In Fig.~\ref{fig:protDist}, we display P$_{rot}$ as a function of mass and age for binaries simulated using both the CPL and CTL model and for single stars.
-We perform this procedure for stars losing angular momentum via magnetic braking according to both the \citet{Matt2015} and \citet{Reiners2012} magnetic braking models.
-
-\subsubsection{\citet{Matt2015} Magnetic Braking} \label{sec:mattDist}
+
+% cut
+%  We perform this procedure for stars losing angular momentum via magnetic braking according to both the \citet{Matt2015} and \citet{Reiners2012} magnetic braking models.
+
+%\subsubsection{\citet{Matt2015} Magnetic Braking} \label{sec:mattDist}
 
@@
 
-\subsubsection{\citet{Reiners2012} Magnetic Braking} \label{sec:reinersDist}
-
-Similar to the \citet{Matt2015} simulations discussed above, the single star simulations display a clear age-dependence of P$_{rot}$.
+% cut as per ref's suggestion
+%\subsubsection{\citet{Reiners2012} Magnetic Braking} \label{sec:reinersDist}
+
+%Similar to the \citet{Matt2015} simulations discussed above, the single star simulations display a clear age-dependence of P$_{rot}$.
 The \citet{Reiners2012} model's magnetic braking transition at $M = 0.35$ M$_{\odot}$ produces a sharp cutoff in the P$_{rot} - M$ distribution between late M dwarfs and more massive stars.
@@
 
-\begin{figure*}[t]
-	\includegraphics[width=\textwidth]{../Plots/protDistReiners.pdf}
-   \caption{Same format as Fig.~\ref{fig:protDist}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
-    \label{fig:protDistReiners}%
-\end{figure*}
-
-The general trends found in the coupled stellar-tidal simulations using the \citet{Matt2015} model persist in the simulations evolved using the \citet{Reiners2012} model, demonstrating that these findings are not magnetic braking model-dependant, but rather a generic result of the interaction between magnetic braking and tidal torques in low-mass binaries.
+%\begin{figure*}[t]
+%	\includegraphics[width=\textwidth]{../Plots/protDistReiners.pdf}
+%   \caption{Same format as Fig.~\ref{fig:protDist}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
+%    \label{fig:protDistReiners}%
+%\end{figure*}
+
+%The general trends found in the coupled stellar-tidal simulations using the \citet{Matt2015} model persist in the simulations evolved using the \citet{Reiners2012} model, demonstrating that these findings are not magnetic braking model-dependant, but rather a generic result of the interaction between magnetic braking and tidal torques in low-mass binaries.
 Both the CPL and CTL simulations display a clear single-star magnetic braking-dominated sequence, but as before, tidally-locked binaries appear across a wide range of P$_{rot}$ and primary star masses.
@@
 
-\subsubsection{Impact of Binarity on Gyrochronology}
+%\subsubsection{Impact of Binarity on Gyrochronology}
 
 Our models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{orb} < 20 $ d.
-Under the \citet{Matt2015} model discussed in $\S$~\ref{sec:mattDist}, the median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
-Similarly for the models evolved with the \citet{Reiners2012} model discussed in $\S$~\ref{sec:reinersDist}, the median ages and $68\%$ interval for fast rotators are $2.6^{+2.8}_{-1.3}$ Gyr and $2.7^{+2.7}_{-1.4}$ Gyr according to the CPL and CTL models, respectfully, compared to single stars with ages of $1.7^{+0.9}_{-0.5}$ Gyr, We highlight this stark dichotomy between the age distribution of single and double stars in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
+\xxx{The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectively}, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
+We highlight this stark dichotomy between the age distribution of single and double stars in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
+
+% cut 
+% Similarly for the models evolved with the \citet{Reiners2012} model discussed in $\S$~\ref{sec:reinersDist}, the median ages and $68\%$ interval for fast rotators are $2.6^{+2.8}_{-1.3}$ Gyr and $2.7^{+2.7}_{-1.4}$ Gyr according to the CPL and CTL models, respectfully, compared to single stars with ages of $1.7^{+0.9}_{-0.5}$ Gyr,
 
@@
 
-Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronlogy.
+Tidal torques pose a fundamental problem for inferring ages of stars via \xxx{gyrochronology}.
 Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
@@
 Fig.~\ref{fig:protAgeHist}.
-This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, e.g.
-Fig.~\ref{fig:protDist} and Fig.~\ref{fig:protDistReiners}.
+This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, \textbf{e.g.
+Fig.~\ref{fig:protDist}}.
 
 In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another.
-Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are potentially subject to systematic errors.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.
 Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
-We caution that any application of gyrochronlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
+We caution that any application of \xxx{gyrochronology} methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
 
@@
 
+% cut
 % In binaries, however, tidal torques can lock stellar P$_{rot}$ to P$_{eq}$, especially at short P$_{orb}$ where tidal torques are stronger, creating a long-lasting population of fast rotators with a wide age range.
 % This feature is not due to long-term spin-down via magnetic braking as this population lies above the old edge of single star sequence. Instead 
+% and Fig.~\ref{fig:protDistReiners}.
 
@@
 In tidally-interacting binaries, gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails.
-We caution that any application of gyrochronlogy methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
+We caution that any application of \xxx{gyrochronology} methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
 

### 2019-04-09 Analysis/locking.py
updating locking cpl and ctl plots

@@ -62,6 +62,7 @@ mpl.rc('font',**{'family':'serif'})
 mpl.rc('text', usetex=True)
 
 ylims = [0, 100]
+xlims = [0.1, 1]
 bins = "auto"
 num = 2500
 seed = 42
@@ -100,11 +101,13 @@ ctl["Free"] = pd.Series(mask, index=ctl.index)
 
 ### CPL ###
 
-fig = plt.figure(figsize=(10,9))
-gs = GridSpec(4,4, wspace=0.45)
+fig = plt.figure(figsize=(9,8.5))
 
-ax = fig.add_subplot(gs[1:4,0:3])
-ax_marg_y = fig.add_subplot(gs[1:4,3])
+gs = GridSpec(3, 3, width_ratios=[1, 0.01, 0.25], wspace=0.05,
+              height_ratios=[0.25, 0.01, 1], hspace=0.05)
+
+# Main panel
+ax = fig.add_subplot(gs[2, 0])
 
 # Not tidally locked binaries: sub and super sync populations
 x = cpl.iloc[inds]["Pri_dMass"][cpl.iloc[inds]["Free"]]
@@ -127,12 +130,13 @@ im = ax.scatter(x, y, color="C0", marker="x", s=50, zorder=3, label="Locked")
 ax.set_xlabel(r"Mass [M$_{\odot}$]", fontsize=30)
 ax.set_xlim(np.min(cpl["Pri_dMass"]), np.max(cpl["Pri_dMass"]))
 ax.set_ylabel("P$_{rot}$ [d]", fontsize=30)
-ax.set_title("CPL", fontsize=30)
 ax.set_ylim(ylims)
 ax.set_rasterization_zorder(0)
 leg = ax.legend(loc="upper left", framealpha=0.7, fontsize=18)
 
-# Plot marginals
+# Plot marginals: y for prot
+ax_marg_y = fig.add_subplot(gs[2, 2])
+
 ax_marg_y.hist(cpl["Pri_ProtAge"][cpl["Free"]], orientation="horizontal",
                bins=bins, histtype="step", lw=3, color="C1",
                label="Not Locked", density=False)
@@ -148,43 +152,67 @@ ax_marg_y.set_ylim(ylims)
 ax_marg_y.legend(loc="upper center", framealpha=0.5, fontsize=12)
 plt.setp(ax_marg_y.get_yticklabels(), visible=False);
 
[…]

### 2019-04-09 Paper/sync.tex
Updates from Overleaf

@@
 
-We improve upon the stellar evolution model used in \citet{Fleming2018}, \stellar, by tracking the evolution of both the stellar radius of gyration, $r_g$, and radius, $R$, using a bicubic interpolation over mass and time of the \citet{Baraffe2015} stellar evolution models.
-We consider two separate magnetic braking models to gauge how sensitive the interaction between tides and magnetic braking is in modifying P$_{rot}$ evolution in stellar binaries.
+We improve upon the \xxx{interpolation of the \citet{Baraffe2015} stellar evolution models employed by \citet{Fleming2018}, \stellar, by additionally performing a bicubic interpolation the stellar radius of gyration, $r_g$, over mass and time of the \citet{Baraffe2015} models.
+This updated version of \stellar now tracks the full moment of interia evoltuion of low-mass stars according to the \citet{Baraffe2015} stellar evolution models, a critical requirement for modeling the angular momentum evolution of low-mass stars \citep[e.g.][]{Matt2015}.} We consider two separate magnetic braking models to gauge how sensitive the interaction between tides and magnetic braking is in modifying P$_{rot}$ evolution in stellar binaries.
 
@@
 
-In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar, using both magnetic braking models.
+In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to \xxx{\stellar and the \citet{Matt2015} magnetic braking model}.
 We assume all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr.
@@
 Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$], consistent with eccentricities of field binaries that likely have not been tidally-processed \citep{Raghavan2010}.
-Although the CTL model is applicable for $e \gsim 0.3$, the CPL model is not, so we restrict $e \leq 0.3$ to allow us to compare both models.
+Although the CTL model is applicable for $e \gsim 0.3$, \xxx{the CPL model is not and can predict qualitatively incorrect evolution in that regime \citep[see Section 4.1 in][]{Leconte2010}}, so we restrict $e \leq 0.3$ to allow us to compare both models.
 We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.

### 2019-04-22 Analysis/porbProtPorbHist.py
updating plots as per ref's comments

@@ -13,7 +13,6 @@ import pandas as pd
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 from matplotlib.gridspec import GridSpec
-from cmap import shiftedColorMap
 
 #Typical plot parameters that make for pretty plots
 mpl.rcParams['font.size'] = 25.0
@@ -105,8 +104,8 @@ for ii in range(len(porbBinEdges)-1):
     for jj in range(len(porbProtBinEdges)-1):
         x = 0.5*(porbProtBinEdges[jj] + porbProtBinEdges[jj+1])
         if not np.isnan(cplTidalQ[ii,jj]):
-            ax1.text(x, y, ("$%.1lf$" % cplTidalQ[ii,jj]),
-                     ha="center", va="center", size=18, color="white", zorder=100)
+            txt = ax1.text(x, y, ("$%.1lf$" % cplTidalQ[ii,jj]),
+                           ha="center", va="center", size=20, color="white", zorder=100)
 
 ax1.set_xlim(0.4, 2.0)
 ax1.set_ylim(0, 100)
@@ -137,8 +136,8 @@ for ii in range(len(porbBinEdges)-1):
         x = 0.5*(porbProtBinEdges[jj] + porbProtBinEdges[jj+1])
         if not np.isnan(ctlTidalTau[ii,jj]):
             # Pick best font color
-            ax2.text(x, y, ("$%.1lf$" % ctlTidalTau[ii,jj]),
-                     ha="center", va="center", size=18, color="white", zorder=100)
+            txt = ax2.text(x, y, ("$%.1lf$" % ctlTidalTau[ii,jj]),
+                           ha="center", va="center", size=20, color="w", zorder=100)
 
 ax2.set_xlim(0.4, 2.0)
 ax2.set_ylim(0, 100)

### 2019-04-22 Paper/sync.tex

@@
 
+\subsubsection{\xxx{Core-Envelope Coupling}} \label{sec:methods:twoLayer}
+
+\xxx{Our simplified model assumes that stars follow solid body rotation, whereas in low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution modeled in this work \citep{MacGregor1991,Allain1998}.
+Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior, i.e.
+into the envelope, for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass \citep{Irwin2007,Bouvier2008,Irwin2009}.
+This core-to-envelope angular momentum transfer can counteract the spin-down caused by magnetic braking on the main sequence.
+For a given magnetic braking model, this angular momentum transfer within the stellar interior would likely lead to slightly faster stellar rotation on the early main sequence and a wider spread of rotation periods on the zero-age main sequence than is predicted by solid body rotation models \citep{Gallet2013}.
+This effect could be more pronounced for low-mass M-dwarfs, potentially accounting for their long spin-down timescales \citep{Gallet2015}.
+Modeling core-envelope coupling is beyond the scope of this work, however.}
+
+\subsubsection{\xxx{Example Stellar Evolution}} \label{sec:methods:exampleStellar}
+
 In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to \xxx{\stellar and the \citet{Matt2015} magnetic braking model}.
@@
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \xxx{\citet{Matt2015} magnetic braking model}.
+   \caption{\xxx{Updated Figure} Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \xxx{\citet{Matt2015} magnetic braking model}.
 Each dot denotes the approximate time when each star reaches the main sequence.}%
@@
 Any constraints we derive as a function $Q$ or $\tau$ can trivially be scaled to other values of $k_2$.
+\xxx{For example, a common re-parameterization of Q is the reduced tidal quality factor, $Q' = 3Q/2k_2$ \citep[e.g.][]{Leconte2010}.
+Given our choice of $k_2 = 0.5$, this reduces to $Q' = 3 Q$.}
 
@@
 We find that this scheme produces physically and numerically accurate results.
+
+\subsubsection{\xxx{The Dynamical Tide}} \label{sec:methods:dynamicalTide}
+
+\xxx{An additional mechanism for tidal dissipation in low-mass stellar binaries not considered in this work is the dynamical tide.
+Dynamical tidal dissipation arises from the turbulent viscous damping of inertial waves that are excited in the stellar convective envelope by a tidal perturber, with Coriolis acceleration serving as the restoring force \citep{Zahn1975, Ogilvie2007}.
+Under the dynamical tide formalism, the stellar mass, evolving stellar structure, rotation rate, and tidal forcing frequency can all strongly impact the strength of tidal dissipation, which can span many orders of magnitude \citep{Ogilvie2007,Ogilvie2013,Mathis2015,Gallet2017}.
+For example, adopting the tidal frequency-averaged model for tidal dissipation of \citet{Ogilvie2013}, both \citet{Mathis2015} and \citet{Gallet2017} show that dynamical tidal dissipation is enhanced on the pre-main sequence via the expansion of the stellar radiative core and rapid rotation, whereas the magnetic braking-driven spin-down on the main sequence decreases the tidal dissipation.
+Dissipation due to the dynamical tide could be important for some of the systems considered in this work since, for coplanar binary stars on circular orbits, inertial waves are excited in the stellar convective envelopes for P$_{orb} > $P$_{rot}/2$ and can drive significant spin and orbital evolution \citep[e.g.][]{Witte2002,Ogilvie2007,Bolmont2016}.
+Although semi-analytic models for dynamical tidal dissipation that account for the evolving stellar structure and rotation exist \citep[e.g.][]{Ogilvie2013,Mathis2015,Bolmont2016,Gallet2017}, we do not consider them here as they are currently limited to circular orbits.
+We instead focus on exploring the impact of the equilibrium tide across a wide range of parameter space and leave an examination of how the combination of the dynamical and equilibrium tide impacts the rotation period evolution of low-mass binary stars for future work.}
 
@@
 	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
-   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model.
+   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model.
 Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d.
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
-   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
+   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
 Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
@@
 	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
-   \caption{Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
+   \caption{\xxx{Updated Figure} Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
 For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, regardless of $Q$, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.
@@
 	\includegraphics[width=\textwidth]{../Plots/porbProtPorbQHist.pdf}
-   \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.
+   \caption{\xxx{Updated Figure} Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.
 In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$ and display these values.
@@
 	\includegraphics[width=\textwidth]{../Plots/porbProtPorbTauHist.pdf}
-   \caption{Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
+   \caption{\xxx{Updated Figure} Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
     \label{fig:taumap}%
@@
 	\includegraphics[width=0.5\textwidth]{../Plots/lockedCPL.pdf}
-   \caption{Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
+   \caption{\xxx{Updated Figure} Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
 Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
@@
 Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.6$ d and $7.4$ d) than unlocked (median CPL, CTL P$_{rot} = 41.3$ d and $40.8$ d) binaries.
-More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale as $R^5$, and $R$ increases with stellar mass.
+More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.
 This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model.
+We highlight this enhanced density of locked binaries in systems with more massive primaries in the marginalized mass distributions in the top panels of Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.}
 
@@
 	\includegraphics[width=0.5\textwidth]{../Plots/lockedCTL.pdf}
-   \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
+   \caption{\xxx{Updated Figure} Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
     \label{fig:lockedCTL}%
@@
 
+\begin{figure*}[t]
+	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
+   \caption{\xxx{Updated Figure} P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model.
+For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
+Right: The P$_{rot}$ distribution for each case, marginalized over stellar mass.}%
+    \label{fig:protDist}%
+\end{figure*}
+
+\xxx{There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
+In stark contrast, both tidal models predict that age does not always strongly correlate with P$_{rot}$ as binaries at a given P$_{rot}$ can assume a range of ages, especially at short and long P$_{rot}$, due to the influence of tidal torques.
+In binaries, tidal torques tend to drive the P$_{rot}$ evolution away from that of single stars and towards P$_{eq}$, either maintaining rapid rotation in tidally-locked short P$_{orb}$ systems, or working with magnetic braking to slow P$_{rot}$ beyond that of single stars of the same age.}
+
+\xxx{We quantify the impact of binarity on gyrochronology in Fig.~\ref{fig:gyro} by computing the percent difference between the mean ages of single and binary stars, for both tidal models, in mass and P$_{rot}$ bins for the populations depicted in Fig.~\ref{fig:protDist}.
+This quantity, referred to here as the ``Relative Age Error", represents the systematic error incurred by assigning a binary star the age expected for single stars at a given mass and P$_{rot}$.
+For this comparison, we select the subset of tidally-locked and tidally-interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, accounting for $31\%$ and $24\%$ of CPL and CTL binaries, respectively, as the remaining binaries weakly tidally-interact, and hence their P$_{rot}$ evolution will predominately follow that of single stars.}
+
+\begin{figure*}[t]
+	\includegraphics[width=\textwidth]{../Plots/gyro.pdf}
+   \caption{\xxx{New Figure: Relative age error between single and binary stars for both the CPL (left) and CTL (right) tidal models.
+The error is computed as the percent difference between the mean ages of single and binary stars in mass and P$_{rot}$ bins.}}%
+    \label{fig:gyro}%
+\end{figure*}
+
+\xxx{Gyrochronology ages, when applied to tidally-interacting binaries systematically underestimate the ages, with the age error increasing with decreasing P$_{rot}$ and increasing with primary star mass where tides are stronger.
+For binaries with P$_{rot} \lsim 30 d$, gyrochronology ages are underestimated by $50\%$, with this error growing to $300\%$ for P$_{rot} \approx 10$ d.
+For slow rotators with P$_{rot} \gsim 40$ d, gyrochronology ages are slightly overestimated by up to $25\%$ for near solar-mass stars.
+Single star-only models fail to produce near-solar mass stars with P$_{rot} \lsim 10$ d nor P$_{rot} \gsim 50$ as only tidally-interacting binaries at short and long P$_{orb}$, respectively, can produce such rotation states, so we can not compare binary star and single star ages there.}
+
+\xxx{...
+This effect is most pronounced at short P$_{orb}$, and hence P$_{rot}$, where tidal torques in binaries are strongest, leading to relative age errors of up to $300\%$ for both tidal models relative to single stars.
+This effect would be more pronounced for the most rapidly-rotating stars, e.g.
+tidally-interacting near-solar mass binaries with P$_{rot} \lsim 10$ d, however our single star model and age prior cannot produce such rapid rotators as only those with ages $\lsim 1$ Gyr could have P$_{rot} \lsim 10$ d.
+In Fig.~\ref{fig:protAgeHist}, we further examine the age distributions of rapid rotators.}
+
+\xxx{The age distribution of tidally-interacting binaries with P$_{rot} < 20 $ d is inconsistent with that of single stars.
+In this range, the median ages and $68\%$ interval are $2.3^{+2.9}_{-0.9}$ Gyr and $2.4^{+3.0}_{-1.1}$ Gyr according to the CPL and CTL models, respectively, compared to the much younger single stars with ages of $1.6^{+0.8}_{-0.4}$ Gyr.
+We highlight this dichotomy in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.}
+
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
+   \caption{\xxx{Updated Figure} Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
+Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
+    \label{fig:protAgeHist}%
+\end{figure}
+
+Tidal torques pose a fundamental problem for inferring ages of stars via \xxx{gyrochronology}.
+Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
+For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages \xxx{$\lsim 1.6$} Gyr.
+If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs, e.g.
+Fig.~\ref{fig:protAgeHist}.
+This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, \xxx{e.g.
+Fig.~\ref{fig:protDist} and Fig.~\ref{fig:gyro}}.
+
+In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.
+Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field \xxx{are consistent with tidally-synchronized photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread}.
+We caution that any application of \xxx{gyrochronology} methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
+
+Furthermore, any study of the angular momentum evolution of a stellar population will be flawed unless one robustly accounts for binarity.
+Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
+Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
+We recommend that the application, or calibration, magnetic braking models to a sample of stellar rotation periods control for binarity.
+
+% cut
+% In binaries, however, tidal torques can lock stellar P$_{rot}$ to P$_{eq}$, especially at short P$_{orb}$ where tidal torques are stronger, creating a long-lasting population of fast rotators with a wide age range.
+% This feature is not due to long-term spin-down via magnetic braking as this population lies above the old edge of single star sequence. Instead 
+% and Fig.~\ref{fig:protDistReiners}.
+
 % cut
@@
 
-\begin{figure*}[t]
-	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
-   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model.
-For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
-Right: The P$_{rot}$ distribution for each case, marginalized over stellar mass.}%
-    \label{fig:protDist}%
-\end{figure*}
-
-There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
+% cut
+% There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
 The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) that have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.
@@
 
-%\subsubsection{Impact of Binarity on Gyrochronology}
-
-Our models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{orb} < 20 $ d.
-\xxx{The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectively}, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr.
-We highlight this stark dichotomy between the age distribution of single and double stars in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
-
 % cut 
@@
 
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
-   \caption{Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
-Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
-    \label{fig:protAgeHist}%
-\end{figure}
-
-Tidal torques pose a fundamental problem for inferring ages of stars via \xxx{gyrochronology}.
-Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
-For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr.
-If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs, e.g.
-Fig.~\ref{fig:protAgeHist}.
-This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, \textbf{e.g.
-Fig.~\ref{fig:protDist}}.
-
-In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another.
-Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.
-Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
-We caution that any application of \xxx{gyrochronology} methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
-
-Furthermore, any study of the angular momentum evolution of a stellar population will be flawed unless one robustly accounts for binarity.
-Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
-Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
-We recommend that the application, or calibration, magnetic braking models to a sample of stellar rotation periods control for binarity.
-
-% cut
-% In binaries, however, tidal torques can lock stellar P$_{rot}$ to P$_{eq}$, especially at short P$_{orb}$ where tidal torques are stronger, creating a long-lasting population of fast rotators with a wide age range.
-% This feature is not due to long-term spin-down via magnetic braking as this population lies above the old edge of single star sequence. Instead 
-% and Fig.~\ref{fig:protDistReiners}.
+%\subsubsection{Impact of Binarity on Gyrochronology}
 
@@
 We compare our results to the P$_{1,min}$ P$_{rot}$ values reported by \citet{Lurie2017} as the authors demonstrated that these values are likely to be close to the equatorial P$_{rot}$ that we track in our simulations.
-In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}, and the \citet{Lurie2017} data.
+In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
-   \caption{P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ according to the CPL model (left) and the CTL model (middle), and \citet{Lurie2017} \kepler EB observations (right).
+   \caption{\xxx{Updated Figure} P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ according to the CPL model (left) and the CTL model (middle), and \citet{Lurie2017} \kepler EB observations (right).
 All points are colored by $e$.
@@
 the lack of long P$_{orb}$ binaries, inhibit our ability to robustly compare our predictions with the data.
-We discuss this point further in $\S$~\ref{sec:inference} and offer observational tests to discriminate between which of the two models better describes tidal interactions in low-mass binaries in $\S$~\ref{sec:whichModel}.
-
-% extra stuff
+
+% extra stuff/cuts
 % Second, In principle, we could correct for such a bias in the \citet{Lurie2017} data and condition our model predictions on the data to infer tidal properties, but that analysis is beyond the scope of this work for reasons we discuss in $\S$~\ref{sec:inference}.
@@
 In comparison, the CPL model (not pictured) predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, producing a population of synchronized binaries similar to what is seen in the \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
+%  We discuss this point further in $\S$~\ref{sec:inference} and offer observational tests to discriminate between which of the two models better describes tidal interactions in low-mass binaries in $\S$~\ref{sec:whichModel}.
 
@@
 Here, we outline three observational tests that can discriminate between the two models.
-The first two require obtaining a series of time-resolved spectra over full orbital phase of short P$_{orb}$ supersynchronous binaries, i.e.
-those identified by \citet{Lurie2017}, to compute radial velocities over the binary's orbit.
-Combined with \kepler photometry, one can derive much more robust $e$ constraints than those computed using the transit approximations in \citet{Lurie2017}.
-
 The first test considers binaries with P$_{orb} < 10$ d that are likely tidally-locked on eccentric orbits, but with $e < 0.23$.
@@
 If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model.
-It is unclear, however, if low-mass binary stars can even lock into spin-orbit resonances, such as those discussed in \citet{Rodriguez2012}, as stellar convective envelopes lack a fixed shape \citep{Burkart2014,Lurie2017}.
-There is no obvious clustering of stars near any spin-orbit resonance, aside from 1:1 spin-orbit synchronization in the \citet{Lurie2017} \kepler EB data.
-The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making any distinction between models conclusive.
-More robust phase-resolved spectroscopic $e$ constrains would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a Gaussian process model as was done by \citet{Angus2018}.
-
 These two tests can fail to discriminate between the CPL and CTL model, however, if the CPL model P$_{eq}$ is a continuous function of $e$, e.g.
@@
 Since there is no obvious clustering of \kepler EBs near any spin-orbit resonance, the above two tests are likely insufficient.
-We offer a third observational test that provides the best method to discriminate between models.
-
-As shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 80$ d would provide strong evidence in favor of the CPL model, as the CTL model is unlikely to tidally-lock binaries past P$_{orb} \approx 20$ d and cannot tidally-lock stars beyond P$_{orb} \approx 80$ d, even with extreme $\tau$, e.g.
+
+ A third test, the detection of tidally-locked binaries \xxx{with solar-mass primaries and P$_{rot} \gsim 60$ d, would provide strong evidence in favor of the CPL model as the CTL model cannot tidally-lock stars beyond P$_{orb} \approx 60$ d, regardless of $\tau$, e.g.
 Fig.~\ref{fig:lockedCTL}.
-The CPL model, however, readily tidally-locks binaries out to P$_{orb} \approx 100$ d.
-We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small star spot modulation amplitudes \citep{McQuillan2014,Lurie2017,Reinhold2018}.
-Long term spectroscopic monitoring may be warranted in such cases.
+The CPL model, however, can tidally-lock binaries out to P$_{orb} \gsim 90$ d.
+We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100 $ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small star spot modulation amplitudes \citep{McQuillan2014,Lurie2017,Reinhold2018}.
+Long term spectroscopic monitoring may be warranted in such cases.}
+
+% cut
+% The first two require obtaining a series of time-resolved spectra over full orbital phase of short P$_{orb}$ supersynchronous binaries, i.e.
+those identified by \citet{Lurie2017}, to compute radial velocities over the binary's orbit.
+Combined with \kepler photometry, one can derive much more robust $e$ constraints than those computed using the transit approximations in \citet{Lurie2017}.
+% It is unclear, however, if low-mass binary stars can even lock into spin-orbit resonances, such as those discussed in \citet{Rodriguez2012}, as stellar convective envelopes lack a fixed shape \citep{Burkart2014,Lurie2017}.
+There is no obvious clustering of stars near any spin-orbit resonance, aside from 1:1 spin-orbit synchronization in the \citet{Lurie2017} \kepler EB data.
+The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making any distinction between models conclusive.
+More robust phase-resolved spectroscopic $e$ constrains would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a Gaussian process model as was done by \citet{Angus2018}.
+% We offer a third observational test that provides the best method to discriminate between models.
 
@@
 
-\subsection{Inferring Stellar and Binary Properties from \kepler} \label{sec:inference}
-
-In principle, we could have conditioned our model on the \citet{Lurie2017} data to infer population-level properties of stellar binaries, such as tidal $Q$ and $\tau$ values for stars, perhaps as a function of stellar mass, or the initial binary P$_{orb}$ distribution, using hierarchical probabilistic modeling instead of qualitatively comparing of our simulations with the data.
+%\subsection{Inferring Stellar and Binary Properties from \kepler} \label{sec:inference}
+
+%In principle, we could have conditioned our model on the \citet{Lurie2017} data to infer population-level properties of stellar binaries, such as tidal $Q$ and $\tau$ values for stars, perhaps as a function of stellar mass, or the initial binary P$_{orb}$ distribution, using hierarchical probabilistic modeling instead of qualitatively comparing of our simulations with the data.
 Such an analysis is beyond the scope of this work, however.
@@
 
-Instead, one could use our model to infer tidal properties for individual systems with well-constrained properties.
+%Instead, one could use our model to infer tidal properties for individual systems with well-constrained properties.
 For example, Kepler-47 \citep{Orosz2012} is a low-mass binary whose primary star rotates subsynchronously.
@@
 In this work, we probed the long-term angular momentum evolution of low-mass stellar binaries, with a focus on P$_{rot}$ in short and intermediate P$_{orb}$ binaries.
-We considered the impact of two common equilibrium tidal models, two magnetic braking models, and stellar evolution.
+We considered the impact of two common equilibrium tidal models, \xxx{magnetic braking}, and stellar evolution.
 We performed a large suite of simulations for binaries with physically-motivated initial conditions out to P$_{orb} = 100$ and across a wide range of tidal dissipation parameters to examine the competition between tidal torques and magnetic braking for controlling the stellar P$_{rot}$ evolution.
@@
 The lack of uncertainties on P$_{rot}$, the approximate orbital eccentricities derived by \citet{Lurie2017}, and unconstrained completeness estimates prevent us from discriminating between which tidal model best describes tidal torques in low-mass binaries and from inferring tidal properties of low-mass stars given the \kepler EB data.
-We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
-To perform these tests, we recommend that observers make additional measurements of $e$ and P$_{rot}$ in binaries, especially at long P$_{orb}$ out to 100 d, although we note that these are non-trivial observations.
+
+\xxx{We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
+We primarily suggest that observers measure stellar P$_{rot}$ in binaries with solar-mass primaries at P$_{orb} 60-100$ d.
+If any tidally-locked binaries are identified at long P$_{orb}$, this would be evidence in favor of the CPL model as we found that only binaries tidally-interacting via the CPL model could tidally-lock at such long P$_{orb}$.
+At shorter P$_{orb}$, precise measurements of P$_{rot}$ and binary $e$ and P$_{orb}$ could distinguish between the CPL and CTL model in tidally-locked systems, e.g.
+identifying if P$_{eq}$ follows Eq.~(\ref{eqn:cpl:eqPer}) vs.
+Eq.~(\ref{eqn:ctl:eqPer}), especially if the CPL P$_{eq}$ is in fact a discrete function of $e$.
+The observations required by these tests, however, are non-trivial.
+Beyond these tests, our model could be used to infer the tidal properties of binary stars, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties and reasonable prior probability distributions for parameters like the initial binary $e$.
+This analysis, however, is also beyond the scope of this work and we leave it for future endeavors.}
 
@@
 \acknowledgments
-This work was facilitated though the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system and funded by the Student Technology Fund at the University of Washington.
+\xxx{We thank the anonymous reviewer for helpful comments that improved the quality of this manuscript.} This work was facilitated though the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system and funded by the Student Technology Fund at the University of Washington.
 DPF was supported by NASA Headquarters under the NASA Earth and Space Science Fellowship Program - Grant 80NSSC17K0482.
@@
 %% SOFTWARE %%
-%\software{matplotlib: \citet{Hunter2007}, numpy: \citet{vanderWalt2011}, pandas: \citet{Mckinney2010}}
+\software{matplotlib: \citet{Hunter2007}, numpy: \citet{vanderWalt2011}, pandas: \citet{Mckinney2010}, \vplanet: \citet{Barnes2016,vplanet2018}}
 
@@
 \end{equation}
-where P$_{rot} = 2 \pi / \omega$ and $C_{MB} = 6.3 \times 10^{30}$ ergs \citep{Matt2015,Matt2019}.
+where P$_{rot} = 2 \pi / \omega$ and \xxx{$C_{MB} = 6.3 \times 10^{30}$ ergs \citep{Matt2015,Matt2019}}.
 

### 2019-04-25 Analysis/analyticTorqueBalance.py
updating figures

@@ -63,8 +63,8 @@ def ctlMatt2015(Porb, logK2Tau):
 # end function
 
 nbins = 50
-logK2Taus = np.log10(0.5 * np.logspace(-1.8, 0.3, nbins))[::-1]
-Porbs = np.linspace(20, 80, nbins)
+logK2Taus = np.log10(0.5 * np.logspace(-1.8, -0.3, nbins))[::-1]
+Porbs = np.linspace(10, 80, nbins)
 res = np.zeros((nbins, nbins)) # = Prot/Peq
 
 for ii in range(nbins):
@@ -76,8 +76,12 @@ for ii in range(nbins):
 fig, ax = plt.subplots()
 
 extent = [logK2Taus[0], logK2Taus[-1], Porbs[0], Porbs[-1]]
-im = plt.imshow(res, cmap="viridis", aspect="auto", interpolation="nearest",
-                origin="lower", extent=extent, vmin=1, vmax=1.5)
+im = ax.imshow(res, cmap="viridis", aspect="auto", interpolation="nearest",
+               origin="lower", extent=extent, vmin=1)
+cont = ax.contour(res, colors="white", levels=[1.02, 1.05, 1.1, 1.2, 1.4, 1.6],
+                  extent=extent, origin="lower", extend="both")
+plt.clabel(cont, inline=True, fontsize=12, fmt="%.02lf")
+
 
 ax.set_xlabel(r"log$_{10}$($k_2 \tau$)")
 ax.set_ylabel("Orbital Period [d]")

### 2019-04-25 Analysis/eqPerMatt.py

@@ -169,7 +169,7 @@ for ii in range(len(ctlDirs)):
 ax.plot([100], [100], ls="-", lw=3, color="grey", label="CPL")
 ax.plot([100], [100], ls="--", lw=3, color="grey", label="CTL")
 
-ax.legend(loc="lower right", framealpha=0.0, fontsize=17)
+ax.legend(loc="upper left", framealpha=0.0, fontsize=15)
 ax.set_ylabel("P$_{rot}$/P$_{eq}$", fontsize=30)
 ax.set_xlabel("Time [yr]", fontsize=30)
 ax.set_xscale("log")

### 2019-04-25 Paper/sync.tex
Updates from Overleaf

@@
 Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
-gyrochronology methods can fail unless one accounts for binarity.
+gyrochronology \xxx{ages can err by up to $300\%$} unless one accounts for binarity.
 We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
@@
 Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
-Right: Marginalized P$_{rot}$ distribution for each case.}%
+Right: Marginalized P$_{rot}$ distribution for each case.
+\xxx{Top: Marginalized mass distributions.}}%
     \label{fig:lockedCPL}%
@@
 
-We compare the P$_{rot}$ distribution for tidally-interacting stellar binaries from our CPL and CTL simulations with the P$_{rot}$ distribution of single stars subject to stellar evolution and magnetic braking to gauge the impact of tidal torques on driving P$_{rot}$ distributions away from that of single stars.
+We compare the \xxx{P$_{rot}$ and age distributions of tidally-interacting stellar binaries from our CPL and CTL simulations with that of single stars to gauge the impact of tidal torques on driving P$_{rot}$ distributions away from that of single stars and what implications that may have for estimating stellar ages using gyrochronology}.
 We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar} with initial conditions sampled from the same mass and P$_{rot}$ distributions used for the binary simulations described $\S$~\ref{sec:methods}.
@@
 
-\xxx{There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
-In stark contrast, both tidal models predict that age does not always strongly correlate with P$_{rot}$ as binaries at a given P$_{rot}$ can assume a range of ages, especially at short and long P$_{rot}$, due to the influence of tidal torques.
-In binaries, tidal torques tend to drive the P$_{rot}$ evolution away from that of single stars and towards P$_{eq}$, either maintaining rapid rotation in tidally-locked short P$_{orb}$ systems, or working with magnetic braking to slow P$_{rot}$ beyond that of single stars of the same age.}
-
-\xxx{We quantify the impact of binarity on gyrochronology in Fig.~\ref{fig:gyro} by computing the percent difference between the mean ages of single and binary stars, for both tidal models, in mass and P$_{rot}$ bins for the populations depicted in Fig.~\ref{fig:protDist}.
-This quantity, referred to here as the ``Relative Age Error", represents the systematic error incurred by assigning a binary star the age expected for single stars at a given mass and P$_{rot}$.
-For this comparison, we select the subset of tidally-locked and tidally-interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, accounting for $31\%$ and $24\%$ of CPL and CTL binaries, respectively, as the remaining binaries weakly tidally-interact, and hence their P$_{rot}$ evolution will predominately follow that of single stars.}
+\xxx{In binaries, tidal torques tend to drive the P$_{rot}$ evolution away from that of single stars and towards P$_{eq}$, either maintaining rapid rotation in tidally-locked short P$_{orb}$ systems, or working with magnetic braking to slow P$_{rot}$ beyond that of single stars of the same age.
+The impact of tidal torques on the binary P$_{rot}$ distribution is clear: strong tidal torques in short P$_{orb}$ binaries produce a substantial population of rapid rotators with P$_{rot} \lsim 20$ d, with the influence of tides extending to longer P$_{orb}$ systems as demonstrated in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.
+Except for stars with ages $\lsim 1$ Gyr, or young late M-dwarfs who are either contracting along the pre-main sequence or have just reached the main sequence, our single star simulations fail to produce a population of rapid rotators.
+This theoretical result is consistent with \citet{Simonian2018} who find that the population of \kepler stars with P$_{rot} < 7.5$ d is likely dominated by tidally-interacting binaries.}
+
+\xxx{In the single star population, there is a clear monotonic age-dependence of P$_{rot}$, with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
+This trend is a generic outcome of magnetic braking and is not specific to our choice of magnetic braking model.
+In stark contrast, both tidal models predict that age does not always strongly correlate with P$_{rot}$ as tidally-interacting binaries, at a given primary star mass and P$_{rot}$, can assume a wide range of ages, especially for P$_{rot} \lsim 20$ d.}
+
+\xxx{We quantify the impact of binarity on gyrochronology age estimates in Fig.~\ref{fig:gyro} by computing the percent difference between the mean ages of single and binary stars, for both tidal models, in mass and P$_{rot}$ bins for the populations depicted in Fig.~\ref{fig:protDist}.
+This quantity, referred to here as the ``Relative Age Error", represents the systematic error incurred by assigning a tidally-interacting binary star the age expected for single stars at a given mass and P$_{rot}$.
+For this comparison, we select the subset of tidally-locked and tidally-interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, accounting for $31\%$ and $24\%$ of CPL and CTL binaries, respectively.}
 
@@
    \caption{\xxx{New Figure: Relative age error between single and binary stars for both the CPL (left) and CTL (right) tidal models.
-The error is computed as the percent difference between the mean ages of single and binary stars in mass and P$_{rot}$ bins.}}%
+The error is computed as the percent difference between the mean ages of single and tidally-interacting binary stars in mass and P$_{rot}$ bins.}}%
     \label{fig:gyro}%
@@
 
-\xxx{Gyrochronology ages, when applied to tidally-interacting binaries systematically underestimate the ages, with the age error increasing with decreasing P$_{rot}$ and increasing with primary star mass where tides are stronger.
-For binaries with P$_{rot} \lsim 30 d$, gyrochronology ages are underestimated by $50\%$, with this error growing to $300\%$ for P$_{rot} \approx 10$ d.
-For slow rotators with P$_{rot} \gsim 40$ d, gyrochronology ages are slightly overestimated by up to $25\%$ for near solar-mass stars.
-Single star-only models fail to produce near-solar mass stars with P$_{rot} \lsim 10$ d nor P$_{rot} \gsim 50$ as only tidally-interacting binaries at short and long P$_{orb}$, respectively, can produce such rotation states, so we can not compare binary star and single star ages there.}
-
-\xxx{...
-This effect is most pronounced at short P$_{orb}$, and hence P$_{rot}$, where tidal torques in binaries are strongest, leading to relative age errors of up to $300\%$ for both tidal models relative to single stars.
-This effect would be more pronounced for the most rapidly-rotating stars, e.g.
-tidally-interacting near-solar mass binaries with P$_{rot} \lsim 10$ d, however our single star model and age prior cannot produce such rapid rotators as only those with ages $\lsim 1$ Gyr could have P$_{rot} \lsim 10$ d.
-In Fig.~\ref{fig:protAgeHist}, we further examine the age distributions of rapid rotators.}
-
-\xxx{The age distribution of tidally-interacting binaries with P$_{rot} < 20 $ d is inconsistent with that of single stars.
+\xxx{For most values of P$_{rot}$, gyrochronology methods systematically underestimate the ages of tidally-interacting binaries, with the relative age error increasing with decreasing P$_{rot}$ and increasing with primary star mass, where tides are stronger.
+For binaries with P$_{rot} \lsim 30$ d, gyrochronology ages are underestimated by $50\%$, with this error growing to $300\%$ for P$_{rot} \approx 10$ d.
+For slow rotators with P$_{rot} \gsim 40$ d, gyrochronology ages are slightly overestimated by up to $25\%$, with the largest errors occurring for near solar-mass primary stars where tides and magnetic braking combine to spin down binary stars.
+The relative age errors would be more pronounced for the most rapidly-rotating stars, e.g.
+tidally-interacting near-solar mass binaries with P$_{rot} \lsim 10$ d, however, our single star-only models fail to produce such rotation states.}
+
+\xxx{The age distribution of binaries with P$_{rot} < 20$ d is inconsistent with that of single stars.
 In this range, the median ages and $68\%$ interval are $2.3^{+2.9}_{-0.9}$ Gyr and $2.4^{+3.0}_{-1.1}$ Gyr according to the CPL and CTL models, respectively, compared to the much younger single stars with ages of $1.6^{+0.8}_{-0.4}$ Gyr.
@@
 Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
-For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages \xxx{$\lsim 1.6$} Gyr.
-If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs, e.g.
-Fig.~\ref{fig:protAgeHist}.
-This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, \xxx{e.g.
+For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages \xxx{$\lsim 1.5$ Gyr.
+If the star is actually an unresolved binary, as could be the case for many \kepler rapid rotators \citep{Simonian2018}, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail}.
+This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a \xxx{contaminating signal, e.g.
 Fig.~\ref{fig:protDist} and Fig.~\ref{fig:gyro}}.
 
-In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another.
-Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.
+In general, it is difficult to accurately determine if a source is single star or a stellar binary via \xxx{longterm photometric monitoring, e.g.
+via \kepler or \textit{TESS},} as only a small fraction of stars in binaries will occult one another.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.
 Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field \xxx{are consistent with tidally-synchronized photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread}.
@@
 \end{equation}
-By specifying P$_{orb}$ and the product, $k_2 \tau$, we can numerically solve Eqn.~(\ref{eqn:appendixBalance}) for the P$_{rot}$ at which torques due to magnetic braking and tides balance, often producing subsynchronous rotation as seen in Fig.~\ref{fig:analyticBalance} and our simulations in $\S$~\ref{sec:eq}.
+By specifying P$_{orb}$ and $k_2 \tau$, we can numerically solve Eqn.~(\ref{eqn:appendixBalance}) for the P$_{rot}$ at which torques due to magnetic braking and tides balance, often producing subsynchronous rotation as seen in Fig.~\ref{fig:analyticBalance} and our simulations in $\S$~\ref{sec:eq}.
 

### 2019-04-27 Paper/sync.tex
Updates from Overleaf

@@
 When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$, or more generally P$_{rot} =$ P$_{eq}$ for eccentric orbits.
-In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2017}.
+In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2018}.
 Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[e.g.
@@
 This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations.
-
-The influence of tides is not necessarily limited to short P$_{orb}$ systems, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered.
-The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.
-
+\xxx{The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.}
+
+% cut
+% The influence of tides is not necessarily limited to short P$_{orb}$ systems, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered.
 %The large sample of $816$ P$_{rot}$ measurements derived using star spot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
@@
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.
-Tidal torques can modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, or compete against magnetic braking to form subsychronous rotators, imparting a contaminating signal that is not currently accounted for by models.
+Tidal torques can modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, \xxx{and} compete against magnetic braking to form subsychronous rotators, imparting a contaminating signal that is not currently accounted for by models.
 Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology could be incorrect owing to the influence of tidal torques.
@@
 Here, we present a model for the angular momentum evolution of low-mass stellar binaries over their full premain and main sequence lifetimes using a realistic treatment of stellar evolution, magnetic braking, and tidal torques.
-We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models and two magnetic braking models.
+We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.
 We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of \xxx{gyrochronology} models to potentially fail in such systems.
@@
 
-We improve upon the \xxx{interpolation of the \citet{Baraffe2015} stellar evolution models employed by \citet{Fleming2018}, \stellar, by additionally performing a bicubic interpolation the stellar radius of gyration, $r_g$, over mass and time of the \citet{Baraffe2015} models.
-This updated version of \stellar now tracks the full moment of interia evoltuion of low-mass stars according to the \citet{Baraffe2015} stellar evolution models, a critical requirement for modeling the angular momentum evolution of low-mass stars \citep[e.g.][]{Matt2015}.} We consider two separate magnetic braking models to gauge how sensitive the interaction between tides and magnetic braking is in modifying P$_{rot}$ evolution in stellar binaries.
+We improve upon the \xxx{interpolation of the \citet{Baraffe2015} stellar evolution models employed by \citet{Fleming2018}, \stellar, by additionally performing a bicubic interpolation of the stellar radius of gyration, $r_g$, over mass and time of the \citet{Baraffe2015} models.
+This updated version of \stellar now tracks the full moment of inertia evolution of low-mass stars according to the \citet{Baraffe2015} stellar evolution models, a critical requirement for modeling the angular momentum evolution of low-mass stars.}
 
@@
 This model is strongly dependent on the stellar Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
-The \citet{Matt2015} model predicts that below a certain $Ro$ for rapidly rotating stars, stellar magnetic activity saturates at a constant value, producing a magnetic braking torque that is directly proportional to the stellar rotation rate.
+The \citet{Matt2015} model predicts that below a certain $Ro$ for rapidly-rotating stars, stellar magnetic activity saturates at a constant value, producing a magnetic braking torque that is directly proportional to the stellar rotation rate.
 The angular momentum loss for rapidly-rotating saturated stars is given by
@@
 
-\xxx{Our simplified model assumes that stars follow solid body rotation, whereas in low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution modeled in this work \citep{MacGregor1991,Allain1998}.
-Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior, i.e.
-into the envelope, for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass \citep{Irwin2007,Bouvier2008,Irwin2009}.
-This core-to-envelope angular momentum transfer can counteract the spin-down caused by magnetic braking on the main sequence.
-For a given magnetic braking model, this angular momentum transfer within the stellar interior would likely lead to slightly faster stellar rotation on the early main sequence and a wider spread of rotation periods on the zero-age main sequence than is predicted by solid body rotation models \citep{Gallet2013}.
-This effect could be more pronounced for low-mass M-dwarfs, potentially accounting for their long spin-down timescales \citep{Gallet2015}.
+\xxx{Our simplified model assumes that stars follow solid body rotation, whereas in low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution \citep{MacGregor1991,Allain1998,Bouvier2008,Irwin2009}.
+Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior and into the envelope for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass.
+We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal locking timescales.
+Internal angular momentum transport torques could potentially balance both tidal and magnetic braking torques near the tidally-locked state, producing slight supersynchronous rotation, analogous to the subsynchronous case examined in $\S$~\ref{sec:eq}.
 Modeling core-envelope coupling is beyond the scope of this work, however.}
@@
 
-\subsubsection{Subsynchronous Rotation at Short P$_{orb}$}
-
-In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a P$_{orb} = 5$ d binary for various tidal dissipation parameters.
-Subsynchronous rotation occurs in general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
-Previous theoretical studies have also predicted subsynchronous rotation in short P$_{orb}$ binaries arising from the balance between tidal torques and magnetic braking \citep[e.g.][]{Habets1989,Zahn1994,Keppens1997} suggesting that this behavior is not an artifact of our choice of tidal or magnetic braking models, but rather a general outcome of the competition between magnetic braking and tidal evolution in low-mass binaries.
-Short P$_{orb}$ subsynchronous rotators can eventually tidally-lock via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
-
-\begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
-   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
-Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
-}%
-    \label{fig:eqPerShortPorb}%
-\end{figure}
-
-% caption extra:
-% In the long-term, magnetic braking weakens relative to tidal torques for longer P$_{rot}$, allowing tides to drive the stars back towards the tidally-locked state.
-
-Short P$_{orb}$ subsynchronous binaries exist in nature, such as many \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012}, EPIC 219394517 \citep{Torres2018}, and in ``Binary 6211" observed by \citet{Meibom2006}, suggesting that this theoretical observation is real and borne out in nature.
-Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries and sunsynchronous rotation should be expected in many tidally-interacting binaries.
-We explore these effect further and compare our theory to observations of \kepler EBs in $\S$~\ref{sec:kepler}.
-
 \subsection{Influence of P$_{orb}$, $Q$ and $\tau$} \label{sec:qTauMaps}
@@
 This population corresponds to the evolution described in $\S$~\ref{sec:eq} in which magnetic braking initially spins down stars past the tidally-locked state, but in the long-term, tidal torques spin up the stars, shepherding them towards the tidally-locked state.
-This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ and $Q$ or $\tau$ (see Figs.~\ref{fig:eqPer},\ref{fig:eqPerShortPorb}).
+This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ and $Q$ or $\tau$ (see Figs.~\ref{fig:eqPer},$\S$~\ref{sec:shortPorbSubSync}).
 
@@
 % The CPL model predicts that stellar binaries can tidally-lock out to P$_{orb} = 30$ d for most values of $Q$, with some locking all the way up to P$_{orb} = 100$ d for small $Q < 10^5$.
-
-We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$.
-In each bin, we compute the median $Q$ and $\tau$, and the $16\%$, and $84\%$ percentiles, a measure of the spread of the distribution, to quantify the distribution of tidal strength parameters that are consistent with each spin-orbital state.
 
@@
 
+\xxx{We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$}.
 For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems, especially for P$_{orb} > 20$ d.
@@
 According to the CTL model, many binaries tidally-lock for P$_{orb} \lsim 20$ d.
-Tidally-locking can occur for P$_{orb} < 50$ d and $\tau \gsim 0.25$ s.
-For P$_{orb} < 50$ d, subsynchronous rotation occurs for stars with $\tau < 0.1$ s as magnetic braking dominates over tidal torques.
-For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s has upper and lower limits that encompass most of the range we considered for $\tau$, suggesting that tides are not important in this region of parameter space.
+\xxx{for P$_{orb} < 50$ d, tidal-locking can occur when $\tau \gsim 0.25$ s  whereas subsynchronous rotation occurs for stars with $\tau < 0.1$ s}.
+For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, typically insufficient for tides given the wide orbital separations}.
 The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
-We do not observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).
+We do not \xxx{often} observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).
+
+%% Short Porb subsync
+
+\subsubsection{Subsynchronous Rotation at Short P$_{orb}$} \label{sec:shortPorbSubSync}
+
+\xxx{As seen in Fig.~\ref{fig:qTauLock}, subsynchronous rotation can even occur for short P$_{orb}$ binaries, where tidal-locking is the expectation, if the tidal torques are sufficiently weak, i.e.
+$Q > 10^7$ and $\tau > 0.1$ s.} In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a \xxx{P$_{orb} = 7.5$ d} binary for various tidal dissipation parameters.
+Subsynchronous rotation occurs in general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
+Previous theoretical studies have also predicted subsynchronous rotation in short P$_{orb}$ binaries arising from the balance between tidal torques and magnetic braking \citep[e.g.][]{Habets1989,Zahn1994,Keppens1997} suggesting that this behavior is not an artifact of our choice of tidal or magnetic braking models, but rather a general outcome of the competition between magnetic braking and tidal evolution in low-mass binaries.
+Short P$_{orb}$ subsynchronous rotators can eventually tidally-lock \xxx{after several Gyrs, e.g.
+the $Q=10^8$ case in Fig.~\ref{fig:eqPerShortPorb},} via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
+
+\begin{figure}
+	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
+   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 7.5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
+Systems with strong tidal torques tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
+}%
+    \label{fig:eqPerShortPorb}%
+\end{figure}
+
+% caption extra:
+% In the long-term, magnetic braking weakens relative to tidal torques for longer P$_{rot}$, allowing tides to drive the stars back towards the tidally-locked state.
+
+Short P$_{orb}$ subsynchronous binaries exist in nature, such as many \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012}, EPIC 219394517 \citep{Torres2018}, and in ``Binary 6211" observed by \citet{Meibom2006}, suggesting that this theoretical observation is real and borne out in nature.
+Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries and sunsynchronous rotation should be expected in many tidally-interacting binaries.
+We explore these effect further and compare our theory to observations of \kepler EBs in $\S$~\ref{sec:kepler}.
 
@@
 Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we find stellar binaries can tidally-lock over a wide range of P$_{rot}$ up to P$_{rot} =$ P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
-This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d and with the observations of \citet{Abt2004}.
+This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
 Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d, in stark contrast to the CPL model's predictions.

### 2019-04-29 Analysis/analyticTorqueBalance.py
updated lurie plot colorbar

@@ -64,7 +64,7 @@ def ctlMatt2015(Porb, logK2Tau):
 
 nbins = 50
 logK2Taus = np.log10(0.5 * np.logspace(-1.8, -0.3, nbins))[::-1]
-Porbs = np.linspace(10, 80, nbins)
+Porbs = np.linspace(15, 80, nbins)
 res = np.zeros((nbins, nbins)) # = Prot/Peq
 
 for ii in range(nbins):

### 2019-04-29 Analysis/lurie.py

@@ -154,7 +154,7 @@ ax2.set_title("Lurie et al. (2017)")
 
 ### Colorbar ###
 cbaxes = fig.add_subplot(gs[-1])
-cb = plt.colorbar(im, cax=cbaxes)
+cb = plt.colorbar(im, cax=cbaxes, extend='max')
 cb.set_label(label="Eccentricity")
 
 fig.savefig("../Plots/lurieFig7.pdf", bbox_inches="tight", dpi=200)

### 2019-04-29 Paper/sync.tex
Updates from Overleaf

@@
 Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
-gyrochronology \xxx{ages can err by up to $300\%$} unless one accounts for binarity.
+gyrochronology \xxx{can systematically underpredict ages by up to $300\%$} unless one accounts for binarity.
 We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
@@
 
-We examine stellar angular momentum evolution in low-mass binaries by simulating four sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism, using both magnetic braking models.
+We examine stellar angular momentum evolution in low-mass binaries by simulating \xxx{two} sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism.
 We simulate both stars' spin evolution but mainly consider the P$_{rot}$ evolution for the primary, i.e.
@@
 (2) from \citet{Heller2011}.
-\xxx{All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.}
-
-We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them.
+\xxx{All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.} We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them.
 All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}.
@@
 
-In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation.
-In this case when P$_{rot} >$ P$_{eq}$ for long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated magnetic braking regime \citep{Matt2015}.
+In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation\xxx{, P$_{rot} >$ P$_{eq}$.
+For long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated magnetic braking regime \citep{Matt2015}.
 Since magnetic braking scales as P$_{rot}^{-3}$ for unsaturated rotators, e.g.
-Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as the stellar rotation slows down, so at some P$_{rot}$, tidal torques that try to spin-up subsynchronous rotators back towards P$_{eq}$ will balance the spin-down torque of magnetic braking, producing a long-lasting state of subsynchronous rotation.
-
-We compute the P$_{rot}$ at which this balance occurs by by setting the sum of Eqn.~(\ref{eqn:ctl:omega}) and Eqn.~(\ref{eqn:mattUnSat}) equal to 0, noting that $d\omega/dt = (1/\mathrm{I}) dJ/dt$ for fixed moment of inertia, I.
-In this calculation, we consider tidal torques under the CTL formalism and magnetic braking under the \citet{Matt2015} model.
-For simplicity, we assume both stars are solar-mass with 0 obliquity, a circular binary orbit, and that the torque balance occurs while the stars are on the main sequence where stellar properties change slowly, allowing us to set $R = 1 R_{\odot}$ and assume constant moments of inertia.
-We find below in simulations that these assumptions are well-justified.
-Given these assumptions, we can specify P$_{orb}$, $k_2$, and $\tau$ to compute the equilibrium P$_{rot}$ at which tidal and magnetic braking torques balance.
-We derive this balance, given by Eqn.~\ref{eqn:appendixBalance}, in $\S$~\ref{sec:appendix:balance} and visualize it in Fig.~\ref{fig:analyticBalance}.
-We normalize P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$ (see Eqn.~(\ref{eqn:ctl:eqPer})).
+Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as the stellar rotation slows down, so at some P$_{rot}$, tidal torques will balance magnetic braking, producing a long-lasting state of subsynchronous rotation.
+We compute the P$_{rot}$ at which this balance occurs as a function of P$_{orb}$, $k_2$, and $\tau$ in $\S$~\ref{sec:appendix:balance} by setting the sum of Eqn.~(\ref{eqn:ctl:omega}) and Eqn.~(\ref{eqn:mattUnSat}) equal to 0, considering tidal torques under the CTL formalism.
+For simplicity, we assume both stars are solar-mass with 0 obliquity, a circular binary orbit, and that the torque balance occurs while the stars are on the main sequence where stellar properties change slowly.
+Although solar mass stars are the most massive stars we consider in this work, and hence will have the strongest tidal torque for a given tidal dissipation parameter and P$_{orb}$, they can still exhibit subsynchronous rotation and serve as a useful pedagogical case to examine here and in simulations below.
+We display the results of this calculation in Fig.~\ref{fig:analyticBalance}, normalizing P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$.}
 
@@
 	\includegraphics[width=0.5\textwidth]{../Plots/analyticTorque.pdf}
-   \caption{The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}.
+   \caption{\xxx{Updated Figure} The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}.
 The white dashed line indicates our fiducial values for $k_2$ and $\tau$, 0.5 and 0.1 s, respectively, that we adopt in the simulations in $\S$~\ref{sec:mattBalance}.}%
@@
 
-In general as tides weaken, i.e.
-increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where magnetic braking torques are weaker, or equivalently at larger P$_{rot}/$P$_{eq}$.
-Conversely, in the strong tides regime, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for most orbital periods, tidally-locking binaries into synchronous rotation.
-Given this calculation, we expect that for many orbital periods and tidal time lags, stellar binaries will reach a state of subsynchronous rotation.
-For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries on low eccentricity orbits reach the subsynchronous state with P$_{rot}/$P$_{eq} \approx 1.1-1.4$, with this value increasing with P$_{orb}$.
-This simple analytic calculation, however, does not account for stellar evolution or secular orbital evolution due to tides, e.g.
-tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we cannot precisely determine when the binaries reach the subsynchronous state nor how long it will last.
-For that, we turn to simulations.
-
-Below, we verify the results of our analytic calculations by examining how stellar binaries evolve towards subsynchronous rotation, under what conditions, and when it occurs in the weak tides regime.
-We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a combination of both tidal models and both magnetic braking models.
-In Fig.~\ref{fig:eqPer}, we plot the simulated evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models.
+\xxx{In general as tides weaken, i.e.
+increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where tides are weaker.
+Conversely, for strong tides, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for most orbital periods, tidally-locking binaries into synchronous rotation.
+For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries will rotate subsynchronously for P$_{orb} \gsim 20$ d, with more severe subsynchronism for longer P$_{orb}$.
+This simple calculation, however, does not account for stellar evolution or secular tidal orbital evolution, e.g.
+tidal friction that will shrink the orbit, gradually strengthening tidal torques.}
+
+% cut
+% , noting that $d\omega/dt = (1/\mathrm{I}) dJ/dt$ for fixed moment of inertia, I.  In this calculation, we
 
@@
 
+\xxx{We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits to examine how stellar binaries evolve towards subsynchronous rotation.
+In Fig.~\ref{fig:eqPer}, we plot P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models.
 Both tidal models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
-The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.
-The CTL model, however, predicts that for P$_{orb} \gsim 10$ d, magnetic braking overpowers weak tidal torques, spinning down the stars past the tidally-locked state.
-Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.05$ with the maximum value set by the torque balance between tidal torques and magnetic braking, in good agreement with the analytic calculations in $\S$~\ref{sec:analytic}.
+The CPL model predicts that all binaries tidally-lock, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.
+As anticipated by our analytic calculations, the CTL model predicts subsynchronous rotation for P$_{orb} \geq 20$ d as magnetic braking overpowers tidal torques.
+For P$_{orb} = 20$ d, Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.05$, with the maximum value set by the torque balance.
 As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
-Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution, driving the binaries towards subsynchronous rotation.
+Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution.}
 
@@
 Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d.
-The black dotted line indicates the tidally-locked state where P$_{rot} = $ P$_{eq}$.
+The black dotted line indicates the tidally-locked state.
 Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.
@@
 For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
-For all P$_{orb}$, the CPL models tidally-lock into synchronous rotation whereas the CTL model predicts subsynchronous (P$_{rot} > $ P$_{eq}$) rotation that persists for Gyrs.}%
+For all P$_{orb}$, the CPL models tidally-lock whereas the CTL model predicts subsynchronous rotation that persists for Gyrs.}%
     \label{fig:eqPer}%
@@
 Subsynchronous rotation does not persist indefinitely, however, as P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
-First, as P$_{rot}$ continue to increase due to magnetic braking, magnetic braking itself weakens as its torque scales as P$_{rot}^{-3}$ for unsaturated rotators \citep{Matt2015}.
-Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example).
-Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$ and strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.
-These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long term.
-
-We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to stellar evolution, magnetic braking, and tidal torques as a function of time in our simulations.
-Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 60 Myr when the stars reach the main sequence.
-After $R$ contraction, $\dot{\mathrm{P}}_{rot} > 0$ as tides and magnetic braking combine to spin down stars towards the tidally-locked state.
-For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$ driving the stars towards subsynchronous rotation.
-$\ddot{\mathrm{P}}_{rot} < 0$, however, as the three processes described above gradually strengthen tidal torques relative to magnetic braking.
-In the long-run, tidal torques overpower magnetic braking, with a slightly negative P$_{rot}$ derivative, slowly driving P$_{rot}$ back towards P$_{eq}$ (see the P$_{orb} = 20, 40$ d cases in Fig.~\ref{fig:eqPer}).
-This slowly shifting torque balance produces a population of subsynchronous rotators with P$_{rot} > $ P$_{eq}$ that can persist for Gyrs.
+First, \xxx{magnetic braking weakens at long P$_{rot}$ as its torque scales as P$_{rot}^{-3}$ for unsaturated rotators \citep{Matt2015}.
+Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega})).
+Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$, strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.
+These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long-term.}
+
+\xxx{We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to tidal torques, stellar evolution, and magnetic braking.
+Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 60 Myr when the stars reach the zero age main sequence.
+Tides and magnetic braking then combine to spin down stars towards the tidally-locked state.
+For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$ as magnetic braking dominates, driving the stars into subsynchronous rotation.
+In the long-term, $\ddot{\mathrm{P}}_{rot} < 0$, however, as the three processes described above gradually strengthen tidal torques relative to magnetic braking.
+Tidal torques eventually overpower magnetic braking, seen as a slight negative P$_{rot}$ derivative, slowly driving P$_{rot}$ back towards P$_{eq}$, producing a population of subsynchronous rotators that can persist for Gyrs.
+We explore this point further in $\S$~\ref{sec:shortPorbSubSync}.}
 
@@
 Spin-orbit synchronization is the typical outcome for binaries with P$_{orb} < 10$ d according to the CPL model for most values of $Q$.
-The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to P$_{orb} \approx 100 $ d for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
+The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to \xxx{P$_{orb} \gsim 80 $ d} for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
 According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \gsim 3$ s.
@@
 \xxx{We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$}.
-For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems, especially for P$_{orb} > 20$ d.
-For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 70$ d as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
+For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems.
+For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for \xxx{P$_{orb} < 60$ d} as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
 Eqn.~(\ref{eqn:cpl:eqPer}).
 
-Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking that spin-down the stars past the tidally-locked state.
-This effect can be seen for P$_{orb} < 50$ d where the median $Q$ gradually increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state.
+Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking.
+This effect can be seen for P$_{orb} < 50$ d where the median $Q$ \xxx{tends to} increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state.
 This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.
@@
 
-According to the CTL model, many binaries tidally-lock for P$_{orb} \lsim 20$ d.
-\xxx{for P$_{orb} < 50$ d, tidal-locking can occur when $\tau \gsim 0.25$ s  whereas subsynchronous rotation occurs for stars with $\tau < 0.1$ s}.
-For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, typically insufficient for tides given the wide orbital separations}.
+\xxx{According to the CTL model, many binaries tidally-lock for P$_{orb} \lsim 20$ d when $\tau \gsim 0.1$ s, with some tidally-locking up to P$_{orb} \approx 50$ d when $\tau \gsim 1$ s.
+Subsynchronous rotation typically occurs for stars with $\tau < 0.1$ s}.
+For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, typically insufficient for tide to strongly influence the evolution given the wide orbital separations}.
 The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
@@
 
-\xxx{As seen in Fig.~\ref{fig:qTauLock}, subsynchronous rotation can even occur for short P$_{orb}$ binaries, where tidal-locking is the expectation, if the tidal torques are sufficiently weak, i.e.
-$Q > 10^7$ and $\tau > 0.1$ s.} In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a \xxx{P$_{orb} = 7.5$ d} binary for various tidal dissipation parameters.
+\xxx{As seen in Fig.~\ref{fig:qTauLock}, subsynchronous rotation can even occur for short P$_{orb}$ binaries, where tidal-locking is the expectation, if the tidal torques are sufficiently weak} In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a \xxx{P$_{orb} = 7.5$ d} binary for various tidal dissipation parameters.
 Subsynchronous rotation occurs in general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
@@
 Here we examine how the competition between tidal torques and magnetic braking shape the P$_{rot}$ distribution of low-mass stellar binaries.
-We consider two cases where tidal torques dominate: ``Locked" where P$_{rot} = $ P$_{eq}$ and ``Interacting" where P$_{rot}$ is within $10\%$ of P$_{eq}$ as in this regime, tides are likely shepherding P$_{rot}$ towards the tidally-locked state keeping P$_{rot}$ close to P$_{eq}$, typically within about $10\%$ as we demonstrated in $\S$~\ref{sec:eq}, e.g.
+We consider two cases where tidal torques dominate: ``Locked", where P$_{rot} = $ P$_{eq}$, and ``Interacting", where P$_{rot}$ is within $10\%$ of P$_{eq}$ as in this regime, tides are likely shepherding P$_{rot}$ towards the tidally-locked state keeping P$_{rot}$ close to P$_{eq}$, typically within about $10\%$ as we demonstrated in $\S$~\ref{sec:eq}, e.g.
 Fig.~\ref{fig:eqPer}.
@@
 
-Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.6$ d and $7.4$ d) than unlocked (median CPL, CTL P$_{rot} = 41.3$ d and $40.8$ d) binaries.
+Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster \xxx{(median CPL, CTL P$_{rot} = 22.6$ d and $8.8$ d)} than unlocked \xxx{(median CPL, CTL \xxx{both} P$_{rot} = 32.4$ d)} binaries.
 More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.
@@
 This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d, in stark contrast to the CPL model's predictions.
-Some binaries can tidally-lock or tidally-interact out to $P_{rot} \approx 80$ d, but not nearly as many as the CPL model, nor can the CTL model tidally-lock binaries out to P$_{orb} = 100$ d, regardless of $\tau$.
-Only $7.5\%$ of binaries tidally-lock according to the CTL model, compared with $25\%$ of CPL model binaries, as tidal torques are generally stronger in the CPL formalism.
+Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d\xxx{, although binaries with more massive primaries can lock, or at least tidally-interact, out to P$_{orb} \approx 70$ d, in stark contrast to the CPL model's predictions of binaries tidally-locking out to $P_{rot} \gsim 90$ d.} Only \xxx{$8\%$} of binaries tidally-lock according to the CTL model, compared with \xxx{$24\%$} of CPL model binaries, as tidal torques are generally stronger in the CPL formalism \xxx{and due to our choice of prior}.
 
@@
 At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthening enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
-The CPL and CTL models predict that $34\%$ and $25\%$ of stars, respectively, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
+The CPL and CTL models predict that \xxx{$31\%$ and $24\%$} of stars, respectively, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
 The P$_{rot}$ - mass distribution for unlocked binaries resembles the single star sequence observed in the center right panel of Fig.~\ref{fig:protDist} indicating that magnetic braking and stellar evolution dictate their angular momentum evolution as expected.
@@
 As shown in Fig.~\ref{fig:lockedProtHist}, both models predict that tidal torques can influence the P$_{rot}$ evolution out to P$_{orb} \approx 80$ d, causing a significant departure from the long-term expected spin down of single stars due to magnetic braking.
-The CTL model predicts that the majority tidally-locked binaries, $86\%$, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL P$_{orb} = 7.6$ d.
-The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states with only $39\%$ of locking binaries have P$_{rot} < 20$ d, while many lock out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
+The CTL model predicts that the majority tidally-locked binaries, \xxx{$83\%$}, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL \xxx{P$_{orb} = 8.8$ d}.
+The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states with only \xxx{$46\%$ of locked} binaries have P$_{rot} < 20$ d, while \xxx{the rest can} lock out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
 The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
@@
 \xxx{In binaries, tidal torques tend to drive the P$_{rot}$ evolution away from that of single stars and towards P$_{eq}$, either maintaining rapid rotation in tidally-locked short P$_{orb}$ systems, or working with magnetic braking to slow P$_{rot}$ beyond that of single stars of the same age.
-The impact of tidal torques on the binary P$_{rot}$ distribution is clear: strong tidal torques in short P$_{orb}$ binaries produce a substantial population of rapid rotators with P$_{rot} \lsim 20$ d, with the influence of tides extending to longer P$_{orb}$ systems as demonstrated in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.
+The impact of tidal torques on the binary P$_{rot}$ distribution is clear: strong tidal torques in short P$_{orb}$ binaries produce a substantial population of rapid rotators with P$_{rot} \lsim 20$ d.
 Except for stars with ages $\lsim 1$ Gyr, or young late M-dwarfs who are either contracting along the pre-main sequence or have just reached the main sequence, our single star simulations fail to produce a population of rapid rotators.
-This theoretical result is consistent with \citet{Simonian2018} who find that the population of \kepler stars with P$_{rot} < 7.5$ d is likely dominated by tidally-interacting binaries.}
-
-\xxx{In the single star population, there is a clear monotonic age-dependence of P$_{rot}$, with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
+This theoretical result is consistent with \citet{Simonian2018} who find that the population of \kepler stars with P$_{rot} < 7.5$ d is likely dominated by tidally-interacting binaries.
+The influence of tides extends to longer P$_{orb}$ systems producing a slowly-rotating population above the upper envelope of the single star sequence for M$ \gsim 0.6$ M$_{\odot}$, e.g.
+Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, and is a population that single-star models fail to produce.}
+
+\xxx{In the single star population, there is a clear monotonic relation between P$_{rot}$ and age, with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
 This trend is a generic outcome of magnetic braking and is not specific to our choice of magnetic braking model.
@@
 Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
-For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages \xxx{$\lsim 1.5$ Gyr.
+For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages \xxx{$\lsim 1.6$ Gyr.
 If the star is actually an unresolved binary, as could be the case for many \kepler rapid rotators \citep{Simonian2018}, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail}.
@@
 These young, supersynchronous rotators are unlikely to be tidally-locked, usually have P$_{orb}/$P$_{rot} > 1.5$, and do not stay supersynchronous for long given that solar mass pre-main sequence lifetimes are $\lsim 100$ Myr, distinguishing them from tidally-locked binaries (see Fig.~\ref{fig:lurie7}).
-If supersynchronous rotation is observed in binaries with P$_{orb} < 10$ d, P$_{orb}/$P$_{rot} < 1.5$, and $0 < e \lsim 0.23$, it is evidence in favor of the CTL model over our discrete CPL model.
+If supersynchronous rotation is observed in binaries with P$_{orb} < 10$ d, P$_{orb}/$P$_{rot} < 1.5$, and $0 < e \lsim 0.23$, it is evidence in favor of the CTL model over the CPL model.
 
@@
 Tidal interactions can cause P$_{rot}$ evolution in stellar binaries to differ from the long-term spin down due to magnetic braking experienced by single stars, decoupling P$_{rot}$ from age.
-In tidally-interacting binaries, gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails.
+In tidally-interacting binaries, gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails\xxx{, potentially underestimating stellar ages by up to $300\%$}.
 We caution that any application of \xxx{gyrochronology} methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
@@
 \xxx{We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
-We primarily suggest that observers measure stellar P$_{rot}$ in binaries with solar-mass primaries at P$_{orb} 60-100$ d.
+We primarily suggest that observers measure stellar P$_{rot}$ in binaries with solar-mass primaries for P$_{orb}$ between $60-100$ d.
 If any tidally-locked binaries are identified at long P$_{orb}$, this would be evidence in favor of the CPL model as we found that only binaries tidally-interacting via the CPL model could tidally-lock at such long P$_{orb}$.
@@
 Beyond these tests, our model could be used to infer the tidal properties of binary stars, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties and reasonable prior probability distributions for parameters like the initial binary $e$.
-This analysis, however, is also beyond the scope of this work and we leave it for future endeavors.}
+This analysis, however, is beyond the scope of this work and we leave it for future endeavors.}
 

### 2019-05-15 Paper/sync.tex
proofread after addressing ref report

@@
 Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
-gyrochronology \xxx{can systematically underpredict ages by up to $300\%$} unless one accounts for binarity.
+gyrochronology \xxx{can underpredict ages by up to $300\%$} unless one accounts for binarity.
 We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
@@
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.
-Tidal torques can modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, \xxx{and} compete against magnetic braking to form subsychronous rotators, imparting a contaminating signal that is not currently accounted for by models.
+\xxx{Tidally-interacting binaries impart a} contaminating signal that is not currently accounted for by models.
 Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology could be incorrect owing to the influence of tidal torques.
@@
 We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.
-We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of \xxx{gyrochronology} models to potentially fail in such systems.
+We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of \xxx{gyrochronology} models fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
@@
 \xxx{We simulate magnetic braking using the model derived by \citet{Matt2015} as this formalism} has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
-This model is strongly dependent on the stellar Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
+This model \xxx{depends} on the stellar Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
 The \citet{Matt2015} model predicts that below a certain $Ro$ for rapidly-rotating stars, stellar magnetic activity saturates at a constant value, producing a magnetic braking torque that is directly proportional to the stellar rotation rate.
@@
 Saturated magnetic braking occurs for $Ro \leq Ro_{\odot}/\chi$ for $\chi = 10$ where \citet{Matt2015} defines $\chi = Ro_{\odot}/Ro_{sat}$.
-We adopt all model parameters given in Table 1 from \citet{Matt2015}, with the correction from \citet{Matt2019}, and compute $\tau_{cz}$ using Eqn.
+We adopt all model parameters given in Table 1 from \citet{Matt2015}\xxx{, with the correction from \citet{Matt2019},} and compute $\tau_{cz}$ using Eqn.
 (36) from \citet{Cranmer2011}.
@@
 Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior and into the envelope for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass.
-We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal locking timescales.
+We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal-locking timescales.
 Internal angular momentum transport torques could potentially balance both tidal and magnetic braking torques near the tidally-locked state, producing slight supersynchronous rotation, analogous to the subsynchronous case examined in $\S$~\ref{sec:eq}.
@@
 The ``Constant Phase Lag" (CPL) \citep[][]{FerrazMello2008,Heller2011} equilibrium tidal model assumes that the tidal torque on one body due to its companion arises from a linear combination of several discrete, uncoupled tidal bulges, each with its own associated frequency, that maintain a fixed phase offset with respect to the line connecting the two stars' centers of mass.
-We use the \eqtide implementation of the CPL model in VPLanet following the derivation of \citet{FerrazMello2008}.
+We use the \eqtide implementation of the CPL model in \vplanet following the derivation of \citet{FerrazMello2008}.
 The equations that govern the secular change in $e$ and $a$ are as follows:
@@
 In this formalism, unlike the CPL model, the CTL model is continuous over a range of tidal wave frequencies and applicable for large $e$.
-However if the assumption of linearity is relaxed, i.e.
+\xxx{However,} if the assumption of linearity is relaxed, i.e.
 frequencies associated with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}.
-We use the \eqtide implementation of the CTL model in VPLanet following the derivation of \citet{Leconte2010}.
+We use the \eqtide implementation of the CTL model in \vplanet following the derivation of \citet{Leconte2010}.
 The equations that govern the secular changes in $e$, $a$, $\omega$, and $\psi$ are as follows:
@@
 
-We plot the tidal evolution for $a$, $e$, and $\omega$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively, in Fig.~\ref{fig:tidalExample}.
-Both the CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally-locked state, transferring rotational angular momentum into the orbit in the process.
+We plot the tidal evolution for $a$, $e$, and $P_{rot}$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively, in Fig.~\ref{fig:tidalExample}.
+\xxx{The} CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally-locked state, transferring rotational angular momentum into the orbit in the process.
 At late times, both the CPL and CTL drive the binaries towards orbital circularization, with tidal dissipation decreasing P$_{orb}$.
@@
 \dot{a}_{coupled}^{(2)} = \frac{-\dot{J}_{mb} - 2 \omega \left( \sum_{i=1}^{2} m_i r_{g,i}^2 R_i \dot{R_i} + m_i r_{g,i} \dot{r}_{g,i} R_i^2 \right)}
-{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} \left( m_1 r_{g,1}^2 R_1^2 + m_2 r_{g,2}^2 R_2^2 \right)}
+{\frac{\mu^2 G M (1-e^2)}{2J_{orb}} - \frac{3 \omega}{2a} \left( m_1 r_{g,1}^2 R_1^2 + m_2 r_{g,2}^2 R_2^2 \right)},
 \end{split}
@@
 Although the CTL model is applicable for $e \gsim 0.3$, \xxx{the CPL model is not and can predict qualitatively incorrect evolution in that regime \citep[see Section 4.1 in][]{Leconte2010}}, so we restrict $e \leq 0.3$ to allow us to compare both models.
-We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.
+We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Fabrycky2007,Munoz2015,Martin2015b,Hamers2016,Moe2018}.
 
@@
 
-Here we focus on binaries in weak tides regime, i.e.
+Here we focus on binaries in \xxx{the ``weak tides"} regime, i.e.
 long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking via analytic calculations and simulations.
@@
 
-\xxx{In general as tides weaken, i.e.
+\xxx{Subsynchronous rotation can occur across a wide range of tidal parameters and P$_{orb}$.
+In general as tides weaken, i.e.
 increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where tides are weaker.
-Conversely, for strong tides, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for most orbital periods, tidally-locking binaries into synchronous rotation.
-For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries will rotate subsynchronously for P$_{orb} \gsim 20$ d, with more severe subsynchronism for longer P$_{orb}$.
+Conversely, for strong tides, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for P$_{orb} \lsim 40$ d, tidally-locking binaries into synchronous rotation.
+For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries will rotate subsynchronously for P$_{orb} \gsim 20$ d, with more severe subsynchronism at longer P$_{orb}$.
 This simple calculation, however, does not account for stellar evolution or secular tidal orbital evolution, e.g.
-tidal friction that will shrink the orbit, gradually strengthening tidal torques.}
+tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we turn to simulations to characterize this evolution.}
 
@@
 As anticipated by our analytic calculations, the CTL model predicts subsynchronous rotation for P$_{orb} \geq 20$ d as magnetic braking overpowers tidal torques.
-For P$_{orb} = 20$ d, Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.05$, with the maximum value set by the torque balance.
+For P$_{orb} = 20$ d, magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.05$, with the maximum value set by the torque balance.
 As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
@@
 Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega})).
-Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$, strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.
+Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$, gradually strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.
 These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long-term.}
@@
 According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \gsim 3$ s.
-Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e.
-stars with large $Q$ or small $\tau$, that form via the mechanism described in $\S$~\ref{sec:eq}, with weaker tidal torques allowing more subsynchronous rotation.
+Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques.
 The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) with P$_{orb} > 80$ d does not in general correspond to binaries tidally-locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally-locked state.
@@
    \caption{\xxx{Updated Figure} Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
-For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, regardless of $Q$, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.
-Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spun down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
-For large $Q$ ($> 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
+For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.
+Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$ d, that \xxx{does} not in general reflect tidally-locking, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spun down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
+For large $Q$ ($> 10^7$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
     \label{fig:qTauLock}%
@@
 \xxx{We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$}.
-For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems.
-For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for \xxx{P$_{orb} < 60$ d} as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
+For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock \xxx{these binaries}.
+For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$, nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for \xxx{P$_{orb} < 60$ d}, as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
 Eqn.~(\ref{eqn:cpl:eqPer}).
@@
 Subsynchronous rotation typically occurs for stars with $\tau < 0.1$ s}.
-For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, typically insufficient for tide to strongly influence the evolution given the wide orbital separations}.
+For P$_{orb} > 50$ d, magnetic braking dominates the evolution \xxx{seen in} the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, a value that is typically insufficient for tides to strongly influence the evolution given the wide orbital separations}.
 The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
@@
 
-\xxx{As seen in Fig.~\ref{fig:qTauLock}, subsynchronous rotation can even occur for short P$_{orb}$ binaries, where tidal-locking is the expectation, if the tidal torques are sufficiently weak} In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a \xxx{P$_{orb} = 7.5$ d} binary for various tidal dissipation parameters.
+\xxx{As seen in Fig.~\ref{fig:qTauLock}, subsynchronous rotation can even occur for short P$_{orb}$ binaries, where tidal-locking is the expectation, if the tidal torques are sufficiently weak.} In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a \xxx{P$_{orb} = 7.5$ d} binary for various tidal dissipation parameters.
 Subsynchronous rotation occurs in general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
@@
    \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 7.5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
-Systems with strong tidal torques tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
+Systems with strong tidal torques tidally-lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
 }%
@@
 Here we examine how the competition between tidal torques and magnetic braking shape the P$_{rot}$ distribution of low-mass stellar binaries.
-We consider two cases where tidal torques dominate: ``Locked", where P$_{rot} = $ P$_{eq}$, and ``Interacting", where P$_{rot}$ is within $10\%$ of P$_{eq}$ as in this regime, tides are likely shepherding P$_{rot}$ towards the tidally-locked state keeping P$_{rot}$ close to P$_{eq}$, typically within about $10\%$ as we demonstrated in $\S$~\ref{sec:eq}, e.g.
-Fig.~\ref{fig:eqPer}.
-We refer to the remaining binaries as ``Unlocked" as magnetic braking and stellar evolution likely dominate their angular momentum evolution.
-In Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, we plot P$_{rot}$ as a function of mass for the primary stars in stellar binaries for both the CPL and CTL model, respectively, integrated to system ages uniformly sampled over $1-7$ Gyr.
+We consider two cases where tidal torques dominate: ``Locked", where P$_{rot} = $ P$_{eq}$, and ``Interacting", where P$_{rot}$ is within $10\%$ of P$_{eq}$ as in this regime, tides are likely shepherding P$_{rot}$ towards the tidally-locked state as we demonstrated in $\S$~\ref{sec:eq}.
+We refer to the remaining binaries as ``not locked" as magnetic braking and stellar evolution likely dominate their angular momentum evolution.
+In Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, we plot P$_{rot}$ as a function of mass for the primary stars in stellar binaries for both the CPL and CTL model, respectively, integrated to system ages uniformly sampled over $1-7$ Gyr\xxx{, consistent with ages of stars in the \kepler field \citep{Chaplin2014}}.
 
@@
 
-Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster \xxx{(median CPL, CTL P$_{rot} = 22.6$ d and $8.8$ d)} than unlocked \xxx{(median CPL, CTL \xxx{both} P$_{rot} = 32.4$ d)} binaries.
-More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.
+Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster \xxx{(median CPL, CTL P$_{rot} = 22.6$ d and $8.8$ d)} than not locked \xxx{(median CPL, CTL both P$_{rot} = 32.4$ d)} binaries.
+\xxx{The CTL model predicts that the majority tidally-locked binaries, \xxx{$83\%$}, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest.
+The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states as only $46\%$ of locked binaries have P$_{rot} < 20$ d, while the rest can lock out to P$_{rot} \approx 100$ d in long P$_{orb}$ binaries.} More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.
 This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model.
@@
 
-Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we find stellar binaries can tidally-lock over a wide range of P$_{rot}$ up to P$_{rot} =$ P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of single stars.
-This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d\xxx{, although binaries with more massive primaries can lock, or at least tidally-interact, out to P$_{orb} \approx 70$ d, in stark contrast to the CPL model's predictions of binaries tidally-locking out to $P_{rot} \gsim 90$ d.} Only \xxx{$8\%$} of binaries tidally-lock according to the CTL model, compared with \xxx{$24\%$} of CPL model binaries, as tidal torques are generally stronger in the CPL formalism \xxx{and due to our choice of prior}.
-
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{../Plots/lockedCTL.pdf}
+	\includegraphics[width=0.48\textwidth]{../Plots/lockedCTL.pdf}
    \caption{\xxx{Updated Figure} Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
@@
 
-The interacting population tends to rotate more slowly than the unlocked population as at short P$_{orb}$, and hence P$_{rot}$, binaries preferentially tidally-lock due to stronger tidal torques.
+\xxx{The interacting population tends to rotate more slowly than the not locked population as at short P$_{orb}$, and hence P$_{rot}$, binaries preferentially tidally-lock due to stronger tidal torques.
 At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthening enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
 The CPL and CTL models predict that \xxx{$31\%$ and $24\%$} of stars, respectively, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
-The P$_{rot}$ - mass distribution for unlocked binaries resembles the single star sequence observed in the center right panel of Fig.~\ref{fig:protDist} indicating that magnetic braking and stellar evolution dictate their angular momentum evolution as expected.
+The P$_{rot}$ - mass distribution for not locked binaries resembles the single star sequence as magnetic braking and stellar evolution dictate their angular momentum evolution.}
+
+Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we find stellar binaries can tidally-lock over a wide range of P$_{rot}$ up to P$_{rot} =$ P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of \xxx{solar-mass} single stars.
+This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
+Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d\xxx{, although binaries with more massive primaries can occasionally lock, or at least tidally-interact, out to P$_{orb} \approx 80$ d, in contrast to the CPL model's predictions.
+\xxx{We highlight this behavior with a histogram of locked and interacting binaries over P$_{rot}$ for both tidal models in Fig.~\ref{fig:lockedProtHist}.} The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
+We discuss this point further in $\S$~\ref{sec:whichModel}}.
 
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{../Plots/lockedProtHist.pdf}
+	\includegraphics[width=0.48\textwidth]{../Plots/lockedProtHist.pdf}
    \caption{P$_{rot}$ distribution for tidally-locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.}%
@@
 \end{figure}
-
-As shown in Fig.~\ref{fig:lockedProtHist}, both models predict that tidal torques can influence the P$_{rot}$ evolution out to P$_{orb} \approx 80$ d, causing a significant departure from the long-term expected spin down of single stars due to magnetic braking.
-The CTL model predicts that the majority tidally-locked binaries, \xxx{$83\%$}, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest as the median locked CTL \xxx{P$_{orb} = 8.8$ d}.
-The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states with only \xxx{$46\%$ of locked} binaries have P$_{rot} < 20$ d, while \xxx{the rest can} lock out to P$_{rot} \approx 100$ d, with the latter state corresponding to tidal synchronization in long P$_{orb}$ binaries.
-The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
-We discuss this point further in $\S$~\ref{sec:whichModel}.
 
@@
 This quantity, referred to here as the ``Relative Age Error", represents the systematic error incurred by assigning a tidally-interacting binary star the age expected for single stars at a given mass and P$_{rot}$.
-For this comparison, we select the subset of tidally-locked and tidally-interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, accounting for $31\%$ and $24\%$ of CPL and CTL binaries, respectively.}
+For this comparison, we select the subset of tidally-locked and tidally-interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.}
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
-   \caption{\xxx{Updated Figure} Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
+   \caption{\xxx{Updated Figure} Histogram of rapidly-rotating (P$_{rot} < 20$ d) \xxx{star} ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
 Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
@@
 We caution that any application of \xxx{gyrochronology} methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
-
-Furthermore, any study of the angular momentum evolution of a stellar population will be flawed unless one robustly accounts for binarity.
-Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
+\xxx{Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
 Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
-We recommend that the application, or calibration, magnetic braking models to a sample of stellar rotation periods control for binarity.
+We recommend that the application, or calibration, magnetic braking models to a sample of stellar rotation periods control for binarity.}
 
@@
 
-We compare our simulation results using both tidal models and the \citet{Matt2015} magnetic braking model to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing stellar binaries by \citet{Lurie2017}.
-The purpose of this comparison is to see if our model predictions, which by design populate a wide, but physically-plausible, region of parameter space, can reproduce features observed in the data.
+\xxx{We compare our simulation results to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing binaries by \citet{Lurie2017}} to gauge if our model predictions, which by design populate a wide, but physically-plausible, region of parameter space, can reproduce features observed in the data.
 \citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with star spot modulations and visually inspected each light curve to ensure their accuracy.
@@
 
-Both tidal models qualitatively reproduce most features seen in the \citet{Lurie2017} data, especially at short P$_{orb}$.
-Nearly all simulated binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in agreement with the \citet{Lurie2017} observations.
-At very short P$_{orb}$, in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries, in good agreement with theoretical predictions \citep{Counselman1973}.
-For P$_{orb} \gsim 4$ d, our models produce large populations of subsynchronous rotators.
-Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces subsynchronous rotators.
-For example under the CTL model, our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
+\xxx{Qualitatively, the CTL model appears to do a better job of reproducing features seen in the \citet{Lurie2017} data than the CPL model.
+The CPL model, for example, cannot produce the observed cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.2$ for P$_{orb} < 10$ d whereas the CTL model can.
+Instead, owing to the its discrete P$_{eq}$, the CPL model predicts that all tidally-locked supersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.}  This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the \citet{Lurie2017} data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
+\xxx{Neither model reproduces the very supersynchronous, P$_{orb}/$P$_{rot} > 1.6$, binaries in the \citet{Lurie2017} as they typically have orbital eccentricities in excess of 0.3 and are outside of the region of parameter space we consider.
+The CTL model, however, could in principle reproduce these points if they are tidally-locked binaries as its P$_{eq}$ is a continuous function of $e$ and is applicable for large $e$, in contrast to the discrete P$_{eq}$ predicted by the CPL model that is valid for smaller $e$.
+Both tidal models predict that nearly all binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in agreement with the \citet{Lurie2017} observations.
+At very short P$_{orb}$, in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries \citep{Counselman1973}.}
+
 \begin{figure*}[t]
@@
 
-The CPL model can produce subsynchronous rotators similar to the observed population with $10 \lsim$ P$_{orb} \lsim 20$ d, but struggles to populate the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
-The CTL model, on the other hand, populates a much wider swath of parameter space than the CPL model can, qualitatively reproducing many clusters seen in the \citet{Lurie2017} data.
-\citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, the cluster of stars those authors identified near $P_{orb}/$P$_{rot} = 0.9$, compared with $9\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries.
-Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$, compared with $72\%$ of our CTL population, in good agreement.
-In comparison, the CPL model predicts $68\%$ of the population has P$_{orb}/$P$_{rot} \in [0.92, 1.2]$ and $2\%$ has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$ for 2 $< P_{orb} < $ 10 d, producing a population of synchronized binaries similar to what is seen in the \kepler EB population, but not the subsynchronous cluster identified by \citet{Lurie2017}.
-
-The CPL model cannot produce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d present in the \citet{Lurie2017} data.
-Instead, owing to the its discrete P$_{eq}$, the CPL model predicts that all tidally-locked supersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.
-This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the \citet{Lurie2017} data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless coupling occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
-On the other hand, the CTL model produces a large number of supersynchronous binaries with eccentric orbits.
-These systems have tidally-locked but their orbits have not yet circularized, even though they likely experience strong tidal torques at short separations, because the tidal locking time scale is 2-3 orders of magnitude less than the circularization timescale \citep{Mazeh2008}.
-
-Both models predict a large number of extremely subsynchronous rotators with P$_{orb}/$P$_{rot} < 0.7$ across all P$_{orb}$ that is not seen in the \citet{Lurie2017} data.
-Magnetic braking and age combine to produce the lower limit of this subsynchronous population.
-The limit is a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down via magnetic braking in the weak tides regime over 7 Gyr, the longest age considered in our simulations.
-Our choice of prior distributions for both $Q$ and $\tau$ permit very weak tidal interactions, e.g.
-$Q \gsim 10^7$ and $\tau < 0.1$ s, that likely gives rise to this extremely subsynchronous population and suggests that our prior does not reflect the underlying distribution of stellar tidal parameters in nature.
-Alternatively, the data could be incomplete where our models predict slowly-rotating subsynchronous rotators as the photometric amplitude of star spot modulations tends to decrease with increasing P$_{rot}$, making reliable rotation periods difficult to detect \citep{McQuillan2014,Lurie2017,Reinhold2018}.
-
-Both tidal models reproduce many features observed in the \kepler EB distribution, suggesting that our model is a fair representation of the dynamical interactions of tidally-evolving, low-mass stellar binaries.
-Our comparison between theory and observations is limited, however.
-The \citet{Lurie2017} P$_{rot}$ data lack uncertainties and \citet{Lurie2017} used transit durations and ingress/egress times to approximate the EB orbital $e$, leading to inaccurate $e$ determinations.
-Furthermore, biases in the data, e.g.
-the lack of long P$_{orb}$ binaries, inhibit our ability to robustly compare our predictions with the data.
+\xxx{For P$_{orb} \gsim 4$ d, our models produce a substantial number of subsynchronous rotators.
+Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces this population.
+The CPL model, however, struggles to populate the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
+\citet{Lurie2017} find that $15\%$ of their sample with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $8\%$ of our CTL population and only $2\%$ of the CPL population.}
+
+\xxx{Both models predict a large number of extremely subsynchronous rotators with P$_{orb}/$P$_{rot} < 0.7$ across all P$_{orb}$ that is not present in the \citet{Lurie2017} data.
+Magnetic braking creates the lower limit of this subsynchronous population, a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down over 7 Gyr, the longest age considered in our simulations.
+Our choice of prior distributions for both $Q$ and $\tau$ permit very weak tidal interactions that likely gives rise to this population and suggests that our prior does not reflect the underlying distribution of stellar tidal parameters in nature.
+Alternatively, the data could be incomplete where our models predict slowly-rotating subsynchronous rotators as the photometric amplitude of star spot modulations tends to decrease with increasing P$_{rot}$, making reliable rotation periods difficult to detect \citep{McQuillan2014,Lurie2017,Reinhold2018}.}
+
+\xxx{Although the CTL model seems to better reproduce the \citet{Lurie2017} data, both tidal models can reproduce features observed in the \kepler EB distribution, e.g.
+the synchronized population and subsynchronous rotators, suggesting that our models reasonably approximate the dynamical interactions of tidally-evolving, low-mass stellar binaries.
+Our comparison between theory and observations is limited, however, because the \citet{Lurie2017} P$_{rot}$ data lack uncertainties and \citet{Lurie2017} approximated the EB orbital $e$ via transit durations and ingress/egress times, potentially leading to inaccurate $e$ determinations.
+Unconstrained biases in the data, e.g.
+the lack of long P$_{orb}$ binaries, further inhibit our ability to compare our predictions with the data.
+Moreover, our prior distributions were chosen to be plausible, but wide in order to examine our model predictions over parameter space and are not suited for a robust statistical inference to select between which equilibrium tidal model best describes tidal interactions in low-mass binaries stars.
+Below, we offer observational tests that could discriminate between models.}
 
 % extra stuff/cuts
+% On the other hand, the CTL model produces a large number of supersynchronous binaries with eccentric orbits similar to what is seen in the \citet{Lurie2017} data.
+These systems have tidally-locked, but their orbits have not yet circularized, because the tidal locking time scale is 2-3 orders of magnitude less than the circularization timescale \citep{Mazeh2008}.
+% For example under the CTL model, our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
 % Second, In principle, we could correct for such a bias in the \citet{Lurie2017} data and condition our model predictions on the data to infer tidal properties, but that analysis is beyond the scope of this work for reasons we discuss in $\S$~\ref{sec:inference}.
@@
 Our predictions are not strongly dependant on the choice of magnetic braking model, but rather are generic outcomes of the interaction between magnetic braking and tidal torques.
-Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact at longer P$_{orb}$ than have previously been considered, out to P$_{orb} \approx 80$ d.
+Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact at longer P$_{orb}$ than have previously been considered, out to P$_{orb} \approx 60-100$ d.
 Many binaries with P$_{orb} \lsim 20$ d tidally-lock according to both models, in good agreement with previous results, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.

### 2019-05-22 Paper/sync.tex
final edits after 1st review

@@
 In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2018}.
-Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[e.g.
-P$_{rot} >$ P$_{eq}$, ][]{Habets1989,Zahn1994,Keppens1997}.
+Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[P$_{rot} >$ P$_{eq}$, ][]{Habets1989,Zahn1994,Keppens1997}.
 This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations.
@@
 We implement our model in the open-source code VPLanet\footnote{VPLanet is publicly available
-at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep[][Barnes et al., in prep]{Barnes2016,vplanet2018}.
-We integrate all model equations (see $\S$~\ref{sec:methods:stellar} and $\S$~\ref{sec:methods:eqtide}) using using the $4^{th}$ order Runge-Kutta scheme with adaptive timestepping described in \citet{Fleming2018}.
+at \href{https://github.com/VirtualPlanetaryLaboratory/vplanet}{{https://github.com/VirtualPlanetaryLaboratory/vplanet}}.} \citep{Barnes2019}.
+We integrate all model equations (see $\S$~\ref{sec:methods:stellar} and $\S$~\ref{sec:methods:eqtide}) using the $4^{th}$ order Runge-Kutta scheme with adaptive timestepping described in \citet{Fleming2018}.
 
@@
 
-\xxx{Our simplified model assumes that stars follow solid body rotation, whereas in low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution \citep{MacGregor1991,Allain1998,Bouvier2008,Irwin2009}.
-Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior and into the envelope for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass.
-We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal-locking timescales.
+\xxx{Our simplified model assumes that stars follow solid body rotation, whereas in real low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution \citep{MacGregor1991,Allain1998,Bouvier2008,Irwin2009}.
+Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior and into the envelope for up to 1 Gyr, depending on the adopted magnetic braking model, the initial rotation rate, and the stellar mass.
+We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal locking timescales.
 Internal angular momentum transport torques could potentially balance both tidal and magnetic braking torques near the tidally-locked state, producing slight supersynchronous rotation, analogous to the subsynchronous case examined in $\S$~\ref{sec:eq}.
-Modeling core-envelope coupling is beyond the scope of this work, however.}
+Modeling core-envelope coupling is beyond the scope of this work, however.
+}
 
@@
 For example, \citet{Rodriguez2012} show that tidally-interacting bodies can get captured into many spin-orbit resonances states, e.g.
-2:1, 5:2, 4:3, etc.
-Although our models do not resolve capture into such states, we search for evidence of them in data of the spin-orbital states \kepler EBs.
+2:1, 5:2, 4:3, etc\xxx{, and below, we search for evidence of them in data of the spin-orbital states of \kepler EBs.
+Note that our model does not resolve capture into such states as the CPL model, owing to its inclusion of only 4 discrete tidal lags, only allows a body to enter into 3:2 and 1:1 spin - orbit commensurabilities.
+The CTL model predicts a continuous equilibrium period as a function of the P$_{orb}$, $e$, and obliquity, for tidally-locked bodies, only resolving capture into 1:1 synchronous rotation.}
 
@@
 
-\xxx{An additional mechanism for tidal dissipation in low-mass stellar binaries not considered in this work is the dynamical tide.
-Dynamical tidal dissipation arises from the turbulent viscous damping of inertial waves that are excited in the stellar convective envelope by a tidal perturber, with Coriolis acceleration serving as the restoring force \citep{Zahn1975, Ogilvie2007}.
+\xxx{An additional mechanism for tidal dissipation in low-mass stellar binaries is the dynamical tide.
+This effect arises from the turbulent viscous damping of inertial waves that are excited in the stellar convective envelope by a tidal perturber, with Coriolis acceleration serving as the restoring force \citep{Zahn1975,Ogilvie2007}.
 Under the dynamical tide formalism, the stellar mass, evolving stellar structure, rotation rate, and tidal forcing frequency can all strongly impact the strength of tidal dissipation, which can span many orders of magnitude \citep{Ogilvie2007,Ogilvie2013,Mathis2015,Gallet2017}.
-For example, adopting the tidal frequency-averaged model for tidal dissipation of \citet{Ogilvie2013}, both \citet{Mathis2015} and \citet{Gallet2017} show that dynamical tidal dissipation is enhanced on the pre-main sequence via the expansion of the stellar radiative core and rapid rotation, whereas the magnetic braking-driven spin-down on the main sequence decreases the tidal dissipation.
-Dissipation due to the dynamical tide could be important for some of the systems considered in this work since, for coplanar binary stars on circular orbits, inertial waves are excited in the stellar convective envelopes for P$_{orb} > $P$_{rot}/2$ and can drive significant spin and orbital evolution \citep[e.g.][]{Witte2002,Ogilvie2007,Bolmont2016}.
-Although semi-analytic models for dynamical tidal dissipation that account for the evolving stellar structure and rotation exist \citep[e.g.][]{Ogilvie2013,Mathis2015,Bolmont2016,Gallet2017}, we do not consider them here as they are currently limited to circular orbits.
+For example, adopting the tidal frequency-averaged model for tidal dissipation of \citet{Ogilvie2013}, both \citet{Mathis2015} and \citet{Gallet2017} show that dynamical tidal dissipation is enhanced during the pre-main sequence due to the expansion of the stellar radiative core and rapid rotation, whereas the magnetic braking-driven spin-down on the main sequence decreases the tidal dissipation.
+Dissipation due to the dynamical tide could be important for some of the systems considered in this work since, for binary stars on circular orbits, inertial waves are excited in the stellar convective envelopes for P$_{orb} > $P$_{rot}/2$ and can drive significant spin and orbital evolution \citep[e.g.][]{Witte2002,Ogilvie2007,Bolmont2016}.
+Although semi-analytic models for dynamical tidal dissipation that account for the evolving stellar structure and rotation exist \citep[e.g.][]{Mathis2015,Bolmont2016,Gallet2017}, we do not consider them here as they are currently limited to circular orbits.
 We instead focus on exploring the impact of the equilibrium tide across a wide range of parameter space and leave an examination of how the combination of the dynamical and equilibrium tide impacts the rotation period evolution of low-mass binary stars for future work.}
@@
 \xxx{All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.} We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them.
-All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}.
+All code used to run simulations and generate figures is available online.\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}
 
@@
 In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation\xxx{, P$_{rot} >$ P$_{eq}$.
-For long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated magnetic braking regime \citep{Matt2015}.
+For long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated regime \citep{Matt2015}.
 Since magnetic braking scales as P$_{rot}^{-3}$ for unsaturated rotators, e.g.
@@
 For simplicity, we assume both stars are solar-mass with 0 obliquity, a circular binary orbit, and that the torque balance occurs while the stars are on the main sequence where stellar properties change slowly.
-Although solar mass stars are the most massive stars we consider in this work, and hence will have the strongest tidal torque for a given tidal dissipation parameter and P$_{orb}$, they can still exhibit subsynchronous rotation and serve as a useful pedagogical case to examine here and in simulations below.
+Although solar mass stars are the most massive stars we consider in this work, and hence will have the strongest tidal torque for a given tidal dissipation parameter and P$_{orb}$, they can still exhibit subsynchronous rotation and serve as a useful end member case to examine here and in simulations below.
 We display the results of this calculation in Fig.~\ref{fig:analyticBalance}, normalizing P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$.}
@@
 
-\xxx{Subsynchronous rotation can occur across a wide range of tidal parameters and P$_{orb}$.
+\xxx{Our calculations show that subsynchronous rotation occurs across a wide range of tidal parameters and P$_{orb}$.
 In general as tides weaken, i.e.
-increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where tides are weaker.
-Conversely, for strong tides, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for P$_{orb} \lsim 40$ d, tidally-locking binaries into synchronous rotation.
+increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$.
+For strong tides, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for P$_{orb} \lsim 40$ d, tidally-locking binaries into synchronous rotation.
 For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries will rotate subsynchronously for P$_{orb} \gsim 20$ d, with more severe subsynchronism at longer P$_{orb}$.
@@
 Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.
-We truncate each curve when the binary tidally locks.
+We truncate each curve when the binary tidally-locks.
 For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
-   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 7.5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
+   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer})), for initial circular binary orbits with initial P$_{orb} = 7.5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
 Systems with strong tidal torques tidally-lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
@@
 Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster \xxx{(median CPL, CTL P$_{rot} = 22.6$ d and $8.8$ d)} than not locked \xxx{(median CPL, CTL both P$_{rot} = 32.4$ d)} binaries.
-\xxx{The CTL model predicts that the majority tidally-locked binaries, \xxx{$83\%$}, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest.
+\xxx{The CTL model predicts that the majority of tidally-locked binaries, \xxx{$83\%$}, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest.
 The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states as only $46\%$ of locked binaries have P$_{rot} < 20$ d, while the rest can lock out to P$_{rot} \approx 100$ d in long P$_{orb}$ binaries.} More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.
@@
 This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d\xxx{, although binaries with more massive primaries can occasionally lock, or at least tidally-interact, out to P$_{orb} \approx 80$ d, in contrast to the CPL model's predictions.
+Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d\xxx{, although binaries with more massive primaries can occasionally lock, or at least tidally-interact, out to P$_{orb} \approx 80$ d.
 \xxx{We highlight this behavior with a histogram of locked and interacting binaries over P$_{rot}$ for both tidal models in Fig.~\ref{fig:lockedProtHist}.} The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
@@
 
-\xxx{For most values of P$_{rot}$, gyrochronology methods systematically underestimate the ages of tidally-interacting binaries, with the relative age error increasing with decreasing P$_{rot}$ and increasing with primary star mass, where tides are stronger.
+\xxx{For most values of P$_{rot}$, gyrochronology methods systematically underestimate the ages of tidally-interacting binaries, with the relative age error increasing with decreasing P$_{rot}$ and increasing with primary star mass.
 For binaries with P$_{rot} \lsim 30$ d, gyrochronology ages are underestimated by $50\%$, with this error growing to $300\%$ for P$_{rot} \approx 10$ d.
@@
 Instead, owing to the its discrete P$_{eq}$, the CPL model predicts that all tidally-locked supersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.}  This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the \citet{Lurie2017} data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
-\xxx{Neither model reproduces the very supersynchronous, P$_{orb}/$P$_{rot} > 1.6$, binaries in the \citet{Lurie2017} as they typically have orbital eccentricities in excess of 0.3 and are outside of the region of parameter space we consider.
+\xxx{Neither model reproduces the very supersynchronous, P$_{orb}/$P$_{rot} > 1.6$, binaries in the \citet{Lurie2017} data as they typically have orbital eccentricities in excess of 0.3 and are outside of the region of parameter space we consider.
 The CTL model, however, could in principle reproduce these points if they are tidally-locked binaries as its P$_{eq}$ is a continuous function of $e$ and is applicable for large $e$, in contrast to the discrete P$_{eq}$ predicted by the CPL model that is valid for smaller $e$.
@@
 the lack of long P$_{orb}$ binaries, further inhibit our ability to compare our predictions with the data.
-Moreover, our prior distributions were chosen to be plausible, but wide in order to examine our model predictions over parameter space and are not suited for a robust statistical inference to select between which equilibrium tidal model best describes tidal interactions in low-mass binaries stars.
+Moreover, our prior distributions were chosen to be plausible, but wide, in order to examine our model predictions over parameter space and are not suited for a robust statistical inference to select between which equilibrium tidal model best describes tidal interactions in low-mass binaries stars.
 Below, we offer observational tests that could discriminate between models.}
@@
 the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permitted, compared to the continuum of supersynchronous rotation states in eccentric tidally-locked rotators predicted by the CTL model.
-If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model.
+If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model\xxx{, but there is no obvious clustering of \kepler EBs near any spin-orbit resonance}.
 These two tests can fail to discriminate between the CPL and CTL model, however, if the CPL model P$_{eq}$ is a continuous function of $e$, e.g.
@@
 In practice, this is extremely observationally expensive as it requires extensive photometric and spectroscopic observations of many binaries.
-Since there is no obvious clustering of \kepler EBs near any spin-orbit resonance, the above two tests are likely insufficient.
 
@@
 Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
-Understanding the long-term angular momentum evolution of stellar binaries out to P$_{orb} = 100$ d requires detailed modeling of its coupled-stellar tidal evolution, and characterizing tidal dissipation parameters, e.g.
-how $Q$ and $\tau$ vary from star-to-star.
+Understanding the long-term angular momentum evolution of stellar binaries out to P$_{orb} = 100$ d requires detailed modeling of its coupled-stellar tidal evolution, and characterizing tidal dissipation parameters.
 Many new eclipsing stellar binaries will be discovered by TESS \citep[e.g.][]{Sullivan2015,Matson2018} and in analysis of K2 data.
@@
 %% SOFTWARE %%
-\software{matplotlib: \citet{Hunter2007}, numpy: \citet{vanderWalt2011}, pandas: \citet{Mckinney2010}, \vplanet: \citet{Barnes2016,vplanet2018}}
+\software{matplotlib: \citet{Hunter2007}, numpy: \citet{vanderWalt2011}, pandas: \citet{Mckinney2010}, \vplanet: \citet{Barnes2019}}
 

### 2019-07-01 Paper/Response1/response.txt
finished response to ref #2

Dear Dr. Shafter,

Thank you for considering our manuscript. Below you will find our response to the referee's report where I list the issues raised in the report, denoted by REF with the referee's text in quotes for clarity, and how we addressed them in our revised manuscript, denoted by REPLY. Following the referee's comments and suggestions, we have substantially edited portions of the Results section to improve the concision and clarity of our work. These edits merely change the language and phrasing and do not alter the scientific content or our conclusions that were present in the original submitted manuscript. Note that in the revised manuscript, all text changes, except for deletions, are in bold text for emphasis. For each updated figure, we included "Updated Figure" in bold in the figure caption for clarity.

Additionally, we added citations for VPLanet, the software used to perform all of our simulations, to the "Software" section at the end of the manuscript. We did not included STELLAR here as it is a submodule of VPLanet and can only be used as a component of the VPLanet executable when performing our simulations.

Best,

David Fleming, Rory Barnes, Jim Davenport, and Rodrigo Luger

Response to Reviewer
--------------------

Dear Reviewer,

Thank you very much for your careful review of our manuscript.  Your in-depth comments and critiques pointed out weaknesses in our work that, once addressed, improved the quality of our work. Below, I list the issues raised in your report, denoted by REF with your text in quotes, and how we addressed them in our revised manuscript, denoted by REPLY. Note that in the revised manuscript, all text changes, except for deletions, are in bold text for emphasis. For each updated figure, we included "Updated Figure" in bold in the figure caption for clarity.

During the review process, we noticed that an erratum was recently released for the Matt et al. (2015) magnetic braking model, Matt et al., (2019). The erratum impacted the torque normalization in equation 3 in our paper, overestimating it by a factor of 3/2. This error impacted every simulation in our paper that used the Matt+2015 model. We implemented this fix in our code, re-ran all of the simulations, and found that this error had no qualitative impact on our results: we still find that binaries can rotate subsynchronously at short Porb and tidally-lock at long Porb, even out to 100 d. The error weakened the torque due to magnetic braking relative to tidal torques, making it slightly easier for tides to tidally-lock binaries, strengthening our results, in a sense. The number of binaries tidally-locking at long Porb did decrease somewhat, however, as strong spin-down torques due to the combination of magnetic braking and tides are required to spin down long Porb (Porb > 60 d) stars enough to get captured into the tidally-locked state. This decrease, however, was small and did not significantly impact our results. After re-running all our simulations, we re-generated every figure that displayed simulation results that used the Matt+2015 model and updated every quantity derived from our simulation results in the revised manuscript.

Referee Report + Responses
--------------------------

Major points

REF: "As mentioned above, the science presented in the paper is good enough to be published, the lack of concision and proof-reading makes difficult for the reader to estimate the paper at its true value. Many parts can be reduced to make the point of the paper clearer and the reading more convenient. I believe this will help a lot for the acceptance of this work in the community."

REPLY: We apologize for the deficient proofreading and lack of concision in the submitted manuscript. After making the edits to respond to your comments, we thoroughly proofread the manuscript and worked to make our language more direct and concise, making many edits throughout the paper, all in bold.

REF: "Even though the whole range of parameter is explored for equilibrium tides, nothing is said about the importance that dynamical tides could have on the dissipation. Yet, Ogilvie (2007,2013) showed they are a major element of tidal evolution of rotating stars and Mathis (2015) gave a formalism that could be easily implemented in this type of study. In particular, in the case of low-mass stars, it has been shown by eg Gallet & al. (2017) that the tidal dissipation factor (in the CPL formalism) will be strongly affected. I do not expect that the authors re-do all the models for this paper (though I would highly recommend that they do such calculations for a new paper), but the authors need to acknowledge the limitation of their model, at least in the opening section at the end, at best in the description of their model."

REPLY: Thank you for pointing this out.  Tidal dissipation due to the dynamical tide is an important physical consideration that can impact the rotation rates of low mass stars, e.g. in the star-planet tidal interactions examined in Bolmont et al. (2016). To address this point, we added the following text in bold in a new subsubsection, 2.2.5, in the description of our model:

An additional mechanism for tidal dissipation in low-mass stellar binaries not considered in this work is the dynamical tide. Dynamical tidal dissipation arises from the turbulent viscous damping of inertial waves that are excited in the stellar convective envelope by a tidal perturber, with Coriolis acceleration serving as the restoring force (Zahn 1975; Ogilvie+2007). Under the dynamical tide formalism, the stellar mass, evolving stellar structure, rotation rate, and tidal forcing frequency can all strongly impact the strength of tidal dissipation, which can span many orders of magnitude (Ogilvie+2007, Ogilvie 2013, Mathis 2015, Gallet+2017}. For example, adopting the tidal frequency-averaged model for tidal dissipation of Ogilvie (2013), both Mathis (2015) and Gallet+2017 show that dynamical tidal dissipation is enhanced on the pre-main sequence via the expansion of the stellar radiative core and rapid rotation, whereas the magnetic braking-driven spin-down on the main sequence decreases the tidal dissipation. Dissipation due to the dynamical tide could be important for some of the systems considered in this work since, for coplanar binary stars on circular orbits, inertial waves are excited in the stellar convective envelopes for P$_{orb} > $P$_{rot}/2$ and can drive significant spin and orbital evolution (e.g. Witte+2002, Ogilvie+2007, Bolmont+2016). Although semi-analytic models for dynamical tidal dissipation that account for the evolving stellar structure and rotation exist (e.g. Mathis 2015, Bolmont+2016, Gallet+2017), we do not consider them here as they are currently limited to circular orbits. We instead focus on exploring the impact of the equilibrium tide across a wide range of parameter space and leave an examination of how the combination of the dynamical and equilibrium tide impacts the rotation period evolution of low-mass binary stars for future work.

REF: "It can be interesting to compare the effects of different magnetic braking models, and I understand that it allows a connection to the computations done in Fleming et al. (2018) using the same torque. But the rotation evolution trends provided by Reiners & Mohanty (2012)'s paper have been discussed at length in the literature as not realistic. They did a great work and were the first to show a direct dependence to the Rossby number and a cluster comparison for different mass at the same time but their converged sequence for solar-like stars (1.0-0.4 M_\odot) is evolving in the wrong way with higher mass stars spinning slower once converged (see eg their fig 4 for Hyades/Praesepe). Moreover, this comparison is not mentioned in the abstract nor discussed in the conclusions. I would remove this part that doesn't seem to be needed for this paper and just add a sentence about it with something like "we tested another braking law and the conclusion's were no different"."

REPLY: Thank you for pointing this discussion out and explaining the short-comings of the model. As per your suggestion, we removed our examination and discussion of simulations using the Reiners & Mohanty (2012) model. Specifically, we removed Sections 4.1.3 and 4.4.2 and Figures 5 and 14 in the original submitted manuscript.

REF: "Abstract : CPL is not define"

REPLY: The CPL model is now properly defined in the abstract.

1.Introduction

REF: "-Reiners & Mohanty (2012) didn't explore the Kepler field but did comparison with Praesepe"

REPLY: We corrected this errant citation.

REF: "Gyrochronlogy => Gyrochronology (A lot of typos are present all over the text, a bit of proof reading would be nice)"

REPLY: We corrected these typos in the text and apologize for the lack of sufficient proofreading in the initial submission.

REF: "Abt & Booyarak (2004) found that long periods binary's Prot can deviate from that of single stars was applicable for A and B type stars. Is that expected to be valid for lower masses ?"

REPLY: Thank you for pointing this out.  After reviewing the literature, we believe that that connection is a bit too speculative as tidal processes in early and late type stars are likely not similar due to the different stellar structures, e.g. radiative vs convective envelope. Mazeh (2008) explained how Zahn (1975) showed that the tidal orbital circularization timescales differ between stars with radiative and convective envelopes, so we would anticipate that such a difference exists between the synchronization timescales as well. We have therefore removed the reference to Abt & Booyarak (2004) in both the introduction and in Section 4.3 as it is not applicable as we had originally believed.

REF: "Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyro chronology could be incorrect owing to the influence of tidal torques. No previous study has quantified this effect". Indeed, it would be a really nice output of that paper or a following one."

REF: "4.4 A lot can be cleaned here. It would have been more relevant to mimic a realistic population of binaries in the field to estimate the impact of binarity on gyrochronology."

REPLY: We grouped these comments as they touched on a similar point. Thank you for pointing out this deficiency in our results: we did not properly quantify the impact of tidally-interacting binaries on age estimates using gyrochronology. To address this deficiency, we created a new figure, Figure 13 in the revised manuscript, that quantifies the systematic error introduced by applying gyrochronology methods to tidally-interacting binaries. To quantify this error, we computed the the difference between the mean ages of single stars and tidally-interacting binaries in mass - Prot bins for both the CPL and CTL model simulation suites. We defined "tidally-interacting binaries" as binaries that were tidally-locked or had Prots within 10% of Peq, i.e. we combined the "Interacting" and "Locked" populations examined in Figures 9 and 10 in the revised manuscript. We found that gyrochronology methods can underpredict ages by up to 300% at short Prot and overpredict ages by up to 25% at long Prot. Only tidally-interacting binaries populated the Prot < 10 d and > 50 d bins for solar mass stars, so we could not compare to the ages of single stars in these cases. We expanded on these points in more detail and significantly edited Section 4.4, "Deviations From Single Star Prot Evolution: Implications for Gyrochronology", in the revised manuscript to accommodate these changes and clean up the section to improve the concision of our writing.

2. Methods

REF: "STELLAR is not a stellar evolution model. Be careful with the vocabulary used here, at best it can be considered as an interpolation tool in Baraffe+98's tracks. (Same comment at the beginning of the last paragraph of 2.1)"

REPLY: Thank you for pointing out this error in our description of the model. We have corrected this error in both locations in the text.

REF: "I assume it is only a typo in the text but an exponent is missing in the descriptions of Matt+15 magnetic braking in eq. 2 for the convective turnover timescale."

REPLY: We corrected this typo and confirmed that our code used the correct exponent in all simulations.

REF: "2.2.3 (end) What prevents your model from capturing other spin-orbit resonances ?"

REPLY: The CPL model, owing to its inclusion of only 4 discrete tidal lags, only allows a body to enter into 3:2 and 1:1 spin - orbit commensurabilities. The CTL model does not resolve capture into spin-orbit resonances besides 1:1 since it predicts a continuous equilibrium period as a function of the orbital period, eccentricity, and obliquity, for tidally-locked bodies. See Sections 2.1 and 2.2 in Barnes (2017) for an expanded discussion. Models that do resolve captures into spin-orbit resonances, e.g. Rodriguez et al. (2012), instead directly integrate the equations of motion and torques for tidally-interacting bodies, whereas our model is purely secular.

3. Simulations

REF: "I know it is not expected to evolve very much in that direction but what happened in the eccentricity increases and overcomes 0.3 in the CPL case ?"

REPLY: In the CPL model, the eccentricity can grow for certain orbital and rotational states (see Equation 8 in Barnes 2017), such that the eccentricity could increase to beyond 0.3. In such a case, the CPL model can still be applied, however, it is not strictly valid as it is in effect derived by truncating the equations for tidal evolution to second order in eccentricity, whereas the CTL model is 8th order in eccentricity. Leconte et al. (2010) argued that this truncation renders the predictions of the CPL model incorrect in this eccentricity regime. Specifically, Leconte at al. (2010) found that for e > 0.32, the higher-order eccentricity terms dominate the constant and second order terms, meaning the the CPL model does not accurately capture the physics of the tidal interaction for moderate to large eccentricities. In this case, Leconte et al. (2010) showed that this can lead to semi-major axis growth under the 2nd order CPL model instead of the decrease predicted by the CTL model. Moreover, Leconte et al. (2010) found that under the CPL model for pseudo-synchronously (supersynchronously) binaries with eccentricities > 0.326, the eccentricities can grow, whereas the CTL model predicts that the eccentricity should damp. Leconte et al. (2010) argue that the latter is the correct behavior. We added a brief reference to this discussion in Leconte et al. (2010) in the text to clarify this point.

REF: "Respectfully" -> Respectively (and elsewhere in the text)"

REPLY: We have corrected all of these errors in the text.

4. Results

REF: "4.1.1 Low-mass stars can be in the saturated regime for very slow rotation rates + In your range of parameter, a binary 1Msun-1Msun is an extreme case with maximal tidal effect. So it doesn't seem very appropriate for a representative example as fig. 3."

REPLY: Those are good points, however, we chose to use a 1Msun - 1 Msun binary as a pedagogical case because even with the extreme tidal torques in such a binary, tides and magnetic braking can still balance to produce subsynchronous rotation, similar to how we consider short Porb systems where tides are strong and in addition to long Porb cases up to 100 d. Furthermore, given our assumptions and under the Matt+2015 magnetic braking model, both stars would be in the unsaturated regime, justifying our assumptions. We clarified these assumptions in Section 4.1.1 in the revised manuscript and by adding the following text:

Although solar mass stars are the most massive stars we consider in this work, and hence will have the strongest tidal torque for a given tidal dissipation parameter and P$_{orb}$, they can still exhibit subsynchronous rotation and serve as a useful pedagogical case to examine here and in simulations below.

Additionally, we took this opportunity to update Fig. 3 by making the range of k2 * tau values more symmetric about our fiducial values and added contours for clarity.

REF: "4.1 is very interesting and the core of the paper. It can however be improved with more concision."

REPLY: We edited and trimmed down Section 4.1 to improve both the concision and focus. We bolded nearly all the text in Section 4.1 as most portions were edited to improve concision without any substantial nor qualitative change to the scientific content of the text.

REF: "4.1.4 would be more appropriate as part of section 4.2 in my opinion"

REPLY: Upon further reflection, we agree that 4.1.4 would be better suited as part of Section 4.2. It is now Section 4.2.1 in the revised manuscript and we have slightly edited the text near the end of Section 4.2 and beginning of 4.2.1 to accommodate its inclusion.

REF: "4.2 Fig 8 and 9 needs to be adjusted and reshaped, legend, white is not very visible on yellow, etc... They are both really interesting and I spent a lot of time understanding each part of the diagram but I was disappointed that no discussions were made about the narrowness and the wideness of some constraining intervals, particular around for the CPL model. Thus I advise either to remove the interval in the figure, that would allow it to be smaller and clearer and there is already a lot to say about the figure without them. Either to add some comment in the text in a short and concise way."

REPLY: We removed the intervals from the figure since displaying the median values still capture the physics of the various spin-orbital states examined in said figure. We significantly increased the font size to improve the figure's legibility.

REF: "4.3 It would be interesting given the type of diagram of figure 10 and 11, to compare to the Kepler rotation period field by McQuillan+2014 for example and to look at what could be some missed binaries."

REPLY: We are glad you think this as we believe it would be an interesting study as well! We presently are pursuing this very point in a separate project using the same model presented in the submitted manuscript. We presented this work-in-progress at the 2019 meeting of the American Astronomical Society (Here's an ADS link to the abstract for reference: http://adsabs.harvard.edu/abs/2019AAS...23334101F) and leave this for a future publication.

REF: "Also, I don't think anyone ever gave an age to a star with gyrochronology using Reiners&Mohanty (2012). I wouldn't recommend the use of the two different saturations value in a single population model."

REPLY: We have removed the discussion and analysis involving the Reiners & Mohanty (2012) model.

REF: "4.4.1 has actually been already discussed in 4.3."

REPLY: We have removed this subsubsection as it mainly repeated content from 4.3.

REF: "4.5 This section has to be present but should be twice smaller in my opinion and merged with 4.6 and 4.7, the message is very diluted here. Also, I'm surprised by the conclusion of the authors, while they claim that it is hard to make a clear statement, it seems very clear to me that the CTL model is doing a better job a reproducing the field by Lurie+2017. Particularly because the CPL models seems enable to produce super-synchronous rotators. But this is only an eye judgement. I'm also surprise to see that nothing is said about the highly super synchronous rotators with high eccentricity in Lurie+2017 sample. Are they still in the modelled range of parameters ? Should the authors expect their models to reproduce these binary ?"

REPLY: As addressed in your point below, we removed section 4.7 and compressed its point to a few sentences added to the Discussion. We wish to keep 4.5 and 4.6 separate as 4.5 deals with examining how our model predictions compare to the Lurie+2017 data whereas 4.6 offers general observational tests to distinguish between the CPL and CTL model.

We agree that qualitatively, the CTL model does a better job reproducing the Lurie+2017 data than the CPL model as the former populates regions of parameter space that the CPL model fails to do, e.g. the supersynchronous population noted by the reviewer. We choose to be cautious with our wording and stick with a more qualitative interpretation as our goal in this work was not to perform inference, but instead to examine the model predictions across a wide range of parameter space. To this end, we chose our tidal dissipation parameter and mass prior distributions to be intentionally broad such that they span a plausible region of parameter space. For inference and proper statistical model selection, we would have instead used a more proper mass prior, perhaps sampling the primary star masses, with replacement, from the masses in the Kepler Input Catalogue, or by using a reasonable initial mass function, but still accounting for Kepler's bias towards predominantly detecting FGK stars. Furthermore, since the underlying distribution of tidal dissipation parameters as a function of stellar mass and age is not well known, we did not believe that any inference between the CPL and CTL models we made would be robust. When performing inference, we could instead test the impact of different distributions of tidal dissipation parameters, or adopt a model for tidal dissipation as a function of stellar mass and age, like the Mathis (2015) model. Moreover, even though the Lurie+2017 data is currently the best available for such studies, its lack of uncertainties on Prot and eccentricties, and the approximate means in which the eccentricities were derived, prevent a robust comparison. We have significantly edited section 4.5, with most of the text reflecting the changes in bold in the revised manuscript, to compress the section, clarify the message, and address the differences between the CTL and CPL model predictions when compared to the Lurie+2017 data. We additionally worked to shorten the section and note that these edits did not significantly alter the scientific content nor conclusions.

As for the highly supersynchronous rotators, our CTL model could reproduce them if they are tidally-locked binaries with eccentricities > 0.3, beyond the range we considered. In this case for large eccentricities, the CTL model equilibrium period, our equation 22, predicts that binaries on more eccentric orbits rotate more supersynchronously, e.g. Fig. 1 in Barnes (2017), so if we considered more eccentric binaries, our model could reproduce that observation. We have updated the colorbar for Fig. 15 to note that the highly supersynchronous rotators in the Lurie+2017 data have eccentricities > 0.3 and added the following text discussing why our model does not reproduces these systems in the revised manuscripts:

Neither model reproduces the very supersynchronous, P$_{orb}/$P$_{rot} > 1.6$, binaries in the \citet{Lurie2017} as these binaries have orbital eccentricities in excess of 0.3 and are outside of the region of parameter space we consider. The CTL model, however, could in principle reproduce these points if they are tidally-locked binaries as its P$_{eq}$ is a continuous function of $e$ and is applicable for large $e$, in contrast to the discrete P$_{eq}$ predicted by the CPL model that is valid for smaller $e$.

REF: "4.6 The discussion around spin-orbit resonances as the ones by Rodriguez et al (2012) has already been discussed in much more positive terms in Sect. 2.2.3. + See comment above about the conclusion on CTL or CPL and the comparison to Lurie+2017 data."

REPLY: We removed the repeated discussion of spin-orbit resonances from this section. Furthermore, we cut out additional unnecessary text in this section to make its points more concise.

REF: "4.7 I'm sorry, I don't really understand the point of this section, at best it could be added in the conclusion in a sentence or two about future perspectives."

REPLY: We removed this section and compressed the point to the following few sentences in the discussion as suggested by the reviewer:

Beyond these tests, our model could be used to infer the tidal properties of binary stars, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties and reasonable prior probability distributions for parameters like the initial binary $e$. This analysis, however, is also beyond the scope of this work and we leave it for future endeavors.

5. Discussion

REF: "We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries." => Which observational tests ? The discussion would be a good place to summarise them."

REPLY: We added the following text in the discussion to briefly summarize our proposed observational tests:

We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries. We primarily suggest that observers measure stellar P$_{rot}$ in binaries with solar-mass primaries at P$_{orb} 60-100$ d. If any tidally-locked binaries are identified at long P$_{orb}$, this would be evidence in favor of the CPL model as we found that only binaries tidally-interacting via the CPL model could tidally-lock at such long P$_{orb}$. At shorter P$_{orb}$, precise measurements of P$_{rot}$ and binary $e$ and P$_{orb}$ could distinguish between the CPL and CTL model in tidally-locked systems, e.g. identifying if P$_{eq}$ follows Eq.~(\ref{eqn:cpl:eqPer}) vs. Eq.~(\ref{eqn:ctl:eqPer}), especially if the CPL P$_{eq}$ is in fact a discrete function of $e$. The observations required by these tests, however, are non-trivial.

Additional comments

REF: "These low-masse stars are expected to have their core and envelope rotation evolution decoupled and to re-couple on a long time-scale, see eg Gallet & Bouvier (2015). How would that affect the results of the authors ? In what direction ? In the aforementioned paper, they seem to have some fast rotating stars at late ages that are non-explained by their single star models. Do the authors think their binary model could help to reproduce these ?"

REPLY: As explained by Gallet et al. (2013) and Gallet et al. (2015), the radiative core could act as an angular momentum reservoir for the envelope for up to 1 Gyr. This effect would work to spin up stars early on the main sequence against the spin-down induced by tidal torques and magnetic braking. We anticipate that this would increase the tidal locking timescales, but not significantly since the transfer lasts for only up to 1 Gyr. One interesting situation that could occur, however, is if the torque from the convective envelope's spin-up from internal angular momentum transportation balances the spin-down from tidal and magnetic braking torques. We would expect that in this case, a binary star could be rotating slightly faster than its expected equilibrium rotation period from tidal locking, analogous to the subsynchronous cases examined in Section 4.1 in this work. This state would likely not persist for long, however, as the internal angular momentum transport does not persist for the age of many systems.

To address this, we added the following text to a new subsubsection, 2.1.1: Core-Envelope Coupling, in the Methods section:

Our simplified model assumes that stars follow solid body rotation, whereas in low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution \citep{MacGregor1991,Allain1998,Bouvier2008,Irwin2009}. Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior and into the envelope for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass. We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal locking timescales. Internal angular momentum transport torques could potentially balance both tidal and magnetic braking torques near the tidally-locked state, producing slight supersynchronous rotation, analogous to the subsynchronous case examined in $\S$~\ref{sec:eq}. Modeling core-envelope coupling is beyond the scope of this work, however.

The fast rotating stars at late ages examined by Gallet & Bouvier (2015) could be explained by our model if they are in fact unresolved binaries at relatively short orbital periods. For example, for sufficiently strong tidal parameters, our theory shows that binaries can tidally-lock in good agreement with previous results, e.g. Zahn & Bouchet (1989). In this case, the tidal locking maintains rapid rotation against the spin-down caused by magnetic braking at the expense of orbital angular momentum, often keeping Prot <= 20 d. We find that this state of tidally-induced rapid rotation can persist for Gyrs across a wide range of primary star masses, e.g. Figs. 9, 10, and 14 in the revised manuscript.

Other changes not suggested by the reviewer:

-Throughout the revised manuscript, we corrected typos, grammatical errors, and sentence structure with the corrected text in bold.

- Near the end of the introduction, we performed some minor text edits to improve the concision and removed a redundant phrase. The edited sentence appears in bold in the revised manuscript.

-At the end of Section 2.2.2 in the revised manuscript, we added the following text to further clarify how our results can be interpreted in the context of other similar works as a function of tidal Q:

"For example, a common re-parameterization of Q is the reduced tidal quality factor, $Q' = 3Q/2k_2$ \citep[e.g.][]{Leconte2010}. Given our choice of $k_2 = 0.5$, this reduces to $Q' = 3 Q$.""

-Given that the Matt+2015 model erratum effectively strengthened tides relative to magnetic braking, all the CPL simulations in Fig. 5 in the original manuscript actually tidally-locked after we implemented the correction to the Matt+2015 model instead of rotating subsynchronously, e.g. in the Q = 10^8 case. We re-ran the simulations in Fig. 5 in the revised manuscript, keeping all initial conditions the same, but assuming a slightly longer initial Porb = 7.5 d to get weaker tidal torques. The new simulations now again displays how subsynchronous rotation can occur for short Porb systems and highlights how subsynchronous rotators can transition from subsynchronous to synchronous rotation via the mechanism described in the text. The new simulations qualitatively replicated the evolution seen in original Figure 5 in the submitted manuscript and do not change our results. We noted these changes in bold in the text in Section 4.1.3 and in the Fig. 5 caption in the revised manuscript.

-At the end of the 1st paragraph in Section 4.3 in the revised manuscript, we added the text "consistent with ages of stars in the \kepler field \citep{Chaplin2014}" to clarify our choice of age distribution.

-Following the referee's emphasis on concision, we consolidated much of Section 4.3 to remove redundant points and improve the readability of this section. The scientific content and conclusions remain unchanged, but the language has been edited for clarity. We highlighted the changes in bold text in the revised manuscript.

-------

Again, thank you for the detailed review of our manuscript and we look forward to the next correspondence.

Best,

David Fleming, Rory Barnes, Jim Davenport, and Rodrigo Luger

### 2019-07-01 Analysis/tidalExample.py

@@ -20,6 +20,13 @@ mpl.rcParams['font.size'] = 22.0
 mpl.rc('font',**{'family':'serif','serif':['Computer Modern']})
 mpl.rc('text', usetex=True)
 
+# Define physical constants
+BIGG = 6.674e-8 # Universal gravitational constant in cgs
+MSUN = 1.988416e30 # Mass of sun in g
+RSUN = 6.957e10 # Radius of sun in cm
+AUCM = 1.49597870700e11 # cm per AU
+DAYSEC = 24.0 * 3600.0 # seconds per day
+
 # Path to sim results
 cplDir = "../Sims/TidalExample/CPL"
 ctlDir = "../Sims/TidalExample/CTL"
@@ -34,9 +41,9 @@ cpl = np.genfromtxt(os.path.join(cplDir,"bintides.secondary.forward"))
 ctl = np.genfromtxt(os.path.join(ctlDir,"bintides.secondary.forward"))
 
 # Plot e, Porb, Prot
-fig, axes = plt.subplots(nrows=3, figsize=(6,16), sharex=True)
+fig, axes = plt.subplots(nrows=3, figsize=(6,15), sharex=True)
 
-# Left: e
+# Top: e
 axes[0].plot(cpl[:,0], cpl[:,5], lw=3, ls="-", color="C0", label="CPL")
 axes[0].plot(ctl[:,0], ctl[:,5], lw=3, ls="-", color="C1", label="CTL")
 
@@ -49,7 +56,7 @@ axes[1].plot(ctl[:,0], ctl[:,6], lw=3, ls="-", color="C1", label="CTL")
 
 axes[1].set_ylabel("Orbital Period [d]")
 
-# Right: Prot
+# Bottom: Prot
 axes[2].plot(cpl[:,0], cpl[:,4], lw=3, ls="-", color="C0", label="CPL")
 axes[2].plot(ctl[:,0], ctl[:,4], lw=3, ls="-", color="C1", label="CTL")
 
@@ -61,7 +68,6 @@ for ax in axes:
     ax.set_xscale("log")
     ax.axvline(cplLock, lw=3, ls="--", color="C0")
     ax.axvline(ctlLock, lw=3, ls="--", color="C1")
-
 
 axes[2].set_xlabel("Time [yr]")
 

### 2019-07-01 Paper/sync.tex
Updates from Overleaf

@@
 We examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries up to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models.
-We find that many binaries with P$_{orb} \lsim 20$ d tidally-lock, and most with $P_{orb} \lsim 4$ d tidally-lock into synchronous rotation on circularized orbits.
+We find that many binaries with P$_{orb} \lsim 20$ d tidally lock, and most with $P_{orb} \lsim 4$ d tidally lock into synchronous rotation on circularized orbits.
 At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.
 In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, qualitatively reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}.
-Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the \xxx{Constant Phase Lag tidal} model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d.
+Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the Constant Phase Lag tidal model predicts that binaries can tidally lock out to P$_{orb} \approx 100$ d.
 Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
-gyrochronology \xxx{can underpredict ages by up to $300\%$} unless one accounts for binarity.
+gyrochronology can underpredict ages by up to $300\%$ unless one accounts for binarity.
 We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.
@@
 Furthermore, at short P$_{orb}$, tidal locking naturally producing a population of fast rotators that cannot be reproduced by single-star-only models of magnetic braking.
-Many binaries can tidally-lock even out to P$_{orb} \approx 100$ d, fixing P$_{rot}$ to P$_{orb}$, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
+Many binaries can tidally lock even out to P$_{orb} \approx 100$ d, fixing P$_{rot}$ to P$_{orb}$, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e.
 gyrochronology methods must be applied carefully.
@@
 Once stars reach the main sequence, stellar radii remain mostly constant while magnetic braking removes angular momentum from the stars, gradually spinning them down over time \citep{Skumanich1972}.
-Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \textbf{\citep[e.g.
-Praesepe, ][]{Reiners2012,Matt2015,Douglas2017}} and field stars \citep[e.g.
+Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \citep[e.g.
+Praesepe, ][]{Reiners2012,Matt2015,Douglas2017} and field stars \citep[e.g.
 the \kepler field, ][]{Matt2015,vanSaders2018}.
-Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as \xxx{gyrochronology} \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.
+Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as gyrochronology \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.
 
@@
 Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially for stellar spins.
-Tidal torques drive binaries towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
+Tidal torques drive binaries towards the tidally locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
 Tidal-locking occurs much earlier than orbital circularization with the tidal-locking timescale estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is typically much less angular momentum in stellar spins than the binary orbit.
@@
 When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$, or more generally P$_{rot} =$ P$_{eq}$ for eccentric orbits.
-In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2018}.
-Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[P$_{rot} >$ P$_{eq}$, ][]{Habets1989,Zahn1994,Keppens1997}.
+In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2018}.
+Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally locked state into subsynchronous rotation \citep[P$_{rot} >$ P$_{eq}$, ][]{Habets1989,Zahn1994,Keppens1997}.
 This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations.
-\xxx{The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.}
+The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.
 
@@
 % The influence of tides is not necessarily limited to short P$_{orb}$ systems, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered.
-%The large sample of $816$ P$_{rot}$ measurements derived using star spot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
+%The large sample of $816$ P$_{rot}$ measurements derived using star spot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d.
 Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking.
-These findings suggest that synchronization is not a given for tidally-interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$.
+These findings suggest that synchronization is not a given for tidally interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$.
 Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.
@@
 In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.
-\xxx{Tidally-interacting binaries impart a} contaminating signal that is not currently accounted for by models.
+Tidally-interacting binaries impart a contaminating signal that is not currently accounted for by models.
 Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology could be incorrect owing to the influence of tidal torques.
@@
 We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.
-We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of \xxx{gyrochronology} models fail in such systems.
+We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronology models fail in such systems.
 We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.
@@
 
-We improve upon the \xxx{interpolation of the \citet{Baraffe2015} stellar evolution models employed by \citet{Fleming2018}, \stellar, by additionally performing a bicubic interpolation of the stellar radius of gyration, $r_g$, over mass and time of the \citet{Baraffe2015} models.
-This updated version of \stellar now tracks the full moment of inertia evolution of low-mass stars according to the \citet{Baraffe2015} stellar evolution models, a critical requirement for modeling the angular momentum evolution of low-mass stars.}
-
-\xxx{We simulate magnetic braking using the model derived by \citet{Matt2015} as this formalism} has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
-This model \xxx{depends} on the stellar Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
+We improve upon the interpolation of the \citet{Baraffe2015} stellar evolution models employed by \citet{Fleming2018}, \stellar, by additionally performing a bicubic interpolation of the stellar radius of gyration, $r_g$, over mass and time of the \citet{Baraffe2015} models.
+This updated version of \stellar now tracks the full moment of inertia evolution of low-mass stars according to the \citet{Baraffe2015} stellar evolution models, a critical requirement for modeling the angular momentum evolution of low-mass stars.
+
+We simulate magnetic braking using the model derived by \citet{Matt2015} as this formalism has been shown to successfully model the spin-down of low-mass stars across many ages in both the Praesepe cluster and in the \kepler field.
+This model depends on the stellar Rossby number, $Ro = $ P$_{rot}/\tau_{cz}$, the ratio of the stellar P$_{rot}$ to the stellar convective turnover timescale, $\tau_{cz}$.
 The \citet{Matt2015} model predicts that below a certain $Ro$ for rapidly-rotating stars, stellar magnetic activity saturates at a constant value, producing a magnetic braking torque that is directly proportional to the stellar rotation rate.
@@
 \end{equation}
-\xxx{while for more slowly-rotating unsaturated stars,}
+while for more slowly-rotating unsaturated stars,
 \begin{equation} \label{eqn:mattUnSat}
@@
 Saturated magnetic braking occurs for $Ro \leq Ro_{\odot}/\chi$ for $\chi = 10$ where \citet{Matt2015} defines $\chi = Ro_{\odot}/Ro_{sat}$.
-We adopt all model parameters given in Table 1 from \citet{Matt2015}\xxx{, with the correction from \citet{Matt2019},} and compute $\tau_{cz}$ using Eqn.
+We adopt all model parameters given in Table 1 from \citet{Matt2015}, with the correction from \citet{Matt2019}, and compute $\tau_{cz}$ using Eqn.
 (36) from \citet{Cranmer2011}.
@@
 
-\subsubsection{\xxx{Core-Envelope Coupling}} \label{sec:methods:twoLayer}
-
-\xxx{Our simplified model assumes that stars follow solid body rotation, whereas in real low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution \citep{MacGregor1991,Allain1998,Bouvier2008,Irwin2009}.
+\subsubsection{Core-Envelope Coupling} \label{sec:methods:twoLayer}
+
+Our simplified model assumes that stars follow solid body rotation, whereas in real low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution \citep{MacGregor1991,Allain1998,Bouvier2008,Irwin2009}.
 Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior and into the envelope for up to 1 Gyr, depending on the adopted magnetic braking model, the initial rotation rate, and the stellar mass.
 We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal locking timescales.
-Internal angular momentum transport torques could potentially balance both tidal and magnetic braking torques near the tidally-locked state, producing slight supersynchronous rotation, analogous to the subsynchronous case examined in $\S$~\ref{sec:eq}.
+Internal angular momentum transport torques could potentially balance both tidal and magnetic braking torques near the tidally locked state, producing slight supersynchronous rotation, analogous to the subsynchronous case examined in $\S$~\ref{sec:eq}.
 Modeling core-envelope coupling is beyond the scope of this work, however.
-}
-
-\subsubsection{\xxx{Example Stellar Evolution}} \label{sec:methods:exampleStellar}
-
-In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to \xxx{\stellar and the \citet{Matt2015} magnetic braking model}.
+
+\subsubsection{Example Stellar Evolution} \label{sec:methods:exampleStellar}
+
+In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to \stellar and the \citet{Matt2015} magnetic braking model.
 We assume all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr.
@@
 	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
-   \caption{\xxx{Updated Figure} Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \xxx{\citet{Matt2015} magnetic braking model}.
+   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} magnetic braking model.
 Each dot denotes the approximate time when each star reaches the main sequence.}%
@@
  Equilibrium tidal models, first introduced by \citet{Darwin1880}, track the secular evolution of an orbiter's semi-major axis, $a$, eccentricity, $e$, and the rotation rates, $\omega_i$, and obliquities $\psi_i$, of both gravitating bodies due to tidal torques.
-Equilibrium tidal models assume that tidally-interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the bodies' centers of mass due to friction within each body.
+Equilibrium tidal models assume that tidally interacting bodies raise tidal bulges on their companions that remain offset from the line connecting the bodies' centers of mass due to friction within each body.
 This assumption is typically referred to as the ``weak friction approximation" \citep{Zahn2008}.
@@
 \end{equation}
-where if the $i^{th}$ body is tidally-locked in a synchronous orbit,
+where if the $i^{th}$ body is tidally locked in a synchronous orbit,
 \begin{equation} \label{eqn:cpl:dadt_locked}
@@
 
-The ``Constant Time Lag" (CTL) \citep[][]{Hut1981,Leconte2010} equilibrium tidal model assumes a constant time interval between the body's tidal bulge and the passage of the tidally-interacting companion.
+The ``Constant Time Lag" (CTL) \citep[][]{Hut1981,Leconte2010} equilibrium tidal model assumes a constant time interval between the body's tidal bulge and the passage of the tidally interacting companion.
 In this formalism, unlike the CPL model, the CTL model is continuous over a range of tidal wave frequencies and applicable for large $e$.
-\xxx{However,} if the assumption of linearity is relaxed, i.e.
+However, if the assumption of linearity is relaxed, i.e.
 frequencies associated with tidal bulges are allowed to depend on a spin or orbital forcing frequency, then this model is only valid over a small range of frequencies \citep{Greenberg2009}.
@@
 Any constraints we derive as a function $Q$ or $\tau$ can trivially be scaled to other values of $k_2$.
-\xxx{For example, a common re-parameterization of Q is the reduced tidal quality factor, $Q' = 3Q/2k_2$ \citep[e.g.][]{Leconte2010}.
-Given our choice of $k_2 = 0.5$, this reduces to $Q' = 3 Q$.}
+For example, a common re-parameterization of Q is the reduced tidal quality factor, $Q' = 3Q/2k_2$ \citep[e.g.][]{Leconte2010}.
+Given our choice of $k_2 = 0.5$, this reduces to $Q' = 3 Q$.
 
@@
 
-Tidal torques drive a body's rotation rate towards the tidally-locked state.
+Tidal torques drive a body's rotation rate towards the tidally locked state.
 When a body tidally locks, tidal torques fix P$_{rot}$ to the equilibrium P$_{rot}$, P$_{eq}$.
@@
 when P$_{rot} = $ P$_{eq} = $ P$_{orb}$.
-Although spin-orbit synchronization is an expected outcome of tidal evolution \citep{Counselman1973}, in general for tidally-locked bodies on non-circular orbits, both the CPL and CTL model predict pseudosynchronous, or supersynchronous rotation, e.g.
+Although spin-orbit synchronization is an expected outcome of tidal evolution \citep{Counselman1973}, in general for tidally locked bodies on non-circular orbits, both the CPL and CTL model predict pseudosynchronous, or supersynchronous rotation, e.g.
 Mercury's 3:2 spin-orbit resonance \citep[P$_{rot} = 2/3$ P$_{orb}$,][]{GoldreichPeale1966}.
@@
 \end{equation}
-Therefore, the CPL model predicts synchronous rotation for $e \lsim 0.23$, and a supersychronous 3:2 spin-orbit state otherwise for tidally-locked rotators.
-
-We note that two discrete rotation states are not the only permitted ones for tidally-locked systems under the CPL formalism.
+Therefore, the CPL model predicts synchronous rotation for $e \lsim 0.23$, and a supersychronous 3:2 spin-orbit state otherwise for tidally locked rotators.
+
+We note that two discrete rotation states are not the only permitted ones for tidally locked systems under the CPL formalism.
 For example, an alternate derivation of P$_{eq}$ for orbiters with rotation axes perpendicular to the orbital plane under the CPL model predicts
@@
 \end{equation}
-The CTL model predicts that bodies on eccentric orbits tidally-lock into supersyncronous rotation, and only bodies with aligned spins on circular orbits are synchronous rotators.
-
-In general, a continuous P$_{eq}$ and the discrete 1:1 and 3:2 spin-orbit commensurabilities are not the only equilibrium rotation states for tidally-locked rotators predicted by equilibrium tidal models.
-For example, \citet{Rodriguez2012} show that tidally-interacting bodies can get captured into many spin-orbit resonances states, e.g.
-2:1, 5:2, 4:3, etc\xxx{, and below, we search for evidence of them in data of the spin-orbital states of \kepler EBs.
+The CTL model predicts that bodies on eccentric orbits tidally lock into supersyncronous rotation, and only bodies with aligned spins on circular orbits are synchronous rotators.
+
+In general, a continuous P$_{eq}$ and the discrete 1:1 and 3:2 spin-orbit commensurabilities are not the only equilibrium rotation states for tidally locked rotators predicted by equilibrium tidal models.
+For example, \citet{Rodriguez2012} show that tidally interacting bodies can get captured into many spin-orbit resonances states, e.g.
+2:1, 5:2, 4:3, etc, and below, we search for evidence of them in data of the spin-orbital states of \kepler EBs.
 Note that our model does not resolve capture into such states as the CPL model, owing to its inclusion of only 4 discrete tidal lags, only allows a body to enter into 3:2 and 1:1 spin - orbit commensurabilities.
-The CTL model predicts a continuous equilibrium period as a function of the P$_{orb}$, $e$, and obliquity, for tidally-locked bodies, only resolving capture into 1:1 synchronous rotation.}
+The CTL model predicts a continuous equilibrium period as a function of the P$_{orb}$, $e$, and obliquity, for tidally locked bodies, only resolving capture into 1:1 synchronous rotation.
 
@@
 To ensure that tidal torques dominate over torques due to magnetic braking and stellar evolution when forcing tidal-locking, we additionally require that the P$_{rot}$ derivative points towards $P_{eq}$ on both sides of P$_{eq}$, i.e.
-when the gradient of P$_{rot}$ points towards the tidally-locked state, before fixing P$_{rot} = $ P$_{eq}$.
+when the gradient of P$_{rot}$ points towards the tidally locked state, before fixing P$_{rot} = $ P$_{eq}$.
 We find that this scheme produces physically and numerically accurate results.
 
-\subsubsection{\xxx{The Dynamical Tide}} \label{sec:methods:dynamicalTide}
-
-\xxx{An additional mechanism for tidal dissipation in low-mass stellar binaries is the dynamical tide.
+\subsubsection{The Dynamical Tide} \label{sec:methods:dynamicalTide}
+
+An additional mechanism for tidal dissipation in low-mass stellar binaries is the dynamical tide.
 This effect arises from the turbulent viscous damping of inertial waves that are excited in the stellar convective envelope by a tidal perturber, with Coriolis acceleration serving as the restoring force \citep{Zahn1975,Ogilvie2007}.
@@
 Although semi-analytic models for dynamical tidal dissipation that account for the evolving stellar structure and rotation exist \citep[e.g.][]{Mathis2015,Bolmont2016,Gallet2017}, we do not consider them here as they are currently limited to circular orbits.
-We instead focus on exploring the impact of the equilibrium tide across a wide range of parameter space and leave an examination of how the combination of the dynamical and equilibrium tide impacts the rotation period evolution of low-mass binary stars for future work.}
+We instead focus on exploring the impact of the equilibrium tide across a wide range of parameter space and leave an examination of how the combination of the dynamical and equilibrium tide impacts the rotation period evolution of low-mass binary stars for future work.
 
@@
 We plot the tidal evolution for $a$, $e$, and $P_{rot}$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively, in Fig.~\ref{fig:tidalExample}.
-\xxx{The} CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally-locked state, transferring rotational angular momentum into the orbit in the process.
+The CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally locked state, transferring rotational angular momentum into the orbit\xxx{, increasing the orbital angular momentum by ${\sim}1\%$} in the process.
 At late times, both the CPL and CTL drive the binaries towards orbital circularization, with tidal dissipation decreasing P$_{orb}$.
-The predictions of the CPL and CTL model, differ, however, when the binaries tidally-lock.
-Under the CPL model, the binary tidally-locks into a synchronous orbit when $e < \sqrt{1/19}$, e.g.
+The predictions of the CPL and CTL model, differ, however, when the binaries tidally lock.
+Under the CPL model, the binary tidally locks into a synchronous orbit when $e < \sqrt{1/19}$, e.g.
 Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence, e.g.
@@
 	\includegraphics[width=0.4\textwidth]{../Plots/tidalExample.pdf}
-   \caption{Tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary's $e$ (top), P$_{orb}$ (middle) and P$_{rot}$ (bottom) for the CPL (blue) and CTL (orange) model.
+   \caption{Tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary's $e$ (top), P$_{orb}$ (middle), and P$_{rot}$ (bottom) for the CPL (blue) and CTL (orange) model.
 The blue (CPL) and orange (CTL) vertical dashed lines denote when the stellar binary tidally locks.
 Both the CPL and CTL model predict the same qualitative evolution.
-The rotational evolution differs, however, as under the CPL model, the binary tidally-locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
+The rotational evolution differs, however, as under the CPL model, the binary tidally locks into a synchronous orbit as $e < \sqrt{1/19}$, e.g.
 Eq.~(\ref{eqn:cpl:eqPer}), while the CTL model predicts supersyncronous rotation due to the CTL model's equilibrium period eccentricity dependence (see Eq.~(\ref{eqn:ctl:eqPer})).}%
@@
 
-\subsection{Coupled Stellar-Tidal Evolution For Tidally-Locked Systems} \label{sec:coupled}
-
-Following \citet{Fleming2018}, when one or both binary stars are tidally locked, tidal forces prevent magnetic braking from spinning down the tidally-locked star(s), and any angular momentum lost comes at the expense of the binary orbit, decreasing $a$ as a result \citep{Verbunt1981}.
-Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify the $a$ decay equations due to stellar evolution and magnetic braking in tidally-locked binaries from \citet{Fleming2018}, their Eqs.
-(18) and (20), to additionally account for $r_g$ evolution when one or both stars tidally-lock, respectively, assuming conservation of angular momentum:
+\subsection{Coupled Stellar-Tidal Evolution For tidally locked Systems} \label{sec:coupled}
+
+Following \citet{Fleming2018}, when one or both binary stars are tidally locked, tidal forces prevent magnetic braking from spinning down the tidally locked star(s), and any angular momentum lost comes at the expense of the binary orbit, decreasing $a$ as a result \citep{Verbunt1981}.
+Below in Eq.~(\ref{eqn:tidal_locked_one}) and Eq.~(\ref{eqn:tidal_locked_two}), we modify the $a$ decay equations due to stellar evolution and magnetic braking in tidally locked binaries from \citet{Fleming2018}, their Eqs.
+(18) and (20), to additionally account for $r_g$ evolution when one or both stars tidally lock, respectively, assuming conservation of angular momentum:
 \small
@@
 
-We examine stellar angular momentum evolution in low-mass binaries by simulating \xxx{two} sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism.
+We examine stellar angular momentum evolution in low-mass binaries by simulating two sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism.
 We simulate both stars' spin evolution but mainly consider the P$_{rot}$ evolution for the primary, i.e.
@@
 Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$], consistent with eccentricities of field binaries that likely have not been tidally-processed \citep{Raghavan2010}.
-Although the CTL model is applicable for $e \gsim 0.3$, \xxx{the CPL model is not and can predict qualitatively incorrect evolution in that regime \citep[see Section 4.1 in][]{Leconte2010}}, so we restrict $e \leq 0.3$ to allow us to compare both models.
+Although the CTL model is applicable for $e \gsim 0.3$, the CPL model is not and can predict qualitatively incorrect evolution in that regime \citep[see Section 4.1 in][]{Leconte2010}, so we restrict $e \leq 0.3$ to allow us to compare both models.
 We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Fabrycky2007,Munoz2015,Martin2015b,Hamers2016,Moe2018}.
@@
 These parameters can also vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass stars' evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}.
-Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, \xxx{respectively} \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a range of values exist in the literature.
+Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, respectively \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a range of values exist in the literature.
 Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s log-uniformly over $[10^4,10^8]$ and $\tau$ log-uniformly over $[10^{-2},10]$ s.
@@
 (2) from \citet{Heller2011}.
-\xxx{All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.} We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them.
+All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.
+We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them.
 All code used to run simulations and generate figures is available online.\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}
@@
 
-Here we focus on binaries in \xxx{the ``weak tides"} regime, i.e.
+Here we focus on binaries in the ``weak tides" regime, i.e.
 long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking via analytic calculations and simulations.
@@
 
-In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation\xxx{, P$_{rot} >$ P$_{eq}$.
+In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation, P$_{rot} >$ P$_{eq}$.
 For long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated regime \citep{Matt2015}.
@@
 Although solar mass stars are the most massive stars we consider in this work, and hence will have the strongest tidal torque for a given tidal dissipation parameter and P$_{orb}$, they can still exhibit subsynchronous rotation and serve as a useful end member case to examine here and in simulations below.
-We display the results of this calculation in Fig.~\ref{fig:analyticBalance}, normalizing P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$.}
+We display the results of this calculation in Fig.~\ref{fig:analyticBalance}, normalizing P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$.
 
@@
 	\includegraphics[width=0.5\textwidth]{../Plots/analyticTorque.pdf}
-   \caption{\xxx{Updated Figure} The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}.
+   \caption{The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}.
 The white dashed line indicates our fiducial values for $k_2$ and $\tau$, 0.5 and 0.1 s, respectively, that we adopt in the simulations in $\S$~\ref{sec:mattBalance}.}%
@@
 
-\xxx{Our calculations show that subsynchronous rotation occurs across a wide range of tidal parameters and P$_{orb}$.
+Our calculations show that subsynchronous rotation occurs across a wide range of tidal parameters and P$_{orb}$.
 In general as tides weaken, i.e.
-increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$.
-For strong tides, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for P$_{orb} \lsim 40$ d, tidally-locking binaries into synchronous rotation.
+increasing P$_{orb}$ and/or \xxx{decreasing $\log_{10}(k_2 \tau)$}, tidal and magnetic braking torques balance at longer P$_{rot}$.
+\xxx{For strong tides, $\log_{10}(k_2 \tau) \gsim -1$}, tidal torques overpower magnetic braking for P$_{orb} \lsim 40$ d, tidally locking binaries into synchronous rotation.
 For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries will rotate subsynchronously for P$_{orb} \gsim 20$ d, with more severe subsynchronism at longer P$_{orb}$.
 This simple calculation, however, does not account for stellar evolution or secular tidal orbital evolution, e.g.
-tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we turn to simulations to characterize this evolution.}
+tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we turn to simulations to characterize this evolution.
 
@@
 
-\subsubsection{\xxx{Torque Balance}} \label{sec:mattBalance}
-
-\xxx{We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits to examine how stellar binaries evolve towards subsynchronous rotation.
+\subsubsection{Torque Balance} \label{sec:mattBalance}
+
+We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits to examine how stellar binaries evolve towards subsynchronous rotation.
 In Fig.~\ref{fig:eqPer}, we plot P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models.
-Both tidal models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
-The CPL model predicts that all binaries tidally-lock, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.
+Both tidal models predict that binaries with P$_{orb} < 10$ d will tidally lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}.
+The CPL model predicts that all binaries tidally lock, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.
 As anticipated by our analytic calculations, the CTL model predicts subsynchronous rotation for P$_{orb} \geq 20$ d as magnetic braking overpowers tidal torques.
@@
 As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g.
-Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution.}
+Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
-   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model.
+   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model.
 Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d.
-The black dotted line indicates the tidally-locked state.
+The black dotted line indicates the tidally locked state.
 Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.
-We truncate each curve when the binary tidally-locks.
-For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
-For all P$_{orb}$, the CPL models tidally-lock whereas the CTL model predicts subsynchronous rotation that persists for Gyrs.}%
+We truncate each curve when the binary tidally locks.
+\xxx{The legend denotes the initial binary orbital period and we note that the orbital periods do not vary by more than a few percent over the course of the simulations.} For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
+For all P$_{orb}$, the CPL models tidally lock whereas the CTL model predicts subsynchronous rotation that persists for Gyrs.}%
     \label{fig:eqPer}%
@@
 %In the CTL cases, magnetic braking overpowers tidal torques, spinning down the stars, preventing tidal-locking.
-Tidal torques eventually overpower magnetic braking for long P$_{rot}$ \citep{Kawaler1988,Matt2015}, slowly driving the stellar rotation back towards the tidally-locked state.
+Tidal torques eventually overpower magnetic braking for long P$_{rot}$ \citep{Kawaler1988,Matt2015}, slowly driving the stellar rotation back towards the tidally locked state.
 % Initially, all systems spin-up due to stellar radius contraction.
-Once the stars reach the main sequence at ${\sim} 60$ Myr, tidal torques and magnetic braking combine to spin down the stars towards the tidally-locked state.
+Once the stars reach the main sequence at ${\sim} 60$ Myr, tidal torques and magnetic braking combine to spin down the stars towards the tidally locked state.
 For many CTL models, $\dot{\mathrm{P}}_{rot} > 0$ as magnetic braking overpowers tidal torques causing sustained subsynchronous rotation.
@@
 
-Subsynchronous rotation does not persist indefinitely, however, as P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.
-First, \xxx{magnetic braking weakens at long P$_{rot}$ as its torque scales as P$_{rot}^{-3}$ for unsaturated rotators \citep{Matt2015}.
-Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega})).
+Subsynchronous rotation does not persist indefinitely, however, as P$_{rot}$ eventually decreases back towards the tidally locked state in the long-term due to a combination of three simultaneous physical effects.
+First, magnetic braking weakens at long P$_{rot}$ as its torque scales as P$_{rot}^{-3}$ for unsaturated rotators \citep{Matt2015}.
+Second, as P$_{rot}$ increases further from the tidally locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega})).
 Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$, gradually strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.
-These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long-term.}
-
-\xxx{We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to tidal torques, stellar evolution, and magnetic braking.
+These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long-term.
+
+We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to tidal torques, stellar evolution, and magnetic braking.
 Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 60 Myr when the stars reach the zero age main sequence.
-Tides and magnetic braking then combine to spin down stars towards the tidally-locked state.
+Tides and magnetic braking then combine to spin down stars towards the tidally locked state.
 For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$ as magnetic braking dominates, driving the stars into subsynchronous rotation.
@@
 Tidal torques eventually overpower magnetic braking, seen as a slight negative P$_{rot}$ derivative, slowly driving P$_{rot}$ back towards P$_{eq}$, producing a population of subsynchronous rotators that can persist for Gyrs.
-We explore this point further in $\S$~\ref{sec:shortPorbSubSync}.}
+We explore this point further in $\S$~\ref{sec:shortPorbSubSync}.
 
@@
 %The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but is a generic outcome of the coupling of magnetic braking and weak tidal torques.
-The CTL model can only tidally-lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the longer P$_{orb}$ binaries into subsynchronous rotation.
+The CTL model can only tidally lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the longer P$_{orb}$ binaries into subsynchronous rotation.
 Curiously, all CTL subsynchronous rotators seem to converge to P$_{rot}/$P$_{eq} \approx 1.1$.
@@
 
-We next examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, \xxx{respectively}.
+We next examine how P$_{rot}$ evolution in stellar binaries depends on P$_{orb}$ and the strength of tidal dissipation, parameterized by $Q$ and $\tau$ for the CPL and CTL models, respectively.
 In Fig.~\ref{fig:qTauLock}, we bin our simulation results after the full 7 Gyr evolution by P$_{orb}$ and $Q$ or $\tau$ and compute the median P$_{orb}/$P$_{rot}$ in each bin, marginalizing over all other parameters.
@@
 Spin-orbit synchronization is the typical outcome for binaries with P$_{orb} < 10$ d according to the CPL model for most values of $Q$.
-The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to \xxx{P$_{orb} \gsim 80 $ d} for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
-According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \gsim 3$ s.
+The strong tidal torques predicted by the CPL model can even tidally lock binaries out to P$_{orb} \gsim 80 $ d for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.
+According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally lock for $\tau \gsim 0.1$ s, and seldomly tidally lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \gsim 3$ s.
 Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques.
-The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) with P$_{orb} > 80$ d does not in general correspond to binaries tidally-locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally-locked state.
+The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) \xxx{with P$_{orb} > 60$ d} does not in general correspond to binaries tidally locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally locked state.
+\xxx{At a given age, longer P$_{orb}$ binaries will tend to rotate faster as they experience weaker tidal torques, and hence require longer to spin down towards the tidally locked state.}
 
 Both tidal models predict a population of nearly synchronous rotators near P$_{orb} \approx 60$ d.
-This population corresponds to the evolution described in $\S$~\ref{sec:eq} in which magnetic braking initially spins down stars past the tidally-locked state, but in the long-term, tidal torques spin up the stars, shepherding them towards the tidally-locked state.
-This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ and $Q$ or $\tau$ (see Figs.~\ref{fig:eqPer},$\S$~\ref{sec:shortPorbSubSync}).
+This population corresponds to the evolution described in $\S$~\ref{sec:eq} in which magnetic braking initially spins down stars past the tidally locked state, but in the long-term, tidal torques spin up the stars, shepherding them towards the tidally locked state.
+This process can keep stellar P$_{rot} \gsim$ P$_{eq}$ for several Gyrs or longer, depending on the P$_{orb}$ and $Q$ or $\tau$ (see Fig.~\ref{fig:eqPer}, $\S$~\ref{sec:shortPorbSubSync}).
 
@@
 	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
-   \caption{\xxx{Updated Figure} Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
-For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.
-Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$ d, that \xxx{does} not in general reflect tidally-locking, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spun down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution.
-For large $Q$ ($> 10^7$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
+   \caption{Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
+For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, whereas the CTL model requires $\tau \geq 0.1$ s to tidally lock.
+For large $Q$ ($> 10^7$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
     \label{fig:qTauLock}%
@@
 % caption extra:
-% The CPL model predicts that stellar binaries can tidally-lock out to P$_{orb} = 30$ d for most values of $Q$, with some locking all the way up to P$_{orb} = 100$ d for small $Q < 10^5$.
+% The CPL model predicts that stellar binaries can tidally lock out to P$_{orb} = 30$ d for most values of $Q$, with some locking all the way up to P$_{orb} = 100$ d for small $Q < 10^5$.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/porbProtPorbQHist.pdf}
-   \caption{\xxx{Updated Figure} Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.
-In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$ and display these values.
-Narrow constraints imply that certain tidal parameters are required to produce that rotation state, e.g.
-supersynchronous rotators near P$_{orb}/$P$_{rot} = 1.5$ require $Q \lsim 10^{5.5}$.
-}%
+   \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model. }%
     \label{fig:qmap}%
@@
 
-\xxx{We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$}.
-For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock \xxx{these binaries}.
-For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$, nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for \xxx{P$_{orb} < 60$ d}, as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e.
-Eqn.~(\ref{eqn:cpl:eqPer}).
+We isolate the impact of $Q$ and $\tau$ \xxx{on the spin-orbital state of tidally interacting stellar binaries} by binning our \xxx{CPL and CTL} simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$ \xxx{in Figures \ref{fig:qmap} and \ref{fig:taumap}, respectively.
+In these figures, we estimate the typical strength of tidal torques, using $Q$ and $\tau$ as a proxy, that can produce various spin-orbital states.
+For the CPL simulations depicted in Fig.~\ref{fig:qmap},} synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally lock these binaries.
+For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$, nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 60$ d, as in the CPL model, binaries with eccentric orbits can only tidally lock into a 1:1 or 3:2 spin-orbit commensurability, see Eqn.~(\ref{eqn:cpl:eqPer}).
 
 Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking.
-This effect can be seen for P$_{orb} < 50$ d where the median $Q$ \xxx{tends to} increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state.
-This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.
+\xxx{Subsynchronous rotation can occur under the CPL model for binaries with P$_{orb} < 50$ d.
+In this regime, the median $Q$ tends to increase with decreasing P$_{orb}/$P$_{rot}$, except near the tidally locked state, as magnetic braking dominates weaker tidal torques, yielding longer P$_{rot}$.}  This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally locked state by the end of the simulation.
 In this case, the more supersynchronous the rotation, the weaker the tidal torques must be, and hence the larger the $Q$ must be.
@@
 	\includegraphics[width=\textwidth]{../Plots/porbProtPorbTauHist.pdf}
-   \caption{\xxx{Updated Figure} Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
+   \caption{Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
     \label{fig:taumap}%
@@
 
-\xxx{According to the CTL model, many binaries tidally-lock for P$_{orb} \lsim 20$ d when $\tau \gsim 0.1$ s, with some tidally-locking up to P$_{orb} \approx 50$ d when $\tau \gsim 1$ s.
-Subsynchronous rotation typically occurs for stars with $\tau < 0.1$ s}.
-For P$_{orb} > 50$ d, magnetic braking dominates the evolution \xxx{seen in} the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, a value that is typically insufficient for tides to strongly influence the evolution given the wide orbital separations}.
+According to the CTL model \xxx{simulations, depicted in Fig.~\ref{fig:taumap}}, many binaries tidally lock for P$_{orb} \lsim 20$ d when $\tau \gsim 0.1$ s, with some tidally locking up to P$_{orb} \approx 50$ d when $\tau \gsim 1$ s.
+Subsynchronous rotation typically occurs for stars with $\tau < 0.1$ s.
+\xxx{Similar to the behavior depicted in Fig.~\ref{fig:qmap}, longer P$_{rot}$ are produced by binaries with weaker tidal interactions since P$_{orb}/$P$_{rot}$ decreases monotonically with $\tau$ for P$_{orb} < 40$ d.}  For P$_{orb} > 50$ d, magnetic braking dominates the evolution \xxx{seen in} the diagonal sequence with a median $\tau \approx 0.1$ s, a value that is typically insufficient for tides to strongly influence the evolution given the wide orbital separations.
 The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
-We do not \xxx{often} observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).
+\xxx{In this P$_{orb}$ regime, most binaries rotate supersynchronously as tides and magnetic braking fail to sufficiently spin down the stars by the age of the system.
+At longer P$_{orb}$, some binaries can strongly tidally-interact, but these systems require extreme tidal $\tau \gsim 10$ s.
+Unlike the CPL simulations depicted in Fig.~\ref{fig:qmap}, these binaries with P$_{orb} \approx 90$ d are not tidally locked as the tides are still not strong enough to lock the system.
+We explore this point further in $\S$~\ref{sec:protDist}.} We do not often observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).
 
@@
 
-\xxx{As seen in Fig.~\ref{fig:qTauLock}, subsynchronous rotation can even occur for short P$_{orb}$ binaries, where tidal-locking is the expectation, if the tidal torques are sufficiently weak.} In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a \xxx{P$_{orb} = 7.5$ d} binary for various tidal dissipation parameters.
+As seen in Fig.~\ref{fig:qTauLock}, subsynchronous rotation can even occur for short P$_{orb}$ binaries, where tidal-locking is the expectation, if the tidal torques are sufficiently weak.
+In Fig.~\ref{fig:eqPerShortPorb}, we examine subsynchronous rotation in short P$_{orb}$ binaries by displaying the P$_{rot}$ evolution for a P$_{orb} = 7.5$ d binary for various tidal dissipation parameters.
 Subsynchronous rotation occurs in general for weak tidal torques, $Q>10^7$ or $\tau < 0.1$ s in these cases, and is not restricted to long P$_{orb}$ binaries.
 Previous theoretical studies have also predicted subsynchronous rotation in short P$_{orb}$ binaries arising from the balance between tidal torques and magnetic braking \citep[e.g.][]{Habets1989,Zahn1994,Keppens1997} suggesting that this behavior is not an artifact of our choice of tidal or magnetic braking models, but rather a general outcome of the competition between magnetic braking and tidal evolution in low-mass binaries.
-Short P$_{orb}$ subsynchronous rotators can eventually tidally-lock \xxx{after several Gyrs, e.g.
-the $Q=10^8$ case in Fig.~\ref{fig:eqPerShortPorb},} via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
+Short P$_{orb}$ subsynchronous rotators can eventually tidally lock after several Gyrs, e.g.
+the $Q=10^8$ case in Fig.~\ref{fig:eqPerShortPorb}, via the mechanism described above where tidal torques gradually strengthen relative to magnetic braking.
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
-   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer})), for initial circular binary orbits with initial P$_{orb} = 7.5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
-Systems with strong tidal torques tidally-lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation.
+   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer})), for initial circular binary orbits with initial P$_{orb} = 7.5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
+Systems with strong tidal torques tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally locked state, resulting in subsynchronous rotation.
 }%
@@
 % caption extra:
-% In the long-term, magnetic braking weakens relative to tidal torques for longer P$_{rot}$, allowing tides to drive the stars back towards the tidally-locked state.
+% In the long-term, magnetic braking weakens relative to tidal torques for longer P$_{rot}$, allowing tides to drive the stars back towards the tidally locked state.
 
 Short P$_{orb}$ subsynchronous binaries exist in nature, such as many \kepler EBs (\citet{Lurie2017}, see $\S$~\ref{sec:kepler} for further discussion), Kepler-47 \citep{Orosz2012}, EPIC 219394517 \citep{Torres2018}, and in ``Binary 6211" observed by \citet{Meibom2006}, suggesting that this theoretical observation is real and borne out in nature.
-Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries and sunsynchronous rotation should be expected in many tidally-interacting binaries.
+Spin-orbit synchronization should therefore not be assumed for short P$_{orb}$ binaries and sunsynchronous rotation should be expected in many tidally interacting binaries.
 We explore these effect further and compare our theory to observations of \kepler EBs in $\S$~\ref{sec:kepler}.
@@
 Here we examine how the competition between tidal torques and magnetic braking shape the P$_{rot}$ distribution of low-mass stellar binaries.
-We consider two cases where tidal torques dominate: ``Locked", where P$_{rot} = $ P$_{eq}$, and ``Interacting", where P$_{rot}$ is within $10\%$ of P$_{eq}$ as in this regime, tides are likely shepherding P$_{rot}$ towards the tidally-locked state as we demonstrated in $\S$~\ref{sec:eq}.
+We consider two cases where tidal torques dominate: ``Locked", where P$_{rot} = $ P$_{eq}$, and ``Interacting", where P$_{rot}$ is within $10\%$ of P$_{eq}$ as in this regime, tides are likely shepherding P$_{rot}$ towards the tidally locked state as we demonstrated in $\S$~\ref{sec:eq}.
 We refer to the remaining binaries as ``not locked" as magnetic braking and stellar evolution likely dominate their angular momentum evolution.
-In Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, we plot P$_{rot}$ as a function of mass for the primary stars in stellar binaries for both the CPL and CTL model, respectively, integrated to system ages uniformly sampled over $1-7$ Gyr\xxx{, consistent with ages of stars in the \kepler field \citep{Chaplin2014}}.
+In Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, we plot P$_{rot}$ as a function of mass for the primary stars in stellar binaries for both the CPL and CTL model, respectively, integrated to system ages uniformly sampled over $1-7$ Gyr, consistent with ages of stars in the \kepler field \citep{Chaplin2014}.
 
@@
 	\includegraphics[width=0.5\textwidth]{../Plots/lockedCPL.pdf}
-   \caption{\xxx{Updated Figure} Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
+   \caption{Rotation state for tidally locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
 Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr.
 Right: Marginalized P$_{rot}$ distribution for each case.
-\xxx{Top: Marginalized mass distributions.}}%
+Top: Marginalized mass distributions.}%
     \label{fig:lockedCPL}%
@@
 
-Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster \xxx{(median CPL, CTL P$_{rot} = 22.6$ d and $8.8$ d)} than not locked \xxx{(median CPL, CTL both P$_{rot} = 32.4$ d)} binaries.
-\xxx{The CTL model predicts that the majority of tidally-locked binaries, \xxx{$83\%$}, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest.
-The CPL model, however predicts that binaries can tidally-lock into a wide range of rotation states as only $46\%$ of locked binaries have P$_{rot} < 20$ d, while the rest can lock out to P$_{rot} \approx 100$ d in long P$_{orb}$ binaries.} More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.
+Both models predict a substantial population of tidally locked fast rotators with P$_{rot} \lsim 20$ d, with tidally locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 22.6$ d and $8.8$ d) than not locked (median CPL, CTL both P$_{rot} = 32.4$ d) binaries.
+The CTL model predicts that the majority of tidally locked binaries, $83\%$, lock into rapid rotation with P$_{rot} \lsim 20$ d, typically in short P$_{orb}$ binaries where tidal torques are strongest.
+The CPL model, however predicts that binaries can tidally lock into a wide range of rotation states as only $46\%$ of locked binaries have P$_{rot} < 20$ d, while the rest can lock out to P$_{rot} \approx 100$ d in long P$_{orb}$ binaries.
+More massive stars are more likely to tidally lock compared to less massive stars as tidal torques scale with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.
 This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model.
-We highlight this enhanced density of locked binaries in systems with more massive primaries in the marginalized mass distributions in the top panels of Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.}
+We highlight this enhanced density of locked binaries in systems with more massive primaries in the marginalized mass distributions in the top panels of Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.
 
@@
 	\includegraphics[width=0.48\textwidth]{../Plots/lockedCTL.pdf}
-   \caption{\xxx{Updated Figure} Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
+   \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations.}%
     \label{fig:lockedCTL}%
@@
 
-\xxx{The interacting population tends to rotate more slowly than the not locked population as at short P$_{orb}$, and hence P$_{rot}$, binaries preferentially tidally-lock due to stronger tidal torques.
+The interacting population tends to rotate more slowly than the not locked population as at short P$_{orb}$, and hence P$_{rot}$, binaries preferentially tidally lock due to stronger tidal torques.
 At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthening enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
-The CPL and CTL models predict that \xxx{$31\%$ and $24\%$} of stars, respectively, are either tidally-locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
-The P$_{rot}$ - mass distribution for not locked binaries resembles the single star sequence as magnetic braking and stellar evolution dictate their angular momentum evolution.}
-
-Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we find stellar binaries can tidally-lock over a wide range of P$_{rot}$ up to P$_{rot} =$ P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of \xxx{solar-mass} single stars.
-This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally-lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
-Under the CTL model, however, binaries predominantly tidally-lock out to only P$_{orb} \approx 20$ d\xxx{, although binaries with more massive primaries can occasionally lock, or at least tidally-interact, out to P$_{orb} \approx 80$ d.
-\xxx{We highlight this behavior with a histogram of locked and interacting binaries over P$_{rot}$ for both tidal models in Fig.~\ref{fig:lockedProtHist}.} The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
-We discuss this point further in $\S$~\ref{sec:whichModel}}.
+The CPL and CTL models predict that \xxx{$31\%$ and $24\%$} of stars, respectively, are either tidally locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
+The P$_{rot}$ - mass distribution for not locked binaries resembles the single star sequence as magnetic braking and stellar evolution dictate their angular momentum evolution.
+
+Tidal locking is not limited to P$_{orb} \lsim 20$ d, however, as we find stellar binaries can tidally lock over a wide range of P$_{rot}$ up to P$_{rot} =$ P$_{orb} \approx 100$ d according to the CPL model, producing a slow-rotating population above the P$_{rot}$ distribution envelop of solar-mass single stars.
+This behavior is consistent with observations of P$_{rot}$ in \kepler eclipsing binaries by \citet{Lurie2017} who find tentative evidence that binaries can tidally lock up to their detection limit of P$_{orb} = $ P$_{rot} = 45$ d.
+Under the CTL model, however, binaries predominantly tidally lock out to only P$_{orb} \approx 20$ d, although binaries with more massive primaries can occasionally lock, or at least tidally-interact, out to P$_{orb} \approx 80$ d.
+We highlight this behavior with a histogram of locked and interacting binaries over P$_{rot}$ for both tidal models in Fig.~\ref{fig:lockedProtHist}.
+\xxx{The CTL model predicts fewer tidally locked binaries at longer P$_{rot}$, concentrating most of the locked CTL distribution's density at short P$_{rot}$, whereas the CPL distribution has a heavy tail extending towards longer P$_{rot}$.
+The CPL model, however, predicts larger tidal-locking rates than the CTL model as seen in the enhanced numbers of tidally locked binaries at low P$_{rot}$ Fig.~\ref{fig:lockedCPL} compared with Fig.~\ref{fig:lockedCTL}.} The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
+We discuss this point further in $\S$~\ref{sec:whichModel}.
 
@@
 	\includegraphics[width=0.48\textwidth]{../Plots/lockedProtHist.pdf}
-   \caption{P$_{rot}$ distribution for tidally-locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.}%
+   \caption{P$_{rot}$ distribution for tidally locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.}%
     \label{fig:lockedProtHist}%
@@
 
-We compare the \xxx{P$_{rot}$ and age distributions of tidally-interacting stellar binaries from our CPL and CTL simulations with that of single stars to gauge the impact of tidal torques on driving P$_{rot}$ distributions away from that of single stars and what implications that may have for estimating stellar ages using gyrochronology}.
+We compare the P$_{rot}$ and age distributions of tidally interacting stellar binaries from our CPL and CTL simulations with that of single stars to gauge the impact of tidal torques on driving P$_{rot}$ distributions away from that of single stars and what implications that may have for estimating stellar ages using gyrochronology.
 We simulate 10,000 single star systems according to the evolution described in $\S$~\ref{sec:methods:stellar} with initial conditions sampled from the same mass and P$_{rot}$ distributions used for the binary simulations described $\S$~\ref{sec:methods}.
@@
 	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
-   \caption{\xxx{Updated Figure} P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model.
+   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model.
 For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions.
@@
 
-\xxx{In binaries, tidal torques tend to drive the P$_{rot}$ evolution away from that of single stars and towards P$_{eq}$, either maintaining rapid rotation in tidally-locked short P$_{orb}$ systems, or working with magnetic braking to slow P$_{rot}$ beyond that of single stars of the same age.
+In binaries, tidal torques tend to drive the P$_{rot}$ evolution away from that of single stars and towards P$_{eq}$, either maintaining rapid rotation in tidally locked short P$_{orb}$ systems, or working with magnetic braking to slow P$_{rot}$ beyond that of single stars of the same age.
 The impact of tidal torques on the binary P$_{rot}$ distribution is clear: strong tidal torques in short P$_{orb}$ binaries produce a substantial population of rapid rotators with P$_{rot} \lsim 20$ d.
 Except for stars with ages $\lsim 1$ Gyr, or young late M-dwarfs who are either contracting along the pre-main sequence or have just reached the main sequence, our single star simulations fail to produce a population of rapid rotators.
-This theoretical result is consistent with \citet{Simonian2018} who find that the population of \kepler stars with P$_{rot} < 7.5$ d is likely dominated by tidally-interacting binaries.
+This theoretical result is consistent with \citet{Simonian2018} who find that the population of \kepler stars with P$_{rot} < 7.5$ d is likely dominated by tidally interacting binaries.
 The influence of tides extends to longer P$_{orb}$ systems producing a slowly-rotating population above the upper envelope of the single star sequence for M$ \gsim 0.6$ M$_{\odot}$, e.g.
-Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, and is a population that single-star models fail to produce.}
-
-\xxx{In the single star population, there is a clear monotonic relation between P$_{rot}$ and age, with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
+Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, and is a population that single-star models fail to produce.
+
+In the single star population, there is a clear monotonic relation between P$_{rot}$ and age, with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
 This trend is a generic outcome of magnetic braking and is not specific to our choice of magnetic braking model.
-In stark contrast, both tidal models predict that age does not always strongly correlate with P$_{rot}$ as tidally-interacting binaries, at a given primary star mass and P$_{rot}$, can assume a wide range of ages, especially for P$_{rot} \lsim 20$ d.}
-
-\xxx{We quantify the impact of binarity on gyrochronology age estimates in Fig.~\ref{fig:gyro} by computing the percent difference between the mean ages of single and binary stars, for both tidal models, in mass and P$_{rot}$ bins for the populations depicted in Fig.~\ref{fig:protDist}.
-This quantity, referred to here as the ``Relative Age Error", represents the systematic error incurred by assigning a tidally-interacting binary star the age expected for single stars at a given mass and P$_{rot}$.
-For this comparison, we select the subset of tidally-locked and tidally-interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.}
+In stark contrast, both tidal models predict that age does not always strongly correlate with P$_{rot}$ as tidally interacting binaries, at a given primary star mass and P$_{rot}$, can assume a wide range of ages, especially for P$_{rot} \lsim 20$ d.
+
+We quantify the impact of binarity on gyrochronology age estimates in Fig.~\ref{fig:gyro} by computing the percent difference between the mean ages of single and binary stars, for both tidal models, in mass and P$_{rot}$ bins for the populations depicted in Fig.~\ref{fig:protDist}.
+This quantity, referred to here as the ``Relative Age Error", represents the systematic error incurred by assigning a tidally interacting binary star the age expected for single stars at a given mass and P$_{rot}$.
+For this comparison, we select the subset of tidally locked and tidally interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/gyro.pdf}
-   \caption{\xxx{New Figure: Relative age error between single and binary stars for both the CPL (left) and CTL (right) tidal models.
-The error is computed as the percent difference between the mean ages of single and tidally-interacting binary stars in mass and P$_{rot}$ bins.}}%
+   \caption{Relative age error between single and binary stars for both the CPL (left) and CTL (right) tidal models.
+The error is computed as the percent difference between the mean ages of single and tidally interacting binary stars in mass and P$_{rot}$ bins.}%
     \label{fig:gyro}%
@@
 
-\xxx{For most values of P$_{rot}$, gyrochronology methods systematically underestimate the ages of tidally-interacting binaries, with the relative age error increasing with decreasing P$_{rot}$ and increasing with primary star mass.
+For most values of P$_{rot}$, gyrochronology methods systematically underestimate the ages of tidally interacting binaries, with the relative age error increasing with decreasing P$_{rot}$ and increasing with primary star mass.
 For binaries with P$_{rot} \lsim 30$ d, gyrochronology ages are underestimated by $50\%$, with this error growing to $300\%$ for P$_{rot} \approx 10$ d.
@@
 The relative age errors would be more pronounced for the most rapidly-rotating stars, e.g.
-tidally-interacting near-solar mass binaries with P$_{rot} \lsim 10$ d, however, our single star-only models fail to produce such rotation states.}
-
-\xxx{The age distribution of binaries with P$_{rot} < 20$ d is inconsistent with that of single stars.
+tidally interacting near-solar mass binaries with P$_{rot} \lsim 10$ d, however, our single star-only models fail to produce such rotation states.
+
+The age distribution of binaries with P$_{rot} < 20$ d is inconsistent with that of single stars.
 In this range, the median ages and $68\%$ interval are $2.3^{+2.9}_{-0.9}$ Gyr and $2.4^{+3.0}_{-1.1}$ Gyr according to the CPL and CTL models, respectively, compared to the much younger single stars with ages of $1.6^{+0.8}_{-0.4}$ Gyr.
-We highlight this dichotomy in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.}
+We highlight this dichotomy in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.
 
@@
 	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
-   \caption{\xxx{Updated Figure} Histogram of rapidly-rotating (P$_{rot} < 20$ d) \xxx{star} ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
-Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
+   \caption{Histogram of rapidly-rotating (P$_{rot} < 20$ d) star ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
+Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally locked rapidly-rotating binaries exhibit a wide range of ages.}%
     \label{fig:protAgeHist}%
@@
 
-Tidal torques pose a fundamental problem for inferring ages of stars via \xxx{gyrochronology}.
-Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
-For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages \xxx{$\lsim 1.6$ Gyr.
-If the star is actually an unresolved binary, as could be the case for many \kepler rapid rotators \citep{Simonian2018}, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail}.
-This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a \xxx{contaminating signal, e.g.
-Fig.~\ref{fig:protDist} and Fig.~\ref{fig:gyro}}.
-
-In general, it is difficult to accurately determine if a source is single star or a stellar binary via \xxx{longterm photometric monitoring, e.g.
-via \kepler or \textit{TESS},} as only a small fraction of stars in binaries will occult one another.
-Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.
-Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field \xxx{are consistent with tidally-synchronized photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread}.
-We caution that any application of \xxx{gyrochronology} methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
-\xxx{Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
+Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronology.
+Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age.
+For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 1.6$ Gyr.
+If the star is actually an unresolved binary, as could be the case for many \kepler rapid rotators \citep{Simonian2018}, it would likely be tidally locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail.
+This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a contaminating signal, e.g.
+Fig.~\ref{fig:protDist} and Fig.~\ref{fig:gyro}.
+
+In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring, e.g.
+via \kepler or \textit{TESS}, as only a small fraction of stars in binaries will occult one another.
+Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronology are potentially subject to systematic errors.
+Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are consistent with tidally-synchronized photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread.
+We caution that any application of gyrochronology methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.
+Tidal torques do not just produce spin-orbit synchronization at short P$_{orb}$, but can produce a rich variety of rotation states that deviate from the expected long-term spin-down experienced by single stars, e.g.
 Fig.~\ref{fig:qmap} and Fig.~\ref{fig:taumap}.
-We recommend that the application, or calibration, magnetic braking models to a sample of stellar rotation periods control for binarity.}
+We recommend that the application, or calibration, magnetic braking models to a sample of stellar rotation periods control for binarity.
 
@@
 The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) that have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.
-In stark contrast, both tidal models predict a substantial population of tidally-locked fast rotators (P$_{rot} < 20$ d).
+In stark contrast, both tidal models predict a substantial population of tidally locked fast rotators (P$_{rot} < 20$ d).
 Under the CPL and CTL models, $18\%$ and $19\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.
@@
 %The general trends found in the coupled stellar-tidal simulations using the \citet{Matt2015} model persist in the simulations evolved using the \citet{Reiners2012} model, demonstrating that these findings are not magnetic braking model-dependant, but rather a generic result of the interaction between magnetic braking and tidal torques in low-mass binaries.
-Both the CPL and CTL simulations display a clear single-star magnetic braking-dominated sequence, but as before, tidally-locked binaries appear across a wide range of P$_{rot}$ and primary star masses.
-Both the CPL and CTL model tidally-lock many short P$_{orb}$ stars into rapid rotation across all masses, with both models predicting $23\%$ of stars have P$_{rot} < 20$ d, compared to $11\%$ of single-stars, mostly for late M dwarfs in the latter case.
-As shown above, the CTL model predicts that binaries can strongly tidally-interact out to long P$_{orb}$, seen out to P$_{rot} = 80$ d, while the CPL model can tidally-lock stars out to P$_{orb} \approx 100$ d.
+Both the CPL and CTL simulations display a clear single-star magnetic braking-dominated sequence, but as before, tidally locked binaries appear across a wide range of P$_{rot}$ and primary star masses.
+Both the CPL and CTL model tidally lock many short P$_{orb}$ stars into rapid rotation across all masses, with both models predicting $23\%$ of stars have P$_{rot} < 20$ d, compared to $11\%$ of single-stars, mostly for late M dwarfs in the latter case.
+As shown above, the CTL model predicts that binaries can strongly tidally-interact out to long P$_{orb}$, seen out to P$_{rot} = 80$ d, while the CPL model can tidally lock stars out to P$_{orb} \approx 100$ d.
 These strong tidal interactions seen both here and in $\S$~\ref{sec:mattDist} decouple P$_{rot}$ from age in binaries, especially for rapid rotators.
@@
 
-\xxx{We compare our simulation results to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing binaries by \citet{Lurie2017}} to gauge if our model predictions, which by design populate a wide, but physically-plausible, region of parameter space, can reproduce features observed in the data.
+We compare our simulation results to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing binaries by \citet{Lurie2017} to gauge if our model predictions, which by design populate a wide, but physically-plausible, region of parameter space, can reproduce features observed in the data.
 \citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with star spot modulations and visually inspected each light curve to ensure their accuracy.
@@
 
-\xxx{Qualitatively, the CTL model appears to do a better job of reproducing features seen in the \citet{Lurie2017} data than the CPL model.
+Qualitatively, the CTL model appears to do a better job of reproducing features seen in the \citet{Lurie2017} data than the CPL model.
 The CPL model, for example, cannot produce the observed cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.2$ for P$_{orb} < 10$ d whereas the CTL model can.
-Instead, owing to the its discrete P$_{eq}$, the CPL model predicts that all tidally-locked supersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.}  This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the \citet{Lurie2017} data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
-\xxx{Neither model reproduces the very supersynchronous, P$_{orb}/$P$_{rot} > 1.6$, binaries in the \citet{Lurie2017} data as they typically have orbital eccentricities in excess of 0.3 and are outside of the region of parameter space we consider.
-The CTL model, however, could in principle reproduce these points if they are tidally-locked binaries as its P$_{eq}$ is a continuous function of $e$ and is applicable for large $e$, in contrast to the discrete P$_{eq}$ predicted by the CPL model that is valid for smaller $e$.
+Instead, owing to the its discrete P$_{eq}$, the CPL model predicts that all tidally locked supersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.
+This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the \citet{Lurie2017} data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.
+Neither model reproduces the very supersynchronous, P$_{orb}/$P$_{rot} > 1.6$, binaries in the \citet{Lurie2017} data as they typically have orbital eccentricities in excess of 0.3 and are outside of the region of parameter space we consider.
+The CTL model, however, could in principle reproduce these points if they are tidally locked binaries as its P$_{eq}$ is a continuous function of $e$ and is applicable for large $e$, in contrast to the discrete P$_{eq}$ predicted by the CPL model that is valid for smaller $e$.
 Both tidal models predict that nearly all binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in agreement with the \citet{Lurie2017} observations.
-At very short P$_{orb}$, in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries \citep{Counselman1973}.}
+At very short P$_{orb}$, in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries \citep{Counselman1973}.
 
@@
 	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
-   \caption{\xxx{Updated Figure} P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ according to the CPL model (left) and the CTL model (middle), and \citet{Lurie2017} \kepler EB observations (right).
+   \caption{P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ according to the CPL model (left) and the CTL model (middle), and \citet{Lurie2017} \kepler EB observations (right).
 All points are colored by $e$.
@@
 
-\xxx{For P$_{orb} \gsim 4$ d, our models produce a substantial number of subsynchronous rotators.
+For P$_{orb} \gsim 4$ d, our models produce a substantial number of subsynchronous rotators.
 Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces this population.
 The CPL model, however, struggles to populate the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.
-\citet{Lurie2017} find that $15\%$ of their sample with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $8\%$ of our CTL population and only $2\%$ of the CPL population.}
-
-\xxx{Both models predict a large number of extremely subsynchronous rotators with P$_{orb}/$P$_{rot} < 0.7$ across all P$_{orb}$ that is not present in the \citet{Lurie2017} data.
+\citet{Lurie2017} find that $15\%$ of their sample with 2 $< P_{orb} <$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $8\%$ of our CTL population and only $2\%$ of the CPL population.
+
+Both models predict a large number of extremely subsynchronous rotators with P$_{orb}/$P$_{rot} < 0.7$ across all P$_{orb}$ that is not present in the \citet{Lurie2017} data.
 Magnetic braking creates the lower limit of this subsynchronous population, a line of nearly constant P$_{rot} \approx 60$ d set by how much a star can spin down over 7 Gyr, the longest age considered in our simulations.
 Our choice of prior distributions for both $Q$ and $\tau$ permit very weak tidal interactions that likely gives rise to this population and suggests that our prior does not reflect the underlying distribution of stellar tidal parameters in nature.
-Alternatively, the data could be incomplete where our models predict slowly-rotating subsynchronous rotators as the photometric amplitude of star spot modulations tends to decrease with increasing P$_{rot}$, making reliable rotation periods difficult to detect \citep{McQuillan2014,Lurie2017,Reinhold2018}.}
-
-\xxx{Although the CTL model seems to better reproduce the \citet{Lurie2017} data, both tidal models can reproduce features observed in the \kepler EB distribution, e.g.
+Alternatively, the data could be incomplete where our models predict slowly-rotating subsynchronous rotators as the photometric amplitude of star spot modulations tends to decrease with increasing P$_{rot}$, making reliable rotation periods difficult to detect \citep{McQuillan2014,Lurie2017,Reinhold2018}.
+
+Although the CTL model seems to better reproduce the \citet{Lurie2017} data, both tidal models can reproduce features observed in the \kepler EB distribution, e.g.
 the synchronized population and subsynchronous rotators, suggesting that our models reasonably approximate the dynamical interactions of tidally-evolving, low-mass stellar binaries.
@@
 Moreover, our prior distributions were chosen to be plausible, but wide, in order to examine our model predictions over parameter space and are not suited for a robust statistical inference to select between which equilibrium tidal model best describes tidal interactions in low-mass binaries stars.
-Below, we offer observational tests that could discriminate between models.}
+Below, we offer observational tests that could discriminate between models.
 
@@
 % On the other hand, the CTL model produces a large number of supersynchronous binaries with eccentric orbits similar to what is seen in the \citet{Lurie2017} data.
-These systems have tidally-locked, but their orbits have not yet circularized, because the tidal locking time scale is 2-3 orders of magnitude less than the circularization timescale \citep{Mazeh2008}.
-% For example under the CTL model, our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.
+These systems have tidally locked, but their orbits have not yet circularized, because the tidal locking time scale is 2-3 orders of magnitude less than the circularization timescale \citep{Mazeh2008}.
+% For example under the CTL model, our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally locked supersynchronous population.
 % Second, In principle, we could correct for such a bias in the \citet{Lurie2017} data and condition our model predictions on the data to infer tidal properties, but that analysis is beyond the scope of this work for reasons we discuss in $\S$~\ref{sec:inference}.
@@
 
-%We derive a wide $Q$ constraint, log$_{10}(Q) = 5.55^{+0.98}_{-1.06}$, that is consistent with our flat prior distribution, suggesting that tidally-interacting stellar binaries can have a wide range of $Q$.
+%We derive a wide $Q$ constraint, log$_{10}(Q) = 5.55^{+0.98}_{-1.06}$, that is consistent with our flat prior distribution, suggesting that tidally interacting stellar binaries can have a wide range of $Q$.
 Our $\tau$ constraint, log$_{10}(\tau) = -0.73^{+0.52}_{-0.82}$, however, displays a noticeable increase in density at larger $\tau$ near 1 s.
 This is a significant deviation from our flat prior distribution suggesting that time lags in low-mass stars are larger than the typical value of $\tau \approx 0.1$ s.
-Still, a wide range of values for both $Q$ and $\tau$ are consistent with observations indicating that tidally-interacting binaries can exhibit a wide range of tidal parameters, albeit with a slightl preference for larger time lags under the CTL formalism.
+Still, a wide range of values for both $Q$ and $\tau$ are consistent with observations indicating that tidally interacting binaries can exhibit a wide range of tidal parameters, albeit with a slightl preference for larger time lags under the CTL formalism.
 
@@
 Here, we outline three observational tests that can discriminate between the two models.
-The first test considers binaries with P$_{orb} < 10$ d that are likely tidally-locked on eccentric orbits, but with $e < 0.23$.
-In this $e$ regime, the CPL model predicts that the majority of systems are tidally-locked into synchronous rotation and does not permit a supersynchronous rotation state, e.g.
+The first test considers binaries with P$_{orb} < 10$ d that are likely tidally locked on eccentric orbits, but with $e < 0.23$.
+In this $e$ regime, the CPL model predicts that the majority of systems are tidally locked into synchronous rotation and does not permit a supersynchronous rotation state, e.g.
 Eqn.~(\ref{eqn:cpl:eqPer}).
@@
 Supersynchronous rotation that is not due to tidal interactions can occur in extremely young, rapidly rotating systems that are still contracting along the pre-main sequence, or that have recently reached the main sequence.
-These young, supersynchronous rotators are unlikely to be tidally-locked, usually have P$_{orb}/$P$_{rot} > 1.5$, and do not stay supersynchronous for long given that solar mass pre-main sequence lifetimes are $\lsim 100$ Myr, distinguishing them from tidally-locked binaries (see Fig.~\ref{fig:lurie7}).
+These young, supersynchronous rotators are unlikely to be tidally locked, usually have P$_{orb}/$P$_{rot} > 1.5$, and do not stay supersynchronous for long given that solar mass pre-main sequence lifetimes are $\lsim 100$ Myr, distinguishing them from tidally locked binaries (see Fig.~\ref{fig:lurie7}).
 If supersynchronous rotation is observed in binaries with P$_{orb} < 10$ d, P$_{orb}/$P$_{rot} < 1.5$, and $0 < e \lsim 0.23$, it is evidence in favor of the CTL model over the CPL model.
 
-Second, for tidally-locked binaries with $e > 0.23$, the CPL model predicts supersynchronous rotation in the form of a 3:2 spin-orbit comensurability, e.g.
-the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permitted, compared to the continuum of supersynchronous rotation states in eccentric tidally-locked rotators predicted by the CTL model.
-If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model\xxx{, but there is no obvious clustering of \kepler EBs near any spin-orbit resonance}.
+Second, for tidally locked binaries with $e > 0.23$, the CPL model predicts supersynchronous rotation in the form of a 3:2 spin-orbit comensurability, e.g.
+the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permitted, compared to the continuum of supersynchronous rotation states in eccentric tidally locked rotators predicted by the CTL model.
+If a substantial clustering of stellar binaries with P$_{orb}/$P$_{rot} = 1.5$ is observed, it would be strong evidence in favor of the CPL model, but there is no obvious clustering of \kepler EBs near any spin-orbit resonance.
 These two tests can fail to discriminate between the CPL and CTL model, however, if the CPL model P$_{eq}$ is a continuous function of $e$, e.g.
 Eqn.~(\ref{eqn:cpl:eqPerCont}), as was argued by \citet{Goldreich1966b} and derived by \citet{Murray1999}.
-In such a case, one would need a large number of accurate and precise measurements P$_{orb}$ and $e$, with robust uncertainties, for tidally-interacting binaries to discriminate between the CPL and CTL continuous P$_{eq}$, e.g.
+In such a case, one would need a large number of accurate and precise measurements P$_{orb}$ and $e$, with robust uncertainties, for tidally interacting binaries to discriminate between the CPL and CTL continuous P$_{eq}$, e.g.
 Eqn.~(\ref{eqn:cpl:eqPerCont}) versus Eqn.~(\ref{eqn:ctl:eqPer}).
@@
 
- A third test, the detection of tidally-locked binaries \xxx{with solar-mass primaries and P$_{rot} \gsim 60$ d, would provide strong evidence in favor of the CPL model as the CTL model cannot tidally-lock stars beyond P$_{orb} \approx 60$ d, regardless of $\tau$, e.g.
+ A third test, the detection of tidally locked binaries with solar-mass primaries and P$_{rot} \gsim 60$ d, would provide strong evidence in favor of the CPL model as the CTL model cannot tidally lock stars beyond P$_{orb} \approx 60$ d, regardless of $\tau$, e.g.
 Fig.~\ref{fig:lockedCTL}.
-The CPL model, however, can tidally-lock binaries out to P$_{orb} \gsim 90$ d.
+The CPL model, however, can tidally lock binaries out to P$_{orb} \gsim 90$ d.
 We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100 $ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small star spot modulation amplitudes \citep{McQuillan2014,Lurie2017,Reinhold2018}.
-Long term spectroscopic monitoring may be warranted in such cases.}
+Long term spectroscopic monitoring may be warranted in such cases.
 
@@
 In this work, we probed the long-term angular momentum evolution of low-mass stellar binaries, with a focus on P$_{rot}$ in short and intermediate P$_{orb}$ binaries.
-We considered the impact of two common equilibrium tidal models, \xxx{magnetic braking}, and stellar evolution.
+We considered the impact of two common equilibrium tidal models, magnetic braking, and stellar evolution.
 We performed a large suite of simulations for binaries with physically-motivated initial conditions out to P$_{orb} = 100$ and across a wide range of tidal dissipation parameters to examine the competition between tidal torques and magnetic braking for controlling the stellar P$_{rot}$ evolution.
@@
 In our simulations, nearly all binaries with $P_{orb} \lsim 4$ d have tidally-synchronized spins and circularized orbits, in good agreement with observations of \kepler EBs and binaries in the field.
-We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking, or supersynchronously if they tidally-lock on eccentric orbits.
+We showed for P$_{orb} \gsim 4$ d, primary stars in stellar binaries can rotate subsynchronously for Gyrs due to the competition between tidal torques and magnetic braking, or supersynchronously if they tidally lock on eccentric orbits.
 Our predictions are not strongly dependant on the choice of magnetic braking model, but rather are generic outcomes of the interaction between magnetic braking and tidal torques.
 Both the CPL and CTL equilibrium tidal models predict that binaries tidally-interact at longer P$_{orb}$ than have previously been considered, out to P$_{orb} \approx 60-100$ d.
-Many binaries with P$_{orb} \lsim 20$ d tidally-lock according to both models, in good agreement with previous results, but the CPL model predicts that binaries can readily tidally-lock out to P$_{orb} \approx 100$ d.
+Many binaries with P$_{orb} \lsim 20$ d tidally lock according to both models, in good agreement with previous results, but the CPL model predicts that binaries can readily tidally lock out to P$_{orb} \approx 100$ d.
 Tidal interactions can cause P$_{rot}$ evolution in stellar binaries to differ from the long-term spin down due to magnetic braking experienced by single stars, decoupling P$_{rot}$ from age.
-In tidally-interacting binaries, gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails\xxx{, potentially underestimating stellar ages by up to $300\%$}.
-We caution that any application of \xxx{gyrochronology} methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
+In tidally interacting binaries, gyrochronology, the technique of linking stellar P$_{rot}$ to age, likely fails, potentially underestimating stellar ages by up to $300\%$.
+We caution that any application of gyrochronology methods to stars, especially those with P$_{rot} \lsim 20$ d, should account for the possibility of stellar binarity to prevent deriving incorrect ages.
 
@@
 
-\xxx{We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
+We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries.
 We primarily suggest that observers measure stellar P$_{rot}$ in binaries with solar-mass primaries for P$_{orb}$ between $60-100$ d.
-If any tidally-locked binaries are identified at long P$_{orb}$, this would be evidence in favor of the CPL model as we found that only binaries tidally-interacting via the CPL model could tidally-lock at such long P$_{orb}$.
-At shorter P$_{orb}$, precise measurements of P$_{rot}$ and binary $e$ and P$_{orb}$ could distinguish between the CPL and CTL model in tidally-locked systems, e.g.
+If any tidally locked binaries are identified at long P$_{orb}$, this would be evidence in favor of the CPL model as we found that only binaries tidally interacting via the CPL model could tidally lock at such long P$_{orb}$.
+At shorter P$_{orb}$, precise measurements of P$_{rot}$ and binary $e$ and P$_{orb}$ could distinguish between the CPL and CTL model in tidally locked systems, e.g.
 identifying if P$_{eq}$ follows Eq.~(\ref{eqn:cpl:eqPer}) vs.
@@
 Beyond these tests, our model could be used to infer the tidal properties of binary stars, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties and reasonable prior probability distributions for parameters like the initial binary $e$.
-This analysis, however, is beyond the scope of this work and we leave it for future endeavors.}
+This analysis, however, is beyond the scope of this work and we leave it for future endeavors.
 
@@
 \acknowledgments
-\xxx{We thank the anonymous reviewer for helpful comments that improved the quality of this manuscript.} This work was facilitated though the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system and funded by the Student Technology Fund at the University of Washington.
+We thank the anonymous reviewer for helpful comments that improved the quality of this manuscript.
+This work was facilitated though the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system and funded by the Student Technology Fund at the University of Washington.
 DPF was supported by NASA Headquarters under the NASA Earth and Space Science Fellowship Program - Grant 80NSSC17K0482.
@@
 \end{equation}
-where P$_{rot} = 2 \pi / \omega$ and \xxx{$C_{MB} = 6.3 \times 10^{30}$ ergs \citep{Matt2015,Matt2019}}.
+where P$_{rot} = 2 \pi / \omega$ and $C_{MB} = 6.3 \times 10^{30}$ ergs \citep{Matt2015,Matt2019}.
 

### 2019-07-01 Paper/arxiv/sync.tex
added arxiv version of the paper

--- 2019-07-01 Paper/sync.tex
@@
 \begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
+	\includegraphics[width=\textwidth]{stellarExample.pdf}
    \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \citet{Matt2015} magnetic braking model.
@@
 We plot the tidal evolution for $a$, $e$, and $P_{rot}$, ignoring stellar evolution, for a solar-twin binary with an initial P$_{orb} = 10$ d, P$_{rot} = 1$ d, $e = 0.2$ for the CPL model and CTL model, assuming $Q=10^6$ and $\tau = 0.1$ seconds, respectively, in Fig.~\ref{fig:tidalExample}.
-The CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally locked state, transferring rotational angular momentum into the orbit\xxx{, increasing the orbital angular momentum by ${\sim}1\%$} in the process.
+The CPL and CTL model predict the same qualitative evolution: both the binary's $e$ and P$_{orb}$ slightly increase as tides force the spins toward the tidally locked state, transferring rotational angular momentum into the orbit, increasing the orbital angular momentum by ${\sim}1\%$ in the process.
 At late times, both the CPL and CTL drive the binaries towards orbital circularization, with tidal dissipation decreasing P$_{orb}$.
@@
 \begin{figure}
-	\includegraphics[width=0.4\textwidth]{../Plots/tidalExample.pdf}
+	\includegraphics[width=0.4\textwidth]{tidalExample.pdf}
    \caption{Tidal evolution of a 1 M$_{\odot} -$ 1 M$_{\odot}$ stellar binary's $e$ (top), P$_{orb}$ (middle), and P$_{rot}$ (bottom) for the CPL (blue) and CTL (orange) model.
@@
 \begin{figure}[h]
-	\includegraphics[width=0.5\textwidth]{../Plots/analyticTorque.pdf}
+	\includegraphics[width=0.5\textwidth]{analyticTorque.pdf}
    \caption{The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}.
@@
 In general as tides weaken, i.e.
-increasing P$_{orb}$ and/or \xxx{decreasing $\log_{10}(k_2 \tau)$}, tidal and magnetic braking torques balance at longer P$_{rot}$.
-\xxx{For strong tides, $\log_{10}(k_2 \tau) \gsim -1$}, tidal torques overpower magnetic braking for P$_{orb} \lsim 40$ d, tidally locking binaries into synchronous rotation.
+increasing P$_{orb}$ and/or decreasing $\log_{10}(k_2 \tau)$, tidal and magnetic braking torques balance at longer P$_{rot}$.
+For strong tides, $\log_{10}(k_2 \tau) \gsim -1$, tidal torques overpower magnetic braking for P$_{orb} \lsim 40$ d, tidally locking binaries into synchronous rotation.
 For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries will rotate subsynchronously for P$_{orb} \gsim 20$ d, with more severe subsynchronism at longer P$_{orb}$.
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
+	\includegraphics[width=\textwidth]{eqPerTwoPanelMatt.pdf}
    \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model.
@@
 We truncate each curve when the binary tidally locks.
-\xxx{The legend denotes the initial binary orbital period and we note that the orbital periods do not vary by more than a few percent over the course of the simulations.} For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
-For all P$_{orb}$, the CPL models tidally lock whereas the CTL model predicts subsynchronous rotation that persists for Gyrs.}%
+The legend denotes the initial binary orbital period and we note that the orbital periods do not vary by more than a few percent over the course of the simulations.} For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.
+For all P$_{orb}$, the CPL models tidally lock whereas the CTL model predicts subsynchronous rotation that persists for Gyrs.%
     \label{fig:eqPer}%
@@
 Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques.
-The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) \xxx{with P$_{orb} > 60$ d} does not in general correspond to binaries tidally locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally locked state.
-\xxx{At a given age, longer P$_{orb}$ binaries will tend to rotate faster as they experience weaker tidal torques, and hence require longer to spin down towards the tidally locked state.}
+The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) with P$_{orb} > 60$ d does not in general correspond to binaries tidally locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally locked state.
+At a given age, longer P$_{orb}$ binaries will tend to rotate faster as they experience weaker tidal torques, and hence require longer to spin down towards the tidally locked state.
 
@@
 \begin{figure*}[ht]
-	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
+	\includegraphics[width=\textwidth]{qTauPorbRatioHist.pdf}
    \caption{Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{../Plots/porbProtPorbQHist.pdf}
+	\includegraphics[width=\textwidth]{porbProtPorbQHist.pdf}
    \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model. }%
@@
 
-We isolate the impact of $Q$ and $\tau$ \xxx{on the spin-orbital state of tidally interacting stellar binaries} by binning our \xxx{CPL and CTL} simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$ \xxx{in Figures \ref{fig:qmap} and \ref{fig:taumap}, respectively.
+We isolate the impact of $Q$ and $\tau$ on the spin-orbital state of tidally interacting stellar binaries by binning our CPL and CTL simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$ in Figures \ref{fig:qmap} and \ref{fig:taumap}, respectively.
 In these figures, we estimate the typical strength of tidal torques, using $Q$ and $\tau$ as a proxy, that can produce various spin-orbital states.
-For the CPL simulations depicted in Fig.~\ref{fig:qmap},} synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally lock these binaries.
+For the CPL simulations depicted in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally lock these binaries.
 For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$, nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 60$ d, as in the CPL model, binaries with eccentric orbits can only tidally lock into a 1:1 or 3:2 spin-orbit commensurability, see Eqn.~(\ref{eqn:cpl:eqPer}).
@@
 Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking.
-\xxx{Subsynchronous rotation can occur under the CPL model for binaries with P$_{orb} < 50$ d.
-In this regime, the median $Q$ tends to increase with decreasing P$_{orb}/$P$_{rot}$, except near the tidally locked state, as magnetic braking dominates weaker tidal torques, yielding longer P$_{rot}$.}  This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally locked state by the end of the simulation.
+Subsynchronous rotation can occur under the CPL model for binaries with P$_{orb} < 50$ d.
+In this regime, the median $Q$ tends to increase with decreasing P$_{orb}/$P$_{rot}$, except near the tidally locked state, as magnetic braking dominates weaker tidal torques, yielding longer P$_{rot}$.
+This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally locked state by the end of the simulation.
 In this case, the more supersynchronous the rotation, the weaker the tidal torques must be, and hence the larger the $Q$ must be.
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{../Plots/porbProtPorbTauHist.pdf}
+	\includegraphics[width=\textwidth]{porbProtPorbTauHist.pdf}
    \caption{Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
@@
 
-According to the CTL model \xxx{simulations, depicted in Fig.~\ref{fig:taumap}}, many binaries tidally lock for P$_{orb} \lsim 20$ d when $\tau \gsim 0.1$ s, with some tidally locking up to P$_{orb} \approx 50$ d when $\tau \gsim 1$ s.
+According to the CTL model simulations, depicted in Fig.~\ref{fig:taumap}, many binaries tidally lock for P$_{orb} \lsim 20$ d when $\tau \gsim 0.1$ s, with some tidally locking up to P$_{orb} \approx 50$ d when $\tau \gsim 1$ s.
 Subsynchronous rotation typically occurs for stars with $\tau < 0.1$ s.
-\xxx{Similar to the behavior depicted in Fig.~\ref{fig:qmap}, longer P$_{rot}$ are produced by binaries with weaker tidal interactions since P$_{orb}/$P$_{rot}$ decreases monotonically with $\tau$ for P$_{orb} < 40$ d.}  For P$_{orb} > 50$ d, magnetic braking dominates the evolution \xxx{seen in} the diagonal sequence with a median $\tau \approx 0.1$ s, a value that is typically insufficient for tides to strongly influence the evolution given the wide orbital separations.
+Similar to the behavior depicted in Fig.~\ref{fig:qmap}, longer P$_{rot}$ are produced by binaries with weaker tidal interactions since P$_{orb}/$P$_{rot}$ decreases monotonically with $\tau$ for P$_{orb} < 40$ d.
+For P$_{orb} > 50$ d, magnetic braking dominates the evolution seen in the diagonal sequence with a median $\tau \approx 0.1$ s, a value that is typically insufficient for tides to strongly influence the evolution given the wide orbital separations.
 The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution.
-\xxx{In this P$_{orb}$ regime, most binaries rotate supersynchronously as tides and magnetic braking fail to sufficiently spin down the stars by the age of the system.
+In this P$_{orb}$ regime, most binaries rotate supersynchronously as tides and magnetic braking fail to sufficiently spin down the stars by the age of the system.
 At longer P$_{orb}$, some binaries can strongly tidally-interact, but these systems require extreme tidal $\tau \gsim 10$ s.
 Unlike the CPL simulations depicted in Fig.~\ref{fig:qmap}, these binaries with P$_{orb} \approx 90$ d are not tidally locked as the tides are still not strong enough to lock the system.
-We explore this point further in $\S$~\ref{sec:protDist}.} We do not often observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).
+We explore this point further in $\S$~\ref{sec:protDist}.
+We do not often observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).
 
@@
 \begin{figure}
-	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
+	\includegraphics[width=0.45\textwidth]{eqPerShortPorbMatt.pdf}
    \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer})), for initial circular binary orbits with initial P$_{orb} = 7.5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model.
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{../Plots/lockedCPL.pdf}
+	\includegraphics[width=0.5\textwidth]{lockedCPL.pdf}
    \caption{Rotation state for tidally locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries.
@@
 \begin{figure}
-	\includegraphics[width=0.48\textwidth]{../Plots/lockedCTL.pdf}
+	\includegraphics[width=0.48\textwidth]{lockedCTL.pdf}
    \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations.}%
@@
 At longer P$_{orb}$, weaker tidal torques allow magnetic braking to spin down the stars past P$_{eq}$, with tidal torques eventually strengthening enough to shepherd P$_{rot}$ towards P$_{eq}$ via the mechanism discussed in $\S$~\ref{sec:eq}.
-The CPL and CTL models predict that \xxx{$31\%$ and $24\%$} of stars, respectively, are either tidally locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
+The CPL and CTL models predict that $31\%$ and $24\%$ of stars, respectively, are either tidally locked or interacting, demonstrating that tidal torques play a pivotal role in shaping the angular momentum evolution in stellar binaries across a wide range of parameters.
 The P$_{rot}$ - mass distribution for not locked binaries resembles the single star sequence as magnetic braking and stellar evolution dictate their angular momentum evolution.
@@
 We highlight this behavior with a histogram of locked and interacting binaries over P$_{rot}$ for both tidal models in Fig.~\ref{fig:lockedProtHist}.
-\xxx{The CTL model predicts fewer tidally locked binaries at longer P$_{rot}$, concentrating most of the locked CTL distribution's density at short P$_{rot}$, whereas the CPL distribution has a heavy tail extending towards longer P$_{rot}$.
-The CPL model, however, predicts larger tidal-locking rates than the CTL model as seen in the enhanced numbers of tidally locked binaries at low P$_{rot}$ Fig.~\ref{fig:lockedCPL} compared with Fig.~\ref{fig:lockedCTL}.} The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
+The CTL model predicts fewer tidally locked binaries at longer P$_{rot}$, concentrating most of the locked CTL distribution's density at short P$_{rot}$, whereas the CPL distribution has a heavy tail extending towards longer P$_{rot}$.
+The CPL model, however, predicts larger tidal-locking rates than the CTL model as seen in the enhanced numbers of tidally locked binaries at low P$_{rot}$ Fig.~\ref{fig:lockedCPL} compared with Fig.~\ref{fig:lockedCTL}.
+The presence of P$_{orb} > 20$ d locked population, or lack there of, could be a powerful observational discriminant between which equilibrium tidal model acts in low-mass stellar binaries.
 We discuss this point further in $\S$~\ref{sec:whichModel}.
@@
 \begin{figure}
-	\includegraphics[width=0.48\textwidth]{../Plots/lockedProtHist.pdf}
+	\includegraphics[width=0.48\textwidth]{lockedProtHist.pdf}
    \caption{P$_{rot}$ distribution for tidally locked (blue) and interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked) binaries according to the CPL (solid line) and CTL (dashed line) models.}%
@@
 \begin{figure*}[t]
-	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
+	\includegraphics[width=\textwidth]{protDist.pdf}
    \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model.
@@
 \begin{figure*}[t]
-	\includegraphics[width=\textwidth]{../Plots/gyro.pdf}
+	\includegraphics[width=\textwidth]{gyro.pdf}
    \caption{Relative age error between single and binary stars for both the CPL (left) and CTL (right) tidal models.
@@
 \begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
+	\includegraphics[width=0.45\textwidth]{protAgeHist.pdf}
    \caption{Histogram of rapidly-rotating (P$_{rot} < 20$ d) star ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}.
@@
 \begin{figure*}[t]
-	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
+	\includegraphics[width=\textwidth]{lurieFig7.pdf}
    \caption{P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ according to the CPL model (left) and the CTL model (middle), and \citet{Lurie2017} \kepler EB observations (right).
